### 2016-04-12 oblique.tex
first commit

%\documentclass[usenatbib,iop,apj,numberedappendix]{emulateapj}
\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2010}

\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsbsy}

\usepackage[normalem]{ulem}
\usepackage[usenames]{color}

\definecolor{DarkGreen}{rgb}{0.64,0.80,0.35}
\newcommand\ac[1]{{\color{green} \bf #1}} %% AEB color
\newcommand\asout[1]{\color{green}\sout{{\rm  #1}}} %% AEB strikeout of text
\newcommand\qc[1]{{\color{red} \bf #1}}
%\newcommand\qc[1]{{\color{black} #1}}
\newcommand\cc[1]{{\color{blue} \bf #1}}

\newcommand{\mathsc}[1]{\text{\textsc{#1}}}

\def\eV{{\rm eV}} %.........electron volts
\def\MeV{{\rm M}\eV} %......gigaelectron volts
\def\GeV{{\rm G}\eV} %......gigaelectron volts
\def\TeV{{\rm T}\eV} %......teraelectron volts
\def\PeV{{\rm P}\eV} %......petaelectron volts
\def\XeV{{\rm X}\eV} %......exaelectron volts
\def\ZeV{{\rm Z}\eV} %......zetaelectron volts
\def\erg{{\rm erg}} %.......ergs
\def\J{{\rm J}} %...........joules
\def\K{{\rm K}} %...........Kelvin

\def\cm{{\rm cm}}
\def\s{{\rm s}}

%%% Special Symbols
\def\Ms{M_\odot}
\renewcommand{\d}{{\rm d}}
\newcommand{\rmn}{\mathrm}
\newcommand{\e}{{\rm e}}
\newcommand\bmath[1] {\mbox{\boldmath$\rm #1$}}
\def\x{\bmath{x}}
\def\p{\bmath{p}}
\def\E{\bmath{E}}
\def\B{\bmath{B}}
\def\v{\bmath{v}}
\def\j{\bmath{j}}
\def\k{\bmath{k}}
\def\GTS{\Gamma_{\rm TS}}
\def\GPA{\Gamma_{\rm CA}}
\def\GIC{\Gamma_{\rm IC}}
\def\GW{\Gamma_{\rm W}}
\def\GO{\Gamma_{\rm M}}
\def\GM{\Gamma_{\rm M,k}}
\def\Tc{T_{\rm crit}}
\def\Dpp{D_{\rm pp}}
\def\O{\mathcal{O}}
\def\nb{n_{b}}
\def\nIGM{n_{\rm IGM}}
\def\Qcorr{\dot{Q}_{\rm corr}}
\def\Nb{N_{\rm b}}
\def\NbIC{N_{\rm b,IC}}
\def\Eiso{E_{\rm iso}}
\def\NHI{N_{\rm HI}}
\def\FUV{F_{\rm UV}}
\def\LIR{L_{\rm IR}}
\def\SFR{{\rm SFR}}
\def\bk{\bmath{k}}
\def\bp{\bmath{p}}
\def\bv{\bmath{v}}
\def\dtp{{d^3\!p}}
\def\dtv{{d^3\!v}}

%\def\kvec{\bmath{k}}
%\def\vel{\bmath{v}}

\def\Erfi{{\rm Erfi}}
\def\Erf{{\rm Erf}}

\def\gph{\gamma_{\rm ph}}
\def\vph{v_{\rm ph}}
\def\eph{\epsilon_{\rm ph}}
\def\pph{p_{\rm ph}}
\def\cG{\mathcal{G}}

\def\vpl{v_\parallel}
\def\vpp{v_\perp}
\def\spl{\sigma_\parallel}
\def\spp{\sigma_\perp}

%\newcommand{\vel}{\ensuremath{\boldsymbol v}}
%\newcommand{\pmom}{\ensuremath{\boldsymbol p}}
%\newcommand{\Efield}{\ensuremath{\boldsymbol E}}
%\newcommand{\Bfield}{\ensuremath{\boldsymbol B}}
%\newcommand{\grad}{\ensuremath{\boldsymbol \nabla}}
%\newcommand{\gradp}{\ensuremath{\boldsymbol \nabla}_p}
%\newcommand{\ppos}{\ensuremath{\boldsymbol x}}

\newcommand{\vel}{\ensuremath{\bmath{v}}}
\newcommand{\pmom}{\ensuremath{\bmath{p}}}
\newcommand{\Efield}{\ensuremath{\bmath{E}}}
\newcommand{\Bfield}{\ensuremath{\bmath{B}}}
\newcommand{\grad}{\ensuremath{\bmath{\nabla}}}
\newcommand{\gradp}{\ensuremath{\bmath{\nabla}_p}}
\newcommand{\ppos}{\ensuremath{\bmath{x}}}

\newcommand{\fLAB}{\ensuremath{f_{\rm L}}}
\newcommand{\fCM}{\ensuremath{f_{\rm CM}}}

\newcommand{\dVLAB}{\ensuremath{d^3p_{\rm L} d^3x_{\rm L}}}
\newcommand{\dVCM}{\ensuremath{d^3p_{\rm CM} d^3x_{\rm CM}}}

%\newcommand{\betavec}{\ensuremath{\boldsymbol \beta}}
%\newcommand{\betaLAB}{\ensuremath{\boldsymbol \beta_{\rm L}}}
%\newcommand{\betaCM}{\ensuremath{\boldsymbol \beta_{\rm CM}}}
\newcommand{\betaCMpara}{\ensuremath{\beta_{\rm CM,\parallel}}}
\newcommand{\betaCMperp}{\ensuremath{\beta_{\rm CM,\perp}}}

\newcommand{\betavec}{\ensuremath{\bmath{\beta}}}
\newcommand{\betaLAB}{\ensuremath{\bmath{\beta_{\rm L}}}}
\newcommand{\betaCM}{\ensuremath{\bmath{\beta_{\rm CM}}}}

%\newcommand{\betabar}{\ensuremath{\overline{\boldsymbol \beta}}}
\newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
\newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm CM}}}
\newcommand{\betaBOOST}{\ensuremath{\beta_{\rm B}}}

\newcommand{\betabarLABpara}{\ensuremath{\overline{\beta}_{\rm L,\parallel}}}
\newcommand{\betabarLABperp}{\ensuremath{\overline{\beta}_{\rm L,\perp}}}

\newcommand{\dbetabarpara}{\ensuremath{\alpha_{\parallel}}}
\newcommand{\dbetabarperp}{\ensuremath{\alpha_{\perp}}}

\newcommand{\dbetabarLABpara}{\ensuremath{\overline{\Delta\beta^2}_{\rm L, \parallel}}}
\newcommand{\dbetabarLABperp}{\ensuremath{\overline{\Delta\beta^2}_{\rm L, \perp}}}
\newcommand{\dbetabarCMi}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, i}}}
\newcommand{\dbetabarCMpara}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, \parallel}}}
\newcommand{\dbetabarCMperp}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, \perp}}}

\newcommand{\betaMEAN}{\ensuremath{\beta_{\rm M}}}
\newcommand{\epm}{\ensuremath{e^+e^-}}

%\newcommand{\kvec}{\ensuremath{\boldsymbol k}}
%\newcommand{\rvec}{\ensuremath{\boldsymbol r}}
%\newcommand{\rvec}{\ensuremath{\boldsymbol j}}

\newcommand{\kvec}{\ensuremath{\bmath{k}}}
\newcommand{\rvec}{\ensuremath{\bmath{r}}}
\newcommand{\jvec}{\ensuremath{\bmath{j}}}

%\newcommand{\xhat}{\ensuremath{\hat{\boldsymbol x}}}
%\newcommand{\yhat}{\ensuremath{\hat{\boldsymbol y}}}
%\newcommand{\zhat}{\ensuremath{\hat{\boldsymbol z}}}

\newcommand{\xhat}{\ensuremath{\hat{\bmath{x}}}}
\newcommand{\yhat}{\ensuremath{\hat{\bmath{y}}}}
\newcommand{\zhat}{\ensuremath{\hat{\bmath{z}}}}

\newcommand{\xphat}{\ensuremath{\hat{\bmath{x}}'}}
\newcommand{\yphat}{\ensuremath{\hat{\bmath{y}}'}}
\newcommand{\zphat}{\ensuremath{\hat{\bmath{z}}'}}

\newcommand{\Mpc}{\ensuremath{\rm Mpc}}
\newcommand{\Gpc}{\ensuremath{\rm Gpc}}
\newcommand{\Lya}{\ensuremath{{\rm Ly}\alpha}}
\def\G{{\rm G}}
\newcommand{\kI}{\ensuremath{\mathcal{I}}}

\newcommand{\betabeam}{\ensuremath{\bmath{\beta}_{\rm b}}}
\newcommand{\betabeamz}{\ensuremath{\beta_{\rm b,z}}}
\newcommand{\betabeamx}{\ensuremath{\beta_{\rm b,x}}}
\newcommand{\gammabeam}{\ensuremath{\gamma_{\rm b}}}
\newcommand{\gammaphase}{\ensuremath{\gamma_{\rm ph}}}
\newcommand{\betaphase}{\ensuremath{\beta_{\rm ph}}}

\def\Fermi{{\em Fermi\ }}

\begin{document}

\title{The Plasma Physics of Dilute Ultrarelativistic \epm\ Pair Beams I:\\ Linear Theory}

%\author{Some people}
\author{
Avery E.~Broderick\altaffilmark{1,2},
Philip Chang\altaffilmark{3},
Christoph Pfrommer\altaffilmark{4},
Ewald Puchwein\altaffilmark{5},
}
\altaffiltext{1}{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, ON, N2L 2Y5, Canada}
\altaffiltext{2}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
\altaffiltext{3}{Department of Physics, University of Wisconsin-Milwaukee, 1900 E. Kenwood Boulevard, Milwaukee, WI 53211, USA; chang65@uwm.edu}
\altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}

\keywords{
BL Lacertae objects: general -- gamma rays: general -- plasmas -- instabilities -- magnetic fields
}

%\maketitle

\begin{abstract}
The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that cool them.  Starting from an isotropic distribution in the center of mass frame, we show that the velocity dispersions parallel and transverse to the direction of beam propagation become highly anisotropic under Lorentz boosts and scale differently with the Lorentz factor, $\gamma$.  Using these results to construct a simplified distribution function, we then calculate the linear growth rate of the beam plasma and ``oblique'' instability in both the reactive and kinetic regimes.  We reproduce the well-known reactive growth rates for both the beam plasma and ``oblique'' mode and show that they both have similar growth rates in the kinetic regime.  In contrast to the earlier claim by the authors that the oblique mode is dominant, we instead find that both the beam-plasma (parallel) and oblique modes are similarly strong.  In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the ``oblique'' mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.  Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we not only confirm that these beams are unstable to the kinetic ``oblique'' mode.  For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the ``reactive'' beam-plasma instability.  We show that overall growth rates for both the ``oblique'' and beam plasma growth rates are similar and in line with our earlier claims.  These results strengthen the case that powerful plasma instabilities  sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space, which has profound consequences on constraints of the intergalactic magnetic field, the thermal state of the intergalactic medium, and the formation of structure in the universe. 
\end{abstract}

\section{Introduction}\label{sec:introduction}

The \Fermi satellite and ground based imaging atmospheric Cerenkov telescopes such as
% Phil: I added "imaging atmospheric" to distinguish this class of
% telescopes from imaging water CTs, e.g., MILAGRO; I am not aware of
% any space based IACT apart from a planned telescope on the ISS, but
% who knows whether this will ever fly
H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
  Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the ultra-high
energy Universe is teeming with energetic very high-energy gamma-ray 
(VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
blazars with a minority population of other sources  such as radio galaxies
and starburst galaxies.  These VHEGR observations have proven to be useful in constraining mechanisms 
of particle acceleration
\citep[see, e.g., ][]{Pagl_etal:96,Domi-Torr:05,Thom-Quat-Waxm:07,Pers-Reph-Arie:08,deCe-Torr-Rodr:09,Reph-Arie-Pers:10,Lack_etal:10} and
black hole jets
\citep[see, e.g., ][]{Jones+1974,Ghisellini+1989,Ghisellini+2008,Tavecchio+2008,Ghisellini+2009}.
% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
% and intergalactic magnetic fields (IGMF)
% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.

It has recently been argued by a number authors
that the {\it non-observation} of $\sim 1-100\,\GeV$ flux by the \Fermi satellite toward these TeV blazars provides lower bounds on the strength of the  intergalactic magnetic field (IGMF) of $B\gtrsim 10^{-19}-10^{-15}\,\G$ \citep{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
To understand how these claimed limits arise, we begin by revisiting the physics of VHEGR propagation in the universe.

As these VHEGR photon propagate through the universe, they pair produce off the extragalactic background light (EBL) and produce ultrarelativistic electron-positron beams  with typical Lorentz factors of $10^5$--$10^7$.  
Within the current epoch, the mean free path for these photons is $\Dpp\sim 800 \,E_{\TeV}^{-1}\,\Mpc$, where $E_{\TeV} = E/1\,\TeV$ is the energy of the photon in \TeV.  At high redshifts, $z\geq 1$, this drops to $D_{\rmn{pp}}\sim 35\, E_{\TeV}^{-1}\,\Mpc$ due to the larger number density of EBL photons.  Because these typical mean free paths are much shorter than the Hubble length, the universe is optically thick to VHEGRs. This effect is seen in the relative nearness of TeV blazars ($z\sim 0.1$) and the observed softening of their VHE spectra with increasing distance.

Previously, it has been  assumed that the resultant electron-positron pairs lose energy through inverse-Compton scattering the cosmic microwave background (CMB) and EBL. This yields an inverse-Compton cascade (ICC) that reprocesses TeV emission to gamma-rays with energies $\lesssim 100 \GeV$, for which the mean free path is comparable to the size of the universe.    
Hence, associated with every VHEGR source should be a $\sim 100\,\GeV$ source.  It is the {\it non-observation} of this $100\,\GeV$ flux, presumably due to the magnetic deflection of the pairs, that is used to constraint the strength of a putative IGMF.

In addition, \Fermi has also provided the most precise estimate of the unresolved
extragalactic gamma-ray background (EGRB) for energies between
$200\,\MeV$ and $100\,\GeV$.  Since ICCs reprocess the VHEGR emission of
distant sources into this band, this has been used to constrain the
evolution of the luminosity density of VHEGR sources
\citep[see, e.g., ][]{Naru-Tota:06,Knei-Mann:08,Inou-Tota:09,Vent:10}.
These constraints preclude any dramatic rise in numbers of source by $z\sim1$--$2$ that seen in the quasar
distribution.  That is, the comoving number of blazars must have
remained essentially fixed, at odds with both the physical picture
underlying these systems and with the observed evolution of similarly
accreting systems, i.e., quasars.

Ultimately, these conclusions rest upon the assumption that ICCs dominate the evolution of the ultra-relativistic pairs.  
However, we have strongly challenged this assumption in a recent series of papers, \citet[][hereafter paper I, II, and III respectively]{BCP,CBP, PCB}.  In particular, we find that ultrarelativistic pair beam that results from TeV photon annihilation on the EBL is subject to virulent plasma beam instabilities that sap the energy of the pair beam and convert it to heat in the intergalactic medium (IGM) before ICCs have a chance to act.

In Paper I, we showed that despite the extraordinarily dilute nature of this ultra-relativistic pair plasma, a variety of plasma instabilities grow on timescales short in comparison to the inverse-Compton cooling time. Of the three instabilities that we studied --- the beam-plasma or two-stream instability\footnote{In order to follow the plasma literature more closely, henceforth we will refer to what we have called the ``two-stream'' instability as the ``beam-plasma'' instability}, the Weibel instability, and the ``oblique'' instability  ---  the most
important of which is the ``oblique'' instability \citep{Bret-Firp-Deut:04}
Because the pairs lose their energy primarily via depositing it as heat into the IGM, the ICCs critical to the above arguments commonly employed to provide constraints upon the IGMF and high-redshift blazar population are necessarily precluded.  In the latter case in particular, the \Fermi limits on the EGRB and high-energy blazar statistics allow a large and evolving blazar population more similar to that found for other AGN.

Assuming that these plasma-instability mechanisms drive ultra-relativistic pairs to deposit their energy as heat in
the IGM, we explored the impact of this heating on the
thermodynamics of the IGM in Paper II and the effect on this additional heating on structure formation in the universe in Paper III.  
By correcting for the selection effects of the current pencil beam VHEGR observations using
the all-sky monitoring of the \Fermi satellite, we argued that 
luminosity density in VHEGRs is of order $0.2\%$ of the quasar
luminosity density in Paper II.  Because of the greater conversion efficiency of TeV photon energy to heat (as compared to photoheating), we showed that this heating dominates the
photoheating rate at low $z$. This steady volumetric homogeneous heating naturally produces an inverted temperature-density relation and, with only a minor rescaling of the empirically determined TeV blazar
heating normalization, reproduces the inferred
inverted temperature-density relation at $z=2-3$ \citep{Bolton+09,Viel+09}, a feature that standard reionization models have proven unable to reproduce \citep{McQuinn+09,Bolton+09}.

Following the work of Paper II, we studied the effects of blazar heating in a
more detailed hydrodynamic model of structure formation 
\citep{Puchwein+2011}. Using the same blazar heating prescription as in Paper
II, we showed that the optical depth of the \Lya\ forest is reproduced using a
H{\sc i} photoionization rate of $\Gamma_{\rm H\mathsc{i}} \approx 5\times
10^{-13}\,{\rm s}^{-1}$ or equivalently using the inferred ionizing background
from \citet{FG+09}.  We also demonstrated that the low-density IGM again
possesses an inverted temperature-density relation, confirming the results of
Paper II in greater detail.  In addition, we showed that a blazar heated
universe not only matches the one- and two-point statistics of the
  high redshift Ly$\alpha$ forest but also the linewidth and column depth
  distribution. This excellent quantitative agreement is achieved
without tuning the photoionizing background and is due specifically to the
excess energy injection into the low-density IGM from TeV blazars.

Finally in Paper III, we explored the result of this additional IGM heating upon
the formation of structure in the Universe.  In particular, we showed that
the injection of entropy into the IGM by TeV blazars contributes to
developing a redshift dependent entropy floor for galaxy clusters and groups at
$z\lesssim 2$ and suppresses the formation of dwarfs.  The redshift dependent nature of TeV blazar heating suggests a large injection of entropy around $z\sim 1$, boosting the entropy of late forming groups and clusters.  This predicted enhanced entropy
of young groups is consistent with recent observations 
that show optically bright, and therefore young, groups and
clusters are X-ray dim ---  that is having a lower gas density due to a
raised entropy floor.  We also found that TeV
blazar heating suppresses the formation of late forming dwarfs both in
galactic halos, which is potentially important for the missing satellite problem \citep{Krav:10}, and
in voids, which is likewise potentially important for the void phenomenon \citep{Peebles2001}, by raising the
temperature of the IGM such that gas cannot collapse to form galaxies.

These potential implications of blazar heating rely on the assumption that the nonlinear state of these plasma instabilities efficiently convert the kinetic energy of the beam and convert it into heat.  In our previous work, we appealed to nonlinear simulations conducted in a radically different regime to argue for such a conversion.  However, a full nonlinear study is called for.  We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.  We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.  Using this understanding, we then develop a simple description of the distribution function of the beam, which we then use to calculate the unstable modes analytically in both the reactive (hydrodynamic) and kinetic regimes. Here the reactive instability refers to the instability where the entire beam participates in the instability.  In particular, all the beam particles are resonant with the unstable wave on a timescale longer than the growth time of the instability.  The reactive instability is also referred to as the hydrodynamic instability since the instability can be derived from the fluid equations instead of kinetic theory.  On the other hand, in the kinetic regime, only a fraction of the beam particles are resonant with the beam over a the growth time of the instability, which reduces the growth rate compared to the reactive instability for the same beam density and beam Lorentz factor.  We recover the well-known results for the reactive regime for both the beam-plasma and ``oblique'' modes.  We also derive the growth rate for these two instabilities in the kinetic regime and delineate the range of applicability for both the reactive and kinetic cases and apply these results for ultrarelativistic \epm\ pair beams.

This paper is organized as follows.  In Section \ref{sec:setup}, we describe the transformation properties of an ultrarelativistic \epm\ beam in terms of its distribution function.  We then calculate the various linear instabilities that this beam is subject to in section \ref{sec:linear}.  In particular, we pay careful attention to both the reactive (or hydrodynamic) and kinetic regimes of the beam plasma and ``oblique'' instabilities and the transition between the two.  Applying these result to TeV electron-positron pair beams that arise from TeV photon pair production in Section \ref{sec:application},
we demonstrate the despite the extraordinary coldness of the beam that we are always in the kinetic regime for the ``oblique'' mode, but may be in the reactive regime for the beam plasma mode.  However, because the controlling parameter is close to unity, the growth rates calculated in either regime are similar.  We close with a discussion the implications of this work and application of these results for nonlinear theory in Section \ref{sec:conclusions}.

\ac{[There are some notational issues throughout the text, which I'll list here.  Usually it is simply a matter of choosing, and then sticking to, a convention.
\begin{itemize}
\item Some equations assume $c=1$ while others have the $c$'s explicitly present.  Unless there is a compelling reason (e.g., clarity of derivation vs. utility of an answer in physical units) we should choose one and mention the convention here (at the end of the intro) if $c=1$.
\item $\beta$ and $v$ appear to be used interchangeably, again related to the $c$ or not question.  Again, we should choose one.
\item Throughout section 2 and in the Appendix the frame is defined in terms of $\parallel$ and $\perp$ directions (i.e., $\beta_\parallel$ and $\bmath{\beta}_\perp$, note the latter is a {\em vector}!) while in sections 3 and in the figures we refer to $\beta_z$ and $\beta_x$ for these.
\item ApJ style, which I'm assuming is what we are following, has vectors given by bold, roman fonts and scalars by the italic fonts.  Thus, it should be $\bmath{k}\cdot\bmath{\delta E}$ or $\bmath{k}\cdot\delta\bmath{E}$, etc.  I've tried to fix up various definitions in the preamble to do this, but that should be checked.
\end{itemize}
In any equations I've added/modified I will include the $c$'s and use $\beta$ for concreteness, though I anticipate there will be few such examples.]}

\section{Ultrarelativistic Pair Beams from VHEGRs}\label{sec:setup}

As stated in the Introduction, VHEGR photons pair produce on encountering EBL photons as they propagate throughout the universe \citep{Gould+66}, and this attenuation of VHEGR flux has been use as a probe of the EBL \citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}.
The basic requirement of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL photon
  ($E_{\rm ebl}$) exceed the rest mass energy of the $e^\pm$ pair in the center
  of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4 m_e^2 c^4$, where
  $\theta$ is the relative angle of propagation in the lab frame, an $e^\pm$
  pair can be produced with Lorentz factor $\gamma\simeq E/2m_e c^2$
  \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the pair beam that emerges from this process.

%As the mean free path of these VHEGR photons to pair production is short compared to the Hubble length, observations of extragalactic VHEGR sources have been used as probes of the EBL \citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}

\subsection{Distribution Function of the Pair Beam}

In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.  We model this distribution as a relativistic thermal distribution:
\begin{equation}\label{eq:relativistic maxwellian}
f\propto \exp\left(-\frac E T\right),
\end{equation}
 where $E$ and $T$ are the dimensionless energy and temperature in terms of a particles rest mass.  In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution ({\bf REFS}).

The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.  The relationship between the energies of the lab (boosted) frame and the COM frame is (Jackson) 
\begin{equation}
  E_{\rm CM} = \gamma_B\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
\end{equation}
Plugging this into (\ref{eq:relativistic maxwellian}), we find (Wright \& Hadley 1975)
\begin{equation}\label{eq:drifting maxwellian}
f = \frac{1}{4\pi\gamma_B m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right),
\end{equation}
where we have identified the lab frame temperature $T_{\rm L} \equiv T_{\rm CM}/\gamma_{\rm B}$ and $K_2$ is the 2nd order modified Bessel function. The transformation properties of the drifting Maxwell-Juttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-Juttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).

%In Appendix~\ref{sec:lorentz}, we explicitly derive the scaling of the
%  parallel and perpendicular velocity dispersions with the Lorentz factor upon
%  boosting the distribution function to the lab frame.
%In the limit in which the perpendicular velocity dispersion in the COM frame is non-relativistic, i.e.,
%$\left|\betaCM\right| \ll 1$, we find to lowest order in $\betaCM$
%\begin{equation}
% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2}.
% \label{eq:betaperp}
%\end{equation}
%This simple scaling of the perpendicular velocity dispersion is an immediate 
%consequence of how transverse velocities transform under Lorentz boosts, and may be understood
%in terms of time dilation.
%
%This result is also anticipated by the standard transformation of temperature under boosts, $T\rightarrow T/\gamma$ (see Equation~\ref{eq:drifting maxwellian}).
%This can be understood by considering $mv^2 \sim kT$ -- we pick up two factors of $1/\gamma$
%from the perpendicular velocity dispersion that is countered by one factor of
%$\gamma$ from the mass scaling. 
%
%For the parallel component of the velocity
%dispersion in the non-relativistic limit, $\left|\betaCM\right| \ll 1$, we find
%to lowest order in $\betaCM$:
%\begin{equation}
% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}},
% \label{eq:betapara}
%\end{equation}
%Again this scaling follows immediately from the transformation of velocities under Lorentz boosts, and may be simply understood as a consequence of both time dilation and, for the parallel case, length contraction.
%This scaling of the parallel component of the velocity
%dispersion has important consequences below.
%

% To see how this distribution function appears in velocity space, we show four examples of the  Maxwell-Juttner distribution in the center-of-mass frame for the nonrelativistic temperatures ($T=10^{-2}$ and $ 0.1$) in Figure \ref{fig:distribution_non_rel} and for relativistic temperatures ( $T=1$ and $10$) in Figure \ref{fig:distribution_rel}, where $T$ is listed in the natural relativistic units: $m_e c^2/k_B$.
% We plot these in terms of $\beta_z$, which is parallel to the beam and $\beta_x$, which is perpendicular to the beam direction though, in the center-of-mass frame, an isotropic distribution makes no distinction between these directions.  We note that not all combinations of $\beta_x$ and $\beta_z$ are allowed as $\beta_x^2 + \beta_z^2 < 1$.  \ac{[Could you blank out the portions for which $\beta^2>=1$ in the figure to both drive this point home as well has emphasize the scaling difference?]}
% 
% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_B = 10^6$.  Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_B^{-1}$ and the parallel component by $\gamma_B^{-2}$.
% When $T\ll1$, as anticipated above, the parallel and transverse velocity dispersions are well approximated by the scaling in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}).  As seen explicitly in Figure \ref{fig:distribution_non_rel}, the distribution functions remain roughly isotropic in the rescaled and shifted velocity units.
% 
% For relativistic temperatures, the conditions under which the expressions in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}) were derived no longer hold.  Nevertheless, despite the importance of higher-order and anisotropic terms in the distribution function, the magnitudes of the velocity dispersions remain well approximated by these scalings.  This may be seen explicitly in the right panels of Figure \ref{fig:distribution_rel}.
% While these scaling will necessarily break down as $T \rightarrow \gamma_B$, for the problem of interest this is never the case.  Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_{\rm B}^{-1}$ in the perpendicular direction and like $\gamma_{\rm B}^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_B \gg T$).  
% 
% \begin{figure*}
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_01.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_01.pdf}\\
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_1.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_1.pdf}\\
%  \caption{Distribution function in velocity space in the center-of-mass frame (left) and the boosted (lab) frame (right) for nonrelativistic temperatures: $T=0.01$ (top) and $0.1$ (bottom).  Note that while we plot $\beta_z$ and $\beta_x$ from -1 to 1, not all values of $\beta_z$ and $\beta_x$ are allowed. \ac{[Can this be shown explicitly, e.g., by making those values NaN, or some other method that results in no plot in that instance?  Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.  As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_B$ on the x-axis.  We have also multiplied the $\beta$ factors by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. These plots show that a non-relativistic velocity dispersion transforms as $\gamma_B^{-2}$ and $\gamma_B^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
% \end{figure*}
% \begin{figure*}
% \includegraphics[width=0.5\textwidth]{distribution_cm_E_1.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_1.pdf}\\
%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_10.pdf}
%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_10.pdf}
%  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).  Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}. Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.  Nevertheless, higher-order corrections are beginning to spread the distribution function over a velocity range larger than that anticipated by these scalings. 
% \ac{[Can you make the righthand panels with square aspect ratios?  Right now, I think that they extend a factor of two more in the parallel direction, which appears to give the impression that they are more isotropic than they really are!  Also, how are the normlizations obtained?  Specifically, are they unit normalized?]}
% \label{fig:distribution_rel}}
% \end{figure*}
% 
% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_B$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_{\rm B}(E-\beta_{\rm B}p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}: 
% \begin{equation}
%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_{\rm B}\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),
% \end{equation}
% where $\alpha_\perp\equiv T/\gamma_{\rm B}^2$ and $\alpha_\parallel \equiv T/\gamma_{\rm B}^4 = \alpha_\perp/\gamma_{\rm B}^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
% This form of the distribution function has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.

\section{Linear Theory}\label{sec:linear}

The Vlasov equation for each species is
\begin{equation}\label{eq:linearize vlasov}
  \frac {\partial f_s}{\partial t} + \vel_s \cdot \grad f_s + 
  q_s \left({\Efield} + \frac {\vel_s} {c} \times {\Bfield} \right)
 \cdot\grad_p f_s = 0,
\end{equation}
where $\vel_s=\pmom_s/\gamma_s m_e$ and  $\gamma_s = 1/\sqrt{1-v_s^2/c^2} $
Here, $s$ is the species label, $+$ for positrons and $-$ for electrons,
with $q_{\pm} = \pm e$.  \ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
Upon linearizing this in small perturbations about a background distribution, i.e., setting 
$f_s\rightarrow f_{0s} + \delta f_s$, $\Bfield \rightarrow \delta \Bfield$ and $\Efield \rightarrow \delta \Efield$,  we obtain, 
\begin{equation}\label{eq:vlasov}
  \frac {\partial\delta f_s}{\partial t} + {\vel_s} \cdot \grad\delta f_s  
  +  {q_s} \left(\delta \Efield + \frac {\vel_s} {c} \times \delta \Bfield \right)\cdot\grad_p f_{0s} = 0.
 \end{equation}
The plasma couples to the field through the Maxwell equations
\begin{eqnarray}\label{eq:maxwell}
\grad \times \delta \Efield &=& -\frac 1 c \frac {\partial \delta \Bfield} {\partial t},\\
\grad \times \delta \Bfield &=& \frac {4\pi} {c} \delta\jvec + \frac{1}{ c} \frac {\partial \delta \Efield} {\partial t},\label{eq:maxwell2}
\end{eqnarray}
where $\delta\jvec = \sum_s q_s\int \vel_s \delta f_s d^3 p$ is the linear current density perturbation.

Here it is useful to work within the electrostatic approximation ($\kvec \times \delta \Efield = 0$), where we only need to include Coulomb's law for the electric field rather than the full Maxwell equations:
\begin{equation}
 i\kvec\cdot\delta\Efield = 4\pi\delta\rho,
\end{equation}
where $\delta\rho = \sum_s q_s\int \delta f_s \,d^3\!p$ is the perturbed charge density. 
By adopting the electrostatic approximation, we have explicitly ignored electromagnetic modes.  This would preclude, for example, the Weibel instability.
In addition, the electromagnetic terms would introduce corrections to the physics that are not necessarily small in the limit of relativistic particles, i.e., $v/c \rightarrow 1$.  However, we make this approximation for two reasons.  First, a complete calculation of the unstable modes has already been carried out by \citet{Bret-Grem-Beni:10}, who showed that the electrostatic approximation is in excellent agreement with a full calculation (modulo the Weibel instability).  Second, the electrostatic approximation is much simpler than a full calculation and allows us to analytically calculate the unstable growth rates, while permitting a clear exposition of the relevant physics.

We now adopt perturbations of the form $\delta \sim \exp\left(i \kvec\cdot\rvec - i \omega t\right)$ and without loss of generality assume that $\kvec = (k_x, 0, k_z)$, where $k_z$ is along the beam direction.  Linearizing the Vlasov-Maxwell equations then leads to the dispersion relation:
%\begin{equation}\label{eq:dispersion relation}
% \epsilon = 1 - \frac{\omega_{p}^2}{k}\int \frac{\kvec \cdot \gradp F}{\omega - \kvec\cdot\vel} d^3\!p = 0,
%\end{equation}
\begin{equation}\label{eq:dispersion relation}
 \epsilon = 1 + \sum_s \frac{m_e \omega_{p,s}^2}{k^2}\int \frac{\kvec \cdot \gradp F_s}{\omega - \kvec\cdot\vel} d^3\!p = 0,
\end{equation}
where $\epsilon$ is the simplified dielectric function, and for each species $\omega_{p,s}^2\equiv 4\pi e^2 n_s/m_e$ is the plasma frequency, $n_s\equiv\int f_{0s} d^3\!p$ is the number density, and $F_s\equiv f_{0s} / n_s$ is the normalized distribution function.
Upon integrating by parts, Equation (\ref{eq:dispersion relation}) becomes
%\begin{eqnarray}
% 1 - \frac{4\pi q_s^2}{m_s k^2}\int F_s\kvec \cdot \gradp\frac{1}{\omega - \kvec\cdot\vel} d^3p &=& 0\\
% 1 - \frac{4\pi q_s^2}{m_s k^2}\int F_s \frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &=& 0. \label{eq:dispersion relation by parts}
%\end{eqnarray}
\begin{multline}
  \epsilon = 
  1 - \sum_s \frac{m_e\omega_{p,s}^2}{k^2} \int F_s\kvec \cdot \gradp\frac{1}{\omega - \kvec\cdot\vel} d^3p\\ 
  =
  1 - \sum_s \frac{\omega_{p,s}^2}{k^2}\int F_s \frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p 
  =
  0\,.
  \label{eq:dispersion relation by parts}
\end{multline}
\ac{[$\beta$ vs. $v$ issue above.]}

There are two distinct, often qualitatively different, regimes in which we may consider the implications of this dispersion relation.  The first is the cold plasma limit or the hydrodynamic or reactive limit.  The hydrodynamic limit is aptly named because the resulting dispersion relation that is found could have also been calculated directly from the continuity equation and the momentum equation.  In this limit, the internal distribution of the particles of the background or beam are irrelevant to the physics of the instability and it is only the bulk response that is important.  In particular, this means that the beam particles are resonance with the unstable wave over a timescale much longer than the growth them as stated above, i.e., the beam particles do not drift a distance larger than the wavelength of the unstable mode over the growth time of the instability.  The second is the kinetic regime, where the internal distribution of the beam particles are important to the physics of the instability. Here, only a fraction of the beam particles stay within one wavelength of the unstable mode over the growth time of the instability.   Here, the bulk of the plasma (background or beam) do not response to the disturbance, but rather only a fraction of the particles are relevant for driving (instability) or damping (Landau damping).  We discuss below the evaluation of the dispersion relation in these two regimes, which gives two regimes of instability, and the delineation between these two regimes.

\subsection{Hydrodynamic (Reactive) Instability}

Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
Taking the limit of Equation (\ref{eq:distribution function beam}) as $\alpha_{\parallel}^{-1}, \alpha_{\perp}^{-1} \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
\begin{equation}\label{eq:dispersion relation reactive}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
\end{equation}
For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{BCP}.

The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (eq.[\ref{eq:growth rate reactive appendix}]) is 
\begin{equation}\label{eq:growth rate reactive}
 \Gamma = \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\left(\frac{\gamma^2 Z_x^2 + 1}{Z_x^2 + 1}\right)^{1/3}\frac{\omega_{p,t}}{\gamma},
\end{equation}
where $Z_x = k_xv_b/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.

For $k_x = 0 \rightarrow Z_x = 0$, this reduces to the beam-plasma growth rate, which is
\begin{equation}\label{eq:growth rate bp}
 \Gamma = \Gamma_{\rm TS} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma},
\end{equation}
which we denote the beam-plasma or ``two-stream'' growth rate.
For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.  Indeed for $\gamma \gg 1$ and $Z_x \sim 1$, the growth rate approaches oblique growth rate:
\begin{equation}
 \Gamma = \Gamma_{\rm ob} \equiv \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\frac{\omega_{p,t}}{\gamma^{1/3}},
\end{equation}
which is much faster than the beam-plasma growth rate, $\Gamma_{\rm TS}$

We should caution in the derivation above that the resonance condition, which is $\omega_{p,t} - k_z v_b$, implies that $k_z \neq 0$.  For the case where $k_z \rightarrow 0$, the electrostatic approximation no longer holds and the full dispersion relation must be solved.\footnote{We thank Antoine Bret for helping to clarify this point.}  A solution to the full dispersion relation reveals additional modes, including the zero frequency ($k_z = 0$) filamentation or Weibel mode.

\subsection{Kinetic Instability}

The growth rate expressed in Equation (\ref{eq:growth rate reactive}) sits in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could be been derived from the fluid equations.  Here all the particles participate in the instability.  However, kinetic theory masks another regime of the instability, where only a fraction of the particle participate in the instability, i.e., the kinetic regime.  We now derive the growth rate of the instability in the kinetic regime.

We begin first with the distribution function for the target plasma:
\begin{equation}
 F_t = \left(\frac{1}{2\pi m_e k_B T}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T}\right),
\end{equation}
where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
For the beam plasma, we adopt the Maxwell-Juttner distribution 
\begin{equation}\label{eq:distribution function beam}
F_b = \frac{n_b}{4\pi\gamma_B m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
\end{equation}
Plugging these into the dispersion relation (\ref{eq:dispersion relation}), we find
%\begin{equation}\label{eq:dispersion kinetic}
% 1 - \frac{\omega_{p,t}^2}{k}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p - \frac{\omega_{p,t}^2}{k}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
%\end{equation}
\begin{equation}\label{eq:dispersion kinetic}
 1 - \frac{\omega_{p,t}^2}{k^2}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
\end{equation}
where we have integrated by parts only the second term, associated with target plasma.

We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
\left[
\left( \cG'^2 + 2\cG' + 2 \right) 
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right]
\frac{e^{-\cG'}}{\cG'^3}\,.
\label{eq:OGgen}
\end{equation}

In Figure \ref{fig:ObliqueGMZ}, we plot the maximum growth rate (maximized over $\vph$) as function of $v_{bx} = |k_{bx} v_b/k|$ and $T$.  From the figure, it is clear that the growth rate is maximized for $v_{bx}/v_b \gg 1/\gamma_b$.  Moreover the maximum growth rates vary little and are robust for broad range of $v_{bx}/v_b = k_{bx}/k$, i.e., large range in angles between beam and wavevector.  Hence, this suggests that we might estimate the maximal oblique growth rates by taking the limit as $v_{bx}\gg v_b/\gamma_b$ and maximizing over $\vph$.  This is further seen in Figure \ref{fig:OGgen}, we maximize over both $v_{bx}$ and $\vph$.  Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus, we find: 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & T\ll m\\
& 0.34 \Gamma_0 & T\gg m
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.

\begin{figure*}
\includegraphics[width=0.9\textwidth]{growthrates.png}
\caption{Oblique growth rate maximized over $\vph$ as a function of
  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
  cases, the maximum growth rate occurs when
  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
\end{figure*}

\begin{figure}
\includegraphics[width=0.5\textwidth]{oblique_GammaM.eps}
\caption{Oblique kinetic growth rate, maximized over $v_{bx}/v_b$ and
  $\omega/k$, normalized by
  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function
  of $T$ (in units of $m$) and $\gamma_b$.  Note that unlike the beam
  plasma case shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the
  transition between high and low temperature is only marginal,
  constituting a roughly $10\%$ reduction.} \label{fig:OGgen}
\end{figure}

% We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
% \begin{equation}\label{eq:growth-rate-bp}
%  \Gamma = -\frac{\sqrt{\pi} n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\parallel}}\alpha'_{\parallel}}\exp\left(-\frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}}\right),
% \end{equation}
% where $\alpha'_{\parallel} = \gamma^4\alpha_{\parallel}$ is the rescaled factor such that it is order unity.
% 
% In Figure \ref{fig:bp-growth-rate}, we plot the growth rate for the beam-plasma instability (eq.[\ref{eq:growth-rate-bp}]) for $\gamma^4\alpha_{\parallel} = 0.1, 0.3, 1,$ and 3.  The scaling of $\alpha_{\parallel}$ with $\gamma$ removes the overall scaling of $\alpha_{\parallel}$ with $\gamma$.  We have also parametrized the wavevector in terms of $\Delta k = k - k_0$, where $k_0 = \omega_{p,t}/v_b$ is the wavevector of the wave whose phase speed $v_{\rm ph} = v_b$ is equal to the speed of the beam.  Here in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$, so the positive values of $\Delta k$ results in negative values of $\Delta v_{\rm ph}$, which puts these Langmuir waves in the rising part of the distribution function, i.e., $\partial F_2/\partial\Delta\beta_z > 0$, which is required for growth. Note as well that the maximum growth rate scales like $1/\gamma^4\alpha_{\parallel}$ in analogy with the scaling of $\Delta v^{-2}$ scaling in the non-relativistic case, where $\Delta v$ is the velocity dispersion \citep{Boyd}.   
% 
% The general solution for arbitrary $\kvec$ produces a growth rate of 
% \begin{eqnarray}
% \Gamma &=&  -\sqrt{\frac{\pi k_z}{k_x}}\frac{n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\perp}}}\frac{k_z}{k}\left(\frac{1}{\alpha'_{\perp}} + \frac{1}{\alpha'_{\parallel}} - \frac{k_z}{k_x\alpha'_{\parallel}}\right)\nonumber\\
% && \exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right),\label{eq:growth-rate-oblique}
% \end{eqnarray}
% where $\alpha'_{\perp} = \gamma^2\alpha_{\perp}$ is an O(1) factor.  Again, we find that the overall scaling of the kinetic ``oblique'' instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.  Also we have made the assumption that $k_x/k_z \gg 1/\gamma$.  For $k_x/k_z < 1/\gamma$, the dispersion relation reverts back to the beam plasma case given in Equation (\ref{eq:growth-rate-bp})
% 
% % In Figure \ref{fig:oblique-growth-rate} we plot the growth rate for the oblique instability (eq.[\ref{eq:growth-rate-oblique}]) for $\alpha=\alpha'_{\perp} = \alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3.  Here we have assumed that the perpendicular and parallel velocity dispersions are the same once the scaling for $\gamma$ has been removed.  This is true for small values of $\alpha$ that result from a rest-frame nonrelativistic distribution in Figure \ref{fig:distribution_non_rel}, but would not be the case for a rest frame relativistic distribution as shown in Figure \ref{fig:distribution_rel} where larger $v/c$ corrections come into play.  
% 
% Figure \ref{fig:oblique-growth-rate} also demonstrates that the ``oblique'' mode follows its name in that the fastest 
% growing mode is neither perfectly parallel ($k_x/k_z = 0$) nor perpendicular ($k_x/k_z = \infty$), but rather at a k-vector 
% where $k_x/k_z \sim O(1)$. This qualitatively matches the results of a full analysis of \citet{Bret-Grem-Beni:10}, who also found the greatest growth for nontrivial $k_x/k_z$. 
% 
% % \begin{figure}
% %  \includegraphics[width=0.5\textwidth]{bp-growth-rate.pdf}
% %  \caption{Growth rate for the relativistic kinetic beam-plasma instability (or bump on tail instability) for four different $\alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3.  We choose to parametrized $\alpha_{\parallel}$ in this way to remove the scaling of $\alpha_{\rm parallel}$ with $\gamma$.  On the x-axis, we have plotted in terms of $\gamma^2\Delta k/k_0$, where $k_0 = \omega_{p,t}/v_b$ and $\Delta k = k - k_0$.  A Langmuir wave with $k=k_0$ moves at the exact speed of the beam.  For growth, Langmuir waves must move with a phase velocity $v_{\rm ph} < v_b$ to be on the rising part of the distribution function.  Hence, in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$ where the extra factor of $\gamma^2$ is remove the scaling of $\Delta v_{\rm phase}$ with $\gamma$.  
% %  \label{fig:bp-growth-rate}}
% % \end{figure}
% 
% \begin{figure*}
%  \includegraphics[width=0.5\textwidth]{oblique-0_1.pdf}
%  \includegraphics[width=0.5\textwidth]{oblique-0_3.pdf}\\
%  \includegraphics[width=0.5\textwidth]{oblique-1.pdf}
%  \includegraphics[width=0.5\textwidth]{oblique-3.pdf}
%  \caption{Growth rate of the oblique instability for four different $\alpha=\alpha'_{\perp} = \alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3. 1.  As in Figure \ref{fig:bp-growth-rate}, we choose to parametrized $\alpha_{\parallel}$ and $\alpha_{\perp}$ in this way to remove the scaling with $\gamma$.  On the x-axis, we have plotted in terms of $\gamma \Delta k/k_0$, for the same reasons as in Figure \ref{fig:bp-growth-rate}, but scaling with $\gamma$ instead of $\gamma^2$.  \label{fig:oblique-growth-rate}}
% \end{figure*}
%

\subsection{The Transition between the Kinetic and Hydrodynamic Instability}\label{sec:transition}

The oblique instability exists in two different regimes, raising the important question: how are the two regimes related to each other. While this question has been studied by many authors in the context of the beam-plasma or two-stream instability \citep[see for instance][]{Melrose86,Boyd},  a clear exposition of how these two regimes are related to each other is lacking.

To begin let us return to the reactive instability.  For the growth rate of the reactive instability (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion $\rightarrow 0$.   To reach this limit, we demand
\begin{equation}
 \kvec\cdot\Delta \vel \ll \omega - \kvec\cdot\vel_b.
\end{equation}
Taking $\kvec$ to be $\omega_{p,t}/\vel_b$, so that the RHS is $\approx \Gamma$, we find
\begin{equation}\label{eq:oblique reactive regime}
 \frac{\Delta v_{\perp}}{v_b} \ll \left(\frac{n_b}{\gamma_b n_t}\right)^{1/3}
%\left(1 - \beta_b^2\cos^2\Psi\right)^{1/3}, 
\end{equation}
where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.  Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate (\ref{eq:growth rate reactive}).  For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability \citep{Boyd}
\begin{equation}\label{eq:bp reactive regime}
\frac{\Delta v_{\parallel}}{v_b} \ll \gamma_b^{-1}\left(\frac{n_b}{n_t}\right)^{1/3}.
\end{equation}

Applying these results to the case of ultrarelativistic beams and given our results from Section \ref{sec:setup}, we find
\begin{eqnarray}
\frac{\Delta v_{\perp}}{v_b} &\sim&  \sqrt{\alpha_{\perp}} = \sqrt{\alpha'_{\perp}} \gamma^{-1},\\
\frac{\Delta v_{\parallel}}{v_b} &\sim& \sqrt{\alpha_{\parallel}} = \sqrt{\alpha'_{\parallel}}\gamma^{-2},
\end{eqnarray}
where $\alpha'_{\perp}$ and $\alpha'_{\parallel}$ encodes the dependence on the COM velocity dispersion.  Hence, the conditions for the reactive regime for the ``oblique'' mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
\begin{equation}\label{eq:generic reactive regime}
 \displaystyle
1 \ll \left\{
 \begin{array}{cl}
\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right..
\end{equation}

We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:growth-rate-bp}]) and ``oblique'' mode (eq.[\ref{eq:growth-rate-oblique}]).  Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.  Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$  For the beam plasma case, the growth rate is roughly
\begin{equation}
 \Gamma \sim \frac{n_b}{\gamma^3n_t} \left(\frac{c}{\Delta v_{\parallel}}\right)^{2}.
\end{equation}
The bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:bp kinetic regime}
 \frac{\Delta v_{\parallel}}{v_b} \gtrsim \gamma^{-1}\left(\frac{n_b}{n_t}\right)^{1/3},
\end{equation}
which connects with the condition on the reactive beam plasma instability from Equation (\ref{eq:bp reactive regime}).
Similarly for the ``oblique'' mode, the bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:oblique kinetic regime}
\frac{\Delta v_{\perp}}{v_b} \gtrsim \left(\frac{n_b}{\gamma n_t}\right)^{1/3},
\end{equation}
which can similarly compared to the condition on the reactive ``oblique'' mode from Equation (\ref{eq:oblique reactive regime}).

Combining these two kinetic condition and our result again from Section \ref{sec:setup}, we find  
\begin{equation}\label{eq:generic kinetic regime}
1 \ge \left\{
\begin{array}{cl}
\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right.,
\end{equation}
which in combination with Equation (\ref{eq:generic reactive regime}) denotes the transition between the reactive and kinetic regimes.

\section{Application to Ultrarelativistic $\epm$ Beams}\label{sec:application}

As discussed in the Introduction, the annihilation of VHEGRs and EBL photons produce ultrarelativistic \epm\ beams that are unstable to the beam plasma and ``oblique'' modes discussed above.  To apply the above results to the ultrarelativistic \epm\ beams, we now calculate their initial conditions.

\subsection{Average COM Energy of the \epm\ Beam}\label{sec:temperature}

To find the effective velocity dispersion of the ultrarelativistic \epm\ beam, we must first estimate the average COM energy of the beam.  To do so, we consider the process of photon-photon annihilation.  For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is
\begin{eqnarray}
\Gamma_{\pm}(E_{\rm ph}) &=& \int \sigma c dn_{\rm ebl} \nonumber \\
&=& \int \sigma\left(E_{\rm ph}, E_{\rm ebl}\right) c \frac{dn_{\rm ebl}}{dE_{\rm ebl}} dE_{\rm ebl},
\end{eqnarray}
where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm ebl}$ is the number density of EBL photons, and $E_{\rm ebl}$ is the energy of the EBL photons.  There are two important components to this calculation -- the cross section, $\sigma$, and the spectra of the EBL.

For $\sigma$, we use the results from \citet{Nikishov62} and \citet{Gould+67}, who considered a high energy photon with energy $E_{\rm ph}$ moving along the x-axis annihilating on an EBL photon with energy $E_{\rm ebl}$ moving at an angle, $\theta$, with respect to the x-axis.  The total cross section for this process is \citep{Nikishov62,Gould+67}
\begin{equation}\label{eq:cross section}
 \sigma = \frac {1}{2} \pi r_e^2\left(1-\beta_e^2\right)\left[\left(3-\beta_e^4\right)\ln\frac{1+\beta_e}{1-\beta_e} - 2\beta_e\left(2-\beta_e^2\right)\right],
\end{equation}
where $r_e = e^2/m_e c^2$ is the classical electron radius and $\beta_e$ is the electron velocity in the COM frame.  To find $\beta_e$, we use the energy of the electron in the COM frame, $E_e$, which is 
\begin{equation}
\label{eq:threshold}
 E_e =  \frac{m_e c^2}{1-\beta_e^2} = \sqrt{\frac{1}{2}{E_{\rm ebl} E_{\rm ph}\left(1-\cos\theta\right)}}.
\end{equation}
Pair production occurs when $E_e/m_e c^2 \geq 1$.

The second ingredient is the spectra of the EBL, which is not well constrained.  Here we use the constraints from \citet{Ahar_etal:06}, who demonstrated that VHEGR emission from H 2356-309 and 1ES 1101-232 places an upper limit on the EBL that is close to the lower limit of the integrated light from galaxies \citet{Madau+00}. Looking at Figure 1 of \citet{Ahar_etal:06}, we note that the EBL has a flat spectrum, i.e., constant ${dn_{\rm ebl}}/{dE_{\rm ebl}}$ below $1\,\eV$ and a falling spectrum  ${dn_{\rm ebl}}/{dE_{\rm ebl}}\propto E_{\rm ebl}^{-1.5}$ with a spectral index of $\approx 1.5$ above $1\,\eV$ with a rapid cutoff above $10\,\eV$. Thus, we adopt a simplified model:
\begin{equation}
 \frac{dn_{\rm ebl}}{dE_{\rm ebl}} \propto \left\{ 
\begin{array}{cl}
E_{\rm ebl}^0 & E_{\rm ebl} \leq 1\,\eV \\
E_{\rm ebl}^{-1.5} & 1\,\eV < E_{\rm ebl} \leq 10\,\eV \\
 0 & E_{\rm ebl} > 10\,\eV 
\end{array}
\right.
\end{equation}

In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).  We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm CM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
We note several feature of this Figure.  First the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy. This is because different energy photons probe different regimes of the EBL spectra.  Due to the rapid cutoff in the EBL above $10\,\eV$, lower energy VHEGRs produce colder beams.  This is seen in the average COM energies of the produced electrons, which are respectively, $E_e/m_e c^2 \approx 1.3, 1.6, 2$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$ and $10\,\TeV$. Hence we expect that these pairs are in the sub-relativistic to mildly-relativistic regimes in their COM frame.
Second, the effect of the EBL spectra on the distribution of electron COM energies is also seen in the breaks in the distribution.  For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm ebl}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.  Additionally, the break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.

\begin{figure}
 \includegraphics[width=0.5\textwidth]{rates.pdf}
 \caption{Differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).  The effect of the EBL spectra can be seen in different features in this plot.  The cutoff in $d\Gamma_{\pm}/dE_{\rm ebl}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.  The break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$. The average COM energies of the produced electrons are $E_e/m_e c^2 \approx 1.3, 1.6, 2,$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.  
 \label{fig:rates}}
\end{figure}

\subsection{Regime of Instability}

Given that the range of $E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, this corresponds to values of $\alpha_{\perp}'$ and $\alpha_{\parallel}'$ between $\sim 0.3-2$.  To determine whether or not the reactive or kinetic instabilities apply to these beams, it is necessary to determine $n_b/n_t$.  Here the target is the background IGM, so $n_t = \nIGM$, where $\nIGM \approx 2\times 10^{-7}\,(1+\delta)\,(1+z)^3\,\cm^{-3}$ is the mean density of the IGM, $z$ is the redshift, and $\delta$ denotes the overdensity.  The number density of the TeV beam is more complicated as the production rate of pairs must be balanced against their loss due to plasma instabilities or ICC.  This is discussed extensively in Paper I and will not be repeated here.  However, we note that the important issue here is the loss rate due to plasma instabilities, which is a nonlinear process.  In Paper I, we assumed that the nonlinear loss rate was the same as the linear growth rate.  This remains to be shown and is the focus of ongoing work, of which this paper lays the initial foundation.

Still some progress can be made if we assume that the dominant loss rate is ICC so that we can place a lower limit on the various plasma processes.  In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (Paper I):
\begin{eqnarray}
\frac{\nb}{\nIGM} &\simeq&
\frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \nonumber\\
&\simeq &
3\times10^{-18}
\left(\frac{1+z}{1}\right)^{3\zeta-7}
\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)
\left(\frac{E}{\TeV}\right)
\,\cm^{-3}\,,
\label{eq:nbIC}
\end{eqnarray}
where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta$ is.

In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.[\ref{eq:generic reactive regime}]) and kinetic regimes (eq.[\ref{eq:generic kinetic regime}]) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.  Applying these conditions (eqns. [\ref{eq:generic reactive regime}] and [\ref{eq:generic kinetic regime}]) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter: 
\begin{eqnarray}
\frac{\gamma_b^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 7\times 10^{-3} \frac{\gamma_6^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
\end{eqnarray}
and
\begin{eqnarray}
\frac{\gamma_b}{\sqrt{\alpha_{\parallel}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 0.7 \frac{\gamma_6}{\sqrt{\alpha_{\parallel}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam bp condition}
\end{eqnarray}
where $\gamma_6 = \gamma/10^6$. 
We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability can also sit in the reactive regime for sufficiently cold beams $\alpha_{\parallel} \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$.

The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.  However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.  However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.  First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.  This is a part of ongoing work and will be presented in a future publication. Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{\alpha_{\parallel}'}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.  Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).  Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{Puchwein+2011}.

\section{Summary and Conclusion}\label{sec:conclusions}

The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and ``oblique'' instability.  In this work, we have these linear instabilities as they would apply to the ultrarelativistic pair beams.  Our main findings are:
\begin{itemize}
\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.  We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam. We also show that this can be identified with a drifting relativistic Maxwell-Juttner distribution. }
\item{Using this simplified distribution function, we have analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.  We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the ``oblique'' mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.  In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
\item{We have also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.  In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the ``oblique'' mode.  This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.}
\item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectra of the EBL.  We found that the average energy of these beams range from $E_e/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.  This is due to the rapid cutoff in the EBL above $10\,\eV$. }
\item{Using this average COM energy, we presume that the pair beam can be modelled as a relativistic Maxwellian in the COM frame.   We then demonstrate that the ``oblique'' mode is in the kinetic regime. However, the beam plasma mode is marginally in the reactive regime, so that the use of either regime produces roughly the same growth rate.  Moreover, the inclusion of plasma instability losses on the number density of beam particles may place the beam plasma mode firmly in the kinetic regime.}
\end{itemize}

The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.  In particular, by developing a simple distribution function, whose modes can be analytically calculated, we will develop a quasilinear theory for these modes in a forthcoming publication.  Such a quasilinear calculation will capture the dominant wave-particle interaction and capture the initial nonlinear evolution of this instability.

\acknowledgements

We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions. We also thank E. Zweibel for useful discussions. P.C. is supported in part by an UW-Milwaukee RGI grant and a \Fermi GI grant.  C.P. gratefully
acknowledges financial support of the Klaus Tschira Foundation. G.V. acknowledges support from CITA.\\

\appendix

%\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
%\label{sec:lorentz}
%
%Here we explicitly derive the scaling of the parallel and
%  perpendicular velocity dispersions with the Lorentz factor upon boosting the
%  distribution function to the lab frame. Let us begin with a distribution
%function that is isotropic in the COM frame and depends only on energy.
%Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
%function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
%(denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
%i.e., total number, or
%\begin{equation}
% \int \fLAB \dVLAB = \int \fCM \dVCM.
%\end{equation}
%It is well-known that under Lorentz transformations (Landau), 
%\begin{equation}
% \dVLAB = \dVCM,
%\end{equation}
%so therefore,
%\begin{equation}
%  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
%\end{equation}
%
%Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
%\begin{equation}
% \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
%\end{equation}
%We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$. More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame. This gives:
%\begin{eqnarray}\label{eq:betabarLAB}
% \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
%&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
%\end{eqnarray}
%Breaking the components of \betabarLAB\ into components parallel and
%perpendicular to the boost, we find: 
%\begin{eqnarray}
%\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
%\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
%\end{eqnarray}
%where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
%For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
%However, this Galilean result no longer holds for relativistic motion.
%
%Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
%\begin{equation}
% \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
%\end{equation}
%In the lab frame, it is again useful to break it into components -- the parallel component becomes 
%\begin{eqnarray}
%  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
% &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
% &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
%\end{eqnarray}
%while the perpendicular component becomes
%\begin{eqnarray}
% \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
%&=& \gamma_{\rm B}^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
%\label{eq:dbetabarLABperp}
%\end{eqnarray}
%where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.  
%
%It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
%\begin{equation}
% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2}.
%\end{equation}
%This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.  This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.  Let us now consider the parallel component (eq.(\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$: 
%\begin{equation}
% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}},
%\end{equation}
%Here, the scaling of the parallel velocity dispersion can be understood as a double scaling of both time and coordinate (along the boost axis) between same two frames boosted relative to each other, giving an extra scaling of $\gamma^{-2}$.
%This scaling of the parallel component of the velocity dispersion has important consequences that we explore in the main part of the paper.  
%
%

\section{Solution for the Reactive Regime}\label{sec:solution reactive}

This can be rewritten as 
\begin{eqnarray}\label{eq:dispersion reactive}
 \left(\omega^2 - \omega_{p,t}^2\right)\left[\left(\omega - k_z v_b\right)^2 - \frac{\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}\right] = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2},
\end{eqnarray}
where we have added a factor of $\gamma^{-3}\omega_{p,t}^2\omega_{p,b}^2({\gamma^2 k_x^2 + k_z^2})/({k_x^2 + k_z^2})$ to both sides. 
To solve the dispersion relation (\ref{eq:dispersion reactive}), we take $\omega = \omega_{p,t} + \Delta\omega$ and expand to lowest order in
$\Delta\omega$ and $\omega_{p,b}$.  This gives
\begin{equation}
 2\Delta\omega\omega_{p,t}\left(\Delta\omega + \omega_{p,t} - k_z v_b\right)^2 = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}.
\end{equation}\label{eq:expanded dispersion reactive}
For $\Delta\omega \ll \omega_{p,t} - k_z v_b$, $\Delta\omega$ is real and there is no instability.  However, if $k_z = \omega_{p,t}/v_b$, we then have
\begin{equation}\label{eq:approximate dispersion reactive}
 \Delta\omega^3 = \omega_{p,t}^3\frac{\omega_{p,b}^2}{2\gamma^3\omega_{p,t}^2}\frac{\gamma^2 Z_x^2 + 1}{Z_x^2 + 1},
\end{equation}
where we have multiplied the fraction on the right hand side by $(v_b/\omega_{p,t})^2/(v_b/\omega_{p,t})^2$ and $Z_x = k_xv_b/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.  
Equation (\ref{eq:approximate dispersion reactive}) gives three solutions for $\Delta\omega$: one real and two imaginary (one growing and one damping).  The maximum growth rate is then
\begin{equation}\label{eq:growth rate reactive appendix}
 \Gamma = \frac{\sqrt{3}}{2^{4/3}}\left(\frac{n_b}{n_t}\right)^{1/3}\left(\frac{\gamma^2 Z_x^2 + 1}{Z_x^2 + 1}\right)^{1/3}\frac{\omega_{p,t}}{\gamma}
\end{equation}

\section{Solution for the Kinetic Regime}\label{sec:solution kinetic}

As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.  Expanding the denominator in powers of $v$, we find
\begin{eqnarray}
 \int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2
 \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
&\approx& \frac {k^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),\label{eq:dispersion-real-part}
\end{eqnarray}
where the second term is zero because it is odd\cc{, $\lambda_D^2 = k_B T/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.  These wave are not legitimate and result from the Taylor expansion of the denominator of equation (\ref{eq:dispersion-real-part}).  A correct treatment of equation (\ref{eq:dispersion-real-part}) with the appropriate Landau contours will give the correct growing or damping behavoirs for waves with phase speeds approximately that of the electron phase speeds.}}. If we ignore the third term in the kinetic dispersion relation (\ref{eq:dispersion kinetic}), this yields two plasma modes: an undamped plasma oscillation mode with $\omega = \omega_{p,t}$ and a longitudinal electron plasma wave, i.e., Langmuir wave, with
\begin{equation}
\omega = \omega_{p,t}\left(1 + \frac 3 2 k^2\lambda_{D,t}^2\right).
\end{equation}

To compute the contribution from the beam term, we will reorient our coordinate system and define the $z'$-axis along the wavevector, $\kvec$.  In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.   
This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_z=\gph\vph\epsilon_\perp$, $\epsilon=\gph\epsilon_\perp$, and
$\epsilon_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
\begin{eqnarray}
F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
&=&\frac{1}{
\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam\gammaphase(1 - \betabeamz\betaphase)\epsilon_\perp)} {T}\right)\exp\left(\frac{\gammabeam\betabeamx p_{x'}}{T}\right),\label{eq:dist2}
\end{eqnarray}
Plugging the equation into , we find that the this involves the integral of 
\begin{equation}
\int \frac{k\partial F_b/\partial p_{z'}}{\omega - k_{z'}v_{bz'}}d^3p = i 2\pi R,
\end{equation}
where $R$ is the residue for $p_{z'}$ such that $v_{z'} = v_{\rm ph}$.  
   Here we find for the two elements that contribute to the pole:
\begin{equation}
\left.
\frac{\partial}{\partial p_{z'}} \left(\omega-k v_{bz'}\right)
\right|_{\rm pole}
&=
\left.
- k \left(
  \frac{1}{\epsilon}
  -
  \frac{p_{z'}^2}{\epsilon^3}
\right)
\right|_{\rm pole}
=
- \frac{k}{\gph^3 \epsilon_\perp}\,,
\end{equation}
and 
\begin{equation}
\begin{aligned}
\left.
k\frac{\partial F_b}{\partial p_{z'}}
\right|_{\rm pole}
&\qc{\simeq}
%=
\left.
-\frac{k\gamma_b}{T} \left( \frac{p_z}{\epsilon} - v_{bz}\right)
F_b
\right|_{\rm pole}\\
&=
-
\frac{
  k
  (\vph-v_{bz}) 
  e^{-\gamma_b\gph(1-v_{bz}\vph)\epsilon_\perp/T}
  e^{\gamma_b v_{bx} p_x/T}
}{
  4\pi m^2 T^2 K_2(m/T)
}\,,
\end{aligned}
\end{equation}
Note that the evaluation of the derivative is easiest using equation (\ref{eq:dist1}).
%where we have used
%\begin{equation}
%\left.
%\gamma_b(\epsilon-v_{bz} p_z -v_{bx} p_x)/T
%\right|_{\rm pole}
%=
%\gamma_b \gph ( 1-v_{bz}\vph ) \epsilon_\perp/T - \gamma_b v_{bx} p_x/T\,.
%\label{eq:Eopole}
%\end{equation}

%and therefore,
%where $v_b\rightarrow v_{bz}$ and the only new term is the final
%exponential factor which depends upon $p_x$.

Putting this all together, the residue is
\begin{equation}
R
=
\frac{
  \gph^3 (\vph-v_{bz}) 
}{
  4\pi m T^2 K_2(m/T)
}
\kI
,
\label{eq:Rogen}
\end{equation}
where
\begin{equation}
\kI \equiv \int d^2\!p_\perp \epsilon_\perp e^{-\cG(\epsilon_\perp-w p_x)/m},
\end{equation}
and $\cG\equiv\gamma_b\gph(1-v_{bz}\vph)m/T$ and
$w\equiv \gamma_b v_{bx} m/\cG T\le 1$.  This latter inequality is guaranteed as 
\begin{equation}
\cG \epsilon_\perp - \gamma_b v_{bx} p_x = \gamma_b(E - v_{bz} p_z - v_{bx}p_x)|_{v_{bz}=v_{\rm ph}} > 0
\end{equation}
is the energy in beam frame and is therefore positive definite.  Noting that the $\exp(-wp_{x'})$ term appears as a boosted distribution, we boost by $w$
along the $x'$-axis, removing the anisotropic term from the
exponential.

Thus, we define $p_x'=\gamma_w(p_x-w \epsilon_\perp)$ and $p_y'=p_y$.  With this change in varables, we find:
\begin{eqnarray}
%\epsilon'_\perp &=& \gamma_w(\epsilon_\perp - w p_x),\\
\epsilon_\perp &=& \gamma_w(\epsilon'_\perp + w p'_x),\\
d p_x d p_y &=& \frac{\epsilon_\perp}{\epsilon'_\perp} d p'_x d p'_y.
\end{eqnarray}
Therefore, we have
\begin{equation}
\begin{aligned}
\kI
&= 
\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} e^{-\cG'\epsilon'_\perp/m}\\
%&=
%\gamma_w^2 \int d^2\!p'_\perp \left(
 % \epsilon'_\perp + 2 w 
%\qc{p'_x}
%+ w^2 \frac{p'^2_x}{\epsilon'_\perp}
%\right) e^{-\cG'\epsilon'_\perp/m}\\
&=
\pi \gamma_w^2 \int_0^\infty dp'^2_\perp \left(
  \epsilon'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{\epsilon'_\perp}
\right) e^{-\cG'\epsilon'_\perp/m}\,,
\end{aligned}
\end{equation}
where $\cG'\equiv\cG/\gamma_w$.
% and in the second line we used the fact
%that  $\epsilon'_\perp e^{-\cG'\epsilon'_\perp/m}$ is isotropic in
%$\bp'_\perp$.  
Using the following integrals: 
\begin{equation}
\int_0^\infty dx \sqrt{1+x} e^{-a\sqrt{1+x}} = \frac{2 e^{-a}}{a^3}\left(a^2 + 2a + 2\right)
\end{equation}
and 
\begin{equation}
\int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
=
\frac{4(a+1)}{a^3} e^{-a}\,,
\end{equation}
we find
\begin{equation}
\kI
=
\frac{2 \pi \gamma_w^2 m^3}{\cG'^3} \left[
\left(\qc{\cG'}^2+2\qc{\cG'}+2\right)
+
\frac{w^2}{2}
\left(2\qc{\cG'}+2\right)
\right]\,\qc{e^{-\cG'}}.
\end{equation}
Inserting this into (\ref{eq:Rogen}) yields
\begin{equation}
R
=
\frac{
  \gph^3 m^2 (\vph-v_{bz}) 
}{
  2 T^2 K_2(m/T)
}
\gamma_w^2
\left[
\left(\cG'^2 + 2\cG' + 2\right)
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right] \frac{e^{-\cG'}}{\cG'^3}
\end{equation}
and therefore,
\begin{equation}
\Gamma \simeq - \Gamma_0
\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
\left[
\left( \cG'^2 + 2\cG' + 2 \right) 
+
\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
\right]
\frac{e^{-\cG'}}{\cG'^3}\,.
\label{eq:OGgen}
\end{equation}

%The above differs from the beam plasma instability only due to the
%minor difference in the definition of $\cG'$ and the second term.  When
%$v_{bx}=0$, it is precisely the beam plasma instability growth rate.

%With an eye toward expanding to lowest order in $v_{bx}\ll1/\gamma_b\simeq1/\gph$, we find
%\begin{equation}
%\gamma_w 
%\simeq
%1 + \frac{v_{bx}^2}{2 \gph^2(1-v_{bz}\vph)^2}
%=
%1 + \frac{\gamma_b^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
%\quad\Rightarrow\quad
%\cG'\simeq \cG - \frac{\gamma_b^2 v_{bx}^2}{2 \cG}\frac{m^2}{T^2}.
%\end{equation}  
%Thus, to lowest order in $\gamma_b v_{bx}$, equation (\ref{eq:OGgen}) becomes
%\begin{equation}
%\begin{aligned}
%\Gamma 
%&\simeq 
%- \Gamma_0
%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
%\frac{e^{-\cG}}{\cG^3}
%\left(
%  1
%  +
%  \frac{\gamma_b^2 v_{bx}^2}{\cG^2} \frac{m^2}{T^2}
%\right)\\
%&\qquad\times
%\left[
%\left( 
%  \cG^2 + 2\cG + 2 - \gamma_b^2 v_{bx}^2\frac{m^2}{T^2}  - \frac{\gamma_b^2 v_{bx}\qc{^2}}{\cG}\frac{m^2}{T^2}
%\right)
%+
%\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG + 2\right)
%\right]
%\\
%&\qquad\qquad\qquad\qquad\times 
%\left(
%  1 + 3 \frac{\gamma_b^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
%\right)
%\left(
%  1 + \frac{\gamma_b^2 v_{bx}^2}{2\cG}\frac{m^2}{T^2}
%\right)\\
%&\simeq 
%- \Gamma_0
%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
%e^{-\cG}\\
%&\qquad\qquad\times 
%\left( 
%  \frac{\cG^2 + 2\cG + 2}{\cG^3}
%  +
%  \frac{\gamma_b^2 v_{bx}^2}{2}\frac{m^2}{T^2}
%  \frac{
%    \cG^3
%    +
%    5\cG^2
%    +
%    12\cG
%    +
%    12
%  }{\cG^5}
%\right)\,,
%\end{aligned}
%\end{equation}
%where $\Gamma_0\equiv \omega_{Pt} (n_b/n_t) \gamma_b m v_{bz}^2/T$, i.e.,
%again $v_{bz}$.

However, obtaining asymptotic expansions for the growth rate in the high and
low temperature limits that can be easily solved for the maximum
growth rate is difficult.  Even with the tremendous numerical insight
provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
significant progress.  However, some general remarks can be made.  It
appears at the maximum growth rate, we have to leading order:
\begin{equation}
\begin{gathered}
\vph\simeq\sqrt{2\delta Z_x}\,,\quad
v_{bx}=(1-\delta Z_x) v_b\,,\quad
v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
v_{bz}-\vph\simeq\left\{
\begin{aligned}
&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
&\frac{1}{2\gamma_b} & T\gg m
\end{aligned}
\right.\,,\quad
\gph\simeq1+\delta Z_x\,,\quad
\cG' \sim \frac{m}{T}\,,\\
\text{and}\quad w \sim 1\,,
\end{gathered}
\end{equation}
\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
(\ref{eq:OGgen}) suffices to show that there exists growth rates that
are $\propto \Gamma_0$, with 
\begin{equation}
\Gamma_M
\simeq
\left\{
\begin{aligned}
& 0.38 \Gamma_0 & T\ll m\\
& 0.34 \Gamma_0 & T\gg m
\end{aligned}
\right.\,,
\end{equation}
for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).

% We now proceed to the third term.  Without loss of generality, we orient our axes such that the beam is along the z-axis and the wavevector k sits in the x-z plane, i.e., $\kvec = k_x \xhat + k_z \zhat$. The beam distribution function (\ref{eq:distribution function beam}) now becomes
% \begin{equation}
% df=\frac{n_b}{\sqrt{\pi^3\dbetabarpara\dbetabarperp^2}}\exp\left(-\frac{\Delta \beta_{z}^2}{\dbetabarpara} - \frac{\Delta \beta_{x}^2 + \Delta\beta_{y}^2}{\dbetabarperp}\right) d^3\Delta\beta,
% \end{equation}
% We may integrate over $d\Delta\beta_y$ leaving an integral over $d\Delta\beta_z, d\Delta\beta_x$.  
% \begin{equation}\label{eq:integral1}
% \mathcal{I} = \int \frac{(k_x \partial_{p_x} + k_z\partial_{p_z}) F_1}{\omega - \kvec\cdot\vel}d\Delta\beta_z\, d\Delta\beta_x,
% \end{equation}
% where $F_1$ is the distribution function after has been integrated over $d\Delta\beta_y$.  Note that we have integrated the dispersion relation (\ref{eq:dispersion relation}) rather that the one that was integrated by parts (eq.[\ref{eq:dispersion relation by parts}]) because the Landau contour that we perform below is more explicit with this version.  Equation (\ref{eq:integral1}) can be rewritten as 
% \begin{equation}\label{eq:integral2}
% \mathcal{I} \approx \int \frac{(\gamma^2k_x\partial_{\Delta\beta_x} + k_z\partial_{\Delta\beta_z} ) F_1 d\Delta\beta_z\, d\Delta\beta_x}{k_z\gamma^3m_e c^2(\omega/k_z c - \beta_z - k_x\Delta\beta_x/k_z - \Delta\beta_z)}, 
% \end{equation}
% where we have converted the derivatives on $p_x$ and $p_z$ to derivatives on $\Delta \beta_x, \Delta \beta_z$ by assume that they are $ \ll 1/\gamma$ and $\ll 1/\gamma^2$, and take the following expansions:
% \begin{eqnarray}
%  p_x &\approx& 0 + \frac{\partial p_x}{\partial \Delta\beta_x} \Delta\beta_x \approx \gamma_{0} m_e c \Delta\beta_x\\
%  p_z &\approx& p_{z,0} + \frac{\partial p_z}{\partial \Delta\beta_z} \Delta\beta_z \approx p_{z,0} + \gamma_{0}^3 m_e c\Delta\beta_z,
% \end{eqnarray}
% where $\gamma_0 = \sqrt{1 + p_{z,0}^2/m_e^2c^2}$. 
%  
% \ac{[I had some difficulty with the above, in part because I think many approximations are being utilized at the same time.  Thus perhaps it would be better to explain each step in a bit more dtail.  
% The thing I had the most difficulty with was the $\bmath{p}\rightarrow\vel$ transformation in the integrands.  This is because the expansions in C6 \& C7 do not imply the transformation used in C5.  Rather, we should have
% \begin{equation*}
% \frac{\partial}{\partial p^j} 
% =
% \frac{\partial \beta^k}{\partial p^j} \frac{\partial}{\partial \beta^k}
% =
% \frac{\delta^k_j - \beta_j\beta^k}{\gamma m} \frac{\partial}{\partial \beta_k}\,,
% \end{equation*}
% and thus,
% \begin{equation*}
% \begin{aligned}
% \kvec\cdot\nabla_p F_1
% &=
% \frac{k^k - (k^j \beta_j)\beta^k}{\gamma m} \frac{\partial F_1}{\partial \beta_k}\\
% &=
% \frac{1}{\gamma m}
% \bigg\{
% \left[k_x - (k_x\beta_x+k_z\beta_z)\beta_x\right] \frac{\partial F_1}{\partial \beta_x}\\
% &\qquad\qquad
% +
% \left[k_z - (k_x\beta_x+k_z\beta_z)\beta_z\right] \frac{\partial F_1}{\partial \beta_z}
% \bigg\}\,.
% \end{aligned}
% \end{equation*}
% 
% When $k_x=0$, this simplifies a little bit, and we have
% \begin{equation*}
% \kvec\cdot\nabla_p F_1
% =
% \frac{1}{\gamma m}
% \left[
% -k_z\beta_z\beta_x\frac{\partial F_1}{\partial \beta_x}
% +
% k_z \left(1-\beta_z^2\right) \frac{\partial F_1}{\partial \beta_z}
% \right]\,,
% \end{equation*}
% and with $\beta_x\ll\beta_z\sim1$ we have
% \begin{equation*}
% \kvec\cdot\nabla_p F_1
% \simeq
% \frac{k_z}{\gamma^3 m}
% \left[
% -\gamma^2 \beta_z\beta_x\frac{\partial F_1}{\partial \beta_x}
% +
% \frac{\partial F_1}{\partial \beta_z}
% \right]\,,
% \end{equation*}
% with the important difference being the first term.  With the form of $F_1$ we have $\partial F_1/\partial \beta_j = 2\beta_j F_1 / \alpha_j$, and may evaluate each term to lowest order:
% \begin{equation*}
% \begin{aligned}
% &I_{x} \equiv -\int \frac{1}{\omega-k_z \beta_z} \frac{k_z}{\gamma^3 m} \gamma^2 \beta_z \beta_x 
% \left(\frac{2\beta_x}{\alpha_x} F_1\right)
% d\beta_x d\beta_z\\
% &\qquad\qquad= 
% -\int \frac{k_z \beta_z  F_2}{\gamma m (\omega-k_z \beta_z)} d\beta_z
% \simeq
% \left.-\frac{i \pi F_2}{\gamma m}\right|_{\beta_z=\omega/k_z}
% \\
% &I_{z} \equiv \int \frac{1}{\omega-k_z \beta_z} \frac{k_z}{\gamma^3 m} 
% \left[\frac{2(\beta_z-\beta_b)}{\alpha_z} F_1\right]
% d\beta_x d\beta_z\\
% &\qquad\qquad= 
% \int \frac{2 k_z (\beta_z-\beta_b) F_2}{\gamma^3\alpha_z m (\omega-k_z \beta_z)} d\beta_z
% \simeq
% \left.\frac{2 i \pi (\beta_z-\beta_b) F_2}{\gamma^3 \alpha_z m}\right|_{\beta_z=\omega/k_z}\\
% &\qquad\qquad\simeq
% - \frac{2(\omega/k_z-\beta_b)}{\gamma^2 \alpha_z} I_x\,.
% \end{aligned}
% \end{equation*} 
% Note that $I_{z}$ is just your expression C9.
% Now, since $\alpha_z\sim T/\gamma^4$, and $F_2$ is large for $|\omega/k_z-\beta_b|\lesssim\sqrt{\alpha_z}\sim T^{1/2}/\gamma^2$, $|I_z|\sim |I_x|/T^{1/2}$ and thus the neglected term is not generally small.
% 
% More problematic, I would like to understand why the $k_x=0$ case doesn't reproduce the standard cold plasma instability rate when $\alpha_z$ is very small.
% ]}
% 
% 
% 
% 
% 
% To make a connection with the the beam plasma instability, we take $k_x = 0$ and integrate Equation (\ref{eq:integral2}) over $d\Delta\beta_x$ to find:
% \begin{equation}\label{eq:integral-bp}
% \mathcal{I} \approx \frac{1}{\gamma^3 m_e c^2}\int \frac{\partial_{\Delta\beta_z} F_2}{(\omega/k c - \beta_z - \Delta\beta_z)}d\Delta\beta_z, 
%  \end{equation}
% where $F_2 = \sqrt{1/\pi\alpha_{\parallel}}\exp(-\alpha_{\parallel}^{-1} \Delta\beta_z^2)$ is the normalized distribution function after it has been integrated over $d\Delta\beta_x$.  Equation (\ref{eq:integral-bp}) can be evaluated using the Landau prescription, which yields a principle value (which is small and real and hence can be neglected as the background is dominant) and an imaginary component from the contribution of the pole
% \begin{equation}\label{eq:imaginary-bp}
%  \Im(\mathcal{I}) = \frac{i\pi}{\gamma^3 m_e c^2}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)}
% \end{equation}
% 
% Combining equations (\ref{eq:dispersion kinetic}), (\ref{eq:dispersion-real-part}) and (\ref{eq:imaginary-bp}), we find
% \begin{equation}
%  \omega^2 - \omega_{p,t}^2\left(1 + 3k^2\lambda_D^2\right) - \frac{i\pi\omega_{p,b}^2\omega^2}{\gamma^3 k^2c^2}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)} = 0.
% \end{equation}
% Expanding about $\omega = \omega_r - i\Gamma$ \cc{for $kc \approx \omega_r \approx \omega_{p,t}$ and assuming the imaginary parts is $O(\Gamma)$}, we find 
% \begin{eqnarray}\label{eq:growth-rate-bp-appendix}
%  \Gamma &=& \frac{\pi n_b}{2\gamma^3n_t}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)}\omega_{p,t} \nonumber\\
%  &=& -\frac{\sqrt{\pi} n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\parallel}}\alpha'_{\parallel}}\exp\left(-\frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}}\right),
% \end{eqnarray}
% where $\alpha'_{\parallel} = \gamma^4\alpha_{\parallel}$ is the rescaled factor such that it is order unity.
% As $\Delta\beta_z \sim \sqrt{\alpha_{\parallel}}$, Equation (\ref{eq:growth-rate-bp}) has an overall scaling like $\gamma\nb/n_t$ in line scalings found for the kinetic ``oblique'' mode quoted in Paper I (also see below).  
% 
% We now assume $k_x \neq 0$ and perform the integral over $\Delta\beta_z$ in Equation (\ref{eq:integral2}) to find for the imaginary part
% \begin{equation}\label{eq:integral-kinetic}
%  \Im(\mathcal{I}) = \frac{i\pi}{\gamma^3 m_ec^2}\int \left(\gamma^2\frac{k_x}{k_z}\frac{\partial F_1}{\partial\Delta\beta_x} + \frac{\partial F_1}{\partial\Delta\beta_z}\right)_{\Delta\beta_z = \Delta\beta_{\rm res}} d\Delta\beta_x,
% \end{equation}
% where $\Delta\beta_{\rm res} = \omega/k_z c - \beta_z - k_x\Delta\beta_x/k_z$.  
% 
% To perform this integral, we write the exponent in $F_1$ as 
% \begin{eqnarray}
%  F_1 \sim \exp\left[\left(\frac 1 {\alpha_{\perp}} + \frac {(k_x/k_z)^2} {\alpha_{\parallel}}\right)\left(\Delta\beta_x - \left(1 - \frac{k_x^2\alpha_{\parallel}}{k_z^2\alpha_{\perp}}\right)\frac{\Delta\beta_{z,r}}{k_x/k_z}\right)^2 - \frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}} \right],
% \end{eqnarray}
% where the normalization $1/\pi\sqrt{\alpha_{\perp}\alpha_{\parallel}}$ has been left out and $\beta_{z,r} = \omega/k_zc - \beta_z$.  
% Making an appropriate substitution for $\Delta\beta_x \rightarrow \beta_x - \left(k_z/k_x - {k_x\alpha_{\parallel}}/{k_z\alpha_{\perp}}\right)\Delta\beta_{z,r}$, we find for Equation (\ref{eq:integral-kinetic}) 
% \begin{eqnarray}
% \Im(\mathcal{I}) = -\frac{i2}{\gamma^3 c}\sqrt{\frac{\pi k_z}{k_x\alpha_{\perp}}}k_z\left(\gamma^2\frac{1}{\alpha_{\perp}} + \frac{1}{\alpha_{\parallel}} - \frac{k_z}{k_x\alpha_{\parallel}}\right)\Delta\beta_{z,r}\exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right) 
% \end{eqnarray}
% The resulting growth rate is then
% \begin{eqnarray}
% \Gamma = -\sqrt{\frac{\pi k_z}{k_x}}\frac{n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\perp}}}\frac{k_z}{k}\left(\frac{1}{\alpha'_{\perp}} + \frac{1}{\alpha'_{\parallel}} - \frac{k_z}{k_x\alpha'_{\parallel}}\right) \exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right),\label{eq:growth-rate-oblique-appendix}
% \end{eqnarray}
% where $\alpha'_{\perp} = \gamma^2\alpha_{\perp}$ is an O(1) factor.  Again, we find that the overall scaling of the kinetic ``oblique'' instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.  Also we have made the assumption that $k_x/k_z \gg 1/\gamma$.  For $k_x/k_z < 1/\gamma$, the dispersion relation reverts back to the beam plasma case given in Equation (\ref{eq:growth-rate-bp})

\bibliography{bigmh}
\bibliographystyle{apj}

\end{document}

\ac{[The paragraph below is somewhat redundant now as the primary justifications have all been given above and in the appendix.  Plus there is the rather confusing redefinition of $\beta$ as well as the introduction of $v$.  I leave it here because I'm not quite sure what the point of it is, though believing that the main message is conveyed with my suggestion above.  Perhaps this could be an after-the-fact, more rigorous motivation for the above?  Is such a motivation necessary?  Could it be moved efficiently into the footnote (a potential minimal footnote is already now provided)?]
> > > > > > > > > > > > > > > > > >}
\begin{equation}
 f\propto\exp\left(-\frac{\left(v_{\parallel}-v_{\parallel,0}\right)^2}{\dbetabarpara}-\frac{v_{\perp}^2}{\dbetabarperp}\right),
\end{equation}
where $v_{\parallel} = p_{\parallel}/\gamma m_e c$ is the parallel velocity, $v_{\parallel,0} = p_{\parallel,0}/\gamma_0 m_e c \approx c$ is the velocity of the beam, and $v_{\perp} = p_{\perp}/\gamma m_e c$ is the perpendicular velocity.
To make further progress, we make the assumption that the spread in momentum $\ll p_{\parallel,0}$.  In other words, we assume that $\alpha_{\parallel}, \alpha_{\perp} \ll 1$.  In this limit, we can perform a Taylor expansion for $p_{\parallel}$ about $p_{\parallel,0}$ and $p_{\perp}$ about $0$.  
This gives: 
\begin{eqnarray}
 p_{\perp} &\approx& 0 + \frac{\partial p_{\perp}}{\partial v_{\perp}} \Delta v_{\perp} \approx \gamma_{0} m_e \Delta v_{\perp},\\
 p_{\parallel} &\approx& p_{\parallel,0} + \frac{\partial p_{\parallel}}{\partial v_{\parallel}} \Delta v_{\parallel} \approx p_{\parallel,0} + \gamma_{0}^3 m_e \Delta v_{\parallel},\\
 \gamma &\approx& \gamma_0 + \gamma_0^3 \frac{\Delta v_{\parallel}}{c}
\end{eqnarray}
where $\gamma_0 = \sqrt{1 + p_{\parallel,0}^2/m_e^2c^2}$. Hence this gives an approximate distribution function for a beam in terms of the perturbed velocities as 
\begin{equation}\label{eq:distribution function beam}
df=\frac{n_b}{\sqrt{\pi^3\dbetabarpara\dbetabarperp^2}}\exp\left(-\frac{\Delta \beta_{\parallel}^2}{\dbetabarpara} - \frac{\Delta \beta_{\perp}^2}{\dbetabarperp}\right) d\Delta\beta_{\parallel} d^2\Delta\beta_{\perp},
\end{equation}
where $\Delta\beta = \Delta v/c$ and we have assumed a 3-D geometry so that the perpendicular direction spans two dimensions: $\Delta\beta_{\perp}^2 = \Delta\beta_{\perp,1}^2 + \Delta\beta_{\perp,2}^2$.
This form of the distributions function (\ref{eq:distribution function beam}) has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.
\ac{< < < < < < < < < < < < < < <}

### 2016-04-26 oblique.tex
fixed up appendix to B13

@@
 
+\def\gz{\gamma_{z'}}
+\def\vz{v_{z'}}
+\def\ez{\epsilon_{z'}}
+\def\pz{p_{z'}}
+
 \def\vpl{v_\parallel}
@@
 
-\ac{[There are some notational issues throughout the text, which I'll list here.  Usually it is simply a matter of choosing, and then sticking to, a convention.
-\begin{itemize}
-\item Some equations assume $c=1$ while others have the $c$'s explicitly present.
+% \ac{[There are some notational issues throughout the text, which I'll list here.  Usually it is simply a matter of choosing, and then sticking to, a convention.
+% \begin{itemize}
+% \item Some equations assume $c=1$ while others have the $c$'s explicitly present.
 Unless there is a compelling reason (e.g., clarity of derivation vs.
 utility of an answer in physical units) we should choose one and mention the convention here (at the end of the intro) if $c=1$.
-\item $\beta$ and $v$ appear to be used interchangeably, again related to the $c$ or not question.  Again, we should choose one.
-\item Throughout section 2 and in the Appendix the frame is defined in terms of $\parallel$ and $\perp$ directions (i.e., $\beta_\parallel$ and $\bmath{\beta}_\perp$, note the latter is a {\em vector}!) while in sections 3 and in the figures we refer to $\beta_z$ and $\beta_x$ for these.
-\item ApJ style, which I'm assuming is what we are following, has vectors given by bold, roman fonts and scalars by the italic fonts.
+% \item $\beta$ and $v$ appear to be used interchangeably, again related to the $c$ or not question.  Again, we should choose one.
+% \item Throughout section 2 and in the Appendix the frame is defined in terms of $\parallel$ and $\perp$ directions (i.e., $\beta_\parallel$ and $\bmath{\beta}_\perp$, note the latter is a {\em vector}!) while in sections 3 and in the figures we refer to $\beta_z$ and $\beta_x$ for these.
+% \item ApJ style, which I'm assuming is what we are following, has vectors given by bold, roman fonts and scalars by the italic fonts.
 Thus, it should be $\bmath{k}\cdot\bmath{\delta E}$ or $\bmath{k}\cdot\delta\bmath{E}$, etc.
 I've tried to fix up various definitions in the preamble to do this, but that should be checked.
-\end{itemize}
-In any equations I've added/modified I will include the $c$'s and use $\beta$ for concreteness, though I anticipate there will be few such examples.]}
+% \end{itemize}
+% In any equations I've added/modified I will include the $c$'s and use $\beta$ for concreteness, though I anticipate there will be few such examples.]}
 
@@
 Plugging this into (\ref{eq:relativistic maxwellian}), we find (Wright \& Hadley 1975)
-\begin{equation}\label{eq:drifting maxwellian}
-f = \frac{1}{4\pi\gamma_B m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right),
-\end{equation}
-where we have identified the lab frame temperature $T_{\rm L} \equiv T_{\rm CM}/\gamma_{\rm B}$ and $K_2$ is the 2nd order modified Bessel function.
+\begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
+%f = \frac{1}{4\pi\gamma_B m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
+f = \frac{n_b m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - \vel_B\cdot \bmath{p})} {k_B T}\right)
+\end{eqnarray}
+where $K_2$ is the 2nd order modified Bessel function.
 The transformation properties of the drifting Maxwell-Juttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-Juttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
@@
 Here, $s$ is the species label, $+$ for positrons and $-$ for electrons,
-with $q_{\pm} = \pm e$.  \ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
+with $q_{\pm} = \pm e$.  %\ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
 Upon linearizing this in small perturbations about a background distribution, i.e., setting 
@@
   =
-  1 - \sum_s \frac{\omega_{p,s}^2}{k^2}\int F_s \frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p 
+  1 - \sum_s \frac{\omega_{p,s}^2}{k^2c^2}\int F_s \frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p 
   =
@@
 \end{multline}
-\ac{[$\beta$ vs. $v$ issue above.]}
+%\ac{[$\beta$ vs. $v$ issue above.]}
 
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
-For the beam plasma, we adopt the Maxwell-Juttner distribution 
-\begin{equation}\label{eq:distribution function beam}
-F_b = \frac{n_b}{4\pi\gamma_B m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
-\end{equation}
+For the beam plasma, we adopt the Maxwell-Juttner distribution (eq.[\ref{eq:distribution function beam}]).
+%\begin{equation}\label{eq:distribution function beam}
+%F_b = \frac{n_b}{4\pi\gamma_B m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
+%\end{equation}
 Plugging these into the dispersion relation (\ref{eq:dispersion relation}), we find
@@
 \begin{equation}\label{eq:dispersion kinetic}
- 1 - \frac{\omega_{p,t}^2}{k^2}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
+ 1 - \frac{\omega_{p,t}^2}{k^2c^2}\int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
 \end{equation}
@@
 \begin{equation}\label{eq:cross section}
- \sigma = \frac {1}{2} \pi r_e^2\left(1-\beta_e^2\right)\left[\left(3-\beta_e^4\right)\ln\frac{1+\beta_e}{1-\beta_e} - 2\beta_e\left(2-\beta_e^2\right)\right],
-\end{equation}
-where $r_e = e^2/m_e c^2$ is the classical electron radius and $\beta_e$ is the electron velocity in the COM frame.
-To find $\beta_e$, we use the energy of the electron in the COM frame, $E_e$, which is
+ \sigma = \frac {1}{2} \pi r_e^2\left(1-\frac{v_e^2}{c^2}\right)\left[\left(3-(v_e/c)^4\right)\ln\frac{1+v_e/c}{1-v_e/c} - 2\frac{v_e}{c}\left(2-\frac{v_e^2}{c^2}\right)\right],
+\end{equation}
+where $r_e = e^2/m_e c^2$ is the classical electron radius and $v_e$ is the electron velocity in the COM frame.  To find $v_e$, we use the energy of the electron in the COM frame, $E_e$, which is 
 \begin{equation}
 \label{eq:threshold}
- E_e =  \frac{m_e c^2}{1-\beta_e^2} = \sqrt{\frac{1}{2}{E_{\rm ebl} E_{\rm ph}\left(1-\cos\theta\right)}}.
+ E_e =  \frac{m_e c^2}{1-v_e^2/c^2} = \sqrt{\frac{1}{2}{E_{\rm ebl} E_{\rm ph}\left(1-\cos\theta\right)}}.
 \end{equation}
@@
 \begin{eqnarray}
- \int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2
+ \int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2c^2
  \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
-&\approx& \frac {k^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),\label{eq:dispersion-real-part}
+&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),\label{eq:dispersion-real-part}
 \end{eqnarray}
@@
 In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.
-This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_z=\gph\vph\epsilon_\perp$, $\epsilon=\gph\epsilon_\perp$, and
+This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_{z'}=\gz\vz\epsilon_\perp$, $\epsilon=\gz\epsilon_\perp$, and
 $\epsilon_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
 \begin{eqnarray}
-F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
-&=&\frac{1}{
-\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam\gammaphase(1 - \betabeamz\betaphase)\epsilon_\perp)} {T}\right)\exp\left(\frac{\gammabeam\betabeamx p_{x'}}{T}\right),\label{eq:dist2}
+F_B &=& \frac{m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'} p_{x'})} {k_B T}\right)\\
+%F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
+&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})\epsilon_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
 \end{eqnarray}
-Plugging the equation into , we find that the this involves the integral of 
+Plugging the equation into (??), we find that the this involves the integral of 
 \begin{equation}
@@
 \right|_{\rm pole}
-&=
+=
 \left.
 - k \left(
-  \frac{1}{\epsilon}
-  -
-  \frac{p_{z'}^2}{\epsilon^3}
+  \frac{c^2}{\epsilon} - \frac{p_{z'}^2 c^4}{\epsilon^3}
 \right)
@@
 =
-- \frac{k}{\gph^3 \epsilon_\perp}\,,
+- \frac{k c^2}{\gph^3 \epsilon_\perp}\,,
 \end{equation}
@@
 \right|_{\rm pole}
-&\qc{\simeq}
-%=
+%&=
+&=
 \left.
--\frac{k\gamma_b}{T} \left( \frac{p_z}{\epsilon} - v_{bz}\right)
+-\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{\epsilon} - v_{B,z'}\right)
 F_b
@@
   k
-  (\vph-v_{bz}) 
-  e^{-\gamma_b\gph(1-v_{bz}\vph)\epsilon_\perp/T}
-  e^{\gamma_b v_{bx} p_x/T}
-}{
-  4\pi m^2 T^2 K_2(m/T)
-}\,,
+  (\vph-v_{B,z'}) \mu m_e^{-3}c^{-3}}{
+  4\pi k_BT K_2(\mu)
+} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})\epsilon_\perp)} {k_BT c^2}\right)\\
+&\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
 \end{aligned}
 \end{equation}
-Note that the evaluation of the derivative is easiest using equation (\ref{eq:dist1}).
 %where we have used
@@
 \frac{
-  \gph^3 (\vph-v_{bz}) 
-}{
-  4\pi m T^2 K_2(m/T)
+  \gph^3 (\vph-v_{B,z'})\mu}{4\pi k_BT K_2(\mu)
 }
-\kI
+\frac{\kI}{m_e^3c^5}
 ,
@@
 where
-\begin{equation}
-\kI \equiv \int d^2\!p_\perp \epsilon_\perp e^{-\cG(\epsilon_\perp-w p_x)/m},
-\end{equation}
-and $\cG\equiv\gamma_b\gph(1-v_{bz}\vph)m/T$ and
-$w\equiv \gamma_b v_{bx} m/\cG T\le 1$.  This latter inequality is guaranteed as 
-\begin{equation}
-\cG \epsilon_\perp - \gamma_b v_{bx} p_x = \gamma_b(E - v_{bz} p_z - v_{bx}p_x)|_{v_{bz}=v_{\rm ph}} > 0
+\begin{equation}\label{eq:I-integral}
+\kI \equiv \int d^2\!p_\perp \epsilon_\perp \exp\left(-\frac{\cG(\epsilon_\perp-w p_x)}{k_B T}\right)
+\end{equation}
+and $\cG\equiv\gamma_B\gph(1-v_{B,z'}\vph/c^2)$ and
+$w\equiv \gamma_Bv_{B,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
+\begin{equation}
+\cG \epsilon_\perp - \gamma_B v_{B,x'} p_{x'} = \gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'}p_{x'})|_{v_{B,z'}=v_{\rm ph}} > 0
 \end{equation}
@@
 
-Thus, we define $p_x'=\gamma_w(p_x-w \epsilon_\perp)$ and $p_y'=p_y$.  With this change in varables, we find:
+Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w \epsilon_\perp)$ and $p_{y'}'=p_{y'}$ and find:
 \begin{eqnarray}
@@
 \end{eqnarray}
-Therefore, we have
+Inserting this into equation (\ref{eq:I-integral}), we find
 \begin{equation}
 \begin{aligned}
-\kI
-&= 
-\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} e^{-\cG'\epsilon'_\perp/m}\\
+\kI &= 
+\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\\
 %&=
@@
   \epsilon'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{\epsilon'_\perp}
-\right) e^{-\cG'\epsilon'_\perp/m}\,,
+\right) \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\,,
 \end{aligned}
 \end{equation}
-where $\cG'\equiv\cG/\gamma_w$.
+where $\cG'\equiv\cG/\gamma_w$. Note in the second line that we have used isotropy in $\bp_{\perp}'$ to eliminate terms linear in $\bp_{\perp}'$. 
 % and in the second line we used the fact
@@
 =
-\frac{2 \pi \gamma_w^2 m^3}{\cG'^3} \left[
+\frac{2 \pi \gamma_w^2 (k_B T)^3}{\cG'^3} \left[
 \left(\qc{\cG'}^2+2\qc{\cG'}+2\right)
@@
 \left(2\qc{\cG'}+2\right)
-\right]\,\qc{e^{-\cG'}}.
+\right]\,\exp\left(-\frac{\cG'}{k_B T}\right).
 \end{equation}

### 2016-04-26 oblique.tex
fixed up appendix

@@
 \Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
 \left[
-\left( \cG'^2 + 2\cG' + 2 \right) 
+\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
 +
-\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
+\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
-\frac{e^{-\cG'}}{\cG'^3}\,.
-\label{eq:OGgen}
-\end{equation}
+\exp(-\cG'\mu),
+\end{equation}
+where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
 
@@
 \begin{aligned}
-& 0.38 \Gamma_0 & T\ll m\\
-& 0.34 \Gamma_0 & T\gg m
+& 0.38 \Gamma_0 & k_BT/m_ec^2\ll 1\\
+& 0.34 \Gamma_0 & k_BT/m_ec^2\gg 1
 \end{aligned}
@@
 \end{eqnarray}
-Plugging the equation into (??), we find that the this involves the integral of 
-\begin{equation}
-\int \frac{k\partial F_b/\partial p_{z'}}{\omega - k_{z'}v_{bz'}}d^3p = i 2\pi R,
+Plugging the equation into (??), we find that the this gives an associated growth rate of
+\begin{equation}
+\Gamma \simeq -\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R,
+\end{equation}
+which involves the integral of 
+\begin{equation}
+R\equiv \int \frac{k\partial F_b/\partial p_{z'}}{\omega - k_{z'}v_{bz'}}d^3p,
 \end{equation}
@@
 \begin{equation}
-\kI
-=
-\frac{2 \pi \gamma_w^2 (k_B T)^3}{\cG'^3} \left[
-\left(\qc{\cG'}^2+2\qc{\cG'}+2\right)
+\kI =
+\frac{2 \pi \gamma_w^2 m_e^3c^4}{\cG'^3\mu^3} \left[
+\left(\cG'^2\mu^2+2\cG'\mu+2\right)
 +
-\frac{w^2}{2}
-\left(2\qc{\cG'}+2\right)
-\right]\,\exp\left(-\frac{\cG'}{k_B T}\right).
+\frac{w^2}{2c^2}
+\left(2\cG'\mu+2\right)
+\right]\exp\left(-\cG'\mu\right).
 \end{equation}
@@
 \begin{equation}
-R
-=
-\frac{
-  \gph^3 m^2 (\vph-v_{bz}) 
-}{
-  2 T^2 K_2(m/T)
-}
-\gamma_w^2
+R = \frac{
+  \gph^3 \gamma_w^2 (\vph-v_{B,z'}) 
+}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
 \left[
-\left(\cG'^2 + 2\cG' + 2\right)
+\left(\cG'^2\mu^2 + 2\cG'\mu + 2\right)
 +
-\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
-\right] \frac{e^{-\cG'}}{\cG'^3}
+\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
+\right] \exp\left(-\cG'\mu\right)
 \end{equation}
@@
 \Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
 \left[
-\left( \cG'^2 + 2\cG' + 2 \right) 
+\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
 +
-\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG' + 2\right)
+\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
-\frac{e^{-\cG'}}{\cG'^3}\,.
+\exp(-\cG'\mu),
 \label{eq:OGgen}
 \end{equation}
-
+where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
 %The above differs from the beam plasma instability only due to the
@@
 
-However, obtaining asymptotic expansions for the growth rate in the high and
-low temperature limits that can be easily solved for the maximum
-growth rate is difficult.  Even with the tremendous numerical insight
-provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
-significant progress.  However, some general remarks can be made.  It
-appears at the maximum growth rate, we have to leading order:
-\begin{equation}
-\begin{gathered}
-\vph\simeq\sqrt{2\delta Z_x}\,,\quad
-v_{bx}=(1-\delta Z_x) v_b\,,\quad
-v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
-v_{bz}-\vph\simeq\left\{
-\begin{aligned}
-&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
-&\frac{1}{2\gamma_b} & T\gg m
-\end{aligned}
-\right.\,,\quad
-\gph\simeq1+\delta Z_x\,,\quad
-\cG' \sim \frac{m}{T}\,,\\
-\text{and}\quad w \sim 1\,,
-\end{gathered}
-\end{equation}
-\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
-where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
-(\ref{eq:OGgen}) suffices to show that there exists growth rates that
-are $\propto \Gamma_0$, with 
-\begin{equation}
-\Gamma_M
-\simeq
-\left\{
-\begin{aligned}
-& 0.38 \Gamma_0 & T\ll m\\
-& 0.34 \Gamma_0 & T\gg m
-\end{aligned}
-\right.\,,
-\end{equation}
-for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).
+%However, obtaining asymptotic expansions for the growth rate in the high and
+%low temperature limits that can be easily solved for the maximum
+%growth rate is difficult.  Even with the tremendous numerical insight
+%provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
+%significant progress.  However, some general remarks can be made.  It
+%appears at the maximum growth rate, we have to leading order:
+%\begin{equation}
+%\begin{gathered}
+%\vph\simeq\sqrt{2\delta Z_x}\,,\quad
+%v_{bx}=(1-\delta Z_x) v_b\,,\quad
+%v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
+%v_{bz}-\vph\simeq\left\{
+%\begin{aligned}
+%&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
+%&\frac{1}{2\gamma_b} & T\gg m
+%\end{aligned}
+%\right.\,,\quad
+%\gph\simeq1+\delta Z_x\,,\quad
+%\cG' \sim \frac{m}{T}\,,\\
+%\text{and}\quad w \sim 1\,,
+%\end{gathered}
+%\end{equation}
+%\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
+%where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
+%(\ref{eq:OGgen}) suffices to show that there exists growth rates that
+%are $\propto \Gamma_0$, with
+
+%\begin{equation}
+%\Gamma_M
+%\simeq
+%\left\{
+%\begin{aligned}
+%& 0.38 \Gamma_0 & T\ll m\\
+%& 0.34 \Gamma_0 & T\gg m
+%\end{aligned}
+%\right.\,,
+%\end{equation}
+%for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).  
+%
 

### 2016-05-10 oblique.tex
more polish in appendix, completed references

@@
-%\documentclass[usenatbib,iop,apj,numberedappendix]{emulateapj}
-\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2010}
+\documentclass[usenatbib,iop,apj]{emulateapj}
+%\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2010}
 
@@
 
-\title{The Plasma Physics of Dilute Ultrarelativistic \epm\ Pair Beams I:\\ Linear Theory}
+\title{Clarifications of the Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams}
 
@@
 Ewald Puchwein\altaffilmark{5},
+Astrid Lamberts\altaffilmark{6}
+Mohamad Shalaby\altaffilmark{1,2,7},
 }
@@
 \altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
-
+\altaffiltext{5}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
+\altaffiltext{6}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
+\altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
 \keywords{
@@
  where $E$ and $T$ are the dimensionless energy and temperature in terms of a particles rest mass.
-In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution ({\bf REFS}).
+In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution \citep{1911AnP...340..145J}.
 
 The relativistic Maxwellian distribution can be extended to a drifting (or boosted) distribution via an appropriate Lorentz transformation.
-The relationship between the energies of the lab (boosted) frame and the COM frame is (Jackson)
+The relationship between the energies of the lab (boosted) frame and the COM frame is
 \begin{equation}
@@
 \end{equation}
-Plugging this into (\ref{eq:relativistic maxwellian}), we find (Wright \& Hadley 1975)
+Plugging this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-Juttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
@@
 where $K_2$ is the 2nd order modified Bessel function.
-The transformation properties of the drifting Maxwell-Juttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-Juttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
+
+%The transformation properties of the drifting Maxwell-Juttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-Juttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
 
@@
 \acknowledgements
-
-We thank A.
-Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions.
-We also thank E.
-Zweibel for useful discussions.
-P.C.
-is supported in part by an UW-Milwaukee RGI grant and a \Fermi GI grant.
-C.P.
-gratefully
-acknowledges financial support of the Klaus Tschira Foundation. G.V. acknowledges support from CITA.\\
-
-\appendix
+A.E.B.~and M.S.~receive financial support from the Perimeter
+Institute for Theoretical Physics and the Natural Sciences and
+Engineering Research Council of Canada through a Discovery Grant.
+Research at Perimeter Institute is supported by the Government of
+Canada through Industry Canada and by the Province of Ontario through
+the Ministry of Research and Innovation.
+PC is supported by the UWM Research Growth Initiative, the NASA ATP
+program through NASA grant NNX13AH43G, and NSF grant AST-1255469.
+C.P.~gratefully acknowledges
+financial support of the Klaus Tschira Foundation. E.P. acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
+We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions.
+
+%\onecolappendix
+\appendix%[onecolumn]
+%\onecolumn
 
@@
  \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
-&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),\label{eq:dispersion-real-part}
+&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),
+\label{eq:dispersion-real-part}
 \end{eqnarray}
@@
 \begin{equation}
-\omega = \omega_{p,t}\left(1 + \frac 3 2 k^2\lambda_{D,t}^2\right).
+\omega \approx \omega_{p,t}\left(1 + \frac 3 2 k^2\lambda_{D,t}^2\right).
 \end{equation}
@@
 In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.
-This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.  With an eye toward computing the residue in (??), we define $p_{z'}=\gz\vz\epsilon_\perp$, $\epsilon=\gz\epsilon_\perp$, and
-$\epsilon_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
+This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.
+With an eye toward computing the residue that will appear in equation (\ref{eq:dispersion kinetic}), we define $p_{z'}=\gz\vz E_\perp$, $E=\gz E_\perp$, and
+$E_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
 \begin{eqnarray}
@@
 %F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
-&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})\epsilon_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
+&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
 \end{eqnarray}
-Plugging the equation into (??), we find that the this gives an associated growth rate of
-\begin{equation}
-\Gamma \simeq -\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R,
+where we define $\mu = m_ec^2/k_BT$.
+Plugging the equation into (\ref{eq:dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
+\begin{equation}
+ 1 - \frac{\omega_{p,t}^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right) + i\frac{\pi n_b}{n_t}   m_e v_B^2 R = 0, 
 \end{equation}
@@
 where $R$ is the residue for $p_{z'}$ such that $v_{z'} = v_{\rm ph}$.
-   Here we find for the two elements that contribute to the pole:
+
+We can assume that $k\lambda_D \ll 1$ as the thermal velocity of the background plasma is much smaller than the speed of the ultrarelativistic beam.
+We then take $\omega = \omega_r + i\Gamma$, where $\omega_r = \Re(\omega)$ is the real part of $\omega$ and the growth rate $\Gamma \ll \omega_r$, to find:
+\begin{equation}
+\Gamma \simeq -\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R.
+\end{equation}
+
+Here two elements contribute to the pole:
 \begin{equation}
@@
 - k \left(
-  \frac{c^2}{\epsilon} - \frac{p_{z'}^2 c^4}{\epsilon^3}
+  \frac{c^2}{E} - \frac{p_{z'}^2 c^4}{E^3}
 \right)
@@
 =
-- \frac{k c^2}{\gph^3 \epsilon_\perp}\,,
+- \frac{k c^2}{\gph^3 E_\perp}\,,
 \end{equation}
@@
 \left.
--\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{\epsilon} - v_{B,z'}\right)
+-\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{B,z'}\right)
 F_b
@@
   k
-  (\vph-v_{B,z'}) \mu m_e^{-3}c^{-3}}{
-  4\pi k_BT K_2(\mu)
-} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})\epsilon_\perp)} {k_BT c^2}\right)\\
-&\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
+  (\vph-v_{B,z'}) \mu^2}{
+  4\pi m_e^{4}c^{5}K_2(\mu)
+} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
 \end{aligned}
@@
 %\left.
-%\gamma_b(\epsilon-v_{bz} p_z -v_{bx} p_x)/T
+%\gamma_b(E-v_{bz} p_z -v_{bx} p_x)/T
 %\right|_{\rm pole}
 %=
-%\gamma_b \gph ( 1-v_{bz}\vph ) \epsilon_\perp/T - \gamma_b v_{bx} p_x/T\,.
+%\gamma_b \gph ( 1-v_{bz}\vph ) E_\perp/T - \gamma_b v_{bx} p_x/T\,.
 %\label{eq:Eopole}
@@
 \frac{
-  \gph^3 (\vph-v_{B,z'})\mu}{4\pi k_BT K_2(\mu)
+  \gph^3 (\vph-v_{B,z'})\mu^2}{4\pi K_2(\mu)
 }
-\frac{\kI}{m_e^3c^5}
+\frac{\kI}{m_e^4c^7}
 ,
@@
 \begin{equation}\label{eq:I-integral}
-\kI \equiv \int d^2\!p_\perp \epsilon_\perp \exp\left(-\frac{\cG(\epsilon_\perp-w p_x)}{k_B T}\right)
+\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T}\right)
 \end{equation}
@@
 \begin{equation}
-\cG \epsilon_\perp - \gamma_B v_{B,x'} p_{x'} = \gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'}p_{x'})|_{v_{B,z'}=v_{\rm ph}} > 0
+\cG E_\perp - \gamma_B v_{B,x'} p_{x'} = \gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'}p_{x'})|_{v_{B,z'}=v_{\rm ph}} > 0
 \end{equation}
@@
 
-Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w \epsilon_\perp)$ and $p_{y'}'=p_{y'}$ and find:
+Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w E_\perp/c^2)$ and $p_{y'}'=p_{y'}$ and find:
 \begin{eqnarray}
-%\epsilon'_\perp &=& \gamma_w(\epsilon_\perp - w p_x),\\
-\epsilon_\perp &=& \gamma_w(\epsilon'_\perp + w p'_x),\\
-d p_x d p_y &=& \frac{\epsilon_\perp}{\epsilon'_\perp} d p'_x d p'_y.
+%E'_\perp &=& \gamma_w(E_\perp - w p_x),\\
+E_\perp &=& \gamma_w(E'_\perp + w p'_x),\\
+d p_x d p_y &=& \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
 \end{eqnarray}
@@
 \kI &= 
-\int d^2\!p'_\perp \frac{\epsilon_\perp^2}{\epsilon'_\perp} \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\\
+\int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T}\right)\\
 %&=
 %\gamma_w^2 \int d^2\!p'_\perp \left(
- % \epsilon'_\perp + 2 w 
+ % E'_\perp + 2 w 
 %\qc{p'_x}
-%+ w^2 \frac{p'^2_x}{\epsilon'_\perp}
-%\right) e^{-\cG'\epsilon'_\perp/m}\\
+%+ w^2 \frac{p'^2_x}{E'_\perp}
+%\right) e^{-\cG'E'_\perp/m}\\
 &=
 \pi \gamma_w^2 \int_0^\infty dp'^2_\perp \left(
-  \epsilon'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{\epsilon'_\perp}
-\right) \exp\left(-\frac{\cG'\epsilon'_\perp}{k_B T}\right)\,,
+  E'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{E'_\perp}
+\right) \exp\left(-\frac{\cG'E'_\perp}{k_B T}\right)\,,
 \end{aligned}
@@
 % and in the second line we used the fact
-%that  $\epsilon'_\perp e^{-\cG'\epsilon'_\perp/m}$ is isotropic in
+%that  $E'_\perp e^{-\cG'E'_\perp/m}$ is isotropic in
 %$\bp'_\perp$.  

### 2016-05-17 oblique.tex
added text for temperature and angles

@@
 
-\title{Clarifications of the Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams}
+\title{The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams}
 
@@
 \right]
-\exp(-\cG'\mu),
-\end{equation}
-where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
-
-In Figure \ref{fig:ObliqueGMZ}, we plot the maximum growth rate (maximized over $\vph$) as function of $v_{bx} = |k_{bx} v_b/k|$ and $T$.
-From the figure, it is clear that the growth rate is maximized for $v_{bx}/v_b \gg 1/\gamma_b$.
-Moreover the maximum growth rates vary little and are robust for broad range of $v_{bx}/v_b = k_{bx}/k$, i.e., large range in angles between beam and wavevector.
-Hence, this suggests that we might estimate the maximal oblique growth rates by taking the limit as $v_{bx}\gg v_b/\gamma_b$ and maximizing over $\vph$.
-This is further seen in Figure \ref{fig:OGgen}, we maximize over both $v_{bx}$ and $\vph$.
+\exp(-\cG'\mu),\label{eq:kinetic growth rate}
+\end{equation}
+where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{B,x'} = 0$, which gives:
+\begin{equation}
+\Gamma_{\rm bp} \simeq - \Gamma_0
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
+\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
+\exp(-\cG'\mu),\label{eq:bp kinetic growth rate}
+\end{equation}
+
+\begin{figure*}
+\includegraphics[width=0.9\textwidth]{growthrates.png}
+\caption{Oblique growth rate maximized over $\vph$ as a function of
+  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
+  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
+  cases, the maximum growth rate occurs when
+  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
+\end{figure*}
+
+In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of $\cos\theta = \hat{k}\cdot\vel$, i.e., the angle between the beam and the wave vector, and $k_BT/m_e c^2$.
+Here, it is clear that the growth rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$, i.e., at an oblique, but not perpendicular angle.
+Moreover the maximum growth rates vary little and are robust for broad range of angles between the wavevector and the beam direction.
+Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_B - 1$.
 Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.
@@
 for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.
-
-\begin{figure*}
-\includegraphics[width=0.9\textwidth]{growthrates.png}
-\caption{Oblique growth rate maximized over $\vph$ as a function of
-  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
-  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
-  cases, the maximum growth rate occurs when
-  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
-\end{figure*}
 
@@
 \end{figure}
+
+This can be contrasted with Figure \ref{fig:bpgen} where we plot the maximized beam-plasma growth rate  (\ref{eq:bp kinetic growth rate}) as a function of $\gamma_b - 1$ and $k_B T/m_ec^2$.
+Here, we see that the maximum growth rate is somewhat sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for $k_BT/m_ec^2 \ll 1$ to $\Gamma/\Gamma_0 \approx 0.1$ for $k_BT/m_ec^2 \gtrsim 1$.
+Note, however, that that beam-plasma growth rate remains competitive with the oblique growth rate, i.e., it is not orders of magnitude lower.
 

### 2016-06-04 oblique.tex
final changes

@@
 Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
-Taking the limit of Equation (\ref{eq:distribution function beam}) as $\alpha_{\parallel}^{-1}, \alpha_{\perp}^{-1} \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
+Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
 \begin{equation}\label{eq:dispersion relation reactive}
@@
 For the case where $k_z \rightarrow 0$, the electrostatic approximation no longer holds and the full dispersion relation must be solved.\footnote{We thank Antoine Bret for helping to clarify this point.}  A solution to the full dispersion relation reveals additional modes, including the zero frequency ($k_z = 0$) filamentation or Weibel mode.
+
+\subsubsection{Sub-maximal instabilities}
+
+Equation (\ref{eq:dispersion relation reactive}) can also be solved numerically in terms of $k_x$ and $k_z$.
+Here let us specialize to the case of $k_x = 0$, i.e., the beam-plasma case.
+In this case, we have
+\begin{equation}\label{eq:dispersion relation beam-plasma}
+ 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2} = 0,
+\end{equation}
+which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $v_b/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.
+In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $v_b/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.
+For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is plasma oscillation frequency.
+This wave would exist in the absence of a tenuous beam.
+However, as we move away from this frequency toward lower $k_z$, we still find substantial growth, with $\Gamma \approx 0.1 \Gamma_{\rm max}$ as $k_z \approx 0.9\lambda_D^{-1}$.
+Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.
+This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed system.
+
+In a continuous system, these waves do not matter compare to the unstable mode at $k\lambda_D = 1$.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state compare to the physical system.
+
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{off_res.pdf}
+\caption{Beam-plasma growth rate (solid line) and unstable wave frequency (dashed-line) as a function of
+  $k_z$ for $\gamma_b = 100$ and $n_b/n_t=10^{-3}$. \label{fig:OffResonance}}
+\end{figure}
 
@@
 
-\begin{figure*}
-\includegraphics[width=0.9\textwidth]{growthrates.png}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{pp1.pdf}
 \caption{Oblique growth rate maximized over $\vph$ as a function of
@@
   $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
-\end{figure*}
+\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics[width=0.5\textwidth]{oblique_GammaM.eps}
-\caption{Oblique kinetic growth rate, maximized over $v_{bx}/v_b$ and
-  $\omega/k$, normalized by
+\begin{figure*}
+\includegraphics[width=0.5\textwidth]{pp2.pdf}
+\includegraphics[width=0.5\textwidth]{pp3.pdf}
+
+%\plottwo{pp2.pdf}{pp3.pdf}
+\caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right), maximized over $\theta$ and normalized by
   $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function
-  of $T$ (in units of $m$) and $\gamma_b$.  Note that unlike the beam
+  of $k_BT/m_ec^2$ and $\gamma_b$.  Note that unlike the beam
   plasma case shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the
@@
   constituting a roughly $10\%$ reduction.} \label{fig:OGgen}
-\end{figure}
+\end{figure*}
 
@@
 
-Applying these results to the case of ultrarelativistic beams and given our results from Section \ref{sec:setup}, we find
+In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale in a Maxwell-Junter distribution.  The relevant results are: 
+\begin{equation}
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 },
+\end{equation}
+and 
+\begin{equation}
+\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
+\end{equation}
+where $T$ is measured in the COM frame of the beam.
+
+Applying these results, we find
 \begin{eqnarray}
-\frac{\Delta v_{\perp}}{v_b} &\sim&  \sqrt{\alpha_{\perp}} = \sqrt{\alpha'_{\perp}} \gamma^{-1},\\
-\frac{\Delta v_{\parallel}}{v_b} &\sim& \sqrt{\alpha_{\parallel}} = \sqrt{\alpha'_{\parallel}}\gamma^{-2},
+\frac{\Delta v_{\perp}}{v_b} &\sim&  \gamma_b^{-1}\sqrt{\frac{k_BT}{m_ec^2}},\\
+\frac{\Delta v_{\parallel}}{v_b} &\sim& \gamma_b^{-2}\sqrt{\frac{k_BT}{m_ec^2}}.
 \end{eqnarray}
-where $\alpha'_{\perp}$ and $\alpha'_{\parallel}$ encodes the dependence on the COM velocity dispersion.
 Hence, the conditions for the reactive regime for the ``oblique'' mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
@@
  \begin{array}{cl}
-\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
-\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+ ({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
+({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:growth-rate-bp}]) and ``oblique'' mode (eq.[\ref{eq:growth-rate-oblique}]).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:bp kinetic growth rate}]) and ``oblique'' mode (eq.[\ref{eq:kinetic growth rate}]).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.
@@
 \begin{array}{cl}
-\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
-\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
+({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 \end{eqnarray}
-where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta$ is.
+where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta = 4.5$ for $z \leq 1$ and $\zeta = 0$ for $z\geq 1$.
 
@@
 \begin{eqnarray}
-\frac{\gamma_b^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 7\times 10^{-3} \frac{\gamma_6^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
+\frac{\gamma_b^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 \begin{eqnarray}
-\frac{\gamma_b}{\sqrt{\alpha_{\parallel}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 0.7 \frac{\gamma_6}{\sqrt{\alpha_{\parallel}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
+\frac{\gamma_b}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability can also sit in the reactive regime for sufficiently cold beams $\alpha_{\parallel} \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability sits in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
 
@@
 This is a part of ongoing work and will be presented in a future publication.
-Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{\alpha_{\parallel}'}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
+Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
 Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
@@
 \begin{itemize}
-\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.
+%\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.
 We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam.
@@
 }
-\item{Using this simplified distribution function, we have analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.
+\item{We analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.
 We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the ``oblique'' mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
-\item{We have also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.
+\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.
 In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the ``oblique'' mode.
@@
 }
-\item{Using this average COM energy, we presume that the pair beam can be modelled as a relativistic Maxwellian in the COM frame.
-We then demonstrate that the ``oblique'' mode is in the kinetic regime.
-However, the beam plasma mode is marginally in the reactive regime, so that the use of either regime produces roughly the same growth rate.
+\item{We demonstrate that the ``oblique'' mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
+In any case, the use of either regime produces roughly the same growth rate.
 Moreover, the inclusion of plasma instability losses on the number density of beam particles may place the beam plasma mode firmly in the kinetic regime.}
@@
 
-The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.
+%The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.
 In particular, by developing a simple distribution function, whose modes can be analytically calculated, we will develop a quasilinear theory for these modes in a forthcoming publication.
@@
 
-%\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
-%\label{sec:lorentz}
-%
-%Here we explicitly derive the scaling of the parallel and
-%  perpendicular velocity dispersions with the Lorentz factor upon boosting the
-%  distribution function to the lab frame. Let us begin with a distribution
-%function that is isotropic in the COM frame and depends only on energy.
-%Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
-%function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
-%(denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
-%i.e., total number, or
-%\begin{equation}
-% \int \fLAB \dVLAB = \int \fCM \dVCM.
-%\end{equation}
-%It is well-known that under Lorentz transformations (Landau), 
-%\begin{equation}
-% \dVLAB = \dVCM,
-%\end{equation}
-%so therefore,
-%\begin{equation}
-%  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
-%\end{equation}
-%
-%Now let us consider moments of the distribution function.
-For clarity, it is helpful to consider moments of the distribution function first in the COM frame.
-The velocity moment is (unsurprisingly):
-%\begin{equation}
-% \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
-%\end{equation}
-%We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$.
+\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
+\label{sec:lorentz}
+
+Here we explicitly derive the scaling of the parallel and
+perpendicular velocity dispersions with the Lorentz factor upon boosting the
+distribution function to the lab frame. Let us begin with a distribution
+function that is isotropic in the COM frame and depends only on energy.
+Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
+function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
+(denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
+i.e., total number, or
+\begin{equation}
+ \int \fLAB \dVLAB = \int \fCM \dVCM.
+\end{equation}
+It is well-known that under Lorentz transformations (Landau), 
+\begin{equation}
+ \dVLAB = \dVCM,
+\end{equation}
+so therefore,
+\begin{equation}
+  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
+\end{equation}
+
+Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
+\begin{equation}
+ \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
+\end{equation}
+We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$.
 More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame.
 This gives:
-%\begin{eqnarray}\label{eq:betabarLAB}
-% \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
-%&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
-%\end{eqnarray}
-%Breaking the components of \betabarLAB\ into components parallel and
-%perpendicular to the boost, we find: 
-%\begin{eqnarray}
-%\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
-%\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
-%\end{eqnarray}
-%where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
-%For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
-%However, this Galilean result no longer holds for relativistic motion.
-%
-%Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
-%\begin{equation}
-% \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
-%\end{equation}
-%In the lab frame, it is again useful to break it into components -- the parallel component becomes 
-%\begin{eqnarray}
-%  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
-% &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
-% &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
-%\end{eqnarray}
-%while the perpendicular component becomes
-%\begin{eqnarray}
-% \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-%&=& \gamma_{\rm B}^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
-%\label{eq:dbetabarLABperp}
-%\end{eqnarray}
-%where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.  
-%
-%It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
-%\begin{equation}
-% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2}.
-%\end{equation}
-%This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
+\begin{eqnarray}\label{eq:betabarLAB}
+ \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
+&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
+\end{eqnarray}
+Breaking the components of \betabarLAB\ into components parallel and
+perpendicular to the boost, we find: 
+\begin{eqnarray}
+\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
+\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+\end{eqnarray}
+where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
+For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
+However, this Galilean result no longer holds for relativistic motion.
+
+Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
+\begin{equation}
+ \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
+\end{equation}
+In the lab frame, it is again useful to break it into components -- the parallel component becomes 
+\begin{eqnarray}
+  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
+ &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
+ &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
+\end{eqnarray}
+while the perpendicular component becomes
+\begin{eqnarray}
+ \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
+&=& \gamma_{\rm B}^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
+\label{eq:dbetabarLABperp}
+\end{eqnarray}
+where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.
+
+It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
+\begin{equation}
+ \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 }.
+\end{equation}
+This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
 This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
 Let us now consider the parallel component (eq.(\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$:
-%\begin{equation}
-% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}},
-%\end{equation}
-%Here, the scaling of the parallel velocity dispersion can be understood as a double scaling of both time and coordinate (along the boost axis) between same two frames boosted relative to each other, giving an extra scaling of $\gamma^{-2}$.
-%This scaling of the parallel component of the velocity dispersion has important consequences that we explore in the main part of the paper.  
-%
-%
+\begin{equation}
+ \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}}  \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
+\end{equation}
+Here, the scaling of the parallel velocity dispersion can be understood as a double scaling of both time and coordinate (along the boost axis) between same two frames boosted relative to each other, giving an extra scaling of $\gamma^{-2}$.
+This scaling of the parallel component of the velocity dispersion has important consequences that we explore in the main part of the paper.
 

### 2016-06-28 oblique.tex
updated intro

@@
 
+This excess heating of the IGM may resolve a variety
+of cosmological puzzles, including naturally explaining
+  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
+\citep{paperIV} and potentially explaining a number of the X-ray
+  properties of groups and clusters and anomalies in galaxy formation on
+  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
+  have recently shown that \textit{if} the IC halos are ignored, it is possible
+  to quantitatively reproduce the redshift and flux distributions
+  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
+  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
+  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
+  circumstantial evidence for the presence of virulent plasma beam
+  instabilities.
+
 The \Fermi satellite and ground based imaging atmospheric Cerenkov telescopes such as
@@
 blazars with a minority population of other sources  such as radio galaxies
-and starburst galaxies.  These VHEGR observations have proven to be useful in constraining mechanisms 
-of particle acceleration
-\citep[see, e.g., ][]{Pagl_etal:96,Domi-Torr:05,Thom-Quat-Waxm:07,Pers-Reph-Arie:08,deCe-Torr-Rodr:09,Reph-Arie-Pers:10,Lack_etal:10} and
-black hole jets
-\citep[see, e.g., ][]{Jones+1974,Ghisellini+1989,Ghisellini+2008,Tavecchio+2008,Ghisellini+2009}.
-% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
-% and intergalactic magnetic fields (IGMF)
-% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
-
-It has recently been argued by a number authors
-that the {\it non-observation} of $\sim 1-100\,\GeV$ flux by the \Fermi satellite toward these TeV blazars provides lower bounds on the strength of the  intergalactic magnetic field (IGMF) of $B\gtrsim 10^{-19}-10^{-15}\,\G$ \citep{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
-To understand how these claimed limits arise, we begin by revisiting the physics of VHEGR propagation in the universe.
-
-As these VHEGR photon propagate through the universe, they pair produce off the extragalactic background light (EBL) and produce ultrarelativistic electron-positron beams  with typical Lorentz factors of $10^5$--$10^7$.
-Within the current epoch, the mean free path for these photons is $\Dpp\sim 800 \,E_{\TeV}^{-1}\,\Mpc$, where $E_{\TeV} = E/1\,\TeV$ is the energy of the photon in \TeV.
-At high redshifts, $z\geq 1$, this drops to $D_{\rmn{pp}}\sim 35\, E_{\TeV}^{-1}\,\Mpc$ due to the larger number density of EBL photons.
-Because these typical mean free paths are much shorter than the Hubble length, the universe is optically thick to VHEGRs.
-This effect is seen in the relative nearness of TeV blazars ($z\sim 0.1$) and the observed softening of their VHE spectra with increasing distance.
-
-Previously, it has been  assumed that the resultant electron-positron pairs lose energy through inverse-Compton scattering the cosmic microwave background (CMB) and EBL.
-This yields an inverse-Compton cascade (ICC) that reprocesses TeV emission to gamma-rays with energies $\lesssim 100 \GeV$, for which the mean free path is comparable to the size of the universe.
-Hence, associated with every VHEGR source should be a $\sim 100\,\GeV$ source.
-It is the {\it non-observation} of this $100\,\GeV$ flux, presumably due to the magnetic deflection of the pairs, that is used to constraint the strength of a putative IGMF.
+and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
+(EBL) and produce pairs \citep[see, e.g.,][]{Goul-Schr:67,Sala-Stec:98,Nero-Semi:09}.
+
+It has been assumed that these ultrarelativistic pairs produced by VHEGR annihilation 
+lose energy exclusively through inverse-Compton (IC) scattering off of the cosmic microwave
+background (CMB),   transferring the energy of the original VHEGR to gamma-rays with
+energies $\lesssim100\,\GeV$.  The absence of observed secondary IC emission leads a number of
+authors to argue that this \textit{lack} places lower bounds upon the
+intergalactic magnetic field
+\citep[IGMF; see, e.g.,][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Taka_etal:11,Dola_etal:11} with 
+typical numbers ranging from $10^{-19}\,\G$ to $10^{-15}\,\G$.
 
@@
 
-Ultimately, these conclusions rest upon the assumption that ICCs dominate the evolution of the ultra-relativistic pairs.  
-However, we have strongly challenged this assumption in a recent series of papers, \citet[][hereafter paper I, II, and III respectively]{BCP,CBP, PCB}.
-In particular, we find that ultrarelativistic pair beam that results from TeV photon annihilation on the EBL is subject to virulent plasma beam instabilities that sap the energy of the pair beam and convert it to heat in the intergalactic medium (IGM) before ICCs have a chance to act.
-
-In Paper I, we showed that despite the extraordinarily dilute nature of this ultra-relativistic pair plasma, a variety of plasma instabilities grow on timescales short in comparison to the inverse-Compton cooling time.
-Of the three instabilities that we studied --- the beam-plasma or two-stream instability\footnote{In order to follow the plasma literature more closely, henceforth we will refer to what we have called the ``two-stream'' instability as the ``beam-plasma'' instability}, the Weibel instability, and the ``oblique'' instability  ---  the most
-important of which is the ``oblique'' instability \citep{Bret-Firp-Deut:04}
-Because the pairs lose their energy primarily via depositing it as heat into the IGM, the ICCs critical to the above arguments commonly employed to provide constraints upon the IGMF and high-redshift blazar population are necessarily precluded.
-In the latter case in particular, the \Fermi limits on the EGRB and high-energy blazar statistics allow a large and evolving blazar population more similar to that found for other AGN.
-
-Assuming that these plasma-instability mechanisms drive ultra-relativistic pairs to deposit their energy as heat in
-the IGM, we explored the impact of this heating on the
-thermodynamics of the IGM in Paper II and the effect on this additional heating on structure formation in the universe in Paper III.  
-By correcting for the selection effects of the current pencil beam VHEGR observations using
-the all-sky monitoring of the \Fermi satellite, we argued that 
-luminosity density in VHEGRs is of order $0.2\%$ of the quasar
-luminosity density in Paper II.  Because of the greater conversion efficiency of TeV photon energy to heat (as compared to photoheating), we showed that this heating dominates the
-photoheating rate at low $z$.
-This steady volumetric homogeneous heating naturally produces an inverted temperature-density relation and, with only a minor rescaling of the empirically determined TeV blazar
-heating normalization, reproduces the inferred
-inverted temperature-density relation at $z=2-3$ \citep{Bolton+09,Viel+09}, a feature that standard reionization models have proven unable to reproduce \citep{McQuinn+09,Bolton+09}.
-
-Following the work of Paper II, we studied the effects of blazar heating in a
-more detailed hydrodynamic model of structure formation 
-\citep{Puchwein+2011}. Using the same blazar heating prescription as in Paper
-II, we showed that the optical depth of the \Lya\ forest is reproduced using a
-H{\sc i} photoionization rate of $\Gamma_{\rm H\mathsc{i}} \approx 5\times
-10^{-13}\,{\rm s}^{-1}$ or equivalently using the inferred ionizing background
-from \citet{FG+09}.  We also demonstrated that the low-density IGM again
-possesses an inverted temperature-density relation, confirming the results of
-Paper II in greater detail.  In addition, we showed that a blazar heated
-universe not only matches the one- and two-point statistics of the
-  high redshift Ly$\alpha$ forest but also the linewidth and column depth
-  distribution. This excellent quantitative agreement is achieved
-without tuning the photoionizing background and is due specifically to the
-excess energy injection into the low-density IGM from TeV blazars.
-
-Finally in Paper III, we explored the result of this additional IGM heating upon
-the formation of structure in the Universe.  In particular, we showed that
-the injection of entropy into the IGM by TeV blazars contributes to
-developing a redshift dependent entropy floor for galaxy clusters and groups at
-$z\lesssim 2$ and suppresses the formation of dwarfs.
-The redshift dependent nature of TeV blazar heating suggests a large injection of entropy around $z\sim 1$, boosting the entropy of late forming groups and clusters.
-This predicted enhanced entropy
-of young groups is consistent with recent observations 
-that show optically bright, and therefore young, groups and
-clusters are X-ray dim ---  that is having a lower gas density due to a
-raised entropy floor.  We also found that TeV
-blazar heating suppresses the formation of late forming dwarfs both in
-galactic halos, which is potentially important for the missing satellite problem \citep{Krav:10}, and
-in voids, which is likewise potentially important for the void phenomenon \citep{Peebles2001}, by raising the
-temperature of the IGM such that gas cannot collapse to form galaxies.
-
-These potential implications of blazar heating rely on the assumption that the nonlinear state of these plasma instabilities efficiently convert the kinetic energy of the beam and convert it into heat.
-In our previous work, we appealed to nonlinear simulations conducted in a radically different regime to argue for such a conversion.
+These two important conclusions depends on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
+ultrarelativistic pair beams  likely is the dominant cooling mechanism
+\citep{paperI,Schlickeiser+12,Schlickeiser+13}, depositing this energy
+as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
+ IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
+\citep{paperI,Schlickeiser+12,Schlickeiser+13}.
+
+% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
+% and intergalactic magnetic fields (IGMF)
+% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
+
+The deposition of the kinetic energy into the IGM via plasma instabilities produces excess heating, which over cosmological time, may resolve a variety
+of puzzles, including explaining
+anomalies in the statistics of the high-redshift Ly$\alpha$ forest
+\citep{paperIV} and potentially explaining a number of the X-ray
+properties of groups and clusters and anomalies in galaxy formation on
+the scale of dwarfs \citep{paperIII,Lu+2013}.  We
+have recently shown that \textit{if} the IC halos are ignored, it is possible
+to quantitatively reproduce the redshift and flux distributions
+of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
+spectrum above 3~GeV simultaneously with a unified model of AGN evolution
+\citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
+circumstantial evidence for the presence of virulent plasma beam
+instabilities.
+
+These potential implications of blazar heating rely on an understanding of the linear and nonlinear physics of these plasma instabilities.
+Recent work in this area has been inconclusive.
+For instance, \citet{Miniati+12} argued that these instabilities are physically irrelevant for the cooling of these pair beams because they would saturate at a very low level due to nonlinear Landau damping (NLD).
+However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.
+In addition, \citet{Siron+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
+It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g.
+Mohamad's paper).
+
 However, a full nonlinear study is called for.
@@
 In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.
-We model this distribution as a relativistic thermal distribution:
+This equilibrium energy distribution of a relativistic thermal plasma gas is
 \begin{equation}\label{eq:relativistic maxwellian}
@@
 \end{eqnarray}
-where $\delta\jvec = \sum_s q_s\int \vel_s \delta f_s d^3 p$ is the linear current density perturbation.
+where $\delta\jvec = \sum_s q_s\int \vel \delta f_s d^3 p$ is the linear current density perturbation.
 
@@
 which we denote the beam-plasma or ``two-stream'' growth rate.
-For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.  Indeed for $\gamma \gg 1$ and $Z_x \sim 1$, the growth rate approaches oblique growth rate:
+For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.
+Indeed for $\gamma \gg 1$ and $Z_x \rightarrow \infty$, the growth rate approaches oblique growth rate:
 \begin{equation}
@@
 Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.
-This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed system.
+This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed target plasma.
 
@@
 Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate (\ref{eq:growth rate reactive}).
-For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability \citep{Boyd}
+For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability:
 \begin{equation}\label{eq:bp reactive regime}

### 2016-06-30 oblique.tex
updated bib

@@
 However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.
-In addition, \citet{Siron+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
+In addition, \citet{Sironi+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
 It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g.
@@
 \end{equation}
-For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{BCP}.
+For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{paperI}.
 
@@
 Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
-Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{Puchwein+2011}.
+Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
 

### 2016-07-01 oblique.tex
edited through Sect. 1

@@
 \newcommand\ac[1]{{\color{green} \bf #1}} %% AEB color
+\newcommand\cp[1]{{\color{blue} \bf #1}} %% CP color
 \newcommand\asout[1]{\color{green}\sout{{\rm  #1}}} %% AEB strikeout of text
@@
 Starting from an isotropic distribution in the center of mass frame, we show that the velocity dispersions parallel and transverse to the direction of beam propagation become highly anisotropic under Lorentz boosts and scale differently with the Lorentz factor, $\gamma$.
-Using these results to construct a simplified distribution function, we then calculate the linear growth rate of the beam plasma and ``oblique'' instability in both the reactive and kinetic regimes.
-We reproduce the well-known reactive growth rates for both the beam plasma and ``oblique'' mode and show that they both have similar growth rates in the kinetic regime.
+Using these results to construct a simplified distribution function, we then calculate the linear growth rate of the beam plasma and oblique instability in both the reactive and kinetic regimes.
+We reproduce the well-known reactive growth rates for both the beam plasma and oblique mode and show that they both have similar growth rates in the kinetic regime.
 In contrast to the earlier claim by the authors that the oblique mode is dominant, we instead find that both the beam-plasma (parallel) and oblique modes are similarly strong.
-In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the ``oblique'' mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
-Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we not only confirm that these beams are unstable to the kinetic ``oblique'' mode.
-For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the ``reactive'' beam-plasma instability.
-We show that overall growth rates for both the ``oblique'' and beam plasma growth rates are similar and in line with our earlier claims.
+In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
+Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we not only confirm that these beams are unstable to the kinetic oblique mode.
+For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the reactive beam-plasma instability.
+We show that overall growth rates for both the oblique and beam plasma growth rates are similar and in line with our earlier claims.
 These results strengthen the case that powerful plasma instabilities  sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space, which has profound consequences on constraints of the intergalactic magnetic field, the thermal state of the intergalactic medium, and the formation of structure in the universe.
@@
 
-This excess heating of the IGM may resolve a variety
-of cosmological puzzles, including naturally explaining
-  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
-\citep{paperIV} and potentially explaining a number of the X-ray
-  properties of groups and clusters and anomalies in galaxy formation on
-  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
-  have recently shown that \textit{if} the IC halos are ignored, it is possible
-  to quantitatively reproduce the redshift and flux distributions
-  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
-  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
-  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
-  circumstantial evidence for the presence of virulent plasma beam
-  instabilities.
-
-The \Fermi satellite and ground based imaging atmospheric Cerenkov telescopes such as
+%CP: this appears twice - I cut the following:
+%This excess heating of the IGM may resolve a variety
+%of cosmological puzzles, including naturally explaining
+%  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
+%\citep{paperIV} and potentially explaining a number of the X-ray
+%  properties of groups and clusters and anomalies in galaxy formation on
+%  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
+%  have recently shown that \textit{if} the IC halos are ignored, it is possible
+%  to quantitatively reproduce the redshift and flux distributions
+%  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
+%  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
+%  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
+%  circumstantial evidence for the presence of virulent plasma beam
+%  instabilities.
+
+The \Fermi satellite and ground-based imaging atmospheric Cerenkov telescopes such as
 % Phil: I added "imaging atmospheric" to distinguish this class of
@@
   Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
-  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the ultra-high
+  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the \cp{\sout{ultra-}}high
 energy Universe is teeming with energetic very high-energy gamma-ray 
 (VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
-blazars with a minority population of other sources  such as radio galaxies
+blazars with a minority population of other sources  such as radio\cp{- \sout{galaxies}}
 and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
@@
 \citep[see, e.g., ][]{Naru-Tota:06,Knei-Mann:08,Inou-Tota:09,Vent:10}.
-These constraints preclude any dramatic rise in numbers of source by $z\sim1$--$2$ that seen in the quasar
+These constraints preclude any dramatic rise in numbers of source by $z\sim1$--$2$ that is seen in the quasar
 distribution.  That is, the comoving number of blazars must have
@@
 underlying these systems and with the observed evolution of similarly
-accreting systems, i.e., quasars.
-
-These two important conclusions depends on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
-ultrarelativistic pair beams  likely is the dominant cooling mechanism
+accreting systems, i.e., quasars and radio galaxies.
+
+These two important conclusions depend on IC cooling dominating the evolution of the ultra-relativistic pairs.  However, it was recently found that plasma instabilities driven by the
+ultrarelativistic pair beams  likely are the dominant cooling mechanisms
 \citep{paperI,Schlickeiser+12,Schlickeiser+13}, depositing this energy
@@
 
-The deposition of the kinetic energy into the IGM via plasma instabilities produces excess heating, which over cosmological time, may resolve a variety
+The deposition of kinetic energy into the IGM via plasma instabilities produces excess heating, which over cosmological time, may resolve a variety
 of puzzles, including explaining
@@
 In addition, \citet{Sironi+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
-It is unclear, however, if the conclusions of their work is robust as it is unclear if these numerical setup is adequate (e.g.
+It is unclear, however, if the conclusions of their work are robust as it is unclear if these numerical setup is adequate (e.g.
 Mohamad's paper).
-
-However, a full nonlinear study is called for.
+\cp{[This last sentence needs to be weakened or supported with a reference, e.g.
+by Mohamad's paper provided it is out by then.]}
+
+\cp{To this end\sout{However}}, a full nonlinear study is called for.
 We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.
@@
 On the other hand, in the kinetic regime, only a fraction of the beam particles are resonant with the beam over a the growth time of the instability, which reduces the growth rate compared to the reactive instability for the same beam density and beam Lorentz factor.
-We recover the well-known results for the reactive regime for both the beam-plasma and ``oblique'' modes.
+We recover the well-known results for the reactive regime for both the beam-plasma and oblique modes.
 We also derive the growth rate for these two instabilities in the kinetic regime and delineate the range of applicability for both the reactive and kinetic cases and apply these results for ultrarelativistic \epm\ pair beams.
@@
 We then calculate the various linear instabilities that this beam is subject to in section \ref{sec:linear}.
-In particular, we pay careful attention to both the reactive (or hydrodynamic) and kinetic regimes of the beam plasma and ``oblique'' instabilities and the transition between the two.
-Applying these result to TeV electron-positron pair beams that arise from TeV photon pair production in Section \ref{sec:application},
-we demonstrate the despite the extraordinary coldness of the beam that we are always in the kinetic regime for the ``oblique'' mode, but may be in the reactive regime for the beam plasma mode.
+In particular, we pay careful attention to both the reactive (or hydrodynamic) and kinetic regimes of the beam plasma and oblique instabilities and the transition between the two.
+Applying these results to TeV electron-positron pair beams that arise from TeV photon pair production in Section \ref{sec:application},
+we demonstrate that despite the extraordinary coldness of the beam \cp{\sout{that}} we are always in the kinetic regime for the oblique mode, but may be in the reactive regime for the beam plasma mode.
 However, because the controlling parameter is close to unity, the growth rates calculated in either regime are similar.
@@
 % where $\alpha'_{\perp} = \gamma^2\alpha_{\perp}$ is an O(1) factor.
-Again, we find that the overall scaling of the kinetic ``oblique'' instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.
+Again, we find that the overall scaling of the kinetic oblique instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.
 Also we have made the assumption that $k_x/k_z \gg 1/\gamma$.
@@
 % 
-% Figure \ref{fig:oblique-growth-rate} also demonstrates that the ``oblique'' mode follows its name in that the fastest 
+% Figure \ref{fig:oblique-growth-rate} also demonstrates that the oblique mode follows its name in that the fastest 
 % growing mode is neither perfectly parallel ($k_x/k_z = 0$) nor perpendicular ($k_x/k_z = \infty$), but rather at a k-vector 
@@
 \end{eqnarray}
-Hence, the conditions for the reactive regime for the ``oblique'' mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
+Hence, the conditions for the reactive regime for the oblique mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
 \begin{equation}\label{eq:generic reactive regime}
@@
  \begin{array}{cl}
- ({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
+ ({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
 ({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:bp kinetic growth rate}]) and ``oblique'' mode (eq.[\ref{eq:kinetic growth rate}]).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:bp kinetic growth rate}]) and oblique mode (eq.[\ref{eq:kinetic growth rate}]).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.
@@
 which connects with the condition on the reactive beam plasma instability from Equation (\ref{eq:bp reactive regime}).
-Similarly for the ``oblique'' mode, the bandwidth, $\Delta \omega$, is then greater than the growth rate if 
+Similarly for the oblique mode, the bandwidth, $\Delta \omega$, is then greater than the growth rate if 
 \begin{equation}\label{eq:oblique kinetic regime}
@@
 \end{equation}
-which can similarly compared to the condition on the reactive ``oblique'' mode from Equation (\ref{eq:oblique reactive regime}).
+which can similarly compared to the condition on the reactive oblique mode from Equation (\ref{eq:oblique reactive regime}).
 
@@
 \begin{array}{cl}
-({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
+({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
 ({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
@@
 
-As discussed in the Introduction, the annihilation of VHEGRs and EBL photons produce ultrarelativistic \epm\ beams that are unstable to the beam plasma and ``oblique'' modes discussed above.
+As discussed in the Introduction, the annihilation of VHEGRs and EBL photons produce ultrarelativistic \epm\ beams that are unstable to the beam plasma and oblique modes discussed above.
 To apply the above results to the ultrarelativistic \epm\ beams, we now calculate their initial conditions.
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability sits in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
-
-The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.
-However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability sits in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+
+The discussion above confirmed our earlier estimate in Paper I that the oblique instability sits in the kinetic regime.
+However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.
 However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
@@
 Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
-Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
+Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
 
@@
 
-The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and ``oblique'' instability.
+The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and oblique instability.
 In this work, we have these linear instabilities as they would apply to the ultrarelativistic pair beams.
@@
 }
-\item{We analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.
-We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the ``oblique'' mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
+\item{We analytically calculated the beam-plasma and oblique instabilities in both the reactive and kinetic regime.
+We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the oblique mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
-\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.
-In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the ``oblique'' mode.
+\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes for both are different.
+In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the oblique mode.
 This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.}
@@
 }
-\item{We demonstrate that the ``oblique'' mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
+\item{We demonstrate that the oblique mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
 In any case, the use of either regime produces roughly the same growth rate.
@@
 program through NASA grant NNX13AH43G, and NSF grant AST-1255469.
-C.P.~gratefully acknowledges
-financial support of the Klaus Tschira Foundation. E.P. acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
+C.P.~gratefully acknowledges support by the European Research Council under ERC-CoG grant CRAGSMAN-646955 and by the Klaus Tschira Foundation.
+E.P. acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
 We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions.
@@
 % where $\alpha'_{\parallel} = \gamma^4\alpha_{\parallel}$ is the rescaled factor such that it is order unity.
-% As $\Delta\beta_z \sim \sqrt{\alpha_{\parallel}}$, Equation (\ref{eq:growth-rate-bp}) has an overall scaling like $\gamma\nb/n_t$ in line scalings found for the kinetic ``oblique'' mode quoted in Paper I (also see below).
+% As $\Delta\beta_z \sim \sqrt{\alpha_{\parallel}}$, Equation (\ref{eq:growth-rate-bp}) has an overall scaling like $\gamma\nb/n_t$ in line scalings found for the kinetic oblique mode quoted in Paper I (also see below).
 % 

### 2016-07-01 oblique.tex
fixed numbering issue of Appendix and equation formatting

@@
-\documentclass[usenatbib,iop,apj]{emulateapj}
-%\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2010}
+%\documentclass[usenatbib,iop,apj]{emulateapj}
+\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2015}
 
@@
 
-As stated in the Introduction, VHEGR photons pair produce on encountering EBL photons as they propagate throughout the universe \citep{Gould+66}, and this attenuation of VHEGR flux has been use as a probe of the EBL \citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}.
-The basic requirement of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL photon
-  ($E_{\rm ebl}$) exceed the rest mass energy of the $e^\pm$ pair in the center
-  of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4 m_e^2 c^4$, where
-  $\theta$ is the relative angle of propagation in the lab frame, an $e^\pm$
-  pair can be produced with Lorentz factor $\gamma\simeq E/2m_e c^2$
-  \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the pair beam that emerges from this process.
+As stated in the Introduction, VHEGR photons pair produce on encountering EBL
+photons as they propagate throughout the universe \citep{Gould+66}, and this
+attenuation of VHEGR flux has been used as a probe of the EBL
+\citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}.  The basic requirement
+of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL
+photon ($E_{\rm ebl}$) exceed the rest mass energy of the $e^\pm$ pair in the
+center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
+m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
+frame\cp{\sout{,}. As a result,} an $e^\pm$ pair can be produced with Lorentz factor $\gamma\simeq E/2m_e
+c^2$ \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
+pair beam that emerges from this process.
 
@@
  where $E$ and $T$ are the dimensionless energy and temperature in terms of a particles rest mass.
-In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-Juttner distribution \citep{1911AnP...340..145J}.
+In the non-relativistic case, this reduces to the Maxwell-Boltzmann distribution, while the relativistic version is known as the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J}.
 
@@
 \end{equation}
-Plugging this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-Juttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
+\cp{\sout{Plugging}Inserting} this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
@@
 
-%The transformation properties of the drifting Maxwell-Juttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-Juttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
+%The transformation properties of the drifting Maxwell-J{\"u}ttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-J{\"u}ttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
 
@@
 
-% To see how this distribution function appears in velocity space, we show four examples of the  Maxwell-Juttner distribution in the center-of-mass frame for the nonrelativistic temperatures ($T=10^{-2}$ and $ 0.1$) in Figure \ref{fig:distribution_non_rel} and for relativistic temperatures ( $T=1$ and $10$) in Figure \ref{fig:distribution_rel}, where $T$ is listed in the natural relativistic units: $m_e c^2/k_B$.
+% To see how this distribution function appears in velocity space, we show four examples of the  Maxwell-J{\"u}ttner distribution in the center-of-mass frame for the nonrelativistic temperatures ($T=10^{-2}$ and $ 0.1$) in Figure \ref{fig:distribution_non_rel} and for relativistic temperatures ( $T=1$ and $10$) in Figure \ref{fig:distribution_rel}, where $T$ is listed in the natural relativistic units: $m_e c^2/k_B$.
 % We plot these in terms of $\beta_z$, which is parallel to the beam and $\beta_x$, which is perpendicular to the beam direction though, in the center-of-mass frame, an isotropic distribution makes no distinction between these directions.
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
-For the beam plasma, we adopt the Maxwell-Juttner distribution (eq.[\ref{eq:distribution function beam}]).
+For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.[\ref{eq:distribution function beam}]).
 %\begin{equation}\label{eq:distribution function beam}
@@
 \begin{equation}
-\Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
-\left[
+\begin{aligned}
+\Gamma &\simeq - \Gamma_0
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
+&\times\left[
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 \exp(-\cG'\mu),\label{eq:kinetic growth rate}
+\end{aligned}
 \end{equation}
@@
 \frac{\nb}{\nIGM} &\simeq&
-\frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \nonumber\\
+\frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \\
 &\simeq &
@@
 \left(\frac{E}{\TeV}\right)
-\,\cm^{-3}\,,
+\,\cm^{-3}\,,\nonumber
 \label{eq:nbIC}
@@
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
-\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam bp condition}
+\left(\frac{E}{\TeV}\right)^{1/3}\label{eq:beam bp condition}
 \end{eqnarray}
@@
 We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam.
-We also show that this can be identified with a drifting relativistic Maxwell-Juttner distribution.
+We also show that this can be identified with a drifting relativistic Maxwell-J{\"u}ttner distribution.
 }
@@
 %\onecolappendix
-\appendix%[onecolumn]
+%\appendix%[onecolumn]
 %\onecolumn
+\begin{appendix}
 
@@
 
+\end{appendix}
+
 \bibliography{bigmh}

### 2016-07-01 oblique.tex
fixed subscripts B and b -> b

@@
 \newcommand{\betabarCM}{\ensuremath{\betabar_{\rm CM}}}
-\newcommand{\betaBOOST}{\ensuremath{\beta_{\rm B}}}
+\newcommand{\betaBOOST}{\ensuremath{\beta_b}}
 
@@
 \begin{equation}
-  E_{\rm CM} = \gamma_B\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
+  E_{\rm CM} = \gamma_b\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
 \end{equation}
@@
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
-%f = \frac{1}{4\pi\gamma_B m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
-f = \frac{n_b m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - \vel_B\cdot \bmath{p})} {k_B T}\right)
+%f = \frac{1}{4\pi\gamma_b m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
+f = \frac{n_b m_e c^2}{4\pi\gamma_b k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_b(E - \vel_b\cdot \bmath{p})} {k_B T}\right)
 \end{eqnarray}
@@
 %\begin{equation}
-% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2}.
+% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2}.
 % \label{eq:betaperp}
@@
 %\begin{equation}
-% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}},
+% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}},
 % \label{eq:betapara}
@@
 % 
-% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_B = 10^6$.
-Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_B^{-1}$ and the parallel component by $\gamma_B^{-2}$.
+% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_b = 10^6$.
+Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_b^{-1}$ and the parallel component by $\gamma_b^{-2}$.
 % When $T\ll1$, as anticipated above, the parallel and transverse velocity dispersions are well approximated by the scaling in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}).
@@
 This may be seen explicitly in the right panels of Figure \ref{fig:distribution_rel}.
-% While these scaling will necessarily break down as $T \rightarrow \gamma_B$, for the problem of interest this is never the case.
-Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_{\rm B}^{-1}$ in the perpendicular direction and like $\gamma_{\rm B}^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_B \gg T$).
+% While these scaling will necessarily break down as $T \rightarrow \gamma_b$, for the problem of interest this is never the case.
+Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_b^{-1}$ in the perpendicular direction and like $\gamma_b^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_b \gg T$).
 % 
@@
 Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.
-As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_B$ on the x-axis.
-We have also multiplied the $\beta$ factors by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
-These plots show that a non-relativistic velocity dispersion transforms as $\gamma_B^{-2}$ and $\gamma_B^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
+As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_b$ on the x-axis.
+We have also multiplied the $\beta$ factors by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
+These plots show that a non-relativistic velocity dispersion transforms as $\gamma_b^{-2}$ and $\gamma_b^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
 % \end{figure*}
@@
 %  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).
-Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_B^2$ for the z (parallel) component and $\gamma_B$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
+Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
 Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.
@@
 % 
-% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_B$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_{\rm B}(E-\beta_{\rm B}p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}:
+% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_b$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_b(E-\beta_b p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}:
 % \begin{equation}
-%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_{\rm B}\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),
+%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_b\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),
 % \end{equation}
-% where $\alpha_\perp\equiv T/\gamma_{\rm B}^2$ and $\alpha_\parallel \equiv T/\gamma_{\rm B}^4 = \alpha_\perp/\gamma_{\rm B}^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
+% where $\alpha_\perp\equiv T/\gamma_b^2$ and $\alpha_\parallel \equiv T/\gamma_b^4 = \alpha_\perp/\gamma_b^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
 % This form of the distribution function has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.
@@
 %\begin{equation}\label{eq:distribution function beam}
-%F_b = \frac{n_b}{4\pi\gamma_B m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
+%F_b = \frac{n_b}{4\pi\gamma_b m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
 %\end{equation}
@@
 \Gamma &\simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
 &\times\left[
@@
 +
-\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
+\frac{\gamma_b^2 v_{b,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{B,x'} = 0$, which gives:
+where $\Gamma_0 \equiv \omega_p \gamma_b (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
 \begin{equation}
 \Gamma_{\rm bp} \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 Moreover the maximum growth rates vary little and are robust for broad range of angles between the wavevector and the beam direction.
-Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_B - 1$.
+Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_b - 1$.
 Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.
@@
 \begin{equation}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 },
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 },
 \end{equation}
@@
 \begin{equation}
-\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
+\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
 \end{equation}
@@
 \betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
-\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_b\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
 \end{eqnarray}
-where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
+where $\gamma_b = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
 For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
@@
  \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-&=& \gamma_{\rm B}^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
+&=& \gamma_b^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
 \label{eq:dbetabarLABperp}
@@
 \begin{equation}
- \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_{\rm B}^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 }.
+ \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 }.
 \end{equation}
@@
 \begin{equation}
- \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_{\rm B}^{4}}  \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
+ \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}}  \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
 \end{equation}
@@
 \begin{eqnarray}
-F_B &=& \frac{m_e c^2}{4\pi\gamma_B k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'} p_{x'})} {k_B T}\right)\\
-%F_b &=& \frac{1}{4\pi\gamma_B m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
-&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_B K_2(\mu)}  \exp\left(-\frac {\gamma_B\gz(c^2 - v_{B,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
+F_B &=& \frac{m_e c^2}{4\pi\gamma_b k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_b(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T}\right)\\
+%F_b &=& \frac{1}{4\pi\gamma_b m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
+&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_b K_2(\mu)}  \exp\left(-\frac {\gamma_b\gz(c^2 - v_{b,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_bv_{b,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
 \end{eqnarray}
@@
 \left.
--\frac{k\gamma_B}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{B,z'}\right)
+-\frac{k\gamma_b}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{b,z'}\right)
 F_b
@@
   k
-  (\vph-v_{B,z'}) \mu^2}{
+  (\vph-v_{b,z'}) \mu^2}{
   4\pi m_e^{4}c^{5}K_2(\mu)
-} \exp\left(-\frac {\gamma_B\gph(c^2 - v_{B,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_Bv_{B,x'} p_{x'}}{k_BT}\right).
+} \exp\left(-\frac {\gamma_b\gph(c^2 - v_{b,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_bv_{b,x'} p_{x'}}{k_BT}\right).
 \end{aligned}
@@
 \frac{
-  \gph^3 (\vph-v_{B,z'})\mu^2}{4\pi K_2(\mu)
+  \gph^3 (\vph-v_{b,z'})\mu^2}{4\pi K_2(\mu)
 }
@@
 \end{equation}
-and $\cG\equiv\gamma_B\gph(1-v_{B,z'}\vph/c^2)$ and
-$w\equiv \gamma_Bv_{B,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
-\begin{equation}
-\cG E_\perp - \gamma_B v_{B,x'} p_{x'} = \gamma_B(E - v_{B,z'} p_{z'} - v_{B,x'}p_{x'})|_{v_{B,z'}=v_{\rm ph}} > 0
+and $\cG\equiv\gamma_b\gph(1-v_{b,z'}\vph/c^2)$ and
+$w\equiv \gamma_bv_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
+\begin{equation}
+\cG E_\perp - \gamma_b v_{b,x'} p_{x'} = \gamma_b(E - v_{b,z'} p_{z'} - v_{b,x'}p_{x'})|_{v_{b,z'}=v_{\rm ph}} > 0
 \end{equation}
@@
 R = \frac{
-  \gph^3 \gamma_w^2 (\vph-v_{B,z'}) 
+  \gph^3 \gamma_w^2 (\vph-v_{b,z'}) 
 }{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
@@
 +
-\frac{\gamma_B^2 v_{B,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
+\frac{\gamma_b^2 v_{b,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
 \right] \exp\left(-\cG'\mu\right)
@@
 \Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{B,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
 \left[
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gamma_B (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
+where $\Gamma_0 \equiv \omega_p \gamma_b (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
 %The above differs from the beam plasma instability only due to the

### 2016-07-01 oblique.tex
edited through Sect 3.1

@@
 
-There are two distinct, often qualitatively different, regimes in which we may consider the implications of this dispersion relation.
-The first is the cold plasma limit or the hydrodynamic or reactive limit.
-The hydrodynamic limit is aptly named because the resulting dispersion relation that is found could have also been calculated directly from the continuity equation and the momentum equation.
-In this limit, the internal distribution of the particles of the background or beam are irrelevant to the physics of the instability and it is only the bulk response that is important.
-In particular, this means that the beam particles are resonance with the unstable wave over a timescale much longer than the growth them as stated above, i.e., the beam particles do not drift a distance larger than the wavelength of the unstable mode over the growth time of the instability.
-The second is the kinetic regime, where the internal distribution of the beam particles are important to the physics of the instability.
-Here, only a fraction of the beam particles stay within one wavelength of the unstable mode over the growth time of the instability.
-Here, the bulk of the plasma (background or beam) do not response to the disturbance, but rather only a fraction of the particles are relevant for driving (instability) or damping (Landau damping).
-We discuss below the evaluation of the dispersion relation in these two regimes, which gives two regimes of instability, and the delineation between these two regimes.
+There are two distinct, often qualitatively different, regimes in which we may
+consider the implications of this dispersion relation.  The first is the cold
+plasma limit or the hydrodynamic or reactive limit.  The hydrodynamic limit is
+aptly named because the resulting dispersion relation that is found could have
+also been calculated directly from the continuity equation and the momentum
+equation.  In this limit, the internal distribution of the particles of the
+background or beam are irrelevant to the physics of the instability and it is
+only the bulk response that is important.  In particular, this means that the
+beam particles are resonance with the unstable wave over a timescale much longer
+than the growth \cp{\sout{them}time} as stated above, i.e., the beam particles
+do not drift a distance larger than the wavelength of the unstable mode over the
+growth time of the instability.  The second is the kinetic regime, where the
+internal distribution of \cp{\sout{the}} beam particles \cp{\sout{are}is}
+important to the physics of the instability. Here, only a fraction of
+\cp{\sout{the}} beam particles stay within one wavelength of the unstable mode
+over the growth time of the instability.  \cp{\sout{Here}Moreover}, the bulk of
+the plasma (background or beam) does not respond to the disturbance\cp{\sout{, but rather}. Instead,}
+only a fraction of \cp{\sout{the}} particles \cp{\sout{are}is} relevant for
+driving (instability) or damping (Landau damping).  We discuss below the
+evaluation of the dispersion relation in these two regimes, which gives two
+regimes of instability, and the delineation between these two regimes.
 
@@
 
-The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (eq.[\ref{eq:growth rate reactive appendix}]) is
+The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (eq.~\ref{eq:growth rate reactive appendix}) is
 \begin{equation}\label{eq:growth rate reactive}
@@
 
-In a continuous system, these waves do not matter compare to the unstable mode at $k\lambda_D = 1$.
-However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state compare to the physical system.
+In a continuous system, these waves do not matter in comparison to the unstable mode at $k\lambda_D = 1$.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
 
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
-For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.[\ref{eq:distribution function beam}]).
+For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.~\ref{eq:distribution function beam}).
 %\begin{equation}\label{eq:distribution function beam}
@@
 \end{eqnarray}
-Hence, the conditions for the reactive regime for the oblique mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
+Hence, the conditions for the reactive regime for the oblique mode (eq.~\ref{eq:oblique reactive regime}) and beam-plasma mode (eq.~\ref{eq:bp reactive regime}) can be reduced to 
 \begin{equation}\label{eq:generic reactive regime}
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:bp kinetic growth rate}]) and oblique mode (eq.[\ref{eq:kinetic growth rate}]).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.~\ref{eq:bp kinetic growth rate}) and oblique mode (eq.~\ref{eq:kinetic growth rate}).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.
@@
 
-In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.[\ref{eq:generic reactive regime}]) and kinetic regimes (eq.[\ref{eq:generic kinetic regime}]) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
-Applying these conditions (eqns.
-[\ref{eq:generic reactive regime}] and [\ref{eq:generic kinetic regime}]) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter:
+In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.~\ref{eq:generic reactive regime}) and kinetic regimes (eq.~\ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
+Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter:
 \begin{eqnarray}

### 2016-07-01 oblique.tex
edited through the main body of the paper

@@
 
-In a continuous system, these waves do not matter in comparison to the unstable mode at $k\lambda_D = 1$.
-However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
+In a continuous system, these waves do not matter \cp{\sout{compare}in comparison} to the unstable mode at $k\lambda_D = 1$.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state \cp{\sout{compare}in comparison} to the physical system.
 
@@
 
-The growth rate expressed in Equation (\ref{eq:growth rate reactive}) sits in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could be been derived from the fluid equations.
+The growth rate expressed in Equation (\ref{eq:growth rate reactive}) \cp{\sout{sits}is} in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could \cp{\sout{be}have} been derived from the fluid equations.
 Here all the particles participate in the instability.
@@
 \end{equation}
-where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and$T$ is the temperature of the background.
+where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T$ is the temperature of the background.
 For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.~\ref{eq:distribution function beam}).
@@
 %\end{equation}
-Plugging these into the dispersion relation (\ref{eq:dispersion relation}), we find
+\cp{\sout{Plugging}Inserting} these into the dispersion relation (\ref{eq:dispersion relation}), we find
 %\begin{equation}\label{eq:dispersion kinetic}
@@
 \end{equation}
-where we have integrated by parts only the second term, associated with target plasma.
+where we have integrated by parts only the second term, associated with target plasma.  \cp{[Define $\gamma$, $\gamma_w$, $\gamma_b$, $\gamma_{\rm ph}$.]}
 
@@
 \caption{Oblique growth rate maximized over $\vph$ as a function of
-  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
-  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
-  cases, the maximum growth rate occurs when
-  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
+  \cp{\sout{$v_{bx}$ and $T$}$\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector.\sout{ for various $\gamma_b$.
+  In each the green line shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.
+  In all cases, the maximum growth rate occurs when
+  $v_{bx}/v_b\gg1/\gamma_b$.}} \cp{[Which $\gamma_b$ was assumed for this -- $10^6$?
+Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
+In the paper, the notation $T$ is used instead of $T_b$ here -- can you unify this?]}}\label{fig:ObliqueGMZ}
 \end{figure}
 
-In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of $\cos\theta = \hat{k}\cdot\vel$, i.e., the angle between the beam and the wave vector, and $k_BT/m_e c^2$.
-Here, it is clear that the growth rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$, i.e., at an oblique, but not perpendicular angle.
-Moreover the maximum growth rates vary little and are robust for broad range of angles between the wavevector and the beam direction.
-Because the maximal oblique growth rate varies little for a wide range of angles and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$ in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$ and $\gamma_b - 1$.
-Here for relativistic beams, the maximum growth rate varies little with $T$, varying by less than $10\%$ between hot and cold beams.
-Thus, we find:
+In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of
+\cp{$\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$} \sout{[is
+    this correct?] $\cos\theta = \hat{k}\cdot\vel$}, i.e., the angle between the
+beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth
+rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$ \cp{[This is not
+    clear, as the axis shows $\sin\theta$.]}, i.e., at an oblique, but not
+perpendicular angle.  Moreover the maximum growth rates vary little and are
+robust for broad range of angles between the wavevector and the beam direction.
+Because the maximal oblique growth rate varies little for a wide range of angles
+and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$
+in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$
+and $\gamma_b - 1$.  Here for relativistic beams, the maximum growth rate varies
+little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus,
+we find:
 \begin{equation}
@@
 %\plottwo{pp2.pdf}{pp3.pdf}
-\caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right), maximized over $\theta$ and normalized by
-  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function
-  of $k_BT/m_ec^2$ and $\gamma_b$.  Note that unlike the beam
-  plasma case shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the
-  transition between high and low temperature is only marginal,
-  constituting a roughly $10\%$ reduction.} \label{fig:OGgen}
+\caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right),
+  maximized over $\theta$ and normalized by
+  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function of
+  $k_BT/m_ec^2$ and \cp{$\gamma_b-1$}.  \cp{\sout{Note that unlike the beam plasma case
+      shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the transition between
+      high and low temperature is only marginal, constituting a roughly $10\%$
+      reduction.} Unlike the beam-plasma case on the right, in the case of the
+    oblique kinetic growth rate at high $\gamma_b$ the transition between high
+    and low temperature is only marginal, constituting a roughly $10\%$
+    reduction.}  \cp{[Can you please add $k_B$ instead of $k$ to the axis label
+      and $\Gamma/\Gamma_0$ at the color bar? In the paper, the notation $T$ is
+      used instead of $T_b$ here -- can you unify this?]}} \label{fig:OGgen}
 \end{figure*}
 
-This can be contrasted with Figure \ref{fig:bpgen} where we plot the maximized beam-plasma growth rate  (\ref{eq:bp kinetic growth rate}) as a function of $\gamma_b - 1$ and $k_B T/m_ec^2$.
-Here, we see that the maximum growth rate is somewhat sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for $k_BT/m_ec^2 \ll 1$ to $\Gamma/\Gamma_0 \approx 0.1$ for $k_BT/m_ec^2 \gtrsim 1$.
-Note, however, that that beam-plasma growth rate remains competitive with the oblique growth rate, i.e., it is not orders of magnitude lower.
+This can be contrasted with \cp{\sout{Figure~\ref{fig:bpgen}} the right panel of
+  Figure~\ref{fig:OGgen}} where we plot the maximized beam-plasma growth rate
+(\ref{eq:bp kinetic growth rate}) as a function of $\gamma_b - 1$ and $k_B
+T/m_ec^2$.  Here, we see that the maximum growth rate is somewhat \cp{more}
+sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for
+$k_BT/m_ec^2 \ll 1$ to $\Gamma/\Gamma_0 \approx 0.1$ for $k_BT/m_ec^2 \gtrsim
+1$.  Note, however, that that beam-plasma growth rate remains competitive with
+the oblique growth rate, i.e., it is not orders of magnitude lower.
+\cp{[Explain why we found the opposite result in Paper I, i.e., a negligible
+    beam-plasma growth rate in comparison to the oblique growth rate.]}
 
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm CM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm CM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
 We note several feature of this Figure.
@@
 Additionally, the break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
+\cp{[What $\Dpp$ does this correspond to?]}
 
@@
 Given that the range of $E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, this corresponds to values of $\alpha_{\perp}'$ and $\alpha_{\parallel}'$ between $\sim 0.3-2$.
-To determine whether or not the reactive or kinetic instabilities apply to these beams, it is necessary to determine $n_b/n_t$.
+\cp{[Define $\alpha_{\perp}'$ and $\alpha_{\parallel}'$.]}  To determine whether or not the reactive or kinetic instabilities apply to these beams, it is necessary to determine $n_b/n_t$.
 Here the target is the background IGM, so $n_t = \nIGM$, where $\nIGM \approx 2\times 10^{-7}\,(1+\delta)\,(1+z)^3\,\cm^{-3}$ is the mean density of the IGM, $z$ is the redshift, and $\delta$ denotes the overdensity.
@@
 \end{eqnarray}
-where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta = 4.5$ for $z \leq 1$ and $\zeta = 0$ for $z\geq 1$.
+where $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
+\cp{\sout{$\Dpp$ is defined in the Introduction to be}} the mean free path of a VHEGR \cp{is given by
+\begin{equation}
+\Dpp(E,z) =
+35 
+\left(\frac{1+z}{2}\right)^{-\zeta}
+\left(\frac{E}{1~{\rm TeV}}\right)^{-1}
+~{\rm Mpc}\,,
+\label{eq:Dpp}
+\end{equation}
+with $\zeta=4.5$ for $z<1$ and $\zeta=0$ for $z\ge1$ \citep{Knei_etal:04, Nero-Semi:09}.}
 
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability sits in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
-
-The discussion above confirmed our earlier estimate in Paper I that the oblique instability sits in the kinetic regime.
-However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability \cp{\sout{sits}is} in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs \cp{for} $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+
+The discussion above confirmed our earlier estimate in Paper I that the oblique instability \cp{\sout{sits}is} in the kinetic regime.
+However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also \cp{\sout{sit}exist} in the reactive regime.
 However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
@@
 This is a part of ongoing work and will be presented in a future publication.
-Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
-Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
+Second, while it seems that the beam plasma mode may \cp{\sout{sit}be} in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
+Thus, both the reactive and the kinetic growth rates are similar and \cp{it} likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
 Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
@@
 The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and oblique instability.
-In this work, we have these linear instabilities as they would apply to the ultrarelativistic pair beams.
+In this work, we \cp{\sout{have}examined} these linear instabilities as they would apply to the ultrarelativistic pair beams.
 Our main findings are:
@@
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
-\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes for both are different.
+\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes \cp{\sout{for both}} are different.
 In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the oblique mode.

### 2016-07-02 oblique.tex
small change
dereddened most of christoph's changes

@@
   Stereoscopic System, Major Atmospheric Gamma Imaging Cerenkov Telescope, Very
-  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the \cp{\sout{ultra-}}high
+  Energetic Radiation Imaging Telescope Array System.} have demonstrated that the high
 energy Universe is teeming with energetic very high-energy gamma-ray 
 (VHEGR, $E > 100\,\GeV$) sources, the extragalactic component of which mainly consists of TeV
-blazars with a minority population of other sources  such as radio\cp{- \sout{galaxies}}
+blazars with a minority population of other sources  such as radio
 and starburst galaxies. These extragalactic VHEGR emitters produce TeV photons that are greatly attenuated via annihilation upon soft photons in the extragalactic background light
@@
 
-\cp{To this end\sout{However}}, a full nonlinear study is called for.
+To this end, a full nonlinear study is called for.
 We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.
@@
 Applying these results to TeV electron-positron pair beams that arise from TeV photon pair production in Section \ref{sec:application},
-we demonstrate that despite the extraordinary coldness of the beam \cp{\sout{that}} we are always in the kinetic regime for the oblique mode, but may be in the reactive regime for the beam plasma mode.
+we demonstrate that despite the extraordinary coldness of the beam we are always in the kinetic regime for the oblique mode, but may be in the reactive regime for the beam plasma mode.
 However, because the controlling parameter is close to unity, the growth rates calculated in either regime are similar.
@@
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
-frame\cp{\sout{,}. As a result,} an $e^\pm$ pair can be produced with Lorentz factor $\gamma\simeq E/2m_e
+frame. As a result, an $e^\pm$ pair can be produced with Lorentz factor $\gamma\simeq E/2m_e
 c^2$ \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
@@
 \end{equation}
-\cp{\sout{Plugging}Inserting} this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
+Inserting this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
@@
 beam particles are resonance with the unstable wave over a timescale much longer
-than the growth \cp{\sout{them}time} as stated above, i.e., the beam particles
+than the growth time as stated above, i.e., the beam particles
 do not drift a distance larger than the wavelength of the unstable mode over the
 growth time of the instability.  The second is the kinetic regime, where the
-internal distribution of \cp{\sout{the}} beam particles \cp{\sout{are}is}
+internal distribution of  beam particles is
 important to the physics of the instability. Here, only a fraction of
-\cp{\sout{the}} beam particles stay within one wavelength of the unstable mode
-over the growth time of the instability.  \cp{\sout{Here}Moreover}, the bulk of
-the plasma (background or beam) does not respond to the disturbance\cp{\sout{, but rather}. Instead,}
-only a fraction of \cp{\sout{the}} particles \cp{\sout{are}is} relevant for
+ beam particles stay within one wavelength of the unstable mode
+over the growth time of the instability.  Moreover, the bulk of
+the plasma (background or beam) does not respond to the disturbance. Instead,
+only a fraction of  particles is relevant for
 driving (instability) or damping (Landau damping).  We discuss below the
@@
 
-In a continuous system, these waves do not matter \cp{\sout{compare}in comparison} to the unstable mode at $k\lambda_D = 1$.
-However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state \cp{\sout{compare}in comparison} to the physical system.
+In a continuous system, these waves do not matter in comparison to the unstable mode at $k\lambda_D = 1$.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
 
@@
 
-The growth rate expressed in Equation (\ref{eq:growth rate reactive}) \cp{\sout{sits}is} in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could \cp{\sout{be}have} been derived from the fluid equations.
+The growth rate expressed in Equation (\ref{eq:growth rate reactive}) is in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could have been derived from the fluid equations.
 Here all the particles participate in the instability.
@@
 %\end{equation}
-\cp{\sout{Plugging}Inserting} these into the dispersion relation (\ref{eq:dispersion relation}), we find
+Inserting these into the dispersion relation (\ref{eq:dispersion relation}), we find
 %\begin{equation}\label{eq:dispersion kinetic}
@@
 In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of
-\cp{$\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$} \sout{[is
-    this correct?] $\cos\theta = \hat{k}\cdot\vel$}, i.e., the angle between the
+$\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$, i.e., the angle between the
 beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth
@@
 
-This can be contrasted with \cp{\sout{Figure~\ref{fig:bpgen}} the right panel of
-  Figure~\ref{fig:OGgen}} where we plot the maximized beam-plasma growth rate
+This can be contrasted with the right panel of
+  Figure~\ref{fig:OGgen} where we plot the maximized beam-plasma growth rate
 (\ref{eq:bp kinetic growth rate}) as a function of $\gamma_b - 1$ and $k_B
-T/m_ec^2$.  Here, we see that the maximum growth rate is somewhat \cp{more}
+T/m_ec^2$.  Here, we see that the maximum growth rate is somewhat more
 sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for
@@
 where $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
-\cp{\sout{$\Dpp$ is defined in the Introduction to be}} the mean free path of a VHEGR \cp{is given by
+ the mean free path of a VHEGR 
 \begin{equation}
@@
 \end{equation}
-with $\zeta=4.5$ for $z<1$ and $\zeta=0$ for $z\ge1$ \citep{Knei_etal:04, Nero-Semi:09}.}
+with $\zeta=4.5$ for $z<1$ and $\zeta=0$ for $z\ge1$ \citep{Knei_etal:04, Nero-Semi:09}.
 
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability \cp{\sout{sits}is} in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs \cp{for} $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
-
-The discussion above confirmed our earlier estimate in Paper I that the oblique instability \cp{\sout{sits}is} in the kinetic regime.
-However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also \cp{\sout{sit}exist} in the reactive regime.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+
+The discussion above confirmed our earlier estimate in Paper I that the oblique instability is in the kinetic regime.
+However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also exist in the reactive regime.
 However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
@@
 This is a part of ongoing work and will be presented in a future publication.
-Second, while it seems that the beam plasma mode may \cp{\sout{sit}be} in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
-Thus, both the reactive and the kinetic growth rates are similar and \cp{it} likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
+Second, while it seems that the beam plasma mode may be in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
+Thus, both the reactive and the kinetic growth rates are similar and it likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
 Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
@@
 The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and oblique instability.
-In this work, we \cp{\sout{have}examined} these linear instabilities as they would apply to the ultrarelativistic pair beams.
+In this work, we examined these linear instabilities as they would apply to the ultrarelativistic pair beams.
 Our main findings are:
@@
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
-\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes \cp{\sout{for both}} are different.
+\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes are different.
 In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the oblique mode.
@@
 where we define $\mu = m_ec^2/k_BT$.
-Plugging the equation into (\ref{eq:dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
+Inserting the equation into (\ref{eq:dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
 \begin{equation}

### 2016-07-04 oblique.tex
included astrid comments

@@
 \newcommand{\gammabeam}{\ensuremath{\gamma_{\rm b}}}
+\newcommand{\vbeam}{\ensuremath{v_{\rm b}}}
 \newcommand{\gammaphase}{\ensuremath{\gamma_{\rm ph}}}
@@
 Ewald Puchwein\altaffilmark{5},
-Astrid Lamberts\altaffilmark{6}
-Mohamad Shalaby\altaffilmark{1,2,7},
+Astrid Lamberts\altaffilmark{6},
+Mohamad Shalaby\altaffilmark{1,2,7}
 }
@@
 extragalactic gamma-ray background (EGRB) for energies between
-$200\,\MeV$ and $100\,\GeV$.  Since ICCs reprocess the VHEGR emission of
+$200\,\MeV$ and $100\,\GeV$.  Since inverse-Compton cascades (ICCs) reprocess the VHEGR emission of
 distant sources into this band, this has been used to constrain the
@@
 anomalies in the statistics of the high-redshift Ly$\alpha$ forest
-\citep{paperIV} and potentially explaining a number of the X-ray
+\citep{paperIV,2015ApJ...811...19L} and potentially explaining a number of the X-ray
 properties of groups and clusters and anomalies in galaxy formation on
@@
 For instance, \citet{Miniati+12} argued that these instabilities are physically irrelevant for the cooling of these pair beams because they would saturate at a very low level due to nonlinear Landau damping (NLD).
-However, \citet{Chang+14} showed that a detailed calculation of NLD showed that these plasma processes remain dominant.
+However, \citet{Chang+14} performed a detailed calculation of NLD to show that these plasma processes remain dominant.
 In addition, \citet{Sironi+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
-It is unclear, however, if the conclusions of their work are robust as it is unclear if these numerical setup is adequate (e.g.
+It is unclear, however, if the conclusions of their work are robust as it is unclear if their numerical setup is adequate (e.g.
 Mohamad's paper).
@@
 To this end, a full nonlinear study is called for.
-We begin this task in this initial work where we study the distribution function of the electron-positron pairs that are produced from VHEGR-EBL photon annihilation.
+We begin this task in this initial work where we study the distribution function of the $e^+e^-$ pairs that are produced from VHEGR-EBL photon annihilation.
 We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.
@@
 In particular, we pay careful attention to both the reactive (or hydrodynamic) and kinetic regimes of the beam plasma and oblique instabilities and the transition between the two.
-Applying these results to TeV electron-positron pair beams that arise from TeV photon pair production in Section \ref{sec:application},
+Applying these results to TeV $e^+e^-$ pair beams that arise from TeV photon pair production in Section \ref{sec:application},
 we demonstrate that despite the extraordinary coldness of the beam we are always in the kinetic regime for the oblique mode, but may be in the reactive regime for the beam plasma mode.
-However, because the controlling parameter is close to unity, the growth rates calculated in either regime are similar.
+However, for the relevant parameters, the growth rates calculated in either regime are similar.
 We close with a discussion the implications of this work and application of these results for nonlinear theory in Section \ref{sec:conclusions}.
@@
 of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL
-photon ($E_{\rm ebl}$) exceed the rest mass energy of the $e^\pm$ pair in the
+photon ($E_{\rm ebl}$) exceed the rest mass energy of the $e^+e^-$ pair in the
 center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
-frame. As a result, an $e^\pm$ pair can be produced with Lorentz factor $\gamma\simeq E/2m_e
-c^2$ \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
+frame. As a result, an $e^\pm$ pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
+c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
 pair beam that emerges from this process.
@@
 \begin{equation}
-  E_{\rm CM} = \gamma_b\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right).
-\end{equation}
+  E_{\rm COM} = \gammabeam\left(E_{\rm L} - \betaBOOST p_{\rm L,\parallel}\right),
+\end{equation}
+where $\gammabeam = \gamma(v_{\rm b}) = \left(1 - v_{\rm b}^2/c^2\right)$ is the Lorentz factor of the beam and $v_{\rm b}$ is the bulk velocity of the pair beam.
 Inserting this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
-%f = \frac{1}{4\pi\gamma_b m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
-f = \frac{n_b m_e c^2}{4\pi\gamma_b k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_b(E - \vel_b\cdot \bmath{p})} {k_B T}\right)
+%f = \frac{1}{4\pi\gammabeam m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
+f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T}\right)
 \end{eqnarray}
@@
 %\begin{equation}
-% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2}.
+% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2}.
 % \label{eq:betaperp}
@@
 %\begin{equation}
-% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}},
+% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}},
 % \label{eq:betapara}
@@
 % 
-% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gamma_b = 10^6$.
-Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gamma_b^{-1}$ and the parallel component by $\gamma_b^{-2}$.
+% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gammabeam = 10^6$.
+Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gammabeam^{-1}$ and the parallel component by $\gammabeam^{-2}$.
 % When $T\ll1$, as anticipated above, the parallel and transverse velocity dispersions are well approximated by the scaling in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}).
@@
 This may be seen explicitly in the right panels of Figure \ref{fig:distribution_rel}.
-% While these scaling will necessarily break down as $T \rightarrow \gamma_b$, for the problem of interest this is never the case.
-Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gamma_b^{-1}$ in the perpendicular direction and like $\gamma_b^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gamma_b \gg T$).
+% While these scaling will necessarily break down as $T \rightarrow \gammabeam$, for the problem of interest this is never the case.
+Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gammabeam^{-1}$ in the perpendicular direction and like $\gammabeam^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gammabeam \gg T$).
 % 
@@
 As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_b$ on the x-axis.
-We have also multiplied the $\beta$ factors by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
-These plots show that a non-relativistic velocity dispersion transforms as $\gamma_b^{-2}$ and $\gamma_b^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
+We have also multiplied the $\beta$ factors by a factor of $\gammabeam^2$ for the z (parallel) component and $\gammabeam$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
+These plots show that a non-relativistic velocity dispersion transforms as $\gammabeam^{-2}$ and $\gammabeam^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
 % \end{figure*}
@@
 %  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).
-Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gamma_b^2$ for the z (parallel) component and $\gamma_b$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
+Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gammabeam^2$ for the z (parallel) component and $\gammabeam$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
 Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.
@@
 % 
-% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gamma_b$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gamma_b(E-\beta_b p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}:
+% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gammabeam$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gammabeam(E-\beta_b p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}:
 % \begin{equation}
@@
 % \end{equation}
-% where $\alpha_\perp\equiv T/\gamma_b^2$ and $\alpha_\parallel \equiv T/\gamma_b^4 = \alpha_\perp/\gamma_b^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
+% where $\alpha_\perp\equiv T/\gammabeam^2$ and $\alpha_\parallel \equiv T/\gammabeam^4 = \alpha_\perp/\gammabeam^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
 % This form of the distribution function has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.
@@
 Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
-Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
+Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
 \begin{equation}\label{eq:dispersion relation reactive}
- 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
+ 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z \vbeam)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
 \end{equation}
@@
 \end{equation}
-where $Z_x = k_xv_b/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.
+where $Z_x = k_x\vbeam/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.
 
@@
 which we denote the beam-plasma or ``two-stream'' growth rate.
-For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by Bret et al.
-Indeed for $\gamma \gg 1$ and $Z_x \rightarrow \infty$, the growth rate approaches oblique growth rate:
+For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and $Z_x \rightarrow \infty$, the growth rate approaches oblique growth rate:
 \begin{equation}
@@
 
-We should caution in the derivation above that the resonance condition, which is $\omega_{p,t} - k_z v_b$, implies that $k_z \neq 0$.
+We should caution in the derivation above that the resonance condition, which is $\omega_{p,t} - k_z \vbeam$, implies that $k_z \neq 0$.
 For the case where $k_z \rightarrow 0$, the electrostatic approximation no longer holds and the full dispersion relation must be solved.\footnote{We thank Antoine Bret for helping to clarify this point.}  A solution to the full dispersion relation reveals additional modes, including the zero frequency ($k_z = 0$) filamentation or Weibel mode.
@@
 \begin{equation}\label{eq:dispersion relation beam-plasma}
- 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2} = 0,
-\end{equation}
-which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $v_b/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.
-In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $v_b/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.
-For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is plasma oscillation frequency.
+ 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z \vbeam)^2} = 0,
+\end{equation}
+which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $\vbeam/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.
+In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $\vbeam/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.
+For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is the plasma oscillation frequency.
 This wave would exist in the absence of a tenuous beam.
@@
 \caption{Beam-plasma growth rate (solid line) and unstable wave frequency (dashed-line) as a function of
-  $k_z$ for $\gamma_b = 100$ and $n_b/n_t=10^{-3}$. \label{fig:OffResonance}}
+  $k_z$ for $\gammabeam = 100$ and $n_b/n_t=10^{-3}$. \label{fig:OffResonance}}
 \end{figure}
@@
 %\begin{equation}\label{eq:distribution function beam}
-%F_b = \frac{n_b}{4\pi\gamma_b m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
+%F_b = \frac{n_b}{4\pi\gammabeam m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
 %\end{equation}
@@
 \end{equation}
-where we have integrated by parts only the second term, associated with target plasma.  \cp{[Define $\gamma$, $\gamma_w$, $\gamma_b$, $\gamma_{\rm ph}$.]}
+where we have integrated by parts only the second term, associated with target plasma.
 
@@
 \Gamma &\simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}\\
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}\\
 &\times\left[
@@
 +
-\frac{\gamma_b^2 v_{b,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gamma_b (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
+where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
 \begin{equation}
 \Gamma_{\rm bp} \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 \caption{Oblique growth rate maximized over $\vph$ as a function of
-  \cp{\sout{$v_{bx}$ and $T$}$\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector.\sout{ for various $\gamma_b$.
-  In each the green line shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.
+  \cp{\sout{$v_{bx}$ and $T$}$\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector.\sout{ for various $\gammabeam$.
+  In each the green line shows the chord along which $v_{bx}/\vbeam = 1/\gammabeam$.
   In all cases, the maximum growth rate occurs when
-  $v_{bx}/v_b\gg1/\gamma_b$.}} \cp{[Which $\gamma_b$ was assumed for this -- $10^6$?
+  $v_{bx}/\vbeam\gg1/\gammabeam$.}} \cp{[Which $\gammabeam$ was assumed for this -- $10^6$?
 Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
@@
 beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth
-rate reaches it maximum value $\cos\theta \gg 1/\gamma_b$ \cp{[This is not
+rate reaches it maximum value $\cos\theta \gg 1/\gammabeam$ \cp{[This is not
     clear, as the axis shows $\sin\theta$.]}, i.e., at an oblique, but not
@@
 in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$
-and $\gamma_b - 1$.  Here for relativistic beams, the maximum growth rate varies
+and $\gammabeam - 1$.  Here for relativistic beams, the maximum growth rate varies
 little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus,
@@
 \end{equation}
-for $\gamma_b\gtrsim10$, as seen in Figure \ref{fig:OGgen}.
+for $\gammabeam\gtrsim10$, as seen in Figure \ref{fig:OGgen}.
 
@@
   maximized over $\theta$ and normalized by
-  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function of
-  $k_BT/m_ec^2$ and \cp{$\gamma_b-1$}.  \cp{\sout{Note that unlike the beam plasma case
-      shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the transition between
+  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gammabeam m \vbeam^2/T$ as a function of
+  $k_BT/m_ec^2$ and \cp{$\gammabeam-1$}.  \cp{\sout{Note that unlike the beam plasma case
+      shown in Figure \ref{fig:Ggen}, at high $\gammabeam$ the transition between
       high and low temperature is only marginal, constituting a roughly $10\%$
       reduction.} Unlike the beam-plasma case on the right, in the case of the
-    oblique kinetic growth rate at high $\gamma_b$ the transition between high
+    oblique kinetic growth rate at high $\gammabeam$ the transition between high
     and low temperature is only marginal, constituting a roughly $10\%$
@@
   Figure~\ref{fig:OGgen} where we plot the maximized beam-plasma growth rate
-(\ref{eq:bp kinetic growth rate}) as a function of $\gamma_b - 1$ and $k_B
+(\ref{eq:bp kinetic growth rate}) as a function of $\gammabeam - 1$ and $k_B
 T/m_ec^2$.  Here, we see that the maximum growth rate is somewhat more
@@
 The scaling of $\alpha_{\parallel}$ with $\gamma$ removes the overall scaling of $\alpha_{\parallel}$ with $\gamma$.
-We have also parametrized the wavevector in terms of $\Delta k = k - k_0$, where $k_0 = \omega_{p,t}/v_b$ is the wavevector of the wave whose phase speed $v_{\rm ph} = v_b$ is equal to the speed of the beam.
+We have also parametrized the wavevector in terms of $\Delta k = k - k_0$, where $k_0 = \omega_{p,t}/\vbeam$ is the wavevector of the wave whose phase speed $v_{\rm ph} = \vbeam$ is equal to the speed of the beam.
 Here in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$, so the positive values of $\Delta k$ results in negative values of $\Delta v_{\rm ph}$, which puts these Langmuir waves in the rising part of the distribution function, i.e., $\partial F_2/\partial\Delta\beta_z > 0$, which is required for growth.
@@
 We choose to parametrized $\alpha_{\parallel}$ in this way to remove the scaling of $\alpha_{\rm parallel}$ with $\gamma$.
-On the x-axis, we have plotted in terms of $\gamma^2\Delta k/k_0$, where $k_0 = \omega_{p,t}/v_b$ and $\Delta k = k - k_0$.
+On the x-axis, we have plotted in terms of $\gamma^2\Delta k/k_0$, where $k_0 = \omega_{p,t}/\vbeam$ and $\Delta k = k - k_0$.
 A Langmuir wave with $k=k_0$ moves at the exact speed of the beam.
-For growth, Langmuir waves must move with a phase velocity $v_{\rm ph} < v_b$ to be on the rising part of the distribution function.
+For growth, Langmuir waves must move with a phase velocity $v_{\rm ph} < \vbeam$ to be on the rising part of the distribution function.
 Hence, in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$ where the extra factor of $\gamma^2$ is remove the scaling of $\Delta v_{\rm phase}$ with $\gamma$.
@@
 \begin{equation}\label{eq:oblique reactive regime}
- \frac{\Delta v_{\perp}}{v_b} \ll \left(\frac{n_b}{\gamma_b n_t}\right)^{1/3}
+ \frac{\Delta v_{\perp}}{\vbeam} \ll \left(\frac{n_b}{\gammabeam n_t}\right)^{1/3}
 %\left(1 - \beta_b^2\cos^2\Psi\right)^{1/3}, 
@@
 \begin{equation}\label{eq:bp reactive regime}
-\frac{\Delta v_{\parallel}}{v_b} \ll \gamma_b^{-1}\left(\frac{n_b}{n_t}\right)^{1/3}.
+\frac{\Delta v_{\parallel}}{\vbeam} \ll \gammabeam^{-1}\left(\frac{n_b}{n_t}\right)^{1/3}.
 \end{equation}
@@
 \begin{equation}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 },
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gammabeam^2 m_e c^2 },
 \end{equation}
@@
 \begin{equation}
-\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
+\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gammabeam^4 m_e c^2 },
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\Delta v_{\perp}}{v_b} &\sim&  \gamma_b^{-1}\sqrt{\frac{k_BT}{m_ec^2}},\\
-\frac{\Delta v_{\parallel}}{v_b} &\sim& \gamma_b^{-2}\sqrt{\frac{k_BT}{m_ec^2}}.
+\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT}{m_ec^2}},\\
+\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT}{m_ec^2}}.
 \end{eqnarray}
@@
  \begin{array}{cl}
- ({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
-({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+ ({k_BT}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
+({k_BT}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.~\ref{eq:bp kinetic growth rate}) and oblique mode (eq.~\ref{eq:kinetic growth rate}).
-Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.
+Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
 Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$  For the beam plasma case, the growth rate is roughly
@@
 \begin{equation}\label{eq:bp kinetic regime}
- \frac{\Delta v_{\parallel}}{v_b} \gtrsim \gamma^{-1}\left(\frac{n_b}{n_t}\right)^{1/3},
+ \frac{\Delta v_{\parallel}}{\vbeam} \gtrsim \gamma^{-1}\left(\frac{n_b}{n_t}\right)^{1/3},
 \end{equation}
@@
 \begin{equation}\label{eq:oblique kinetic regime}
-\frac{\Delta v_{\perp}}{v_b} \gtrsim \left(\frac{n_b}{\gamma n_t}\right)^{1/3},
+\frac{\Delta v_{\perp}}{\vbeam} \gtrsim \left(\frac{n_b}{\gamma n_t}\right)^{1/3},
 \end{equation}
@@
 \begin{array}{cl}
-({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
-({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+({k_BT}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
+({k_BT}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm CM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
 We note several feature of this Figure.
@@
 
-Still some progress can be made if we assume that the dominant loss rate is ICC so that we can place a lower limit on the various plasma processes.
+Still some progress can be made if we assume that the dominant loss rate is IC so that we can place a lower limit on the various plasma processes.
 In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (Paper I):
@@
 \left(\frac{E}{\TeV}\right)
-\,\cm^{-3}\,,\nonumber
+\,\cm^{-3}\,\nonumber
 \label{eq:nbIC}
 \end{eqnarray}
+at the mean density of the IGM, 
 where $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
@@
 \begin{eqnarray}
-\frac{\gamma_b^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+\frac{\gammabeam^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 \begin{eqnarray}
-\frac{\gamma_b}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+\frac{\gammabeam}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 This is a part of ongoing work and will be presented in a future publication.
-Second, while it seems that the beam plasma mode may be in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
+Second, while it seems that the beam plasma mode may be in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gammabeam}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
 Thus, both the reactive and the kinetic growth rates are similar and it likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
-Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{paperIV}.
+Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, \citet{paperIV}, and \citet{2015ApJ...811...19L}.
 
@@
 }
-\item{We analytically calculated the beam-plasma and oblique instabilities in both the reactive and kinetic regime.
+\item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regime.
 We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the oblique mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
@@
 C.P.~gratefully acknowledges support by the European Research Council under ERC-CoG grant CRAGSMAN-646955 and by the Klaus Tschira Foundation.
-E.P. acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
+E.P.
+acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
+Support for AL was provided by an Alfred P.
+Sloan Research Fellowship, NASA ATP Grant NNX14AH35G, and NSF Collaborative Research Grant \#1411920 and CAREER grant \#1455342.
 We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions.
@@
 Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
-function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
+function is  $\fCM(E_{\rm COM}(\ppos_{\rm COM},\pmom_{\rm COM}))$.  When we move to the lab
 (denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
@@
 \begin{equation}
-  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
+  \fLAB(\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
 \end{equation}
@@
 \betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
-\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_b\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+\betabarLABperp &=& \int \frac{\betaCMperp}{\gammabeam\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
 \end{eqnarray}
-where $\gamma_b = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
+where $\gammabeam = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
 For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
@@
  \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-&=& \gamma_b^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
+&=& \gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
 \label{eq:dbetabarLABperp}
@@
 \begin{equation}
- \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gamma_b^2} \approx \frac {k_B T}{\gamma_b^2 m_e c^2 }.
+ \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2} \approx \frac {k_B T}{\gammabeam^2 m_e c^2 }.
 \end{equation}
@@
 \begin{equation}
- \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gamma_b^{4}}  \approx \frac {k_B T}{\gamma_b^4 m_e c^2 },
+ \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}}  \approx \frac {k_B T}{\gammabeam^4 m_e c^2 },
 \end{equation}
@@
 \begin{eqnarray}\label{eq:dispersion reactive}
- \left(\omega^2 - \omega_{p,t}^2\right)\left[\left(\omega - k_z v_b\right)^2 - \frac{\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}\right] = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2},
+ \left(\omega^2 - \omega_{p,t}^2\right)\left[\left(\omega - k_z \vbeam\right)^2 - \frac{\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}\right] = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2},
 \end{eqnarray}
@@
 \begin{equation}
- 2\Delta\omega\omega_{p,t}\left(\Delta\omega + \omega_{p,t} - k_z v_b\right)^2 = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}.
+ 2\Delta\omega\omega_{p,t}\left(\Delta\omega + \omega_{p,t} - k_z \vbeam\right)^2 = \frac{\omega_{p,t}^2\omega_{p,b}^2}{\gamma^3}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2}.
 \end{equation}\label{eq:expanded dispersion reactive}
-For $\Delta\omega \ll \omega_{p,t} - k_z v_b$, $\Delta\omega$ is real and there is no instability.  However, if $k_z = \omega_{p,t}/v_b$, we then have
+For $\Delta\omega \ll \omega_{p,t} - k_z \vbeam$, $\Delta\omega$ is real and there is no instability.  However, if $k_z = \omega_{p,t}/\vbeam$, we then have
 \begin{equation}\label{eq:approximate dispersion reactive}
@@
 \end{equation}
-where we have multiplied the fraction on the right hand side by $(v_b/\omega_{p,t})^2/(v_b/\omega_{p,t})^2$ and $Z_x = k_xv_b/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.
+where we have multiplied the fraction on the right hand side by $(\vbeam/\omega_{p,t})^2/(\vbeam/\omega_{p,t})^2$ and $Z_x = k_x\vbeam/\omega_{p,t}$ is the dimensionless wavevector perpendicular to the beam direction.
 Equation (\ref{eq:approximate dispersion reactive}) gives three solutions for $\Delta\omega$: one real and two imaginary (one growing and one damping).  The maximum growth rate is then
@@
 \begin{eqnarray}
-F_B &=& \frac{m_e c^2}{4\pi\gamma_b k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gamma_b(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T}\right)\\
-%F_b &=& \frac{1}{4\pi\gamma_b m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
-&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gamma_b K_2(\mu)}  \exp\left(-\frac {\gamma_b\gz(c^2 - v_{b,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_bv_{b,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
+F_B &=& \frac{m_e c^2}{4\pi\gammabeam k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T}\right)\\
+%F_b &=& \frac{1}{4\pi\gammabeam m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
+&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gammabeam K_2(\mu)}  \exp\left(-\frac {\gammabeam\gz(c^2 - v_{b,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
 \end{eqnarray}
@@
 \left.
--\frac{k\gamma_b}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{b,z'}\right)
+-\frac{k\gammabeam}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{b,z'}\right)
 F_b
@@
   4\pi m_e^{4}c^{5}K_2(\mu)
-} \exp\left(-\frac {\gamma_b\gph(c^2 - v_{b,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gamma_bv_{b,x'} p_{x'}}{k_BT}\right).
+} \exp\left(-\frac {\gammabeam\gph(c^2 - v_{b,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT}\right).
 \end{aligned}
@@
 %\left.
-%\gamma_b(E-v_{bz} p_z -v_{bx} p_x)/T
+%\gammabeam(E-v_{bz} p_z -v_{bx} p_x)/T
 %\right|_{\rm pole}
 %=
-%\gamma_b \gph ( 1-v_{bz}\vph ) E_\perp/T - \gamma_b v_{bx} p_x/T\,.
+%\gammabeam \gph ( 1-v_{bz}\vph ) E_\perp/T - \gammabeam v_{bx} p_x/T\,.
 %\label{eq:Eopole}
@@
 %and therefore,
-%where $v_b\rightarrow v_{bz}$ and the only new term is the final
+%where $\vbeam\rightarrow v_{bz}$ and the only new term is the final
 %exponential factor which depends upon $p_x$.
@@
 \end{equation}
-and $\cG\equiv\gamma_b\gph(1-v_{b,z'}\vph/c^2)$ and
-$w\equiv \gamma_bv_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
-\begin{equation}
-\cG E_\perp - \gamma_b v_{b,x'} p_{x'} = \gamma_b(E - v_{b,z'} p_{z'} - v_{b,x'}p_{x'})|_{v_{b,z'}=v_{\rm ph}} > 0
+and $\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$ and
+$w\equiv \gammabeam v_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
+\begin{equation}
+\cG E_\perp - \gammabeam v_{b,x'} p_{x'} = \gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'}p_{x'})|_{v_{b,z'}=v_{\rm ph}} > 0
 \end{equation}
@@
 +
-\frac{\gamma_b^2 v_{b,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
 \right] \exp\left(-\cG'\mu\right)
@@
 \Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gamma_b \mu K_2(\mu) \cG'^3}
+\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
 \left[
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gamma_b (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
+where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
 %The above differs from the beam plasma instability only due to the
@@
 
-%With an eye toward expanding to lowest order in $v_{bx}\ll1/\gamma_b\simeq1/\gph$, we find
+%With an eye toward expanding to lowest order in $v_{bx}\ll1/\gammabeam\simeq1/\gph$, we find
 %\begin{equation}
@@
 %=
-%1 + \frac{\gamma_b^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
+%1 + \frac{\gammabeam^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
 %\quad\Rightarrow\quad
-%\cG'\simeq \cG - \frac{\gamma_b^2 v_{bx}^2}{2 \cG}\frac{m^2}{T^2}.
+%\cG'\simeq \cG - \frac{\gammabeam^2 v_{bx}^2}{2 \cG}\frac{m^2}{T^2}.
 %\end{equation}  
-%Thus, to lowest order in $\gamma_b v_{bx}$, equation (\ref{eq:OGgen}) becomes
+%Thus, to lowest order in $\gammabeam v_{bx}$, equation (\ref{eq:OGgen}) becomes
 %\begin{equation}
@@
 %- \Gamma_0
-%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
+%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gammabeam T K_2(m/T)}
 %\frac{e^{-\cG}}{\cG^3}
@@
 %  +
-%  \frac{\gamma_b^2 v_{bx}^2}{\cG^2} \frac{m^2}{T^2}
+%  \frac{\gammabeam^2 v_{bx}^2}{\cG^2} \frac{m^2}{T^2}
 %\right)\\
@@
 %\left( 
-%  \cG^2 + 2\cG + 2 - \gamma_b^2 v_{bx}^2\frac{m^2}{T^2}  - \frac{\gamma_b^2 v_{bx}\qc{^2}}{\cG}\frac{m^2}{T^2}
+%  \cG^2 + 2\cG + 2 - \gammabeam^2 v_{bx}^2\frac{m^2}{T^2}  - \frac{\gammabeam^2 v_{bx}\qc{^2}}{\cG}\frac{m^2}{T^2}
 %\right)
 %+
-%\frac{\gamma_b^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG + 2\right)
+%\frac{\gammabeam^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG + 2\right)
 %\right]
@@
 %\left(
-%  1 + 3 \frac{\gamma_b^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
+%  1 + 3 \frac{\gammabeam^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
 %\right)
 %\left(
-%  1 + \frac{\gamma_b^2 v_{bx}^2}{2\cG}\frac{m^2}{T^2}
+%  1 + \frac{\gammabeam^2 v_{bx}^2}{2\cG}\frac{m^2}{T^2}
 %\right)\\
@@
 %- \Gamma_0
-%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gamma_b T K_2(m/T)}
+%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gammabeam T K_2(m/T)}
 %e^{-\cG}\\
@@
 %  +
-%  \frac{\gamma_b^2 v_{bx}^2}{2}\frac{m^2}{T^2}
+%  \frac{\gammabeam^2 v_{bx}^2}{2}\frac{m^2}{T^2}
 %  \frac{
@@
 %\end{equation}
-%where $\Gamma_0\equiv \omega_{Pt} (n_b/n_t) \gamma_b m v_{bz}^2/T$, i.e.,
+%where $\Gamma_0\equiv \omega_{Pt} (n_b/n_t) \gammabeam m v_{bz}^2/T$, i.e.,
 %again $v_{bz}$.
@@
 %\vph\simeq\sqrt{2\delta Z_x}\,,\quad
-%v_{bx}=(1-\delta Z_x) v_b\,,\quad
-%v_{bz}\simeq\sqrt{2\delta Z_x}\,v_b\,,\quad\\
+%v_{bx}=(1-\delta Z_x) \vbeam\,,\quad
+%v_{bz}\simeq\sqrt{2\delta Z_x}\,\vbeam\,,\quad\\
 %v_{bz}-\vph\simeq\left\{
 %\begin{aligned}
-%&\frac{\sqrt{T/m}}{\gamma_b} & T\ll m\\
-%&\frac{1}{2\gamma_b} & T\gg m
+%&\frac{\sqrt{T/m}}{\gammabeam} & T\ll m\\
+%&\frac{1}{2\gammabeam} & T\gg m
 %\end{aligned}
@@
 %\end{equation}
-%\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gamma_b^2$ for the hot beam case \ldots]}\\
-%where $\delta Z_x\equiv 1 - v_{bx}/v_b$.  Inserting these into Equation
+%\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gammabeam^2$ for the hot beam case \ldots]}\\
+%where $\delta Z_x\equiv 1 - v_{bx}/\vbeam$.  Inserting these into Equation
 %(\ref{eq:OGgen}) suffices to show that there exists growth rates that
@@
 %\end{equation}
-%for $\gamma_b\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).  
+%for $\gammabeam\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).  
 %

### 2016-07-12 oblique.tex
more updates

@@
 \newcommand\cc[1]{{\color{blue} \bf #1}}
-
+\newcommand\phil[1]{{\color{cyan} #1}}
 \newcommand{\mathsc}[1]{\text{\textsc{#1}}}
@@
 In addition, \citet{Sironi+14} performed particle-in-cell simulations of these plasma processes and argued that these processes saturate at a very low level.
-It is unclear, however, if the conclusions of their work are robust as it is unclear if their numerical setup is adequate (e.g.
-Mohamad's paper).
+It is unclear, however, if the conclusions of their work is applicable to the parameter regime of blazar heating.
 \cp{[This last sentence needs to be weakened or supported with a reference, e.g.
-by Mohamad's paper provided it is out by then.]}
+by Mohamad's paper provided it is out by then.]}\phil{[Hopefully this is sufficiently weak]}
 
@@
 \caption{Oblique growth rate maximized over $\vph$ as a function of
-  \cp{\sout{$v_{bx}$ and $T$}$\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector.\sout{ for various $\gammabeam$.
-  In each the green line shows the chord along which $v_{bx}/\vbeam = 1/\gammabeam$.
-  In all cases, the maximum growth rate occurs when
-  $v_{bx}/\vbeam\gg1/\gammabeam$.}} \cp{[Which $\gammabeam$ was assumed for this -- $10^6$?
-Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
+ $\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector for $\gammabeam = 10^6$. The maximum growth rate occurs when
+  $\sin\theta \gg1/\gammabeam$.
+\cp{[ Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
 In the paper, the notation $T$ is used instead of $T_b$ here -- can you unify this?]}}\label{fig:ObliqueGMZ}
@@
 beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth
-rate reaches it maximum value $\cos\theta \gg 1/\gammabeam$ \cp{[This is not
-    clear, as the axis shows $\sin\theta$.]}, i.e., at an oblique, but not
-perpendicular angle.  Moreover the maximum growth rates vary little and are
+rate reaches it maximum value $\sin\theta \gg 1/\gammabeam$, i.e., at an oblique angle.
+Note that at $\sin\theta = 0$, we recover the beam-plasma instability.
+Moreover the maximum growth rates vary little and are
 robust for broad range of angles between the wavevector and the beam direction.
@@
   $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gammabeam m \vbeam^2/T$ as a function of
-  $k_BT/m_ec^2$ and \cp{$\gammabeam-1$}.  \cp{\sout{Note that unlike the beam plasma case
-      shown in Figure \ref{fig:Ggen}, at high $\gammabeam$ the transition between
-      high and low temperature is only marginal, constituting a roughly $10\%$
-      reduction.} Unlike the beam-plasma case on the right, in the case of the
+  $k_BT/m_ec^2$ and $\gammabeam-1$.  Unlike the beam-plasma case on the right, in the case of the
     oblique kinetic growth rate at high $\gammabeam$ the transition between high
     and low temperature is only marginal, constituting a roughly $10\%$
-    reduction.}  \cp{[Can you please add $k_B$ instead of $k$ to the axis label
+    reduction.  \cp{[Can you please add $k_B$ instead of $k$ to the axis label
       and $\Gamma/\Gamma_0$ at the color bar? In the paper, the notation $T$ is
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \rightarrow \cos\theta = -1$\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}, allowing us to ignore the dependence on $\theta$ in Equation (\ref{eq:cross section}).
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \approx \pi$.
+As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}\phil{[hope this clarifies it.]}
 We note several feature of this Figure.
@@
 
-Given that the range of $E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, this corresponds to values of $\alpha_{\perp}'$ and $\alpha_{\parallel}'$ between $\sim 0.3-2$.
-\cp{[Define $\alpha_{\perp}'$ and $\alpha_{\parallel}'$.]}  To determine whether or not the reactive or kinetic instabilities apply to these beams, it is necessary to determine $n_b/n_t$.
+Given that the range of $k_BT/m_ec^2 = E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+First, it is necessary to determine $n_b/n_t$.
 Here the target is the background IGM, so $n_t = \nIGM$, where $\nIGM \approx 2\times 10^{-7}\,(1+\delta)\,(1+z)^3\,\cm^{-3}$ is the mean density of the IGM, $z$ is the redshift, and $\delta$ denotes the overdensity.
@@
 \end{eqnarray}
-where the second term is zero because it is odd\cc{, $\lambda_D^2 = k_B T/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
+where the second term is zero because it is odd, $\lambda_D^2 = k_B T/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
 These wave are not legitimate and result from the Taylor expansion of the denominator of equation (\ref{eq:dispersion-real-part}).
-A correct treatment of equation (\ref{eq:dispersion-real-part}) with the appropriate Landau contours will give the correct growing or damping behavoirs for waves with phase speeds approximately that of the electron phase speeds.}}.
+A correct treatment of equation (\ref{eq:dispersion-real-part}) with the appropriate Landau contours will give the correct growing or damping behavoirs for waves with phase speeds approximately that of the electron phase speeds.}.
 If we ignore the third term in the kinetic dispersion relation (\ref{eq:dispersion kinetic}), this yields two plasma modes: an undamped plasma oscillation mode with $\omega = \omega_{p,t}$ and a longitudinal electron plasma wave, i.e., Langmuir wave, with

### 2016-07-12 oblique.tex
unified description of T_b

@@
 \begin{equation}\label{eq:relativistic maxwellian}
-f\propto \exp\left(-\frac E T\right),
+f\propto \exp\left(-\frac E {k_BT}\right),
 \end{equation}
@@
 %f = \frac{1}{4\pi\gammabeam m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
-f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T}\right)
+f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T_b}\right),
 \end{eqnarray}
-where $K_2$ is the 2nd order modified Bessel function.
+where $K_2$ is the 2nd order modified Bessel function and $T_b$ is the comoving temperature of the beam.
 
@@
 Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
-Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
+Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT_t  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
 \begin{equation}\label{eq:dispersion relation reactive}
@@
 \begin{equation}
- F_t = \left(\frac{1}{2\pi m_e k_B T}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T}\right),
-\end{equation}
-where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T$ is the temperature of the background.
+ F_t = \left(\frac{1}{2\pi m_e k_B T_t}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T_t}\right),
+\end{equation}
+where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T_t$ is the temperature of the target background plasma.
 For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.~\ref{eq:distribution function beam}).
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
+where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
 \begin{equation}
@@
 \caption{Oblique growth rate maximized over $\vph$ as a function of
- $\sin\theta$ and $k_BT/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector for $\gammabeam = 10^6$. The maximum growth rate occurs when
+ $\sin\theta$ and $k_BT_b/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector for $\gammabeam = 10^6$. The maximum growth rate occurs when
   $\sin\theta \gg1/\gammabeam$.
@@
 $\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$, i.e., the angle between the
-beam and the wave vector, and $k_BT/m_e c^2$.  Here, it is clear that the growth
+beam and the wave vector, and $k_BT_b/m_e c^2$.  Here, it is clear that the growth
 rate reaches it maximum value $\sin\theta \gg 1/\gammabeam$, i.e., at an oblique angle.
@@
 Because the maximal oblique growth rate varies little for a wide range of angles
-and relativistic temperatures for $k_B T/m_ec^2 > 1$, we maximize over $\theta$
+and relativistic temperatures for $k_B T_b/m_ec^2 > 1$, we maximize over $\theta$
 in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$
@@
 \begin{aligned}
-& 0.38 \Gamma_0 & k_BT/m_ec^2\ll 1\\
-& 0.34 \Gamma_0 & k_BT/m_ec^2\gg 1
+& 0.38 \Gamma_0 & k_BT_b/m_ec^2\ll 1\\
+& 0.34 \Gamma_0 & k_BT_b/m_ec^2\gg 1
 \end{aligned}
@@
   $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gammabeam m \vbeam^2/T$ as a function of
-  $k_BT/m_ec^2$ and $\gammabeam-1$.  Unlike the beam-plasma case on the right, in the case of the
+  $k_BT_b/m_ec^2$ and $\gammabeam-1$.  Unlike the beam-plasma case on the right, in the case of the
     oblique kinetic growth rate at high $\gammabeam$ the transition between high
@@
 sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for
-$k_BT/m_ec^2 \ll 1$ to $\Gamma/\Gamma_0 \approx 0.1$ for $k_BT/m_ec^2 \gtrsim
+$k_BT_b/m_ec^2 \ll 1$ to $\Gamma/\Gamma_0 \approx 0.1$ for $k_BT_b/m_ec^2 \gtrsim
 1$.  Note, however, that that beam-plasma growth rate remains competitive with
@@
 \begin{equation}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gammabeam^2 m_e c^2 },
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 },
 \end{equation}
@@
 \begin{equation}
-\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gammabeam^4 m_e c^2 },
+\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT}{m_ec^2}},\\
-\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT}{m_ec^2}}.
+\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT_b}{m_ec^2}},\\
+\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.
 \end{eqnarray}
@@
  \begin{array}{cl}
- ({k_BT}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
-({k_BT}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+ ({k_BT_b}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
+({k_BT_b}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 \begin{array}{cl}
-({k_BT}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
-({k_BT}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
+({k_BT_b}/{m_ec^2})^{-1/2}\gammabeam^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{oblique} \\
+({k_BT_b}/{m_ec^2})^{-1/2}\gammabeam\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
 \end{array}
@@
 
-Given that the range of $k_BT/m_ec^2 = E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
 First, it is necessary to determine $n_b/n_t$.
@@
 \begin{eqnarray}
-\frac{\gammabeam^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+\frac{\gammabeam^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 \begin{eqnarray}
-\frac{\gammabeam}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+\frac{\gammabeam}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
 &&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
 
@@
 This is a part of ongoing work and will be presented in a future publication.
-Second, while it seems that the beam plasma mode may be in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gammabeam}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
+Second, while it seems that the beam plasma mode may be in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gammabeam}/{\sqrt{k_BT_b/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.
 Thus, both the reactive and the kinetic growth rates are similar and it likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
@@
 \begin{equation}
- \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2} \approx \frac {k_B T}{\gammabeam^2 m_e c^2 }.
+ \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 }.
 \end{equation}
@@
 \begin{equation}
- \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}}  \approx \frac {k_B T}{\gammabeam^4 m_e c^2 },
+ \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}}  \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
 \end{equation}
@@
 \end{eqnarray}
-where the second term is zero because it is odd, $\lambda_D^2 = k_B T/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
+where the second term is zero because it is odd, $\lambda_D^2 = k_B T_t/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
 These wave are not legitimate and result from the Taylor expansion of the denominator of equation (\ref{eq:dispersion-real-part}).
@@
 \begin{eqnarray}
-F_B &=& \frac{m_e c^2}{4\pi\gammabeam k_B T K_2(m_ec^2/k_B T )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T}\right)\\
+F_B &=& \frac{m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T_b}\right)\\
 %F_b &=& \frac{1}{4\pi\gammabeam m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
-&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gammabeam K_2(\mu)}  \exp\left(-\frac {\gammabeam\gz(c^2 - v_{b,z'}v_{z'})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT}\right),\label{eq:dist2}
+&=& \frac{\mu m_e^{-3}c^{-3}}{4\pi\gammabeam K_2(\mu)}  \exp\left(-\frac {\gammabeam\gz(c^2 - v_{b,z'}v_{z'})E_\perp)} {k_BT_b c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT_b}\right),\label{eq:dist2}
 \end{eqnarray}
-where we define $\mu = m_ec^2/k_BT$.
+where we define $\mu = m_ec^2/k_BT_b$.
 Inserting the equation into (\ref{eq:dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
@@
 \left.
--\frac{k\gammabeam}{k_BT} \left( \frac{p_{z'} c^2}{E} - v_{b,z'}\right)
+-\frac{k\gammabeam}{k_BT_b} \left( \frac{p_{z'} c^2}{E} - v_{b,z'}\right)
 F_b
@@
   4\pi m_e^{4}c^{5}K_2(\mu)
-} \exp\left(-\frac {\gammabeam\gph(c^2 - v_{b,z'}v_{\rm ph})E_\perp)} {k_BT c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT}\right).
+} \exp\left(-\frac {\gammabeam\gph(c^2 - v_{b,z'}v_{\rm ph})E_\perp)} {k_BT_b c^2}\right)\exp\left(\frac{\gammabeam v_{b,x'} p_{x'}}{k_BT_b}\right).
 \end{aligned}
@@
 \begin{equation}\label{eq:I-integral}
-\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T}\right)
+\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right)
 \end{equation}
@@
 \kI &= 
-\int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T}\right)\\
+\int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T_b}\right)\\
 %&=
@@
   E'_\perp + \frac{w^2}{2} \frac{p'^2_\perp}{E'_\perp}
-\right) \exp\left(-\frac{\cG'E'_\perp}{k_B T}\right)\,,
+\right) \exp\left(-\frac{\cG'E'_\perp}{k_B T_b}\right)\,,
 \end{aligned}
@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T)$ is the typical maximum growth rate.
+where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
 %The above differs from the beam plasma instability only due to the

### 2016-07-19 oblique.tex
updated paper -- near final draft

@@
 Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.
-This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed target plasma.
+%This unstable wave corresponds to a destabilizing beam and does not correspond to any wave in the unperturbed target plasma.
 
@@
 \cp{[ Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
-In the paper, the notation $T$ is used instead of $T_b$ here -- can you unify this?]}}\label{fig:ObliqueGMZ}
+In the paper, the notation $T$ is used instead of $T_b$ here -- can you unify this?]}\phil{[Changed $k\rightarrow k_B$, but color bar is self-evident.]}}\label{fig:ObliqueGMZ}
 \end{figure}
@@
     and low temperature is only marginal, constituting a roughly $10\%$
-    reduction.  \cp{[Can you please add $k_B$ instead of $k$ to the axis label
-      and $\Gamma/\Gamma_0$ at the color bar? In the paper, the notation $T$ is
-      used instead of $T_b$ here -- can you unify this?]}} \label{fig:OGgen}
+    reduction. } \label{fig:OGgen}
 \end{figure*}
@@
 the oblique growth rate, i.e., it is not orders of magnitude lower.
+
+In Paper I, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
+While we also noted oblique instability was in the kinetic regime in Paper I, we made no effort to study the regime of instability of the beam-plasma case.
+Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
+We have also shown that the kinetic growth rates for both the oblique and beam-plasma instabilities are similar.
+As a result, both instabilities (beam-plasma and oblique) have similar growth rates for the case of  pair beams from TeV blazers.
 \cp{[Explain why we found the opposite result in Paper I, i.e., a negligible
-    beam-plasma growth rate in comparison to the oblique growth rate.]}
+    beam-plasma growth rate in comparison to the oblique growth rate.]}\phil{[Hope this clarifies it.]}
 

### 2016-07-19 oblique.tex
final draft

@@
 \end{equation}
-where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.  This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the wave perpendicular to the wavevector, $\cG$ is the Lorentz factor of a beam particle in a frame comoving with wave at the phase velocity, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
+This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
 \begin{equation}

### 2016-07-20 oblique.tex
removed comments

@@
 It is unclear, however, if the conclusions of their work is applicable to the parameter regime of blazar heating.
-\cp{[This last sentence needs to be weakened or supported with a reference, e.g.
-by Mohamad's paper provided it is out by then.]}\phil{[Hopefully this is sufficiently weak]}
 
@@
  $\sin\theta$ and $k_BT_b/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector for $\gammabeam = 10^6$. The maximum growth rate occurs when
-  $\sin\theta \gg1/\gammabeam$.
-\cp{[ Can you please add $k_B$ instead of $k$ to the axis label and $\Gamma/\Gamma_0$ at the color bar?
-In the paper, the notation $T$ is used instead of $T_b$ here -- can you unify this?]}\phil{[Changed $k\rightarrow k_B$, but color bar is self-evident.]}}\label{fig:ObliqueGMZ}
+  $\sin\theta \gg1/\gammabeam$. }\label{fig:ObliqueGMZ}
 \end{figure}
@@
 As a result, both instabilities (beam-plasma and oblique) have similar growth rates for the case of  pair beams from TeV blazers.
-\cp{[Explain why we found the opposite result in Paper I, i.e., a negligible
-    beam-plasma growth rate in comparison to the oblique growth rate.]}\phil{[Hope this clarifies it.]}
 
@@
 We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \approx \pi$.
-As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.\cp{ [What do you mean by $\theta\to\cos\theta$: this is not clear!]}\phil{[hope this clarifies it.]}
+As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.
+
 We note several feature of this Figure.
@@
 Additionally, the break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
-\cp{[What $\Dpp$ does this correspond to?]}
 

### 2016-07-22 oblique.tex
added a footnote in regard to Langmuir wave

@@
 
-As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.  Expanding the denominator in powers of $v$, we find
+As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.
+Expanding the denominator in powers of $v$, we find\footnote{An alert reader will note that $\gamma$ will contribute to expansion in power of $v$ to second order.
+These contributions are the result of the minor deviations from one of the Lorentz factor of the nonrelativistic electrons at order $v^2/c^2$ and effectively corrects the mass of the electron.
+We ignore these effects while keeping the $O(v^2/c^2)$ correction from the Langmuir wave because the Lorentz corrections does not depend on the wavevector while the Langmuir corrections do.}
 \begin{eqnarray}

### 2016-07-24 oblique.tex
added footnote for Ewald's comment

--- 2016-07-20 oblique.tex
@@
 
-As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.  Expanding the denominator in powers of $v$, we find
+As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.
+Expanding the denominator in powers of $v$, we find\footnote{An alert reader will note that Lorentz factor, $\gamma$, and the second term in the numerator both contribute to the expansion in powers of $v$ at second order.
+These contributions are the result of the minor deviations from the Lorentz factor of the nonrelativistic electrons and the subtle different between momentum and velocity at order $v^2/c^2$.
+These corrections correct the plasma frequency, $\omega_p$ at order $v^2/c^2$, but do not change the physics of the oscillations, i.e., they are independent of the wavevector.
+Hence, we ignore these effects while keeping the $O(v^2/c^2)$ correction that determine the Langmuir wave because these corrections depend on the wavevector.}
 \begin{eqnarray}

### 2016-07-25 oblique.tex
Marked up through intro

@@
 \newcommand\ac[1]{{\color{green} \bf #1}} %% AEB color
+\newcommand\ab[1]{{\color{green} \bf #1}} %% AEB color
 \newcommand\cp[1]{{\color{blue} \bf #1}} %% CP color
@@
 The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
-Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that cool them.
+Here we study the initial conditions of these \epm\ \ab{[Should this be $e^+$-$e^-$ or $e^\pm$?]} beams and the linear phase of the plasma instabilities that cool them.
 Starting from an isotropic distribution in the center of mass frame, we show that the velocity dispersions parallel and transverse to the direction of beam propagation become highly anisotropic under Lorentz boosts and scale differently with the Lorentz factor, $\gamma$.
@@
 In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
-Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we not only confirm that these beams are unstable to the kinetic oblique mode.
+Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we \ab{\sout{not only}?} confirm that these beams are unstable to the kinetic oblique mode.
 For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the reactive beam-plasma instability.
@@
 These results strengthen the case that powerful plasma instabilities  sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space, which has profound consequences on constraints of the intergalactic magnetic field, the thermal state of the intergalactic medium, and the formation of structure in the universe.
+\ab{[The only issue with the above is that the primary issues people have raised are with nonlinear effects, e.g., saturation or nonlinear coupling to existing background structures.
+Should we be quite as strident then in the concluding sentences?]}
 \end{abstract}
@@
 
-The \Fermi satellite and ground-based imaging atmospheric Cerenkov telescopes such as
+The \Fermi satellite and ground-based imaging atmospheric Cerenkov \ab{[Cherenkov?  I think MAGIC, VERITAS, and HESS all use the ``h'' on their websites.]} telescopes such as
 % Phil: I added "imaging atmospheric" to distinguish this class of
@@
 Using this understanding, we then develop a simple description of the distribution function of the beam, which we then use to calculate the unstable modes analytically in both the reactive (hydrodynamic) and kinetic regimes.
+
+\ab{[New paragraph]}
 Here the reactive instability refers to the instability where the entire beam participates in the instability.

### 2016-07-25 oblique.tex
Comments through sec 2

@@
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
-frame. As a result, an $e^\pm$ pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
+frame. As a result, an $e^\pm$ \ab{[\epm elsewhere]} pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
 c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
@@
 where $K_2$ is the 2nd order modified Bessel function and $T_b$ is the comoving temperature of the beam.
+
+\ab{[Do we want to describe the characteristics of this distribution at any further length?
+I recall you had a detailed discussion of the characteristics of a boosted thermal distribution which was nice because it clarified some obvious misconceptions that seemed to continuously arise.
+Also, do we want to discuss the potential evolution of the distribution?]}
 

### 2016-07-25 oblique.tex
Comments through end of main text.

@@
 \end{equation}
-where $\epsilon$ is the simplified dielectric function, and for each species $\omega_{p,s}^2\equiv 4\pi e^2 n_s/m_e$ is the plasma frequency, $n_s\equiv\int f_{0s} d^3\!p$ is the number density, and $F_s\equiv f_{0s} / n_s$ is the normalized distribution function.
+where $\epsilon$ is the simplified dielectric function, and for each species $\omega_{p,s}^2\equiv 4\pi e^2 n_s/m_e$ is the plasma frequency, $n_s\equiv\int f_{0s} d^3\!p$ is the number density, and $F_s\equiv f_{0s} / n_s$ is the normalized \ab{background} distribution function.
 Upon integrating by parts, Equation (\ref{eq:dispersion relation}) becomes
@@
 only the bulk response that is important.  In particular, this means that the
-beam particles are resonance with the unstable wave over a timescale much longer
-than the growth time as stated above, i.e., the beam particles
+beam particles are \ab{resonant \sout{resonance}} with the unstable wave over a timescale much longer
+than the growth time \ab{\sout{as stated above}}, i.e., the beam particles
 do not drift a distance larger than the wavelength of the unstable mode over the
@@
 over the growth time of the instability.  Moreover, the bulk of
-the plasma (background or beam) does not respond to the disturbance. Instead,
+the plasma (background or beam) does not respond to the disturbance\ab{; instead, \sout{. Instead,}}
 only a fraction of  particles is relevant for
@@
 evaluation of the dispersion relation in these two regimes, which gives two
-regimes of instability, and the delineation between these two regimes.
+regimes of instability, and the delineation between \ab{them. \sout{these two regimes.}}
 
@@
 which we denote the beam-plasma or ``two-stream'' growth rate.
-For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and $Z_x \rightarrow \infty$, the growth rate approaches oblique growth rate:
+For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and \ab{$Z_x\gg1$ \sout{$Z_x \rightarrow \infty$}}, the growth rate approaches oblique growth rate:
 \begin{equation}
@@
 
-\subsubsection{Sub-maximal instabilities}
+\subsubsection{Sub-maximal instabilities \ab{[There should be either more or fewer subsubsections]}}
 
@@
 In a continuous system, these waves do not matter in comparison to the unstable mode at $k\lambda_D = 1$.
-However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these \ab{sub-maximal \sout{sub maximal}} modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
 
@@
 Here all the particles participate in the instability.
-However, kinetic theory masks another regime of the instability, where only a fraction of the particle participate in the instability, i.e., the kinetic regime.
+However, kinetic theory \ab{marks \sout{masks}} another regime of the instability, where only a fraction of the particle participate in the instability, i.e., the kinetic regime.
 We now derive the growth rate of the instability in the kinetic regime.
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T_t$ is the temperature of the target background plasma.
-For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (eq.~\ref{eq:distribution function beam}).
+For the beam plasma, we \ab{again} adopt the Maxwell-J{\"u}ttner distribution (\ab{Equation \sout{eq.}}~(\ref{eq:distribution function beam})).
 %\begin{equation}\label{eq:distribution function beam}
@@
 %\end{equation}
-Inserting these into the dispersion relation (\ref{eq:dispersion relation}), we find
+Inserting these into the dispersion relation\ab{, Equation} (\ref{eq:dispersion relation}), we find
 %\begin{equation}\label{eq:dispersion kinetic}
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the wave perpendicular to the wavevector, $\cG$ is the Lorentz factor of a beam particle in a frame comoving with wave at the phase velocity, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the wave perpendicular to the wavevector, $\cG$ is the Lorentz factor of a beam particle in a frame comoving with wave at the phase velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
 This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
@@
 beam and the wave vector, and $k_BT_b/m_e c^2$.  Here, it is clear that the growth
-rate reaches it maximum value $\sin\theta \gg 1/\gammabeam$, i.e., at an oblique angle.
+rate reaches it maximum value $\sin\theta \gg 1/\gammabeam$, i.e., at an oblique angle \ab{[I thought it was for $\theta\gtrsim1/\gammabeam$]}.
 Note that at $\sin\theta = 0$, we recover the beam-plasma instability.
@@
 
-In Paper I, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
-While we also noted oblique instability was in the kinetic regime in Paper I, we made no effort to study the regime of instability of the beam-plasma case.
+In Paper I \ab{[Paper I is nowhere defined]}, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
+While we also noted \ab{that the} oblique instability was in the kinetic regime in Paper I, we made no effort to study the regime of instability of the beam-plasma case.
 Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
@@
 As a result, both instabilities (beam-plasma and oblique) have similar growth rates for the case of  pair beams from TeV blazers.
+\ab{[Where did we show that the beam-plasma instability is marginally reactive?
+Does this belong in the conclusions?]}
 
@@
 To begin let us return to the reactive instability.
-For the growth rate of the reactive instability (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion $\rightarrow 0$.
-To reach this limit, we demand
+For the growth rate of the reactive instability \ab{in Equation} (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion \ab{must be vanishingly small.
+Quantitatively, this limit requires \sout{$\rightarrow 0$.
+To reach this limit, we demand}}
 \begin{equation}
@@
 Taking $\kvec$ to be $\omega_{p,t}/\vel_b$, so that the RHS is $\approx \Gamma$, we find
+\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
+Thus, this is likely to be a confusing part.
+Also, if $\kvec=\omega_{p,t}/\vel_b$ then isn't hte LHS $\Delta v_\parallel/v_b$.]}
 \begin{equation}\label{eq:oblique reactive regime}
@@
 where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.
-Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate (\ref{eq:growth rate reactive}).
+Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate \ab{Equation} (\ref{eq:growth rate reactive}).
 For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability:
@@
 \end{eqnarray}
-Hence, the conditions for the reactive regime for the oblique mode (eq.~\ref{eq:oblique reactive regime}) and beam-plasma mode (eq.~\ref{eq:bp reactive regime}) can be reduced to 
+Hence, the conditions for the reactive regime for the oblique mode (\ab{\sout{eq.} Equation}~(\ref{eq:oblique reactive regime})) and beam-plasma mode (\ab{\sout{eq.} Equation}~(\ref{eq:bp reactive regime})) can be reduced to
 \begin{equation}\label{eq:generic reactive regime}
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.~\ref{eq:bp kinetic growth rate}) and oblique mode (eq.~\ref{eq:kinetic growth rate}).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (\ab{\sout{eq.} Equation}~(\ref{eq:bp kinetic growth rate})) and oblique mode (\ab{\sout{eq.} Equation}~(\ref{eq:kinetic growth rate})).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
@@
 To do so, we consider the process of photon-photon annihilation.
-For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is
+For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is \ab{[EBL or ebl?
+Should we homogenize the equation subscripts?]}
 \begin{eqnarray}
@@
 \begin{equation}
- \frac{dn_{\rm ebl}}{dE_{\rm ebl}} \propto \left\{ 
-\begin{array}{cl}
+  \frac{dn_{\rm ebl}}{dE_{\rm ebl}} \propto %\left\{
+  \begin{cases}
+    %\begin{array}{cl}
     E_{\rm ebl}^0 & E_{\rm ebl} \leq 1\,\eV \\
     E_{\rm ebl}^{-1.5} & 1\,\eV < E_{\rm ebl} \leq 10\,\eV \\
- 0 & E_{\rm ebl} > 10\,\eV 
-\end{array}
-\right.
+    0 & E_{\rm ebl} > 10\,\eV. 
+    %\end{array}
+  \end{cases}
+  %\right.
 \end{equation}

### 2016-07-26 oblique.tex
initial fixed of Avery's comments

@@
 \newcommand{\betaMEAN}{\ensuremath{\beta_{\rm M}}}
-\newcommand{\epm}{\ensuremath{e^+e^-}}
+\newcommand{\epm}{\ensuremath{e^{\pm}}}
 
@@
 The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
-Here we study the initial conditions of these \epm\ \ab{[Should this be $e^+$-$e^-$ or $e^\pm$?]} beams and the linear phase of the plasma instabilities that cool them.
+Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that cool them.
 Starting from an isotropic distribution in the center of mass frame, we show that the velocity dispersions parallel and transverse to the direction of beam propagation become highly anisotropic under Lorentz boosts and scale differently with the Lorentz factor, $\gamma$.
@@
 In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
-Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we \ab{\sout{not only}?} confirm that these beams are unstable to the kinetic oblique mode.
+Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the kinetic oblique mode.
 For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the reactive beam-plasma instability.
@@
 
-The \Fermi satellite and ground-based imaging atmospheric Cerenkov \ab{[Cherenkov?  I think MAGIC, VERITAS, and HESS all use the ``h'' on their websites.]} telescopes such as
+The \Fermi satellite and ground-based imaging atmospheric Cherenkov telescopes such as
 % Phil: I added "imaging atmospheric" to distinguish this class of
@@
 ultrarelativistic pair beams  likely are the dominant cooling mechanisms
-\citep{paperI,Schlickeiser+12,Schlickeiser+13}, depositing this energy
+(\citealt{paperI}, hereafter ACP12, \citealt{Schlickeiser+12,Schlickeiser+13}), depositing this energy
 as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
  IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
-\citep{paperI,Schlickeiser+12,Schlickeiser+13}.
+(ACP12;\citealt{Schlickeiser+12,Schlickeiser+13}).
 
@@
 To this end, a full nonlinear study is called for.
-We begin this task in this initial work where we study the distribution function of the $e^+e^-$ pairs that are produced from VHEGR-EBL photon annihilation.
+We begin this task in this initial work where we study the distribution function of the $\epm$ pairs that are produced from VHEGR-EBL photon annihilation.
 We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.
@@
 
-\ab{[New paragraph]}
 Here the reactive instability refers to the instability where the entire beam participates in the instability.
@@
 In particular, we pay careful attention to both the reactive (or hydrodynamic) and kinetic regimes of the beam plasma and oblique instabilities and the transition between the two.
-Applying these results to TeV $e^+e^-$ pair beams that arise from TeV photon pair production in Section \ref{sec:application},
+Applying these results to TeV $\epm$ pair beams that arise from TeV photon pair production in Section \ref{sec:application},
 we demonstrate that despite the extraordinary coldness of the beam we are always in the kinetic regime for the oblique mode, but may be in the reactive regime for the beam plasma mode.
@@
 of this process is that the energies of the VHEGR ($E_{\rm ph}$) and the EBL
-photon ($E_{\rm ebl}$) exceed the rest mass energy of the $e^+e^-$ pair in the
+photon ($E_{\rm ebl}$) exceed the rest mass energy of the $\epm$ pair in the
 center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
-frame. As a result, an $e^\pm$ \ab{[\epm elsewhere]} pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
+frame. As a result, an $\epm$ pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
 c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
@@
 \end{equation}
-where $\epsilon$ is the simplified dielectric function, and for each species $\omega_{p,s}^2\equiv 4\pi e^2 n_s/m_e$ is the plasma frequency, $n_s\equiv\int f_{0s} d^3\!p$ is the number density, and $F_s\equiv f_{0s} / n_s$ is the normalized \ab{background} distribution function.
+where $\epsilon$ is the simplified dielectric function, and for each species $\omega_{p,s}^2\equiv 4\pi e^2 n_s/m_e$ is the plasma frequency, $n_s\equiv\int f_{0s} d^3\!p$ is the number density, and $F_s\equiv f_{0s} / n_s$ is the normalized background distribution function.
 Upon integrating by parts, Equation (\ref{eq:dispersion relation}) becomes
@@
 only the bulk response that is important.  In particular, this means that the
-beam particles are \ab{resonant \sout{resonance}} with the unstable wave over a timescale much longer
-than the growth time \ab{\sout{as stated above}}, i.e., the beam particles
+beam particles are resonant with the unstable wave over a timescale much longer
+than the growth time, i.e., the beam particles
 do not drift a distance larger than the wavelength of the unstable mode over the
@@
 over the growth time of the instability.  Moreover, the bulk of
-the plasma (background or beam) does not respond to the disturbance\ab{; instead, \sout{. Instead,}}
+the plasma (background or beam) does not respond to the disturbance; instead,
 only a fraction of  particles is relevant for
@@
 evaluation of the dispersion relation in these two regimes, which gives two
-regimes of instability, and the delineation between \ab{them. \sout{these two regimes.}}
+regimes of instability, and the delineation between them.
 
@@
 \end{equation}
-For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{paperI}.
+For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of ACP12.
 
@@
 which we denote the beam-plasma or ``two-stream'' growth rate.
-For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and \ab{$Z_x\gg1$ \sout{$Z_x \rightarrow \infty$}}, the growth rate approaches oblique growth rate:
+For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and $Z_x\gg1$, the growth rate approaches oblique growth rate:
 \begin{equation}
@@
 In a continuous system, these waves do not matter in comparison to the unstable mode at $k\lambda_D = 1$.
-However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these \ab{sub-maximal \sout{sub maximal}} modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
+However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub-maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state in comparison to the physical system.
 
@@
 Here all the particles participate in the instability.
-However, kinetic theory \ab{marks \sout{masks}} another regime of the instability, where only a fraction of the particle participate in the instability, i.e., the kinetic regime.
+However, kinetic theory marks another regime of the instability, where only a fraction of the particle participate in the instability, i.e., the kinetic regime.
 We now derive the growth rate of the instability in the kinetic regime.
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T_t$ is the temperature of the target background plasma.
-For the beam plasma, we \ab{again} adopt the Maxwell-J{\"u}ttner distribution (\ab{Equation \sout{eq.}}~(\ref{eq:distribution function beam})).
+For the beam plasma, we again adopt the Maxwell-J{\"u}ttner distribution (Equation~(\ref{eq:distribution function beam})).
 %\begin{equation}\label{eq:distribution function beam}
@@
 %\end{equation}
-Inserting these into the dispersion relation\ab{, Equation} (\ref{eq:dispersion relation}), we find
+Inserting these into the dispersion relation (Equation \ref{eq:dispersion relation}), we find
 %\begin{equation}\label{eq:dispersion kinetic}
@@
 
-In Paper I \ab{[Paper I is nowhere defined]}, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
-While we also noted \ab{that the} oblique instability was in the kinetic regime in Paper I, we made no effort to study the regime of instability of the beam-plasma case.
+In ACP12, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
+While we also noted \ab{that the} oblique instability was in the kinetic regime in ACP12, we made no effort to study the regime of instability of the beam-plasma case.
 Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
@@
 To begin let us return to the reactive instability.
-For the growth rate of the reactive instability \ab{in Equation} (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion \ab{must be vanishingly small.
-Quantitatively, this limit requires \sout{$\rightarrow 0$.
-To reach this limit, we demand}}
+For the growth rate of the reactive instability in Equation (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion must be vanishingly small.
+Quantitatively, this limit requires
 \begin{equation}
@@
 where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.
-Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate \ab{Equation} (\ref{eq:growth rate reactive}).
+Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate Equation (\ref{eq:growth rate reactive}).
 For $Z_x \ll \gamma^{-2}$, we recover the condition for the relativistic, reactive beam-plasma instability:
@@
 \end{eqnarray}
-Hence, the conditions for the reactive regime for the oblique mode (\ab{\sout{eq.} Equation}~(\ref{eq:oblique reactive regime})) and beam-plasma mode (\ab{\sout{eq.} Equation}~(\ref{eq:bp reactive regime})) can be reduced to
+Hence, the conditions for the reactive regime for the oblique mode (Equation~(\ref{eq:oblique reactive regime})) and beam-plasma mode (Equation~(\ref{eq:bp reactive regime})) can be reduced to 
 \begin{equation}\label{eq:generic reactive regime}
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (\ab{\sout{eq.} Equation}~(\ref{eq:bp kinetic growth rate})) and oblique mode (\ab{\sout{eq.} Equation}~(\ref{eq:kinetic growth rate})).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (Equation~(\ref{eq:bp kinetic growth rate})) and oblique mode (Equation~(\ref{eq:kinetic growth rate})).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
@@
 For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is \ab{[EBL or ebl?
-Should we homogenize the equation subscripts?]}
+Should we homogenize the equation subscripts?]}\phil{[Homogenized to EBL.]}
 \begin{eqnarray}
-\Gamma_{\pm}(E_{\rm ph}) &=& \int \sigma c dn_{\rm ebl} \nonumber \\
-&=& \int \sigma\left(E_{\rm ph}, E_{\rm ebl}\right) c \frac{dn_{\rm ebl}}{dE_{\rm ebl}} dE_{\rm ebl},
+\Gamma_{\pm}(E_{\rm ph}) &=& \int \sigma c dn_{\rm EBL} \nonumber \\
+&=& \int \sigma\left(E_{\rm ph}, E_{\rm EBL}\right) c \frac{dn_{\rm EBL}}{dE_{\rm EBL}} dE_{\rm EBL},
 \end{eqnarray}
-where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm ebl}$ is the number density of EBL photons, and $E_{\rm ebl}$ is the energy of the EBL photons.
+where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm EBL}$ is the number density of EBL photons, and $E_{\rm EBL}$ is the energy of the EBL photons.
 There are two important components to this calculation -- the cross section, $\sigma$, and the spectra of the EBL.
 
-For $\sigma$, we use the results from \citet{Nikishov62} and \citet{Gould+67}, who considered a high energy photon with energy $E_{\rm ph}$ moving along the x-axis annihilating on an EBL photon with energy $E_{\rm ebl}$ moving at an angle, $\theta$, with respect to the x-axis.
+For $\sigma$, we use the results from \citet{Nikishov62} and \citet{Gould+67}, who considered a high energy photon with energy $E_{\rm ph}$ moving along the x-axis annihilating on an EBL photon with energy $E_{\rm EBL}$ moving at an angle, $\theta$, with respect to the x-axis.
 The total cross section for this process is \citep{Nikishov62,Gould+67}
@@
 \label{eq:threshold}
- E_e =  \frac{m_e c^2}{1-v_e^2/c^2} = \sqrt{\frac{1}{2}{E_{\rm ebl} E_{\rm ph}\left(1-\cos\theta\right)}}.
+ E_e =  \frac{m_e c^2}{1-v_e^2/c^2} = \sqrt{\frac{1}{2}{E_{\rm EBL} E_{\rm ph}\left(1-\cos\theta\right)}}.
 \end{equation}
@@
 Here we use the constraints from \citet{Ahar_etal:06}, who demonstrated that VHEGR emission from H 2356-309 and 1ES 1101-232 places an upper limit on the EBL that is close to the lower limit of the integrated light from galaxies \citet{Madau+00}.
-Looking at Figure 1 of \citet{Ahar_etal:06}, we note that the EBL has a flat spectrum, i.e., constant ${dn_{\rm ebl}}/{dE_{\rm ebl}}$ below $1\,\eV$ and a falling spectrum  ${dn_{\rm ebl}}/{dE_{\rm ebl}}\propto E_{\rm ebl}^{-1.5}$ with a spectral index of $\approx 1.5$ above $1\,\eV$ with a rapid cutoff above $10\,\eV$.
+Looking at Figure 1 of \citet{Ahar_etal:06}, we note that the EBL has a flat spectrum, i.e., constant ${dn_{\rm EBL}}/{dE_{\rm EBL}}$ below $1\,\eV$ and a falling spectrum  ${dn_{\rm EBL}}/{dE_{\rm EBL}}\propto E_{\rm EBL}^{-1.5}$ with a spectral index of $\approx 1.5$ above $1\,\eV$ with a rapid cutoff above $10\,\eV$.
 Thus, we adopt a simplified model:
 \begin{equation}
-  \frac{dn_{\rm ebl}}{dE_{\rm ebl}} \propto %\left\{
+  \frac{dn_{\rm EBL}}{dE_{\rm EBL}} \propto %\left\{
   \begin{cases}
     %\begin{array}{cl}
-    E_{\rm ebl}^0 & E_{\rm ebl} \leq 1\,\eV \\
-    E_{\rm ebl}^{-1.5} & 1\,\eV < E_{\rm ebl} \leq 10\,\eV \\
-    0 & E_{\rm ebl} > 10\,\eV. 
+    E_{\rm EBL}^0 & E_{\rm EBL} \leq 1\,\eV \\
+    E_{\rm EBL}^{-1.5} & 1\,\eV < E_{\rm EBL} \leq 10\,\eV \\
+    0 & E_{\rm EBL} > 10\,\eV. 
     %\end{array}
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm ebl}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm ebl}/E_{\rm ph}} \approx \pi$.
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$.
 As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.
@@
 Second, the effect of the EBL spectra on the distribution of electron COM energies is also seen in the breaks in the distribution.
-For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm ebl}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
-Additionally, the break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
+For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
+Additionally, the break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
 
@@
 The effect of the EBL spectra can be seen in different features in this plot.
-The cutoff in $d\Gamma_{\pm}/dE_{\rm ebl}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
-The break in $d\Gamma_{\pm}/dE_{\rm ebl}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
+The cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
+The break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
 The average COM energies of the produced electrons are $E_e/m_e c^2 \approx 1.3, 1.6, 2,$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
@@
 The number density of the TeV beam is more complicated as the production rate of pairs must be balanced against their loss due to plasma instabilities or ICC.
-This is discussed extensively in Paper I and will not be repeated here.
+This is discussed extensively in ACP12 and will not be repeated here.
 However, we note that the important issue here is the loss rate due to plasma instabilities, which is a nonlinear process.
-In Paper I, we assumed that the nonlinear loss rate was the same as the linear growth rate.
+In ACP12, we assumed that the nonlinear loss rate was the same as the linear growth rate.
 This remains to be shown and is the focus of ongoing work, of which this paper lays the initial foundation.
@@
 Still some progress can be made if we assume that the dominant loss rate is IC so that we can place a lower limit on the various plasma processes.
-In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (Paper I):
+In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (ACP12):
 \begin{eqnarray}
@@
 In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.~\ref{eq:generic reactive regime}) and kinetic regimes (eq.~\ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
-Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter:
+Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $\epm$ pair beams of interest, we find for the controlling parameter:
 \begin{eqnarray}
@@
 
-The discussion above confirmed our earlier estimate in Paper I that the oblique instability is in the kinetic regime.
+The discussion above confirmed our earlier estimate in ACP12 that the oblique instability is in the kinetic regime.
 However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also exist in the reactive regime.
@@
 Thus, both the reactive and the kinetic growth rates are similar and it likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
-Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, \citet{paperIV}, and \citet{2015ApJ...811...19L}.
+Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode) in ACP12 is valid, and the results of this paper buttresses the results of \citet{paperI,paperII,paperIII,paperIV}, and \citet{2015ApJ...811...19L}.
 

### 2016-07-31 oblique.tex
updated with all changes

@@
 Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that cool them.
-Starting from an isotropic distribution in the center of mass frame, we show that the velocity dispersions parallel and transverse to the direction of beam propagation become highly anisotropic under Lorentz boosts and scale differently with the Lorentz factor, $\gamma$.
-Using these results to construct a simplified distribution function, we then calculate the linear growth rate of the beam plasma and oblique instability in both the reactive and kinetic regimes.
-We reproduce the well-known reactive growth rates for both the beam plasma and oblique mode and show that they both have similar growth rates in the kinetic regime.
-In contrast to the earlier claim by the authors that the oblique mode is dominant, we instead find that both the beam-plasma (parallel) and oblique modes are similarly strong.
-In addition, we also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower $\gamma$ than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
-Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the kinetic oblique mode.
-For sufficiently cold beams in the beam direction or for sufficiently larger Lorentz factor, these beams are also unstable to the reactive beam-plasma instability.
-We show that overall growth rates for both the oblique and beam plasma growth rates are similar and in line with our earlier claims.
-These results strengthen the case that powerful plasma instabilities  sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space, which has profound consequences on constraints of the intergalactic magnetic field, the thermal state of the intergalactic medium, and the formation of structure in the universe.
-\ab{[The only issue with the above is that the primary issues people have raised are with nonlinear effects, e.g., saturation or nonlinear coupling to existing background structures.
+Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.
+We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode.
+We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.
+We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
+Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.
+These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
+%\ab{[The only issue with the above is that the primary issues people have raised are with nonlinear effects, e.g., saturation or nonlinear coupling to existing background structures.
 Should we be quite as strident then in the concluding sentences?]}
@@
 ultrarelativistic pair beams  likely are the dominant cooling mechanisms
-(\citealt{paperI}, hereafter ACP12, \citealt{Schlickeiser+12,Schlickeiser+13}), depositing this energy
+(\citealt{paperI}, hereafter BCP12, \citealt{Schlickeiser+12,Schlickeiser+13}), depositing this energy
 as heat in the intergalactic medium \citep{paperII,paperIII}.  Therefore, the lack of an observed
  IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
-(ACP12;\citealt{Schlickeiser+12,Schlickeiser+13}).
+(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}).
 
@@
 
-\ab{[Do we want to describe the characteristics of this distribution at any further length?
+%\ab{[Do we want to describe the characteristics of this distribution at any further length?
 I recall you had a detailed discussion of the characteristics of a boosted thermal distribution which was nice because it clarified some obvious misconceptions that seemed to continuously arise.
 Also, do we want to discuss the potential evolution of the distribution?]}
+
+The Maxwell-J{\"u}ttner distribution leads to an anisotropic velocity spread parallel and perpendicular to the beam's direction.
+In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale.
+The relevant results are:
+\begin{equation}\label{eq:perp dispersion}
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 },
+\end{equation}
+and 
+\begin{equation}\label{eq:parallel dispersion}
+\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
+\end{equation}
+where $T$ is measured in the COM frame of the beam.  
+These simple scalings of the perpendicular velocity dispersion and parallel velocity dispersions can be understood as a results of time dilation between two frames that are boosted relative to each other, giving one factor of $\gamma^{-1}$.
+The coordinates perpendicular to the boost axis remain invariant while the axis along the boost suffers from length contraction and gives an extra scaling of $\gamma^{-1}$ for the parallel case.
+In any case, an ultrarelativistic beam has a small velocity spread is both the parallel and perpendicular directions by factors of $\gamma^{-2}$ and $\gamma^{-1}$, respectively.
+These velocity dispersions will be important in delineating the regime of instability in \S~\ref{sec:transition}.
 
@@
 \end{equation}
-For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of ACP12.
+For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of BCP12.
 
@@
 For the case where $k_z \rightarrow 0$, the electrostatic approximation no longer holds and the full dispersion relation must be solved.\footnote{We thank Antoine Bret for helping to clarify this point.}  A solution to the full dispersion relation reveals additional modes, including the zero frequency ($k_z = 0$) filamentation or Weibel mode.
-
-\subsubsection{Sub-maximal instabilities \ab{[There should be either more or fewer subsubsections]}}
 
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the wave perpendicular to the wavevector, $\cG$ is the Lorentz factor of a beam particle in a frame comoving with wave at the phase velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with wave at its phase velocity, and $\cG'$ is the Lorentz factor of a beam particle in a frame that comoving with wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
 This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
@@
 \caption{Oblique growth rate maximized over $\vph$ as a function of
- $\sin\theta$ and $k_BT_b/m_e c^2$, where $\theta$ is the angle between the beam direction and wave vector for $\gammabeam = 10^6$. The maximum growth rate occurs when
-  $\sin\theta \gg1/\gammabeam$. }\label{fig:ObliqueGMZ}
-\end{figure}
+ $\sin\theta$ and $k_BT_b/m_e c^2$, where $\theta$ is the angle between the beam direction and wavevector for $\gammabeam = 10^6$. The maximum growth rate occurs when
+  $\sin\theta \gg1/\gammabeam$. Note that as $\sin\theta\rightarrow 0$, we asymptote to the beam-plasma growth rate.}\label{fig:ObliqueGMZ}
+\end{figure}\
 
@@
 $\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$, i.e., the angle between the
-beam and the wave vector, and $k_BT_b/m_e c^2$.  Here, it is clear that the growth
-rate reaches it maximum value $\sin\theta \gg 1/\gammabeam$, i.e., at an oblique angle \ab{[I thought it was for $\theta\gtrsim1/\gammabeam$]}.
-Note that at $\sin\theta = 0$, we recover the beam-plasma instability.
+beam and the wavevector, and $k_BT_b/m_e c^2$.  Here, it is clear that the growth
+rate reaches it maximum value $\sin\theta \gtrsim1/\gammabeam$, i.e., at an oblique angle.
+Note that as $\sin\theta \rightarrow 0$, we recover the beam-plasma instability.
 Moreover the maximum growth rates vary little and are
 robust for broad range of angles between the wavevector and the beam direction.
-Because the maximal oblique growth rate varies little for a wide range of angles
-and relativistic temperatures for $k_B T_b/m_ec^2 > 1$, we maximize over $\theta$
-in Figure \ref{fig:OGgen} and plot the maximal growth rate as a function of $T$
-and $\gammabeam - 1$.  Here for relativistic beams, the maximum growth rate varies
-little with $T$, varying by less than $10\%$ between hot and cold beams.  Thus,
+It is clear from this plots that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable.
+
+This does not imply that any individual mode, i.e., a mode with a fixed wavevector, will grow robustly.
+Any individual mode only grows when the phase velcocity of the mode in the direction of the beam are in resonance and this resonant width is narrow.
+However, the growth is robust as for any for any combination of wavevector orientation and beam temperature, there exists some mode that will grow at the maximum rate.
+
+Because there is little variation in the maximal oblique growth rate as a wavevector orientation, we plot the maximum growth rate as a function of $T$
+and $\gammabeam - 1$ in Figure \ref{fig:OGgen}.  Here for relativistic beams, the maximum growth rate varies
+little with $T$, varying by less than $10\%$ between hot and cold beams and 
 we find:
@@
 for $\gammabeam\gtrsim10$, as seen in Figure \ref{fig:OGgen}.
+Hence, unstable modes exists and robustly grow at roughly $\Gamma \approx 0.4\Gamma_0$ for nearly any value of $k_BT_b/m_ec^2$, wavevector orientation, and $\gammabeam \gg 1$.
 
@@
 the oblique growth rate, i.e., it is not orders of magnitude lower.
-
-In ACP12, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
-While we also noted \ab{that the} oblique instability was in the kinetic regime in ACP12, we made no effort to study the regime of instability of the beam-plasma case.
-Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
-We have also shown that the kinetic growth rates for both the oblique and beam-plasma instabilities are similar.
-As a result, both instabilities (beam-plasma and oblique) have similar growth rates for the case of  pair beams from TeV blazers.
-\ab{[Where did we show that the beam-plasma instability is marginally reactive?
-Does this belong in the conclusions?]}
 
@@
 For the growth rate of the reactive instability in Equation (\ref{eq:growth rate reactive}) to be valid, the velocity dispersion must be vanishingly small.
-Quantitatively, this limit requires
-\begin{equation}
- \kvec\cdot\Delta \vel \ll \omega - \kvec\cdot\vel_b.
-\end{equation}
-Taking $\kvec$ to be $\omega_{p,t}/\vel_b$, so that the RHS is $\approx \Gamma$, we find
-\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
+In particular, over the growth time of the unstable wave, the beam particles may not be spread significantly, i.e., their spread is much smaller than one wavelength.
+Quantitatively, this demands
+\begin{equation}
+ \left|\frac{\kvec\cdot\Delta \vel}{\Gamma}\right| \ll 1.
+\end{equation}
+For $k = \omega_p/v_b$, this gives
+%\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
 Thus, this is likely to be a confusing part.
-Also, if $\kvec=\omega_{p,t}/\vel_b$ then isn't hte LHS $\Delta v_\parallel/v_b$.]}
+Also, if $\kvec=\omega_{p,t}/\vel_b$ then isn't the LHS $\Delta v_\parallel/v_b$.]}
 \begin{equation}\label{eq:oblique reactive regime}
@@
 \end{equation}
-where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$.
+where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$ and that the velocity dispersion is dominated by perpendicular (to the beam) component.
 Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate Equation (\ref{eq:growth rate reactive}).
@@
 
-In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale in a Maxwell-Junter distribution.  The relevant results are: 
-\begin{equation}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 },
-\end{equation}
-and 
-\begin{equation}
-\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
-\end{equation}
-where $T$ is measured in the COM frame of the beam.
-
-Applying these results, we find
+Applying the scaling of the perpendicular (equation [\ref{eq:perp dispersion}]) and parallel velocity dispersions (equation [\ref{eq:parallel dispersion}]) to these results, we find
 \begin{eqnarray}
@@
 To do so, we consider the process of photon-photon annihilation.
-For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is \ab{[EBL or ebl?
-Should we homogenize the equation subscripts?]}\phil{[Homogenized to EBL.]}
+For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is
 \begin{eqnarray}
@@
 The number density of the TeV beam is more complicated as the production rate of pairs must be balanced against their loss due to plasma instabilities or ICC.
-This is discussed extensively in ACP12 and will not be repeated here.
+This is discussed extensively in BCP12 and will not be repeated here.
 However, we note that the important issue here is the loss rate due to plasma instabilities, which is a nonlinear process.
-In ACP12, we assumed that the nonlinear loss rate was the same as the linear growth rate.
+In BCP12, we assumed that the nonlinear loss rate was the same as the linear growth rate.
 This remains to be shown and is the focus of ongoing work, of which this paper lays the initial foundation.
@@
 Still some progress can be made if we assume that the dominant loss rate is IC so that we can place a lower limit on the various plasma processes.
-In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (ACP12):
+In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (BCP12):
 \begin{eqnarray}
@@
 
-The discussion above confirmed our earlier estimate in ACP12 that the oblique instability is in the kinetic regime.
-However, our new results suggest that the beam-plasma instability matches the oblique instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also exist in the reactive regime.
-However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
+In BCP12, we compared the cold plasma growth rates of the oblique and beam-plasma instabilities and noted that the oblique cold growth rate is larger.
+While we also noted that the oblique instability was in the kinetic regime in BCP12, which we confirmed above, we made no effort to study the regime of instability of the beam-plasma case.
+Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
+In any case, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
 First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.
@@
 Thus, both the reactive and the kinetic growth rates are similar and it likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).
-Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode) in ACP12 is valid, and the results of this paper buttresses the results of \citet{paperI,paperII,paperIII,paperIV}, and \citet{2015ApJ...811...19L}.
+Therefore, the use of the kinetic growth rate for the oblique mode (and beam-plasma mode) in BCP12 is valid, and the results of this paper buttresses the results of \citet{paperI,paperII,paperIII,paperIV}, and \citet{2015ApJ...811...19L}.
 
@@
 }
-\item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regime.
+\item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regimes.
 We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the oblique mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
-In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
+In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.
+Finally, we have shown that the growth rate of the kinetic oblique instability has broad support.
+Namely, there exists unstable modes the grow at $\approx 0.4\Gamma_0$ for any value of beam temperature and wavevector orientation for relativistic beams.}
 \item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes are different.
 In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the oblique mode.
-This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.}
+This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.
+Applied to TeV $e^{\pm}$ beams, we find that these beams are unstable to the kinetic oblique instability and (marginally to) the reactive beam-plasma instability.}
 \item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectra of the EBL.
@@
 }
-\item{We demonstrate that the oblique mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
+%\item{We demonstrate that the oblique mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
 In any case, the use of either regime produces roughly the same growth rate.
@@
 \begin{equation}
- \int \fLAB \dVLAB = \int \fCM \dVCM.
-\end{equation}
-It is well-known that under Lorentz transformations (Landau), 
+ N \equiv \int \fLAB \dVLAB = \int \fCM \dVCM.
+\end{equation}
+It is well-known that under Lorentz transformations \citep{1975ctf..book.....L}, 
 \begin{equation}
@@
 
-Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
-\begin{equation}
- \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
+Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is:
+\begin{equation}
+ \betabarCM = N^{-1}\int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
 \end{equation}
@@
 \begin{eqnarray}\label{eq:betabarLAB}
- \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
-&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
+ \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
+&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
 \end{eqnarray}
@@
 \begin{eqnarray}
-\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
-\betabarLABperp &=& \int \frac{\betaCMperp}{\gammabeam\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+\betabarLABpara &=& N^{-1}\int \frac{\betaCMpara +\betaBOOST}{1+\betaBOOST\betaCMpara} \fCM \dVCM,\\
+\betabarLABperp &=& N^{-1}\int \frac{\betaCMperp}{\gammabeam\left(1+\betaBOOST\betaCMpara\right)} \fCM \dVCM,
 \end{eqnarray}
 where $\gammabeam = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
-For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
+For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM + \betaBOOST\hat{\boldsymbol x}$.  
 However, this Galilean result no longer holds for relativistic motion.
@@
 \begin{equation}
- \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
+ \dbetabarCMi = N^{-1}\int \beta_i^2 \fCM \dVCM.
 \end{equation}
@@
 \begin{eqnarray}
-  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
- &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
- &=&\int \left(\frac{\betaCMpara -\betaBOOST}{1-\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
+  \dbetabarLABpara &=& N^{-1}\int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
+ &=& N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
+ &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
 \end{eqnarray}
@@
 \begin{eqnarray}
- \dbetabarLABperp &=& \int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-&=& \gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1-\betaMEAN^2\right)^2} \fCM \dVCM,
+ \dbetabarLABperp &=& N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
+&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaMEAN^2\right)^2} \fCM \dVCM,
 \label{eq:dbetabarLABperp}

### 2016-08-01 oblique.tex
added corresponding author

@@
 Avery E.~Broderick\altaffilmark{1,2},
-Philip Chang\altaffilmark{3},
-Christoph Pfrommer\altaffilmark{4},
-Ewald Puchwein\altaffilmark{5},
-Astrid Lamberts\altaffilmark{6},
-Mohamad Shalaby\altaffilmark{1,2,7}
+Philip Chang\altaffilmark{3,4},
+Christoph Pfrommer\altaffilmark{5},
+Ewald Puchwein\altaffilmark{6},
+Astrid Lamberts\altaffilmark{7},
+Mohamad Shalaby\altaffilmark{1,2,8}
 }
@@
 \altaffiltext{2}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
-\altaffiltext{3}{Department of Physics, University of Wisconsin-Milwaukee, 1900 E. Kenwood Boulevard, Milwaukee, WI 53211, USA; chang65@uwm.edu}
-\altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
-\altaffiltext{5}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
-\altaffiltext{6}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
-\altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
+\altaffiltext{3}{Department of Physics, University of Wisconsin-Milwaukee, 3135 North Maryland Avenue, Milwaukee, WI 53211, USA; chang65@uwm.edu}
+\altaffiltext{4}{corresponding author}
+\altaffiltext{5}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
+\altaffiltext{6}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
+\altaffiltext{7}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
+\altaffiltext{8}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
 \keywords{

### 2016-08-02 oblique.tex
fixed up Ewald comments mostly

@@
 \newcommand\cp[1]{{\color{blue} \bf #1}} %% CP color
+\newcommand\ep[1]{{\color{magenta} \bf #1}} %% EP color
 \newcommand\asout[1]{\color{green}\sout{{\rm  #1}}} %% AEB strikeout of text
@@
 However, for the relevant parameters, the growth rates calculated in either regime are similar.
-We close with a discussion the implications of this work and application of these results for nonlinear theory in Section \ref{sec:conclusions}.
+We close with a discussion of the implications of this work and application of these results for nonlinear theory in Section \ref{sec:conclusions}.
 
@@
 
-In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.
+In the COM frame of the beam, we assume that the distribution function is isotropic \ep{[a brief explanation why this is OK might be useful]}, such that $f=f(E)$ is just a function of energy.
 This equilibrium energy distribution of a relativistic thermal plasma gas is
@@
 which we denote the beam-plasma or ``two-stream'' growth rate.
-For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and $Z_x\gg1$, the growth rate approaches oblique growth rate:
+For the more general case where $Z_x\ne 0$, this becomes the oblique instability studied by \citet{Bret-Grem-Beni:10}  Indeed for $\gamma \gg 1$ and $Z_x\gg1$, the growth rate approaches the oblique growth rate:
 \begin{equation}
@@
 which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $\vbeam/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.
-In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $\vbeam/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.
+In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for the representative case of $\vbeam/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.
 For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is the plasma oscillation frequency.
@@
 \end{equation}
-where we have integrated by parts only the second term, associated with target plasma.
+where we have integrated by parts only the second term, associated with the target plasma.
 
@@
 \Gamma &\simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}\\
+\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}\\
 &\times\left[
@@
 Moreover the maximum growth rates vary little and are
-robust for broad range of angles between the wavevector and the beam direction.
+robust for a broad range of angles between the wavevector and the beam direction.
 It is clear from this plots that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable.
@@
 
-Because there is little variation in the maximal oblique growth rate as a wavevector orientation, we plot the maximum growth rate as a function of $T$
+Because there is little variation in the maximal oblique growth rate as a wavevector orientation, we plot the maximum growth rate as a function of $T_b$
 and $\gammabeam - 1$ in Figure \ref{fig:OGgen}.  Here for relativistic beams, the maximum growth rate varies
-little with $T$, varying by less than $10\%$ between hot and cold beams and 
+little with $T_b$, varying by less than $10\%$ between hot and cold beams and 
 we find:
@@
 \caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right),
-  maximized over $\theta$ and normalized by
-  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gammabeam m \vbeam^2/T$ as a function of
-  $k_BT_b/m_ec^2$ and $\gammabeam-1$.  Unlike the beam-plasma case on the right, in the case of the
+  and normalized by
+  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gammabeam m \vbeam^2/k_BT_b$ as a function of
+  $k_BT_b/m_ec^2$ and $\gammabeam-1$.  The oblique kinetic growth rate has been maximized over $\theta$. Unlike the beam-plasma case on the right, in the case of the
     oblique kinetic growth rate at high $\gammabeam$ the transition between high
@@
 This can be contrasted with the right panel of
-  Figure~\ref{fig:OGgen} where we plot the maximized beam-plasma growth rate
+  Figure~\ref{fig:OGgen} where we plot the beam-plasma growth rate
 (\ref{eq:bp kinetic growth rate}) as a function of $\gammabeam - 1$ and $k_B
@@
 \end{equation}
-For $k = \omega_p/v_b$, this gives
+For \ep{\sout{$k = \omega_p/v_b$} $k_\perp \approx \omega_p/v_b$}, this gives
 %\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
@@
 
-Applying the scaling of the perpendicular (equation [\ref{eq:perp dispersion}]) and parallel velocity dispersions (equation [\ref{eq:parallel dispersion}]) to these results, we find
+Applying the scaling of the perpendicular (equation [\ref{eq:perp dispersion}]) and parallel velocity dispersions (equation [\ref{eq:parallel dispersion}]) to these results and assuming $v_b \approx c$, we find
 \begin{eqnarray}
@@
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
-Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$  For the beam plasma case, the growth rate is roughly
+Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$ \ep{[Wouldn't it be more natural to quote a k range that grows?
+Also some more explanation why $\Delta \omega$ and $\Gamma$ are compared would be helpful.]}.
+For the beam plasma case, the growth rate is roughly
 \begin{equation}
@@
 \Gamma_{\pm}(E_{\rm ph}) &=& \int \sigma c dn_{\rm EBL} \nonumber \\
-&=& \int \sigma\left(E_{\rm ph}, E_{\rm EBL}\right) c \frac{dn_{\rm EBL}}{dE_{\rm EBL}} dE_{\rm EBL},
+&=& \int \sigma\left(E_{\rm ph}, E_{\rm EBL}, \theta\right) c \frac{dn_{\rm EBL}}{dE_{\rm EBL}} dE_{\rm EBL},
 \end{eqnarray}
-where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm EBL}$ is the number density of EBL photons, and $E_{\rm EBL}$ is the energy of the EBL photons.
+where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm EBL}$ is the number density of EBL photons, $E_{\rm EBL}$ is the energy of the EBL photons, and $\theta$ is the angle between the momentum of the VHEGR photon and the EBL photon.
 There are two important components to this calculation -- the cross section, $\sigma$, and the spectra of the EBL.
@@
 \label{eq:threshold}
- E_e =  \frac{m_e c^2}{1-v_e^2/c^2} = \sqrt{\frac{1}{2}{E_{\rm EBL} E_{\rm ph}\left(1-\cos\theta\right)}}.
+ E_e =  \frac{m_e c^2}{\sqrt{1-v_e^2/c^2}} = \sqrt{\frac{1}{2}{E_{\rm EBL} E_{\rm ph}\left(1-\cos\theta\right)}}.
 \end{equation}
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$.
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
 As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.
-
-We note several feature of this Figure.
+\ep{[Not sure I understand this.
+The $E_{\rm EBL}, E_{\rm ph}$ we get from the spectra are not in the COM frame.
+So how do we get rid of $\theta$?
+Or is this an additional assumption here that head-on collisions will dominate pair production?
+Was that assumption used in Fig.
+4?]}\phil{[Assumption i $E_{\rm EBL} \ll E_{\rm ph}$ and so in the COM frame the photons have equal energy, but opposite momentum.]}
+
+We note several features of this Figure.
 First the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy.
@@
 
-Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2$ between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2$ falls between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
 First, it is necessary to determine $n_b/n_t$.
@@
 
-Still some progress can be made if we assume that the dominant loss rate is IC so that we can place a lower limit on the various plasma processes.
-In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (BCP12):
+Still some progress can be made if we use the IC rate as a lower limit to the beam cooling rate.
+This allows as to get an upper limit on the beam density.The ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (BCP12):
 \begin{eqnarray}
@@
 &\simeq &
-3\times10^{-18}
-\left(\frac{1+z}{1}\right)^{3\zeta-7}
+2.3\times10^{-16}
+\left(\frac{1+z}{2}\right)^{3\zeta-7}
 \left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)
@@
 at the mean density of the IGM, 
-where $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
+where $L_E$ is the isotropic luminosity \phil{per unit energy}\ep{[What exactly is the deifniton of $L_E$?
+It has to have the dimension 1/time.
+Has the $2^{3\zeta-4}$ factor been absorbed in the constant assuming $\zeta=4.5$?
+This would then however only be valid for $z<1$.
+Or should it be 2 instead of 1 in the denominator?]}\phil{[Previous answer was correct, factors of 2 was correctly absorbed, but changes to 2 in denominator to be consistent with our subsequent definitions.]} of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
  the mean free path of a VHEGR 
@@
 \end{equation}
-with $\zeta=4.5$ for $z<1$ and $\zeta=0$ for $z\ge1$ \citep{Knei_etal:04, Nero-Semi:09}.
+where  $\zeta=4.5$ for $z<1$ and $\zeta=0$ for $z\ge1$ \citep{Knei_etal:04, Nero-Semi:09}.
 
@@
 \end{eqnarray}
-where $\gamma_6 = \gamma/10^6$. 
+\ep{[The numerical factors here depend on what exactly has been done for Eq. (37).]}\phil{[Numerical factors are correct.]} where $\gamma_6 = \gamma/10^6$. 
 We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
@@
 Finally, we have shown that the growth rate of the kinetic oblique instability has broad support.
-Namely, there exists unstable modes the grow at $\approx 0.4\Gamma_0$ for any value of beam temperature and wavevector orientation for relativistic beams.}
+Namely, there exists unstable modes that grow at $\approx 0.4\Gamma_0$ for any value of beam temperature and wavevector orientation for relativistic beams.}
 \item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and oblique mode, the condition for transition between the kinetic and reactive regimes are different.
@@
 C.P.~gratefully acknowledges support by the European Research Council under ERC-CoG grant CRAGSMAN-646955 and by the Klaus Tschira Foundation.
-E.P.
-acknowledges support by the ERC grant ``The Emergence of Structure during the epoch of Reionization''.
+E.P.~gratefully acknowledges support by the Kavli Foundation.
 Support for AL was provided by an Alfred P.
@@
  &=& N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
- &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaMEAN^2}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
+ &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaCMpara\betaBOOST}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
 \end{eqnarray}
@@
  \dbetabarLABperp &=& N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaMEAN^2\right)^2} \fCM \dVCM,
+&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaCMpara\betaBOOST\right)^2} \fCM \dVCM.
 \label{eq:dbetabarLABperp}
 \end{eqnarray}
-where we have defined $\betaMEAN^2 = \betaCMpara\betaBOOST$ as the geometric mean for matters of convenience.
-
-It is easier to look at the perpendicular component first.  For $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$ 
+
+It is easier to look at the perpendicular component first.
+It is also more intuitive to study how velocity dispersions scale between the center of mass frame and the lab from for non-relativistic center of mass velocity dispersion.
+Hence, for $\left|\betaCM\right| \ll 1$ \ep{[This is not the case for the distributions shown in Fig.
+4.
+Why is this good enough?]}\phil{[velocity dispersions are easiest for non-relativistic motion in COM frame.]}, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$
 \begin{equation}
@@
 \end{equation}
-This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
+\ep{[Factor 2 seems missing since perpenticular component has 2 degrees of freedom]}\phil{[Factor of two is absorb in the definition of the average (A8)]} This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
 This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
@@
 To compute the contribution from the beam term, we will reorient our coordinate system and define the $z'$-axis along the wavevector, $\kvec$.
-In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx}\xphat$.
+In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+\ep{v_{bx'}}\xphat$.
 This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.

### 2016-08-04 oblique.tex
fixed up astrid comments

@@
 \newcommand{\fLAB}{\ensuremath{f_{\rm L}}}
-\newcommand{\fCM}{\ensuremath{f_{\rm CM}}}
+\newcommand{\fCM}{\ensuremath{f_{\rm COM}}}
 
 \newcommand{\dVLAB}{\ensuremath{d^3p_{\rm L} d^3x_{\rm L}}}
-\newcommand{\dVCM}{\ensuremath{d^3p_{\rm CM} d^3x_{\rm CM}}}
+\newcommand{\dVCM}{\ensuremath{d^3p_{\rm COM} d^3x_{\rm COM}}}
 
@@
 %\newcommand{\betaLAB}{\ensuremath{\boldsymbol \beta_{\rm L}}}
-%\newcommand{\betaCM}{\ensuremath{\boldsymbol \beta_{\rm CM}}}
-\newcommand{\betaCMpara}{\ensuremath{\beta_{\rm CM,\parallel}}}
-\newcommand{\betaCMperp}{\ensuremath{\beta_{\rm CM,\perp}}}
+%\newcommand{\betaCM}{\ensuremath{\boldsymbol \beta_{\rm COM}}}
+\newcommand{\betaCMpara}{\ensuremath{\beta_{\rm COM,\parallel}}}
+\newcommand{\betaCMperp}{\ensuremath{\beta_{\rm COM,\perp}}}
 
@@
 \newcommand{\betaLAB}{\ensuremath{\bmath{\beta_{\rm L}}}}
-\newcommand{\betaCM}{\ensuremath{\bmath{\beta_{\rm CM}}}}
+\newcommand{\betaCM}{\ensuremath{\bmath{\beta_{\rm COM}}}}
 
@@
 \newcommand{\betabarLAB}{\ensuremath{\betabar_{\rm L}}}
-\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm CM}}}
+\newcommand{\betabarCM}{\ensuremath{\betabar_{\rm COM}}}
 \newcommand{\betaBOOST}{\ensuremath{\beta_b}}
@@
 \newcommand{\dbetabarLABperp}{\ensuremath{\overline{\Delta\beta^2}_{\rm L, \perp}}}
-\newcommand{\dbetabarCMi}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, i}}}
-\newcommand{\dbetabarCMpara}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, \parallel}}}
-\newcommand{\dbetabarCMperp}{\ensuremath{\overline{\Delta\beta^2}_{\rm CM, \perp}}}
+\newcommand{\dbetabarCMi}{\ensuremath{\overline{\Delta\beta^2}_{\rm COM, i}}}
+\newcommand{\dbetabarCMpara}{\ensuremath{\overline{\Delta\beta^2}_{\rm COM, \parallel}}}
+\newcommand{\dbetabarCMperp}{\ensuremath{\overline{\Delta\beta^2}_{\rm COM, \perp}}}
 
@@
 \author{
-Avery E.~Broderick\altaffilmark{1,2},
-Philip Chang\altaffilmark{3,4},
-Christoph Pfrommer\altaffilmark{5},
-Ewald Puchwein\altaffilmark{6},
-Astrid Lamberts\altaffilmark{7},
-Mohamad Shalaby\altaffilmark{1,2,8}
+Philip Chang\altaffilmark{1},
+Avery E.~Broderick\altaffilmark{2,3},
+Christoph Pfrommer\altaffilmark{4},
+Ewald Puchwein\altaffilmark{5},
+Astrid Lamberts\altaffilmark{6},
+Mohamad Shalaby\altaffilmark{2,3,7}
 }
-\altaffiltext{1}{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, ON, N2L 2Y5, Canada}
-\altaffiltext{2}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
-\altaffiltext{3}{Department of Physics, University of Wisconsin-Milwaukee, 3135 North Maryland Avenue, Milwaukee, WI 53211, USA; chang65@uwm.edu}
-\altaffiltext{4}{corresponding author}
-\altaffiltext{5}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
-\altaffiltext{6}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
-\altaffiltext{7}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
-\altaffiltext{8}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
+\altaffiltext{1}{Department of Physics, University of Wisconsin-Milwaukee, 3135 North Maryland Avenue, Milwaukee, WI 53211, USA; chang65@uwm.edu}
+\altaffiltext{2}{Perimeter Institute for Theoretical Physics, 31 Caroline Street North, Waterloo, ON, N2L 2Y5, Canada}
+\altaffiltext{3}{Department of Physics and Astronomy, University of Waterloo, 200 University Avenue West, Waterloo, ON, N2L 3G1, Canada}
+\altaffiltext{4}{Heidelberg Institute for Theoretical Studies, Schloss-Wolfsbrunnenweg 35, D-69118 Heidelberg, Germany; christoph.pfrommer@h-its.org}
+\altaffiltext{5}{Institute of Astronomy and Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge, CB3 0HA, UK}
+\altaffiltext{6}{TAPIR, Mailcode 350-17, California Institute of Technology, Pasadena, CA 91125, USA}
+\altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
 \keywords{
@@
 
-To this end, a full nonlinear study is called for.
-We begin this task in this initial work where we study the distribution function of the $\epm$ pairs that are produced from VHEGR-EBL photon annihilation.
+\phil{While, a full nonlinear study is required, we focus on the nature of the linear instability in this paper, clarifying its robustness and regimes of applicability.}  We begin by studying the distribution function of the $\epm$ pairs that are produced from VHEGR-EBL photon annihilation.
 We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.
@@
 \begin{equation}\label{eq:perp dispersion}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 },
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 },
 \end{equation}
@@
 The coordinates perpendicular to the boost axis remain invariant while the axis along the boost suffers from length contraction and gives an extra scaling of $\gamma^{-1}$ for the parallel case.
-In any case, an ultrarelativistic beam has a small velocity spread is both the parallel and perpendicular directions by factors of $\gamma^{-2}$ and $\gamma^{-1}$, respectively.
+In any case, an ultrarelativistic beam has a small velocity spread in both the parallel and perpendicular directions by factors of $\gamma^{-2}$ and $\gamma^{-1}$, respectively.
 These velocity dispersions will be important in delineating the regime of instability in \S~\ref{sec:transition}.
@@
 However, we make this approximation for two reasons.
-First, a complete calculation of the unstable modes has already been carried out by \citet{Bret-Grem-Beni:10}, who showed that the electrostatic approximation is in excellent agreement with a full calculation (modulo the Weibel instability).
+First, a complete calculation of the unstable modes has already been carried out by \citet{Bret-Grem-Beni:10}, who showed that the oblique mode is mainly electrostatic (modulo the Weibel instability)\phil{so that a electrostatic approximation to the full dispersion relation should recover the essential physics}.
 Second, the electrostatic approximation is much simpler than a full calculation and allows us to analytically calculate the unstable growth rates, while permitting a clear exposition of the relevant physics.
@@
 
-The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (eq.~\ref{eq:growth rate reactive appendix}) is
+The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (Equation (\ref{eq:growth rate reactive appendix})) is
 \begin{equation}\label{eq:growth rate reactive}
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with wave at its phase velocity, and $\cG'$ is the Lorentz factor of a beam particle in a frame that comoving with wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
 This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
@@
 \Gamma_{\rm bp} \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
+\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 beam and the wavevector, and $k_BT_b/m_e c^2$.  Here, it is clear that the growth
-rate reaches it maximum value $\sin\theta \gtrsim1/\gammabeam$, i.e., at an oblique angle.
+rate reaches its maximum value $\sin\theta \gtrsim1/\gammabeam$, i.e., at an oblique angle.
 Note that as $\sin\theta \rightarrow 0$, we recover the beam-plasma instability.
@@
 robust for a broad range of angles between the wavevector and the beam direction.
-It is clear from this plots that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable.
+It is clear from this plot that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable.
 
@@
 Any individual mode only grows when the phase velcocity of the mode in the direction of the beam are in resonance and this resonant width is narrow.
-However, the growth is robust as for any for any combination of wavevector orientation and beam temperature, there exists some mode that will grow at the maximum rate.
+However, the growth is robust as for any combination of wavevector orientation and beam temperature, there exists some mode that will grow at the maximum rate.
 
@@
 The oblique instability exists in two different regimes, raising the important question: how are the two regimes related to each other.
-While this question has been studied by many authors in the context of the beam-plasma or two-stream instability \citep[see for instance][]{Melrose86,Boyd},  a clear exposition of how these two regimes are related to each other is lacking.
+While this question has been studied by many authors in the context of the beam-plasma or two-stream instability \citep[see for instance][]{Melrose86,Boyd}, a clear exposition of how these two regimes are related to each other \phil{for the oblique instability} is lacking.
 
@@
 
-Applying the scaling of the perpendicular (equation [\ref{eq:perp dispersion}]) and parallel velocity dispersions (equation [\ref{eq:parallel dispersion}]) to these results and assuming $v_b \approx c$, we find
+Applying the scaling of the perpendicular (Equation (\ref{eq:perp dispersion})) and parallel velocity dispersions (Equation (\ref{eq:parallel dispersion})) to these results and assuming $v_b \approx c$, we find
 \begin{eqnarray}
@@
 where $\Gamma_{\pm}$ is the rate of pair production, $\sigma$ is the pair-production cross section, $n_{\rm EBL}$ is the number density of EBL photons, $E_{\rm EBL}$ is the energy of the EBL photons, and $\theta$ is the angle between the momentum of the VHEGR photon and the EBL photon.
-There are two important components to this calculation -- the cross section, $\sigma$, and the spectra of the EBL.
+There are two important components to this calculation -- the cross section, $\sigma$, and the spectrum of the EBL.
 
@@
 
-The second ingredient is the spectra of the EBL, which is not well constrained.
+The second ingredient is the spectrum of the EBL, which is not well constrained.
 Here we use the constraints from \citet{Ahar_etal:06}, who demonstrated that VHEGR emission from H 2356-309 and 1ES 1101-232 places an upper limit on the EBL that is close to the lower limit of the integrated light from galaxies \citet{Madau+00}.
@@
 We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
-As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (eq.[\ref{eq:threshold}]) and set $1-\cos\theta \approx 2$.
+As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation (\ref{eq:threshold})) and set $1-\cos\theta \approx 2$.
 \ep{[Not sure I understand this.
@@
 First the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy.
-This is because different energy photons probe different regimes of the EBL spectra.
+This is because different energy photons probe different regimes of the EBL spectrum.
 Due to the rapid cutoff in the EBL above $10\,\eV$, lower energy VHEGRs produce colder beams.
@@
 Hence we expect that these pairs are in the sub-relativistic to mildly-relativistic regimes in their COM frame.
-Second, the effect of the EBL spectra on the distribution of electron COM energies is also seen in the breaks in the distribution.
-For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
-Additionally, the break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
+Second, the effect of the EBL spectrum on the distribution of electron COM energies is also seen in the breaks in the distribution.
+For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
+Additionally, the break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
 
@@
  \caption{Differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-The effect of the EBL spectra can be seen in different features in this plot.
-The cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectra above $10\,\eV$.
-The break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectra at $1\,\eV$.
+The effect of the EBL spectrum can be seen in different features in this plot.
+The cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
+The break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
 The average COM energies of the produced electrons are $E_e/m_e c^2 \approx 1.3, 1.6, 2,$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
@@
 First, it is necessary to determine $n_b/n_t$.
-Here the target is the background IGM, so $n_t = \nIGM$, where $\nIGM \approx 2\times 10^{-7}\,(1+\delta)\,(1+z)^3\,\cm^{-3}$ is the mean density of the IGM, $z$ is the redshift, and $\delta$ denotes the overdensity.
+Here the target is the background IGM, so $n_t = \nIGM$, where $\nIGM \approx 2\times 10^{-7}\,(1+\delta)\,(1+z)^3\,\cm^{-3}$ is the mean density of the IGM, $z$ is the redshift, and $\delta$ is the overdensity.
 The number density of the TeV beam is more complicated as the production rate of pairs must be balanced against their loss due to plasma instabilities or ICC.
@@
 
-In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.~\ref{eq:generic reactive regime}) and kinetic regimes (eq.~\ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
+In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation (\ref{eq:generic reactive regime})) and kinetic regimes (Equation (\ref{eq:generic kinetic regime})) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
 Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $\epm$ pair beams of interest, we find for the controlling parameter:
@@
 Applied to TeV $e^{\pm}$ beams, we find that these beams are unstable to the kinetic oblique instability and (marginally to) the reactive beam-plasma instability.}
-\item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectra of the EBL.
+\item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectrum of the EBL.
 We found that the average energy of these beams range from $E_e/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.
@@
 
-\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
+\section{Lorentz Factor Dependence of the Distribution Function and Velocity Dispersion}
 \label{sec:lorentz}
@@
 function that is isotropic in the COM frame and depends only on energy.
-Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
+Therefore in the COM frame, which we denote with the subscript ``COM'', the distribution
 function is  $\fCM(E_{\rm COM}(\ppos_{\rm COM},\pmom_{\rm COM}))$.  When we move to the lab
@@
 \begin{equation}
-  \fLAB(\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
+  \fLAB(\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM})) = \fCM(E_{\rm COM}).
 \end{equation}
@@
 \begin{equation}
- \betabarCM = N^{-1}\int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
+ \betabarCM = N^{-1}\int \frac {\pmom_{\rm COM}}{\gamma_{\rm COM} mc} \fCM \dVCM = 0.
 \end{equation}
@@
  \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
-&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
+&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) mc} \fCM \dVCM.
 \end{eqnarray}
@@
 \begin{equation}
- \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2} \approx \frac {k_B T_b}{\gammabeam^2 m_e c^2 }.
-\end{equation}
-\ep{[Factor 2 seems missing since perpenticular component has 2 degrees of freedom]}\phil{[Factor of two is absorb in the definition of the average (A8)]} This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
+ \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }.
+\end{equation}
+\ep{[Factor 2 seems missing since perpenticular component has 2 degrees of freedom]}\phil{[Factor of 2]} This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
 This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
-Let us now consider the parallel component (eq.(\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$:
+Let us now consider the parallel component (Equation (\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$:
 \begin{equation}
@@
 To compute the contribution from the beam term, we will reorient our coordinate system and define the $z'$-axis along the wavevector, $\kvec$.
-In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+\ep{v_{bx'}}\xphat$.
+In this case we have the beam taking on a non-$z'$ component,  $\vel_b = v_{bz'}\zphat+v_{bx'}\xphat$.
 This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.

### 2016-08-05 oblique.tex
corrected a few typos in equations, added introductory lines to App. B and C

@@
 \newcommand\ab[1]{{\color{green} \bf #1}} %% AEB color
-\newcommand\cp[1]{{\color{blue} \bf #1}} %% CP color
+\newcommand\cp[1]{{\color{blue} #1}} %% CP color
 \newcommand\ep[1]{{\color{magenta} \bf #1}} %% EP color
@@
 
-The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (Equation (\ref{eq:growth rate reactive appendix})) is
+The solution to Equation (\ref{eq:dispersion relation reactive}) is given in Appendix \ref{sec:solution reactive} where we show that the associated growth rate (Equation \ref{eq:growth rate reactive appendix}) is
 \begin{equation}\label{eq:growth rate reactive}
@@
 
-The growth rate expressed in Equation (\ref{eq:growth rate reactive}) is in the reactive (or hydrodynamic) regime as the dispersion relation (Equation (\ref{eq:dispersion reactive})) could have been derived from the fluid equations.
+The growth rate expressed in Equation (\ref{eq:growth rate reactive}) is in the reactive (or hydrodynamic) regime as the dispersion relation (Equation \ref{eq:dispersion reactive}) could have been derived from the fluid equations.
 Here all the particles participate in the instability.
@@
 where the target plasma is assumed to be nonrelativistic, $\bmath{p}=m_e\vel$ is the nonrelativistic momentum,  and $T_t$ is the temperature of the target background plasma.
-For the beam plasma, we again adopt the Maxwell-J{\"u}ttner distribution (Equation~(\ref{eq:distribution function beam})).
+For the beam plasma, we again adopt the Maxwell-J{\"u}ttner distribution (Equation~\ref{eq:distribution function beam}).
 %\begin{equation}\label{eq:distribution function beam}
@@
 \Gamma &\simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}\\
+%\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3 }\\
+\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}\\
 &\times\left[
@@
 +
-\frac{\gammabeam^2 v_{b,x'}^2}{2\cG^2c^2} \left(2 \cG'\mu+ 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG\cp{'}^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and \cp{\sout{$\cG'$} $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$} is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
 This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
@@
 \Gamma_{\rm bp} \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
+%\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
+\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 
-Applying the scaling of the perpendicular (Equation (\ref{eq:perp dispersion})) and parallel velocity dispersions (Equation (\ref{eq:parallel dispersion})) to these results and assuming $v_b \approx c$, we find
+Applying the scaling of the perpendicular (Equation \ref{eq:perp dispersion}) and parallel velocity dispersions (Equation \ref{eq:parallel dispersion}) to these results and assuming $v_b \approx c$, we find
 \begin{eqnarray}
@@
 \end{eqnarray}
-Hence, the conditions for the reactive regime for the oblique mode (Equation~(\ref{eq:oblique reactive regime})) and beam-plasma mode (Equation~(\ref{eq:bp reactive regime})) can be reduced to 
+Hence, the conditions for the reactive regime for the oblique mode (Equation~\ref{eq:oblique reactive regime}) and beam-plasma mode (Equation~x\ref{eq:bp reactive regime}) can be reduced to 
 \begin{equation}\label{eq:generic reactive regime}
@@
 
-We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (Equation~(\ref{eq:bp kinetic growth rate})) and oblique mode (Equation~(\ref{eq:kinetic growth rate})).
+We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (Equation~\ref{eq:bp kinetic growth rate}) and oblique mode (Equation~\ref{eq:kinetic growth rate}).
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
@@
 We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
-As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation (\ref{eq:threshold})) and set $1-\cos\theta \approx 2$.
+As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta \approx 2$.
 \ep{[Not sure I understand this.
@@
 
-In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation (\ref{eq:generic reactive regime})) and kinetic regimes (Equation (\ref{eq:generic kinetic regime})) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
+In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation \ref{eq:generic reactive regime}) and kinetic regimes (Equation \ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
 Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $\epm$ pair beams of interest, we find for the controlling parameter:
@@
 This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
-Let us now consider the parallel component (Equation (\ref{eq:dbetabarLABpara})) again to lowest order in $\betaCM$:
+Let us now consider the parallel component (Equation \ref{eq:dbetabarLABpara}) again to lowest order in $\betaCM$:
 \begin{equation}
@@
 
-This can be rewritten as 
+\cp{Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), here we consider the instability of a cold plasma beam. 
+Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT_t  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
+\begin{equation}\label{eq:dispersion relation reactive}
+ 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z \vbeam)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
+\end{equation}
+} This can be rewritten as 
 \begin{eqnarray}\label{eq:dispersion reactive}
@@
 
+\cp{To find the groeth rate for th ekinetic regime, we begin first with the distribution function for the target plasma
+\begin{equation}
+ F_t = \left(\frac{1}{2\pi m_e k_B T_t}\right)^{3/2}\exp\left(-\frac{p^2}{2m_e k_B T_t}\right).
+\end{equation}
+We assume that the target plasma is nonrelativistic with a momentum $\bmath{p}=m_e\vel$,  and $T_t$ is the temperature of the target background plasma.
+For the beam plasma, we adopt the Maxwell-J{\"u}ttner distribution (Equation~\ref{eq:distribution function beam}).
+Inserting these into the dispersion relation (Equation \ref{eq:dispersion relation}), we find
+\begin{equation}\label{eq:dispersion kinetic}
+ 1 - \frac{\omega_{p,t}^2}{k^2c^2}\int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
+\end{equation}
+where we have integrated by parts only the second term, associated with the target plasma. 
+}
 As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.
@@
 \begin{equation}
-R = \frac{
-  \gph^3 \gamma_w^2 (\vph-v_{b,z'}) 
-}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
+R =
+%\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
+\cp{\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu  \cG'^3 K_2(\mu) c} }
 \left[
@@
 +
-\frac{\gammabeam^2 v_{b,x'}^2}{2\cG^2 c^2} \left(2 \cG'\mu + 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG\cp{'}^2 c^2} \left(2 \cG'\mu + 2\right)
 \right] \exp\left(-\cG'\mu\right)
@@
 \Gamma \simeq - \Gamma_0
-\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
+%\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
+\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}
 \left[

### 2016-08-07 oblique.tex
latest changes

--- 2016-08-04 oblique.tex
@@
 
-In the COM frame of the beam, we assume that the distribution function is isotropic \ep{[a brief explanation why this is OK might be useful]}, such that $f=f(E)$ is just a function of energy.
+In the COM frame of the beam, we assume that the distribution function is isotropic, such that $f=f(E)$ is just a function of energy.
 This equilibrium energy distribution of a relativistic thermal plasma gas is
@@
 
-This can be rewritten as 
+\phil{We begin with the dispersion relation (Equation (\ref{eq:dispersion relation by parts})), which is
+\begin{equation}
+  \epsilon = 
+  1 - \sum_s \frac{m_e\omega_{p,s}^2}{k^2} \int F_s\kvec \cdot \gradp\frac{1}{\omega - \kvec\cdot\vel} d^3p =
+  1 - \sum_s \frac{\omega_{p,s}^2}{k^2c^2}\int F_s \frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p 
+  =
+  0\,.
+  \label{eq:dispersion relation appendix}
+\end{equation}
+We then take the limit of Equation (\ref{eq:distribution function beam}) as $k_BT_t  \rightarrow 0$, which yields a $\delta$ function.
+For the target plasma, $v_0=0$, and for a beam plasma $v_0=\vbeam$.
+This leads to equation (\ref{eq:dispersion relation reactive}), which we reproduce below.
+\begin{equation}\label{eq:dispersion appendix}
+ 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z \vbeam)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
+\end{equation}
+Equation (\ref{eq:dispersion appendix}) }can be rewritten as 
 \begin{eqnarray}\label{eq:dispersion reactive}
@@
 As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.
-Expanding the denominator in powers of $v$, we find\footnote{An alert reader will note that Lorentz factor, $\gamma$, and the second term in the numerator both contribute to the expansion in powers of $v$ at second order.
+Expanding the denominator \phil{for the target plasma in equation (\ref{eq:dispersion relation appendix})} in powers of $v$, we find\footnote{An alert reader will note that Lorentz factor, $\gamma$, and the second term in the numerator both contribute to the expansion in powers of $v$ at second order.
 These contributions are the result of the minor deviations from the Lorentz factor of the nonrelativistic electrons and the subtle different between momentum and velocity at order $v^2/c^2$.

### 2016-08-07 oblique.tex

--- 2016-08-05 oblique.tex
@@
 %\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3 }\\
-\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}\\
+\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}\\
 &\times\left[
@@
 +
-\frac{\gammabeam^2 v_{b,x'}^2}{2\cG\cp{'}^2c^2} \left(2 \cG'\mu+ 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG'^2c^2} \left(2 \cG'\mu+ 2\right)
 \right]
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and \cp{\sout{$\cG'$} $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$} is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}, and $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
-This specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}.
+Finally, we define the typical maximum growth rate, $\Gamma_0$, as
+\begin{equation}
+\Gamma_0 \equiv \omega_p \gammabeam\frac{n_b}{n_t} \frac{m_e v_b^2}{k_B T_b}. 
+\end{equation}
+Equation (\ref{eq:kinetic growth rate}) specializes to the beam-plasma growth rate if we take $v_{b,x'} = 0$, which gives:
 \begin{equation}
@@
 %\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
-\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}
+\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}
 \left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
@@
 Note that as $\sin\theta \rightarrow 0$, we recover the beam-plasma instability.
-Moreover the maximum growth rates vary little and are
+Moreover, the maximal growth rates\phil{, normalized to $\Gamma_0$,} vary little and are
 robust for a broad range of angles between the wavevector and the beam direction.
-It is clear from this plot that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable.
+It is clear from this plot that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable \phil{and grow at nearly the maximum growth rate, $\Gamma_0$, for the parameters of the system, $T_b$, $\gamma_b$, and $n_b/n_t$, modulo a factor of order unity.}
 
@@
 (\ref{eq:bp kinetic growth rate}) as a function of $\gammabeam - 1$ and $k_B
-T/m_ec^2$.  Here, we see that the maximum growth rate is somewhat more
+T_b/m_ec^2$.  Here, we see that the maximum growth rate is somewhat more
 sensitive to temperature, varying from $\Gamma/\Gamma_0 \approx 0.4$ for
@@
 \end{equation}
-For \ep{\sout{$k = \omega_p/v_b$} $k_\perp \approx \omega_p/v_b$}, this gives
+For $k_\perp \sim \omega_p/v_b$, this gives
 %\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
@@
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
-Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$ \ep{[Wouldn't it be more natural to quote a k range that grows?
-Also some more explanation why $\Delta \omega$ and $\Gamma$ are compared would be helpful.]}.
-For the beam plasma case, the growth rate is roughly
+Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$.
+\phil{For the kinetic growth rate to be valid, the bandwidth, $\Delta\omega$, must be large compared to the growth rate; otherwise, the entire beam contributes to the growth and, hence, the reactive regime applies.} For the beam plasma case, the growth rate is roughly
 \begin{equation}
@@
 
-In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
-As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta \approx 2$.
+In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron), $E_{e,COM}$ for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
+We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm ph, COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta_{\rm COM} \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
+\phil{A rough order of magnitude calculation bears this out.
+We begin by noting that $\tan\theta = p_{\rm EBL,\perp}/p_{\rm EBL, \parallel}$, where $p_{\rm EBL,\perp}$ and $p_{\rm EBL,\parallel}$ is the perpendicular and parallel components of the EBL photon's momentum relative to the TeV photon's momentum.
+Boosting to the COM frame, the parallel component of the EBL's photons momentum is $p_{\rm EBL, COM,\parallel} \approx -E_{\rm ph, COM}/c$, while the perpendicular momentum is conserved, $p_{\rm EBL, COM,\parallel} = p_{\rm EBL, \parallel}$.
+As a result, the angle in the COM frame is
+\begin{equation}
+\tan\theta_{\rm COM} \approx \frac{p_{\rm EBL, COM,\parallel}}{p_{\rm EBL, COM,\parallel}} \sim -\sqrt{\frac{E_{\rm EBL}}{E_{\rm ph}}} \approx -10^{-3} \left(\frac{E_{\rm EBL}} {1\,{\rm eV}}\right)^{1/2}\left(\frac{E_{\rm ph}}{1\,{\rm TeV}}\right)^{-1/2}
+\end{equation}
+}
+As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta = 2$.
 \ep{[Not sure I understand this.
@@
 
-Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2$ falls between $1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2 - 1$ falls between $0.3-2$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
 First, it is necessary to determine $n_b/n_t$.
@@
 
-\cp{Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), here we consider the instability of a cold plasma beam. 
-Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT_t  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
-\begin{equation}\label{eq:dispersion relation reactive}
+\phil{We begin with the dispersion relation (Equation (\ref{eq:dispersion relation by parts})), which is
+\begin{equation}
+  \epsilon = 
+  1 - \sum_s \frac{m_e\omega_{p,s}^2}{k^2} \int F_s\kvec \cdot \gradp\frac{1}{\omega - \kvec\cdot\vel} d^3p =
+  1 - \sum_s \frac{\omega_{p,s}^2}{k^2c^2}\int F_s \frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p 
+  =
+  0\,.
+  \label{eq:dispersion relation appendix}
+\end{equation}
+We then take the limit of Equation (\ref{eq:distribution function beam}) as $k_BT_t  \rightarrow 0$, which yields a $\delta$ function.
+For the target plasma, we set $v_0=0$, and for a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}.
+This leads to equation (\ref{eq:dispersion relation reactive}), which we reproduce below.
+\begin{equation}\label{eq:dispersion appendix}
  1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z \vbeam)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
 \end{equation}
-} This can be rewritten as 
+Equation (\ref{eq:dispersion appendix}) }can be rewritten as 
 \begin{eqnarray}\label{eq:dispersion reactive}
@@
 
-\cp{To find the groeth rate for th ekinetic regime, we begin first with the distribution function for the target plasma
+To find the growth rate for the kinetic regime, we begin first with the distribution function for the target plasma
 \begin{equation}
@@
 where we have integrated by parts only the second term, associated with the target plasma.
-}
+
 As the target plasma is nonrelativistic, we can take $v\ll c$ and $\gamma \rightarrow 1$.
@@
 %\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
-\cp{\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu  \cG'^3 K_2(\mu) c} }
+\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu  \cG'^3 K_2(\mu) c} 
 \left[
@@
 %\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
-\cp{\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}}
+\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}
 \left[

### 2016-08-08 oblique.tex
more updates

@@
 \end{aligned}
-\right.\,,
+\right.\,,\label{eq:kinetic growth rate}
 \end{equation}
@@
 the oblique growth rate, i.e., it is not orders of magnitude lower.
+
+\phil{Finally, the growth rate that we derived here (Equation (\ref{eq:kinetic growth rate})) and that of BCP12 (their equation (16)) which is originally derived from the numerical fit of \citet{Bret-Grem-Beni:10} are exactly the same.
+We must note, however, that the our definition of $T_b$ is in the COM frame whereas BCP12 defines $T_b$ in the ``lab'' frame.
+As a result, there is a factor of $\gamma_b$ that is explicit in equation (\ref{eq:kinetic growth rate}) that is implicit in equation (16) of BCP12.
+}
 

### 2016-08-11 oblique.tex
added new rates plot

--- 2016-08-07 oblique.tex
@@
 Astrid Lamberts\altaffilmark{6},
-Mohamad Shalaby\altaffilmark{2,3,7}
+Mohamad Shalaby\altaffilmark{2,3,7},
+Geoffrey Vasil\altaffilmark{8}
 }
@@
 \altaffiltext{7}{Department of Physics, Faculty of Science, Cairo University, Giza 12613, Egypt}
+\altaffiltext{8}{School of Mathematics \& Statistics, University of Sydney, NSW 2006, Australia}
+
 \keywords{
@@
 The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
-Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that cool them.
+Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that they drive.
 Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.
@@
 energies $\lesssim100\,\GeV$.  The absence of observed secondary IC emission leads a number of
-authors to argue that this \textit{lack} places lower bounds upon the
+authors to argue that this \textit{lack} of emission places lower bounds upon the
 intergalactic magnetic field
@@
 
-\phil{While, a full nonlinear study is required, we focus on the nature of the linear instability in this paper, clarifying its robustness and regimes of applicability.}  We begin by studying the distribution function of the $\epm$ pairs that are produced from VHEGR-EBL photon annihilation.
+While a full nonlinear study is required, we focus on the nature of the linear instability in this paper, clarifying its robustness and regimes of applicability.
+We begin by studying the distribution function of the $\epm$ pairs that are produced from VHEGR-EBL photon annihilation.
 We study the evolution of a distribution function under Lorentz transformations to develop an analytic understanding of how the perpendicular and parallel velocity dispersions transform under boosts.
@@
 However, we make this approximation for two reasons.
-First, a complete calculation of the unstable modes has already been carried out by \citet{Bret-Grem-Beni:10}, who showed that the oblique mode is mainly electrostatic (modulo the Weibel instability)\phil{so that a electrostatic approximation to the full dispersion relation should recover the essential physics}.
+First, a complete calculation of the unstable modes has already been carried out by \citet{Bret-Grem-Beni:10}, who showed that the oblique mode is mainly electrostatic (modulo the Weibel instability).
+Hence   a electrostatic approximation to the full dispersion relation should recover the essential physics.
 Second, the electrostatic approximation is much simpler than a full calculation and allows us to analytically calculate the unstable growth rates, while permitting a clear exposition of the relevant physics.
@@
 
-The growth rate expressed in Equation (\ref{eq:growth rate reactive}) is in the reactive (or hydrodynamic) regime as the dispersion relation (Equation \ref{eq:dispersion reactive}) could have been derived from the fluid equations.
+The growth rate expressed in Equation (\ref{eq:growth rate reactive}) is in the reactive (or hydrodynamic) regime as the dispersion relation (Equation \ref{eq:dispersion relation reactive}) could have been derived from the fluid equations.
 Here all the particles participate in the instability.
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity \ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity %\ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}.
 Finally, we define the typical maximum growth rate, $\Gamma_0$, as 
@@
 
-In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate as a function of
+In Figure \ref{fig:ObliqueGMZ}, we plot the growth rate for the oblique instability (Equation (\ref{eq:kinetic growth rate})) as a function of
 $\sin\theta$, where $\cos\theta = \hat{\kvec}\cdot\hat{\vel}$, i.e., the angle between the
@@
 Note that as $\sin\theta \rightarrow 0$, we recover the beam-plasma instability.
-Moreover, the maximal growth rates\phil{, normalized to $\Gamma_0$,} vary little and are
+Moreover, the maximal growth rates, normalized to $\Gamma_0$, vary little and are
 robust for a broad range of angles between the wavevector and the beam direction.
-It is clear from this plot that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable \phil{and grow at nearly the maximum growth rate, $\Gamma_0$, for the parameters of the system, $T_b$, $\gamma_b$, and $n_b/n_t$, modulo a factor of order unity.}
+It is clear from this plot that for nearly any combination of wavevector orientation and beam temperature that there exists a broad spectrum of modes that are unstable and grow at nearly the maximum growth rate, $\Gamma_0$, for the parameters of the system, $T_b$, $\gamma_b$, and $n_b/n_t$, modulo a factor of order unity.
 
@@
 \end{aligned}
-\right.\,,
+\right.\,,\label{eq:max kinetic growth rate}
 \end{equation}
@@
 the oblique growth rate, i.e., it is not orders of magnitude lower.
+
+Finally, the maximum growth rate that we derived here (Equation (\ref{eq:max kinetic growth rate})) and that of BCP12 (their equation (16)) which is originally derived from the numerical fit of \citet{Bret-Grem-Beni:10} are exactly the same.
+We must note, however, that the our definition of $T_b$ is in the COM frame whereas BCP12 defines $T_b$ in the ``lab'' frame.
+As a result, there is a factor of $\gamma_b$ that is explicit in equation (\ref{eq:max kinetic growth rate}) that is implicit in equation (16) of BCP12.
 
@@
 The oblique instability exists in two different regimes, raising the important question: how are the two regimes related to each other.
-While this question has been studied by many authors in the context of the beam-plasma or two-stream instability \citep[see for instance][]{Melrose86,Boyd}, a clear exposition of how these two regimes are related to each other \phil{for the oblique instability} is lacking.
+While this question has been studied by many authors in the context of the beam-plasma or two-stream instability \citep[see for instance][]{Melrose86,Boyd}, a clear exposition of how these two regimes are related to each other for the oblique instability is lacking.
 
@@
 Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$.
-\phil{For the kinetic growth rate to be valid, the bandwidth, $\Delta\omega$, must be large compared to the growth rate; otherwise, the entire beam contributes to the growth and, hence, the reactive regime applies.} For the beam plasma case, the growth rate is roughly
+For the kinetic growth rate to be valid, the bandwidth, $\Delta\omega$, must be large compared to the growth rate; otherwise, the entire beam contributes to the growth and, hence, the reactive regime applies.
+For the beam plasma case, the growth rate is roughly
 \begin{equation}
@@
 To do so, we consider the process of photon-photon annihilation.
-For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the production rate of \epm\ on EBL photons is
+For a monoenergetic population of VHEGR photons with energy $E_{\rm ph}$, the angle-averaged production rate of \epm\ on EBL photons is
 \begin{eqnarray}
-\Gamma_{\pm}(E_{\rm ph}) &=& \int \sigma c dn_{\rm EBL} \nonumber \\
-&=& \int \sigma\left(E_{\rm ph}, E_{\rm EBL}, \theta\right) c \frac{dn_{\rm EBL}}{dE_{\rm EBL}} dE_{\rm EBL},
+\Gamma_{\pm}(E_{\rm ph}) &=& \frac{1}{4\pi}\int \sigma c dn_{\rm EBL}d\Omega \nonumber \\
+&=& \frac{1}{2}\int \sigma\left(E_{\rm ph}, E_{\rm EBL}, \theta\right) c \frac{dn_{\rm EBL}}{dE_{\rm EBL}} dE_{\rm EBL} d\cos\theta,
 \end{eqnarray}
@@
 \end{equation}
-where $r_e = e^2/m_e c^2$ is the classical electron radius and $v_e$ is the electron velocity in the COM frame.  To find $v_e$, we use the energy of the electron in the COM frame, $E_e$, which is 
+where $r_e = e^2/m_e c^2$ is the classical electron radius and $v_e$ is the electron velocity in the COM frame.
+To find $v_e$, we use the energy of the electron in the COM frame, $E_{\rm e, COM}$, which is
 \begin{equation}
 \label{eq:threshold}
- E_e =  \frac{m_e c^2}{\sqrt{1-v_e^2/c^2}} = \sqrt{\frac{1}{2}{E_{\rm EBL} E_{\rm ph}\left(1-\cos\theta\right)}}.
-\end{equation}
-Pair production occurs when $E_e/m_e c^2 \geq 1$.
+ E_{\rm e, COM} =  \frac{m_e c^2}{\sqrt{1-v_e^2/c^2}} = \sqrt{\frac{1}{2}{E_{\rm EBL} E_{\rm ph}\left(1-\cos\theta\right)}}.
+\end{equation}
+Pair production occurs when $E_{\rm e, COM}/m_e c^2 \geq 1$.
 
@@
 In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron), $E_{e,COM}$ for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm ph, COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta_{\rm COM} \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
+%We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm ph, COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta_{\rm COM} \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
 \phil{A rough order of magnitude calculation bears this out.
 We begin by noting that $\tan\theta = p_{\rm EBL,\perp}/p_{\rm EBL, \parallel}$, where $p_{\rm EBL,\perp}$ and $p_{\rm EBL,\parallel}$ is the perpendicular and parallel components of the EBL photon's momentum relative to the TeV photon's momentum.
-Boosting to the COM frame, the parallel component of the EBL's photons momentum is $p_{\rm EBL, COM,\parallel} \approx -E_{\rm ph, COM}/c$, while the perpendicular momentum is conserved, $p_{\rm EBL, COM,\parallel} = p_{\rm EBL, \parallel}$.
+Boosting to the COM frame, the parallel component of the EBL's photons momentum is $p_{\rm EBL, COM,\parallel} \approx -E_{\rm ph, COM}/c$, while the perpendicular momentum is conserved, $p_{\rm EBL, COM,\perp} = p_{\rm EBL, \perp}$.
 As a result, the angle in the COM frame is
-\begin{equation}
-\tan\theta_{\rm COM} \approx \frac{p_{\rm EBL, COM,\parallel}}{p_{\rm EBL, COM,\parallel}} \sim -\sqrt{\frac{E_{\rm EBL}}{E_{\rm ph}}} \approx -10^{-3} \left(\frac{E_{\rm EBL}} {1\,{\rm eV}}\right)^{1/2}\left(\frac{E_{\rm ph}}{1\,{\rm TeV}}\right)^{-1/2}
-\end{equation}
-}
-As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta = 2$.
+%\begin{equation}
+%\tan\theta_{\rm COM} \approx \frac{p_{\rm EBL, COM,\perp}}{p_{\rm EBL, COM,\parallel}} \sim -\sqrt{\frac{E_{\rm EBL}}{E_{\rm ph}}} \approx -10^{-3} \left(\frac{E_{\rm EBL}} {1\,{\rm eV}}\right)^{1/2}\left(\frac{E_{\rm ph}}{1\,{\rm TeV}}\right)^{-1/2}
+%\end{equation}
+%}
+%As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta = 2$.
 \ep{[Not sure I understand this.
@@
 4?]}\phil{[Assumption i $E_{\rm EBL} \ll E_{\rm ph}$ and so in the COM frame the photons have equal energy, but opposite momentum.]}
-
-We note several features of this Figure.
-First the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy.
+Note the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy.
 This is because different energy photons probe different regimes of the EBL spectrum.
 Due to the rapid cutoff in the EBL above $10\,\eV$, lower energy VHEGRs produce colder beams.
-This is seen in the average COM energies of the produced electrons, which are respectively, $E_e/m_e c^2 \approx 1.3, 1.6, 2$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$ and $10\,\TeV$.
+This is seen in the average COM energies of the produced electrons, which are respectively, $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.3, 1.5, 1.7$ and $2.1$ for $E_{\rm ph} = 0.3, 1, 3$ and $10\,\TeV$.
 Hence we expect that these pairs are in the sub-relativistic to mildly-relativistic regimes in their COM frame.
-Second, the effect of the EBL spectrum on the distribution of electron COM energies is also seen in the breaks in the distribution.
-For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
+%Second, the effect of the EBL spectrum on the distribution of electron COM energies is also seen in the breaks in the distribution.
+For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_{\rm e, COM}/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
 Additionally, the break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
@@
 The effect of the EBL spectrum can be seen in different features in this plot.
-The cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_e/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
+The cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_{\rm e, COM}/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
 The break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
-The average COM energies of the produced electrons are $E_e/m_e c^2 \approx 1.3, 1.6, 2,$ and $3$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
+The average COM energies of the produced electrons are $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.3, 1.5, 1.7,$ and $2.1$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
  \label{fig:rates}}
@@
 
-Given that the range of $k_BT_b/m_ec^2 = E_e/m_e c^2 - 1$ falls between $0.3-2$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+Given that the range of $k_BT_b/m_ec^2 = E_{\rm e, COM}/m_e c^2 - 1$ falls between $0.3-1$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
 First, it is necessary to determine $n_b/n_t$.
@@
 at the mean density of the IGM, 
-where $L_E$ is the isotropic luminosity \phil{per unit energy}\ep{[What exactly is the deifniton of $L_E$?
-It has to have the dimension 1/time.
-Has the $2^{3\zeta-4}$ factor been absorbed in the constant assuming $\zeta=4.5$?
-This would then however only be valid for $z<1$.
-Or should it be 2 instead of 1 in the denominator?]}\phil{[Previous answer was correct, factors of 2 was correctly absorbed, but changes to 2 in denominator to be consistent with our subsequent definitions.]} of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
+where $L_E$ is the isotropic luminosity per unit energy of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and
  the mean free path of a VHEGR 
@@
 \end{eqnarray}
-\ep{[The numerical factors here depend on what exactly has been done for Eq. (37).]}\phil{[Numerical factors are correct.]} where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+where $\gamma_6 = \gamma/10^6$. 
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 1\,\TeV$, or for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
 
@@
 \item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectrum of the EBL.
-We found that the average energy of these beams range from $E_e/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.
+We found that the average energy of these beams range from $E_{\rm e, COM}/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.
 This is due to the rapid cutoff in the EBL above $10\,\eV$.
@@
 \acknowledgements
+We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions. 
 A.E.B.~and M.S.~receive financial support from the Perimeter
@@
 Sloan Research Fellowship, NASA ATP Grant NNX14AH35G, and NSF Collaborative Research Grant \#1411920 and CAREER grant \#1455342.
-We thank A. Bret for sharing his notes of the oblique instability and for extensive and enlightening discussions.
+GV acknowledges support from the Australian Research Council, project number DE140101960.
 
@@
 \begin{equation}
- \betabarCM = N^{-1}\int \frac {\pmom_{\rm COM}}{\gamma_{\rm COM} mc} \fCM \dVCM = 0.
+ \betabarCM = N^{-1}\int \frac {\pmom_{\rm COM}}{\gamma_{\rm COM} m_ec} \fCM \dVCM = 0.
 \end{equation}
@@
 \begin{eqnarray}\label{eq:betabarLAB}
- \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
-&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) mc} \fCM \dVCM.
+ \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} m_ec} \fLAB \dVLAB  \nonumber\\
+&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) m_ec} \fCM \dVCM.
 \end{eqnarray}
@@
 It is also more intuitive to study how velocity dispersions scale between the center of mass frame and the lab from for non-relativistic center of mass velocity dispersion.
-Hence, for $\left|\betaCM\right| \ll 1$ \ep{[This is not the case for the distributions shown in Fig.
-4.
-Why is this good enough?]}\phil{[velocity dispersions are easiest for non-relativistic motion in COM frame.]}, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$
+Hence, for $\left|\betaCM\right| \ll 1$, Equation (\ref{eq:dbetabarLABperp}) becomes to lowest order in $\betaCM$
 \begin{equation}
@@
 \end{equation}
-\ep{[Factor 2 seems missing since perpenticular component has 2 degrees of freedom]}\phil{[Factor of 2]} This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
+This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
 This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
@@
 
-\phil{We begin with the dispersion relation (Equation (\ref{eq:dispersion relation by parts})), which is
+We begin with the dispersion relation (Equation (\ref{eq:dispersion relation by parts})), which is
 \begin{equation}
@@
 \end{equation}
-Equation (\ref{eq:dispersion appendix}) }can be rewritten as 
+Equation (\ref{eq:dispersion appendix}) can be rewritten as 
 \begin{eqnarray}\label{eq:dispersion reactive}
@@
 Inserting these into the dispersion relation (Equation \ref{eq:dispersion relation}), we find
-\begin{equation}\label{eq:dispersion kinetic}
+\begin{equation}\label{eq:ap dispersion kinetic}
  1 - \frac{\omega_{p,t}^2}{k^2c^2}\int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p + \frac{m_e\omega_{p,b}^2}{k^2}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
@@
 where we define $\mu = m_ec^2/k_BT_b$.
-Inserting the equation into (\ref{eq:dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
+Inserting the equation into (\ref{eq:ap dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
 \begin{equation}
@@
 +
-\frac{\gammabeam^2 v_{b,x'}^2}{2\cG\cp{'}^2 c^2} \left(2 \cG'\mu + 2\right)
+\frac{\gammabeam^2 v_{b,x'}^2}{2\cG'^2 c^2} \left(2 \cG'\mu + 2\right)
 \right] \exp\left(-\cG'\mu\right)

### 2016-08-14 oblique.tex
final updates

@@
 
-We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
+We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic oblique instability is  
 \begin{equation}
@@
 %\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
-\frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}
-\left( \cG'^2\mu^2 + 2\cG'\mu + 2 \right) 
-\exp(-\cG'\mu),\label{eq:bp kinetic growth rate}
-\end{equation}
+\frac{\pi \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG^3 c}
+\left( \cG^2\mu^2 + 2\cG\mu + 2 \right) 
+\exp(-\cG\mu),\label{eq:bp kinetic growth rate}
+\end{equation}
+where we have used the fact that $\gamma_w = 1$ for $v_{b,x'} = 0$
 
@@
 Due to the rapid cutoff in the EBL above $10\,\eV$, lower energy VHEGRs produce colder beams.
-This is seen in the average COM energies of the produced electrons, which are respectively, $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.3, 1.5, 1.7$ and $2.1$ for $E_{\rm ph} = 0.3, 1, 3$ and $10\,\TeV$.
+This is seen in the average COM energies of the produced electrons, which are respectively, $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.5, 1.7, 2.2$ and $2.8$ for $E_{\rm ph} = 0.3, 1, 3$ and $10\,\TeV$.
 Hence we expect that these pairs are in the sub-relativistic to mildly-relativistic regimes in their COM frame.
@@
 The break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
-The average COM energies of the produced electrons are $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.3, 1.5, 1.7,$ and $2.1$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
+The average COM energies of the produced electrons are $\bar{E}_{\rm e, COM}/m_e c^2 \approx 1.5, 1.7, 2.2,$ and $2.8$ for $E_{\rm ph} = 0.3, 1, 3$, and $10\,\TeV$, respectively.
  \label{fig:rates}}
@@
 
-Given that the range of $k_BT_b/m_ec^2 = E_{\rm e, COM}/m_e c^2 - 1$ falls between $0.3-1$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
+Given that the range of $k_BT_b/m_ec^2 = E_{\rm e, COM}/m_e c^2 - 1$ falls between $0.5-2$ for $E_{\rm ph} = 1-10\,\TeV$, we now determine whether or not the reactive or kinetic instabilities apply to these beams.
 First, it is necessary to determine $n_b/n_t$.
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 1\,\TeV$, or for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 0.3\,\TeV$, or for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
 
@@
 While we also noted that the oblique instability was in the kinetic regime in BCP12, which we confirmed above, we made no effort to study the regime of instability of the beam-plasma case.
-Here we have shown that while the oblique growth rate is kinetic, the beam-plasma rate is marginally reactive.
-In any case, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.
+Here we have shown that the oblique growth rate is kinetic and the beam-plasma rate is marginally reactive.
+This implies that the growth rate of the beam plasma instability is similar to that of the oblique instability.
+In any case, we do not expect that the beam-plasma mode will have a major effect on our earlier results.
 First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.
@@
 This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.
-Applied to TeV $e^{\pm}$ beams, we find that these beams are unstable to the kinetic oblique instability and (marginally to) the reactive beam-plasma instability.}
+}
 \item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectrum of the EBL.
-We found that the average energy of these beams range from $E_{\rm e, COM}/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.
-This is due to the rapid cutoff in the EBL above $10\,\eV$.
-}
+We found that the average energy of these beams range from $E_{\rm e, COM}/m_e c^2 = 1.5-2.8$ for $E_{\rm ph} = 0.3-10\,\TeV$, with colder beams at lower energies.
+The average COM energies of the generated pairs implies that the oblique instability is in the kinetic regime, validating our results from BCP12.}
 %\item{We demonstrate that the oblique mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.

### 2016-08-16 oblique.tex
Swapped sims for propto and approx, one-lined equations where appropriate, etc.

@@
 \citep[see, e.g., ][]{Naru-Tota:06,Knei-Mann:08,Inou-Tota:09,Vent:10}.
-These constraints preclude any dramatic rise in numbers of source by $z\sim1$--$2$ that is seen in the quasar
+These constraints preclude any dramatic rise in numbers of source by \ab{\sout{$z\sim1$} $z\approx1$}--$2$ that is seen in the quasar
 distribution.  That is, the comoving number of blazars must have
@@
 photon ($E_{\rm ebl}$) exceed the rest mass energy of the $\epm$ pair in the
-center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
+center of momentum (COM \ab{[Should we use CM instead for compactness? Otherwise ``of'' is getting a big promotion!]}) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
-frame. As a result, an $\epm$ pair can be produced with Lorentz factor $\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e
-c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.  Here, we discuss the distribution function of the
+frame.
+As a result, an $\epm$ pair can be produced with Lorentz factor \ab{\sout{$\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e c^2$} $\gamma = \left(1-v^2/c^2\right)^{-1/2} \approx E/2m_e c^2$}, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.
+Here, we discuss the distribution function of the
 pair beam that emerges from this process.
@@
 \begin{equation}\label{eq:perp dispersion}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 },
-\end{equation}
-and 
-\begin{equation}\label{eq:parallel dispersion}
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }\ab{%\sout{,}
+\quad\text{and}\quad}
 \frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
 \end{equation}
+%and 
+%\begin{equation}\label{eq:parallel dispersion}
+%\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
+%\end{equation}
 where $T$ is measured in the COM frame of the beam.  
@@
 
-We now adopt perturbations of the form $\delta \sim \exp\left(i \kvec\cdot\rvec - i \omega t\right)$ and without loss of generality assume that $\kvec = (k_x, 0, k_z)$, where $k_z$ is along the beam direction.
+We now adopt perturbations of the form \ab{\sout{$\delta \sim \exp\left(i \kvec\cdot\rvec - i \omega t\right)$} $\delta \propto \exp\left(i \kvec\cdot\rvec - i \omega t\right)$} and without loss of generality assume that $\kvec = (k_x, 0, k_z)$, where $k_z$ is along the beam direction.
 Linearizing the Vlasov-Maxwell equations then leads to the dispersion relation:
@@
 \begin{aligned}
-\Gamma &\simeq - \Gamma_0
+\Gamma &\ab{\approx} %\simeq
+- \Gamma_0
 %\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3 }\\
@@
 \begin{equation}
-\Gamma_{\rm bp} \simeq - \Gamma_0
+\Gamma_{\rm bp} \ab{\approx} %\simeq
+- \Gamma_0
 %\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
@@
 \Gamma_M
-\simeq
+\ab{\approx}
+%\simeq
 \left\{
@@
 We must note, however, that the our definition of $T_b$ is in the COM frame whereas BCP12 defines $T_b$ in the ``lab'' frame.
-As a result, there is a factor of $\gamma_b$ that is explicit in equation (\ref{eq:max kinetic growth rate}) that is implicit in equation (16) of BCP12.
+As a result, there is a factor of $\gamma_b$ that is explicit in Equation (\ref{eq:max kinetic growth rate}) that is implicit in Equation (16) of BCP12.
 
@@
 \end{equation}
-For $k_\perp \sim \omega_p/v_b$, this gives
+For \ab{\sout{$k_\perp \sim \omega_p/v_b$} $k_\perp \approx \omega_p/v_b$}, this gives
 %\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
@@
 \end{equation}
-where we have dropped constant factors of order unity and assumed that $Z_x \sim O(1)$ and that the velocity dispersion is dominated by perpendicular (to the beam) component.
+where we have dropped constant factors of order unity and assumed that \ab{\sout{$Z_x \sim O(1)$} $Z_x \propto O(1)$} and that the velocity dispersion is dominated by perpendicular (to the beam) component.
 Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate Equation (\ref{eq:growth rate reactive}).
@@
 
-Applying the scaling of the perpendicular (Equation \ref{eq:perp dispersion}) and parallel velocity dispersions (Equation \ref{eq:parallel dispersion}) to these results and assuming $v_b \approx c$, we find
-\begin{eqnarray}
-\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT_b}{m_ec^2}},\\
-\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.
-\end{eqnarray}
-Hence, the conditions for the reactive regime for the oblique mode (Equation~\ref{eq:oblique reactive regime}) and beam-plasma mode (Equation~x\ref{eq:bp reactive regime}) can be reduced to 
+Applying the scaling of the perpendicular \ab{\sout{(Equation \ref{eq:perp dispersion})}} and parallel velocity dispersions \ab{\sout{(Equation \ref{eq:parallel dispersion})}} \ab{(Equations \ref{eq:perp dispersion})} to these results and assuming $v_b \approx c$, we find
+\ab{
+\begin{equation}
+  \frac{\Delta v_{\perp}}{\vbeam} \approx  \gammabeam^{-1}\sqrt{\frac{k_BT_b}{m_ec^2}}
+  \quad\text{and}\quad
+\frac{\Delta v_{\parallel}}{\vbeam} \approx \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.  
+\end{equation}}
+%\begin{eqnarray}
+%\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT_b}{m_ec^2}},\\
+%\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.
+%\end{eqnarray}
+Hence, the conditions for the reactive regime for the oblique mode (Equation~\ref{eq:oblique reactive regime}) and beam-plasma mode (Equation~\ref{eq:bp reactive regime}) can be reduced to 
 \begin{equation}\label{eq:generic reactive regime}
@@
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
-Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$.
+Hence the bandwidth over which the distribution powers grows is \ab{\sout{$\Delta \omega \sim k \Delta v$} $\Delta \omega \approx k \Delta v$}.
 For the kinetic growth rate to be valid, the bandwidth, $\Delta\omega$, must be large compared to the growth rate; otherwise, the entire beam contributes to the growth and, hence, the reactive regime applies.
@@
 \begin{equation}
- \Gamma \sim \frac{n_b}{\gamma^3n_t} \left(\frac{c}{\Delta v_{\parallel}}\right)^{2}.
+\Gamma \ab{\approx}
+%\sim
+\frac{n_b}{\gamma^3n_t} \left(\frac{c}{\Delta v_{\parallel}}\right)^{2}.
 \end{equation}
@@
 \begin{eqnarray}
-\frac{\nb}{\nIGM} &\simeq&
+\frac{\nb}{\nIGM} & \ab{\approx} %\simeq
+&
 \frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \\
-&\simeq &
+&\ab{\approx}
+%\simeq
+&
 2.3\times10^{-16}
@@
 
-In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation \ref{eq:generic reactive regime}) and kinetic regimes (Equation \ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.
+In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation \ref{eq:generic reactive regime}) and kinetic regimes (Equation \ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, \ab{\sout{$\Delta\omega\sim k\Delta v$} $\Delta\omega\approx k\Delta v$}, with the growth rate, $\Gamma$.
 Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $\epm$ pair beams of interest, we find for the controlling parameter:
-\begin{eqnarray}
-\frac{\gammabeam^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
-&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
-\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
-\end{eqnarray}
+\ab{
+\begin{multline}
+  \frac{\gammabeam^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3}= 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\\
+  \qquad\times \left(\frac{1+z}{2}\right)^{\zeta-7/3}
+  \left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
+  \left(\frac{E}{\TeV}\right)^{1/3},
+  \label{eq:beam oblique condition}
+\end{multline}
+}
+%\begin{eqnarray}
+%\frac{\gammabeam^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+%&&\ab{\times}\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
+%\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
+%\end{eqnarray}
 and
-\begin{eqnarray}
-\frac{\gammabeam}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
-&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
-\left(\frac{E}{\TeV}\right)^{1/3}\label{eq:beam bp condition}
-\end{eqnarray}
+\ab{
+\begin{multline}
+  \frac{\gammabeam}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} = 3.2 \frac{\gamma_6}{\sqrt{k_BT_b/m_e c^2}}\\
+  \qquad\times\left(\frac{1+z}{2}\right)^{\zeta-7/3}
+  \left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
+  \left(\frac{E}{\TeV}\right)^{1/3},
+  \label{eq:beam bp condition}
+\end{multline}
+}
+%\begin{eqnarray}
+%\frac{\gammabeam}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
+%&&\ab{\times}\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
+%\left(\frac{E}{\TeV}\right)^{1/3}\label{eq:beam bp condition}
+%\end{eqnarray}
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 0.3\,\TeV$, or for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 0.3\,\TeV$, or for large $\gamma$, which occurs for \ab{\sout{$E_{\rm ph} \sim 10\,\TeV$} $E_{\rm ph} \approx 10\,\TeV$} at $z=0$.
 
@@
 \item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regimes.
-We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the oblique mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.
+We have recovered the reactive scalings for the beam-plasma mode \ab{\sout{$\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$} $\Gamma \approx \gamma^{-1}(n_b/n_t)^{1/3}$} and the oblique mode \ab{\sout{$\Gamma \sim (n_b/\gamma n_t)^{1/3}$} $\Gamma \approx (n_b/\gamma n_t)^{1/3}$}.
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.
@@
 \begin{equation}
-  \fLAB(\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM})) = \fCM(E_{\rm COM}).
+  \fLAB\ab{[}\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM})\ab{]} = \fCM(E_{\rm COM}).
 \end{equation}
@@
 This gives:
-\begin{eqnarray}\label{eq:betabarLAB}
- \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} m_ec} \fLAB \dVLAB  \nonumber\\
-&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) m_ec} \fCM \dVCM.
-\end{eqnarray}
+\ab{
+\begin{equation}\label{eq:betabarLAB}
+ \betabarLAB = N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} m_ec} \fLAB \dVLAB  
+= N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) m_ec} \fCM \dVCM.
+\end{equation}
+}
+%\begin{eqnarray}\label{eq:betabarLAB}
+% \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} m_ec} \fLAB \dVLAB  \nonumber\\
+%&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) m_ec} \fCM \dVCM.
+%\end{eqnarray}
 Breaking the components of \betabarLAB\ into components parallel and
 perpendicular to the boost, we find: 
-\begin{eqnarray}
-\betabarLABpara &=& N^{-1}\int \frac{\betaCMpara +\betaBOOST}{1+\betaBOOST\betaCMpara} \fCM \dVCM,\\
-\betabarLABperp &=& N^{-1}\int \frac{\betaCMperp}{\gammabeam\left(1+\betaBOOST\betaCMpara\right)} \fCM \dVCM,
-\end{eqnarray}
+\ab{
+\begin{equation}
+  \betabarLABpara = N^{-1}\int \frac{\betaCMpara +\betaBOOST}{1+\betaBOOST\betaCMpara} \fCM \dVCM
+  \quad\text{and}\quad
+  \betabarLABperp = N^{-1}\int \frac{\betaCMperp}{\gammabeam\left(1+\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+\end{equation}
+}
+%\begin{eqnarray}
+%\betabarLABpara &=& N^{-1}\int \frac{\betaCMpara +\betaBOOST}{1+\betaBOOST\betaCMpara} \fCM \dVCM,\\
+%\betabarLABperp &=& N^{-1}\int \frac{\betaCMperp}{\gammabeam\left(1+\betaBOOST\betaCMpara\right)} \fCM \dVCM,
+%\end{eqnarray}
 where $\gammabeam = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
@@
 In the lab frame, it is again useful to break it into components -- the parallel component becomes 
-\begin{eqnarray}
-  \dbetabarLABpara &=& N^{-1}\int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
- &=& N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
- &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaCMpara\betaBOOST}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
-\end{eqnarray}
+\ab{
+\begin{equation}
+  \begin{aligned}
+    \dbetabarLABpara &= N^{-1}\int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB
+    = N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2\\
+    &= N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaCMpara\betaBOOST}\right)^2 \fCM \dVCM - \betabarLABpara^2,
+  \end{aligned}
+  \label{eq:dbetabarLABpara}
+\end{equation}
+}
+%\begin{eqnarray}
+%  \dbetabarLABpara &=& N^{-1}\int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
+% &=& N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
+% &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaCMpara\betaBOOST}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
+%\end{eqnarray}
 while the perpendicular component becomes
-\begin{eqnarray}
- \dbetabarLABperp &=& N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaCMpara\betaBOOST\right)^2} \fCM \dVCM.
+\ab{
+\begin{equation}
+  \dbetabarLABperp = N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB
+  = N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaCMpara\betaBOOST\right)^2} \fCM \dVCM.
 \label{eq:dbetabarLABperp}
-\end{eqnarray}
+\end{equation}
+}
+%\begin{eqnarray}
+% \dbetabarLABperp &=& N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
+%&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaCMpara\betaBOOST\right)^2} \fCM \dVCM.
+%\label{eq:dbetabarLABperp}
+%\end{eqnarray}
 
@@
 This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
-This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \sim kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
+This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., \ab{\sout{$mv^2 \sim kT$} $mv^2 \approx kT$} -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
 Let us now consider the parallel component (Equation \ref{eq:dbetabarLABpara}) again to lowest order in $\betaCM$:
@@
 For the target plasma, we set $v_0=0$, and for a beam plasma $v_0=\vbeam$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}.
-This leads to equation (\ref{eq:dispersion relation reactive}), which we reproduce below.
+This leads to Equation (\ref{eq:dispersion relation reactive}), which we reproduce below.
 \begin{equation}\label{eq:dispersion appendix}
@@
 Hence, we ignore these effects while keeping the $O(v^2/c^2)$ correction that determine the Langmuir wave because these corrections depend on the wavevector.}
-\begin{eqnarray}
- \int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2c^2
- \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
-&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),
+\ab{
+\begin{equation}
+  \int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p
+  \approx k^2c^2  \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p
+  \approx \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),
 \label{eq:dispersion-real-part}
-\end{eqnarray}
-where the second term is zero because it is odd, $\lambda_D^2 = k_B T_t/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
-These wave are not legitimate and result from the Taylor expansion of the denominator of equation (\ref{eq:dispersion-real-part}).
-A correct treatment of equation (\ref{eq:dispersion-real-part}) with the appropriate Landau contours will give the correct growing or damping behavoirs for waves with phase speeds approximately that of the electron phase speeds.}.
+\end{equation}
+}
+%\begin{eqnarray}
+% \int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2c^2
+% \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
+%&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),
+%\label{eq:dispersion-real-part}
+%\end{eqnarray}
+where the second term is zero because it is odd, $\lambda_D^2 = k_B T_t/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to Equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
+These wave are not legitimate and result from the Taylor expansion of the denominator of Equation (\ref{eq:dispersion-real-part}).
+A correct treatment of Equation (\ref{eq:dispersion-real-part}) with the appropriate Landau contours will give the correct growing or damping behavoirs for waves with phase speeds approximately that of the electron phase speeds.}.
 If we ignore the third term in the kinetic dispersion relation (\ref{eq:dispersion kinetic}), this yields two plasma modes: an undamped plasma oscillation mode with $\omega = \omega_{p,t}$ and a longitudinal electron plasma wave, i.e., Langmuir wave, with
@@
 This frame moves with a velocity, $\vel_{\rm ph} = \omega_k/k \zphat$.
-With an eye toward computing the residue that will appear in equation (\ref{eq:dispersion kinetic}), we define $p_{z'}=\gz\vz E_\perp$, $E=\gz E_\perp$, and
+With an eye toward computing the residue that will appear in Equation (\ref{eq:dispersion kinetic}), we define $p_{z'}=\gz\vz E_\perp$, $E=\gz E_\perp$, and
 $E_\perp=\sqrt{m^2+p_\perp^2}$ is the perpendicular energy.  In this case, we can rewrite the beam distribution function as
 \begin{eqnarray}
-F_B &=& \frac{m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T_b}\right)\\
+F_B &=& \frac{m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - v_{b,z'} p_{z'} - v_{b,x'} p_{x'})} {k_B T_b}\right)\nonumber\\
 %F_b &=& \frac{1}{4\pi\gammabeam m^2 T K_2(m/T)} \exp\left(-\frac {\gammabeam(E - \betabeamz p_{z'} - \betabeamx p_{x'}}{T}\right),\label{eq:dist1}\\
@@
 where we define $\mu = m_ec^2/k_BT_b$.
-Inserting the equation into (\ref{eq:ap dispersion kinetic}) and using the results of equation (\ref{eq:dispersion-real-part}), we find 
+Inserting the equation into (\ref{eq:ap dispersion kinetic}) and using the results of Equation (\ref{eq:dispersion-real-part}), we find 
 \begin{equation}
@@
 \begin{equation}
-\Gamma \simeq -\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R.
+\Gamma \ab{\approx}
+%\simeq
+-\omega_p\frac{\pi n_b}{2 n_t}   m_e v_B^2 R.
 \end{equation}
@@
 F_b
-\right|_{\rm pole}\\
-&=
+\right|_{\rm pole}
+=
 -
@@
 \frac{\kI}{m_e^4c^7}
-,
-\label{eq:Rogen}
-\end{equation}
-where
-\begin{equation}\label{eq:I-integral}
-\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right)
-\end{equation}
-and $\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$ and
+\ab{\sout{,} \quad\text{where}\quad}
+\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right),
+\label{eq:Rogen I-integral}
+\end{equation}
+\ab{\sout{where}}
+%\begin{equation}\label{eq:I-integral}
+%\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right)
+%\end{equation}
+\ab{\sout{and}} $\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$\ab{,} and
 $w\equiv \gammabeam v_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
@@
 Thus, we define $p_{x'}'=\gamma_w(p_{x'}-w E_\perp/c^2)$ and $p_{y'}'=p_{y'}$ and find:
-\begin{eqnarray}
-%E'_\perp &=& \gamma_w(E_\perp - w p_x),\\
-E_\perp &=& \gamma_w(E'_\perp + w p'_x),\\
-d p_x d p_y &=& \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
-\end{eqnarray}
-Inserting this into equation (\ref{eq:I-integral}), we find
+\begin{equation}
+  E_\perp = \gamma_w(E'_\perp + w p'_x)
+  \ab{\quad\text{and}\quad}
+  d p_x d p_y = \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
+\end{equation}
+%\begin{eqnarray}
+%%E'_\perp &=& \gamma_w(E_\perp - w p_x),\\
+%E_\perp &=& \gamma_w(E'_\perp + w p'_x),\\
+%d p_x d p_y &=& \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
+%\end{eqnarray}
+Inserting this into Equation (\ab{\sout{\ref{eq:I-integral}} \ref{eq:Rogen I-integral}}), we find
 \begin{equation}
@@
 \kI &= 
-\int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T_b}\right)\\
+\int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T_b}\right) %\\
 %&=
@@
 %\right) e^{-\cG'E'_\perp/m}\\
-&=
+%&
+=
 \pi \gamma_w^2 \int_0^\infty dp'^2_\perp \left(
@@
   \int_0^\infty dx \sqrt{1+x} e^{-a\sqrt{1+x}} = \frac{2 e^{-a}}{a^3}\left(a^2 + 2a + 2\right)
-\end{equation}
-and 
-\begin{equation}
+  \quad\text{and}\quad
 \int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
@@
 \end{equation}
+\ab{\sout{and}}
+%\begin{equation}
+%\int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
+%=
+%\frac{4(a+1)}{a^3} e^{-a}\,,
+%\end{equation}
 we find
@@
 \end{equation}
-Inserting this into (\ref{eq:Rogen}) yields
+Inserting this into (\ref{eq:Rogen I-integral}) yields
 \begin{equation}
@@
 \begin{equation}
-\Gamma \simeq - \Gamma_0
+\Gamma \ab{\approx}
+%\simeq
+- \Gamma_0
 %\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}

### 2016-08-16 oblique.tex
decolored changes

@@
 \citep[see, e.g., ][]{Naru-Tota:06,Knei-Mann:08,Inou-Tota:09,Vent:10}.
-These constraints preclude any dramatic rise in numbers of source by \ab{\sout{$z\sim1$} $z\approx1$}--$2$ that is seen in the quasar
+These constraints preclude any dramatic rise in numbers of source by $z\approx1$--$2$ that is seen in the quasar
 distribution.  That is, the comoving number of blazars must have
@@
 frame.
-As a result, an $\epm$ pair can be produced with Lorentz factor \ab{\sout{$\gamma = \left(1-v^2/c^2\right)^{-1/2} \simeq E/2m_e c^2$} $\gamma = \left(1-v^2/c^2\right)^{-1/2} \approx E/2m_e c^2$}, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.
+As a result, an $\epm$ pair can be produced with Lorentz factor$\gamma = \left(1-v^2/c^2\right)^{-1/2} \approx E/2m_e c^2$, where $v$ is the velocity of the pairs \citep{Goul-Schr:67}.
 Here, we discuss the distribution function of the
@@
 \begin{equation}\label{eq:perp dispersion}
-\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }\ab{%\sout{,}
-\quad\text{and}\quad}
+\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {2 k_B T_b}{\gammabeam^2 m_e c^2 }%\sout{,}
+\quad\text{and}\quad
 \frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
@@
 
-We now adopt perturbations of the form \ab{\sout{$\delta \sim \exp\left(i \kvec\cdot\rvec - i \omega t\right)$} $\delta \propto \exp\left(i \kvec\cdot\rvec - i \omega t\right)$} and without loss of generality assume that $\kvec = (k_x, 0, k_z)$, where $k_z$ is along the beam direction.
+We now adopt perturbations of the form  $\delta \propto \exp\left(i \kvec\cdot\rvec - i \omega t\right)$ and without loss of generality assume that $\kvec = (k_x, 0, k_z)$, where $k_z$ is along the beam direction.
 Linearizing the Vlasov-Maxwell equations then leads to the dispersion relation:
@@
 \begin{aligned}
-\Gamma &\ab{\approx} %\simeq
+\Gamma &\approx %\simeq
 - \Gamma_0
@@
 \begin{equation}
-\Gamma_{\rm bp} \ab{\approx} %\simeq
+\Gamma_{\rm bp} \approx %\simeq
 - \Gamma_0
@@
 \Gamma_M
-\ab{\approx}
+\approx
 %\simeq
@@
 \end{equation}
-For \ab{\sout{$k_\perp \sim \omega_p/v_b$} $k_\perp \approx \omega_p/v_b$}, this gives
+For $k_\perp \approx \omega_p/v_b$, this gives
 %\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
@@
 \end{equation}
-where we have dropped constant factors of order unity and assumed that \ab{\sout{$Z_x \sim O(1)$} $Z_x \propto O(1)$} and that the velocity dispersion is dominated by perpendicular (to the beam) component.
+where we have dropped constant factors of order unity and assumed that $Z_x \propto O(1)$ and that the velocity dispersion is dominated by perpendicular (to the beam) component.
 Hence, this defines the upper limit on the velocity dispersion of the plasma for the cold-plasma approximation to hold and, hence, the range of validity for the reactive oblique growth rate Equation (\ref{eq:growth rate reactive}).
@@
 
-Applying the scaling of the perpendicular \ab{\sout{(Equation \ref{eq:perp dispersion})}} and parallel velocity dispersions \ab{\sout{(Equation \ref{eq:parallel dispersion})}} \ab{(Equations \ref{eq:perp dispersion})} to these results and assuming $v_b \approx c$, we find
-\ab{
+Applying the scaling of the perpendicular and parallel velocity dispersions (Equations \ref{eq:perp dispersion}) to these results and assuming $v_b \approx c$, we find
+
 \begin{equation}
@@
 \frac{\Delta v_{\parallel}}{\vbeam} \approx \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.  
-\end{equation}}
+\end{equation}
 %\begin{eqnarray}
@@
 Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $\vbeam - \Delta v < \omega/k < \vbeam$.
-Hence the bandwidth over which the distribution powers grows is \ab{\sout{$\Delta \omega \sim k \Delta v$} $\Delta \omega \approx k \Delta v$}.
+Hence the bandwidth over which the distribution powers grows is  $\Delta \omega \approx k \Delta v$.
 For the kinetic growth rate to be valid, the bandwidth, $\Delta\omega$, must be large compared to the growth rate; otherwise, the entire beam contributes to the growth and, hence, the reactive regime applies.
@@
 \begin{equation}
-\Gamma \ab{\approx}
+\Gamma \approx
 %\sim
@@
 \begin{eqnarray}
-\frac{\nb}{\nIGM} & \ab{\approx} %\simeq
+\frac{\nb}{\nIGM} & \approx %\simeq
 &
 \frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \\
-&\ab{\approx}
+&\approx
 %\simeq
@@
 
-In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation \ref{eq:generic reactive regime}) and kinetic regimes (Equation \ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, \ab{\sout{$\Delta\omega\sim k\Delta v$} $\Delta\omega\approx k\Delta v$}, with the growth rate, $\Gamma$.
+In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (Equation \ref{eq:generic reactive regime}) and kinetic regimes (Equation \ref{eq:generic kinetic regime}) by comparing the frequency spread of resonant waves, $\Delta\omega\approx k\Delta v$, with the growth rate, $\Gamma$.
 Applying these conditions (eqns.~\ref{eq:generic reactive regime} and \ref{eq:generic kinetic regime}) to the ultrarelativistic $\epm$ pair beams of interest, we find for the controlling parameter:
-\ab{
 \begin{multline}
@@
 \end{multline}
-}
 %\begin{eqnarray}
@@
 and
-\ab{
 \begin{multline}
@@
 where $\gamma_6 = \gamma/10^6$. 
-We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 0.3\,\TeV$, or for large $\gamma$, which occurs for \ab{\sout{$E_{\rm ph} \sim 10\,\TeV$} $E_{\rm ph} \approx 10\,\TeV$} at $z=0$.
+We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the oblique instability always exists in the kinetic regime, but the beam plasma instability is in the reactive regime for $z\gtrsim 1$, for sufficiently cold beams $k_BT_b/m_e c^2 \lesssim 0.5$, which occurs for $E_{\rm ph} \lesssim 0.3\,\TeV$, or for large $\gamma$, which occurs for  $E_{\rm ph} \approx 10\,\TeV$ at $z=0$.
 
@@
 \item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regimes.
-We have recovered the reactive scalings for the beam-plasma mode \ab{\sout{$\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$} $\Gamma \approx \gamma^{-1}(n_b/n_t)^{1/3}$} and the oblique mode \ab{\sout{$\Gamma \sim (n_b/\gamma n_t)^{1/3}$} $\Gamma \approx (n_b/\gamma n_t)^{1/3}$}.
+We have recovered the reactive scalings for the beam-plasma mode  $\Gamma \approx \gamma^{-1}(n_b/n_t)^{1/3}$ and the oblique mode $\Gamma \approx (n_b/\gamma n_t)^{1/3}$.
 In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.
@@
 \begin{equation}
-  \fLAB\ab{[}\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM})\ab{]} = \fCM(E_{\rm COM}).
+  \fLAB[\ppos_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM}),\pmom_{\rm L}(\ppos_{\rm COM},\pmom_{\rm COM})] = \fCM(E_{\rm COM}).
 \end{equation}
@@
 This gives:
-\ab{
 \begin{equation}\label{eq:betabarLAB}
@@
 \end{equation}
-}
 %\begin{eqnarray}\label{eq:betabarLAB}
@@
 perpendicular to the boost, we find: 
-\ab{
 \begin{equation}
@@
 \end{equation}
-}
 %\begin{eqnarray}
@@
 In the lab frame, it is again useful to break it into components -- the parallel component becomes 
-\ab{
 \begin{equation}
@@
 while the perpendicular component becomes
-\ab{
 \begin{equation}
@@
 This simple scaling of the perpendicular velocity dispersion can be understood as a scaling with time between two frames boosted relative to each other, where the coordinates perpendicular to the boost axis remain invariant.
-This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., \ab{\sout{$mv^2 \sim kT$} $mv^2 \approx kT$} -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
+This result is also in line with the transformation of temperature as $T\rightarrow T/\gamma$ under a boost, i.e., $mv^2 \approx kT$ -- two factors of $1/\gamma$ from the perpendicular velocity dispersion is countered by one factor of $\gamma$ from the mass.
 Let us now consider the parallel component (Equation \ref{eq:dbetabarLABpara}) again to lowest order in $\betaCM$:
@@
 Hence, we ignore these effects while keeping the $O(v^2/c^2)$ correction that determine the Langmuir wave because these corrections depend on the wavevector.}
-\ab{
 \begin{equation}
@@
 \begin{equation}
-\Gamma \ab{\approx}
+\Gamma \approx
 %\simeq
@@
 \frac{\kI}{m_e^4c^7}
-\ab{\sout{,} \quad\text{where}\quad}
+\quad\text{where}\quad
 \kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right),
@@
 \end{equation}
-\ab{\sout{where}}
 %\begin{equation}\label{eq:I-integral}
@@
 %\end{equation}
-\ab{\sout{and}} $\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$\ab{,} and
+$\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$, and
 $w\equiv \gammabeam v_{b,x'}/\cG \le 1$.  This latter inequality is guaranteed as 
@@
   E_\perp = \gamma_w(E'_\perp + w p'_x)
-  \ab{\quad\text{and}\quad}
+  \quad\text{and}\quad
   d p_x d p_y = \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
@@
 %\end{eqnarray}
-Inserting this into Equation (\ab{\sout{\ref{eq:I-integral}} \ref{eq:Rogen I-integral}}), we find
+Inserting this into Equation (\ref{eq:Rogen I-integral}), we find
 \begin{equation}
@@
 \end{equation}
-\ab{\sout{and}}
 %\begin{equation}

### 2016-09-20 apjRevised/response.txt
revision response

REVIEWER:
The authors investigate the linear kinetic instability of ultrarelativistic pairs against the excitation of oblique electrostatic oscillations in the unmagnetized
intergalactic medium, where the ultrarelativistic pairs result from the annihilation of TeV photons from extragalactic blazars with the extragactic backgound light (EBL). They demonstrate that the linear growth times of the instability are substantially shorter than the inverse Compton (with the cosmic microwave background radiation (CMB)) energy loss time of the pairs, and conclude from that that most of the kinetic energy of the pairs is dissipated into the electrostatic turbulence.

In the literature only the linear kinetic instability for electrostatic oscillations with wave vectors parallel to the beam direction has been calculated (Schlickeiser, Krakau and Supsar 2013, ApJ 777, 49), whereas oblique electrostatic oscillations were only treated in the reactive limit (Schlickeiser, Ibscher and Supsar 2012, ApJ
758, 102). Therefore the analysis of the authors in principle could be a valuable addition to the literature. However, for two reasons, detailed below, the manuscript in its present form is not acceptable for publication.

RESPONSE:

We thank the referee for a thoughtful report.

REVIEWER:

1) The authors calculated the linear growth rate by adopting the inappropriate relativistically drifting Maxwellian distribition function (Eqs. (1) and (3)) for the pair
production spectrum from the double photon annihilation process. But this Maxwellian distribution function does not result from the double photon annihilation process. The correct pair production spectrum has been calculated before in Eq. (154) of Schlickeiser et al. (2012, ApJ 758, 101) -- see also the approximations (3) and (5) in Schlickeiser, Krakau and Supsar (2013, ApJ 777, 49) -- and by Aharonian, Atoyan and Nagapetian (1983, Astrofizika (Astrophysics), 19, 1), depending on the incoming TeV gamma-ray enery spectrum and the EBL target photon spectrum. This results in a power law energy spectrum with a strong exponential cut-off at small electron energies, which is substantially different from a drifting Maxwellian. It is necessary for the authors to redo their linear instability analysis with the correct pair production energy spectrum.

RESPONSE:

We agree that a drifting Maxwellian is not the correct distribution function to model a pair beam. But we adopt it because:
1. It is simpler and allows us to calculate the kinetic instability exactly in the electrostatic approximation.
2. It has all the features that allows us the elucidate the physics (continuous spread of momenta in the parallel and perpendicular directions)
3. It is relatively not ad-hoc — by this we mean that it is a distribution that can be realized in physical situation.

We have added an additional paragraph to the end of section 2.1 elucidating this choice and saying that a correct calculation involves using a distribution function calculated from the \gamma\gamma process such as the ones mentioned by the referee.  We also note that the analytic techniques that we used for calculating the drifting Maxwellian may be useful in the full calculation.

REVIEWER:

2) Claiming that most of the kinetic energy of the pairs is dissipated into electrostatic turbulence on the ground, that the minimum linear instability groth time is shorter
than the inverse Compton energy loss time, is insufficient and not correct. The backreaction of the electrostatic turbulence on the unstable pair distribution function will ultimately lead to a plateaued distribution function in the original beam direction (see Menzler and Schlickeiser 2015, MNRAS 448, 3405) on a time scale much longer than the minimum linear growth time. The pairs are so redistributed in energy but still able to inverse Compton scatter the CMB photons. The authors should address this redistribution time scale and compare it with the inverse Compton energy loss time scale.

RESPONSE:

We completely agree with the referee that this statement is not correct and while we do not make this claim in this paper, we can understand how the discussion in the introduction can give this impression.  We were aware of these subtleties and focus just on the linear physics in this paper avoid any quantitative comparison with the ICC timescales.  We merely confirmed the linear results uses in earlier papers, clarified the physics, especially in regard to the broad support of modes that are unstable, and address the transition between the kinetic and reactive mode for the oblique instability.

To help clarify this discussion, we have included in the 4th paragraph of the introduction:

We note that this effectiveness of this plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.

We have also added a paragraph below the sixth paragraph

### 2016-09-20 apjRevised/ms.tex
revision response

--- 2016-08-16 oblique.tex
@@
 %\documentclass[usenatbib,iop,apj]{emulateapj}
-\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2015}
+%\documentclass[usenatbib,iop,apj,numberedappendix]{aeb_emulateapj_2015}
+\documentclass[usenatbib,iop,apj,numberedappendix]{aastex}
 
@@
 
-\usepackage[normalem]{ulem}
-\usepackage[usenames]{color}
-
-\definecolor{DarkGreen}{rgb}{0.64,0.80,0.35}
-\newcommand\ac[1]{{\color{green} \bf #1}} %% AEB color
-\newcommand\ab[1]{{\color{green} \bf #1}} %% AEB color
-\newcommand\cp[1]{{\color{blue} #1}} %% CP color
-\newcommand\ep[1]{{\color{magenta} \bf #1}} %% EP color
-\newcommand\asout[1]{\color{green}\sout{{\rm  #1}}} %% AEB strikeout of text
-\newcommand\qc[1]{{\color{red} \bf #1}}
-%\newcommand\qc[1]{{\color{black} #1}}
-\newcommand\cc[1]{{\color{blue} \bf #1}}
-\newcommand\phil[1]{{\color{cyan} #1}}
 \newcommand{\mathsc}[1]{\text{\textsc{#1}}}
@@
 
-%\newcommand{\vel}{\ensuremath{\boldsymbol v}}
-%\newcommand{\pmom}{\ensuremath{\boldsymbol p}}
-%\newcommand{\Efield}{\ensuremath{\boldsymbol E}}
-%\newcommand{\Bfield}{\ensuremath{\boldsymbol B}}
-%\newcommand{\grad}{\ensuremath{\boldsymbol \nabla}}
-%\newcommand{\gradp}{\ensuremath{\boldsymbol \nabla}_p}
-%\newcommand{\ppos}{\ensuremath{\boldsymbol x}}
-
 \newcommand{\vel}{\ensuremath{\bmath{v}}}
@@
 
-%\newcommand{\betavec}{\ensuremath{\boldsymbol \beta}}
-%\newcommand{\betaLAB}{\ensuremath{\boldsymbol \beta_{\rm L}}}
-%\newcommand{\betaCM}{\ensuremath{\boldsymbol \beta_{\rm COM}}}
 \newcommand{\betaCMpara}{\ensuremath{\beta_{\rm COM,\parallel}}}
@@
 
-%\newcommand{\betabar}{\ensuremath{\overline{\boldsymbol \beta}}}
 \newcommand{\betabar}{\ensuremath{\overline{\bmath{\beta}}}}
@@
 
-%\newcommand{\kvec}{\ensuremath{\boldsymbol k}}
-%\newcommand{\rvec}{\ensuremath{\boldsymbol r}}
-%\newcommand{\rvec}{\ensuremath{\boldsymbol j}}
-
 \newcommand{\kvec}{\ensuremath{\bmath{k}}}
@@
 \newcommand{\jvec}{\ensuremath{\bmath{j}}}
-
-%\newcommand{\xhat}{\ensuremath{\hat{\boldsymbol x}}}
-%\newcommand{\yhat}{\ensuremath{\hat{\boldsymbol y}}}
-%\newcommand{\zhat}{\ensuremath{\hat{\boldsymbol z}}}
 
@@
 
-\newcommand{\betabeam}{\ensuremath{\bmath{\beta}_{\rm b}}}
-\newcommand{\betabeamz}{\ensuremath{\beta_{\rm b,z}}}
-\newcommand{\betabeamx}{\ensuremath{\beta_{\rm b,x}}}
-\newcommand{\gammabeam}{\ensuremath{\gamma_{\rm b}}}
-\newcommand{\vbeam}{\ensuremath{v_{\rm b}}}
+\newcommand{\betabeam}{\ensuremath{\bmath{\beta}_{b}}}
+\newcommand{\betabeamz}{\ensuremath{\beta_{b,z}}}
+\newcommand{\betabeamx}{\ensuremath{\beta_{b,x}}}
+\newcommand{\gammabeam}{\ensuremath{\gamma_{b}}}
+\newcommand{\vbeam}{\ensuremath{v_{b}}}
 \newcommand{\gammaphase}{\ensuremath{\gamma_{\rm ph}}}
@@
 
-%\author{Some people}
 \author{
@@
 
-%\maketitle
-
 \begin{abstract}
@@
 These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
-%\ab{[The only issue with the above is that the primary issues people have raised are with nonlinear effects, e.g., saturation or nonlinear coupling to existing background structures.
-Should we be quite as strident then in the concluding sentences?]}
 \end{abstract}
@@
 
-%CP: this appears twice - I cut the following:
-%This excess heating of the IGM may resolve a variety
-%of cosmological puzzles, including naturally explaining
-%  anomalies in the statistics of the high-redshift Ly$\alpha$ forest
-%\citep{paperIV} and potentially explaining a number of the X-ray
-%  properties of groups and clusters and anomalies in galaxy formation on
-%  the scale of dwarfs \citep{paperIII,Lu+2013}.  We
-%  have recently shown that \textit{if} the IC halos are ignored, it is possible
-%  to quantitatively reproduce the redshift and flux distributions
-%  of nearby hard gamma-ray blazars and the extragalactic gamma-ray background
-%  spectrum above 3~GeV simultaneously with a unified model of AGN evolution
-%  \citep{Broderick+2013,Broderick+2013b}.  All of these empirical successes provide
-%  circumstantial evidence for the presence of virulent plasma beam
-%  instabilities.
-
 The \Fermi satellite and ground-based imaging atmospheric Cherenkov telescopes such as
-% Phil: I added "imaging atmospheric" to distinguish this class of
-% telescopes from imaging water CTs, e.g., MILAGRO; I am not aware of
-% any space based IACT apart from a planned telescope on the ISS, but
-% who knows whether this will ever fly
 H.E.S.S., MAGIC, and VERITAS\footnote{High Energy
@@
  IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
-(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}).
-
-% Phil: I cut this part because it is written as a fact (while it is not) and the same long author list is repeated in the next paragraph (the \qc command did not work with the \citep command):
-% and intergalactic magnetic fields (IGMF)
-% \citep[IGMF; ][]{Nero-Vovk:10,Tave_etal:10a,Tave_etal:10b,Derm_etal:10,Tayl-Vovk-Nero:11,Dola_etal:11,Taka_etal:11,Vovk+12}.
+(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}). {\bf We note that this effectiveness of this plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.}
 
@@
 
+{\bf Additional nonlinear effects may also be important.
+For instance, for sufficiently powerful blazers, the modulation instability may operate \citep{Schlickeiser+12,Chang+14,2015MNRAS.448.3405M}, allowing for a rapid transfer of electrostatic wave energy into thermal energy.
+For less powerful blazars, the combination of NLD and quasilinear damping, i.e., beam plateauing, will also reduce the rate of damping compared to the linear rate, and alters the resulting IC spectra \citep{2015MNRAS.448.3405M}.
+Further study of these effects will help clarify these points.}
+
 While a full nonlinear study is required, we focus on the nature of the linear instability in this paper, clarifying its robustness and regimes of applicability.
@@
 
-% \ac{[There are some notational issues throughout the text, which I'll list here.  Usually it is simply a matter of choosing, and then sticking to, a convention.
-% \begin{itemize}
-% \item Some equations assume $c=1$ while others have the $c$'s explicitly present.
-Unless there is a compelling reason (e.g., clarity of derivation vs.
-utility of an answer in physical units) we should choose one and mention the convention here (at the end of the intro) if $c=1$.
-% \item $\beta$ and $v$ appear to be used interchangeably, again related to the $c$ or not question.  Again, we should choose one.
-% \item Throughout section 2 and in the Appendix the frame is defined in terms of $\parallel$ and $\perp$ directions (i.e., $\beta_\parallel$ and $\bmath{\beta}_\perp$, note the latter is a {\em vector}!) while in sections 3 and in the figures we refer to $\beta_z$ and $\beta_x$ for these.
-% \item ApJ style, which I'm assuming is what we are following, has vectors given by bold, roman fonts and scalars by the italic fonts.
-Thus, it should be $\bmath{k}\cdot\bmath{\delta E}$ or $\bmath{k}\cdot\delta\bmath{E}$, etc.
-I've tried to fix up various definitions in the preamble to do this, but that should be checked.
-% \end{itemize}
-% In any equations I've added/modified I will include the $c$'s and use $\beta$ for concreteness, though I anticipate there will be few such examples.]}
-
 \section{Ultrarelativistic Pair Beams from VHEGRs}\label{sec:setup}
@@
 photon ($E_{\rm ebl}$) exceed the rest mass energy of the $\epm$ pair in the
-center of momentum (COM \ab{[Should we use CM instead for compactness? Otherwise ``of'' is getting a big promotion!]}) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
+center of momentum (COM) frame, i.e., $2 E E_{\rm ebl}(1-\cos\theta) \geq 4
 m_e^2 c^4$, where $\theta$ is the relative angle of propagation in the lab
@@
 
-%As the mean free path of these VHEGR photons to pair production is short compared to the Hubble length, observations of extragalactic VHEGR sources have been used as probes of the EBL \citep{Stec-deJa-Sala:92,deJa-Stec-Sala:94,Ahar_etal:06}
-
 \subsection{Distribution Function of the Pair Beam}
@@
 \end{equation}
-where $\gammabeam = \gamma(v_{\rm b}) = \left(1 - v_{\rm b}^2/c^2\right)$ is the Lorentz factor of the beam and $v_{\rm b}$ is the bulk velocity of the pair beam.
+where $\gammabeam = \gamma(v_b) = \left(1 - v_b^2/c^2\right)$ is the Lorentz factor of the beam and $v_b$ is the bulk velocity of the pair beam.
 Inserting this into (\ref{eq:relativistic maxwellian}), we find the Maxwell-J{\"u}ttner distribution \citep{1911AnP...340..145J,1975PhRvA..12..686W}
 \begin{eqnarray}\label{eq:distribution function beam} %{eq:drifting maxwellian}
-%f = \frac{1}{4\pi\gammabeam m^2 T_{\rm L} K_2(m/T_{\rm L})} \exp\left(-\frac {E_{\rm L} - \betaBOOST p_{\rm L,\parallel}} {T_{\rm L}}\right), \\
 f = \frac{n_b m_e c^2}{4\pi\gammabeam k_B T_b K_2(m_ec^2/k_B T_b )m_e^3c^3} \exp\left(-\frac {\gammabeam(E - \vel_b\cdot \bmath{p})} {k_B T_b}\right),
@@
 where $K_2$ is the 2nd order modified Bessel function and $T_b$ is the comoving temperature of the beam.
-
-%\ab{[Do we want to describe the characteristics of this distribution at any further length?
-I recall you had a detailed discussion of the characteristics of a boosted thermal distribution which was nice because it clarified some obvious misconceptions that seemed to continuously arise.
-Also, do we want to discuss the potential evolution of the distribution?]}
 
@@
 \end{equation}
-%and 
-%\begin{equation}\label{eq:parallel dispersion}
-%\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T_b}{\gammabeam^4 m_e c^2 },
-%\end{equation}
 where $T$ is measured in the COM frame of the beam.  
@@
 
-%The transformation properties of the drifting Maxwell-J{\"u}ttner distribution coupled with the Lorentz invariance of the distribution function (as discussed in Appendix~\ref{sec:lorentz}) demonstrates that the drifting Maxwell-J{\"u}ttner distribution is the correct description of a boosted isotropic Maxwellian (relativistic or otherwise).
-
-%In Appendix~\ref{sec:lorentz}, we explicitly derive the scaling of the
-%  parallel and perpendicular velocity dispersions with the Lorentz factor upon
-%  boosting the distribution function to the lab frame.
-%In the limit in which the perpendicular velocity dispersion in the COM frame is non-relativistic, i.e.,
-%$\left|\betaCM\right| \ll 1$, we find to lowest order in $\betaCM$
-%\begin{equation}
-% \dbetabarLABperp \approx \frac{\dbetabarCMperp}{\gammabeam^2}.
-% \label{eq:betaperp}
-%\end{equation}
-%This simple scaling of the perpendicular velocity dispersion is an immediate 
-%consequence of how transverse velocities transform under Lorentz boosts, and may be understood
-%in terms of time dilation.
-%
-%This result is also anticipated by the standard transformation of temperature under boosts, $T\rightarrow T/\gamma$ (see Equation~\ref{eq:drifting maxwellian}).
-%This can be understood by considering $mv^2 \sim kT$ -- we pick up two factors of $1/\gamma$
-%from the perpendicular velocity dispersion that is countered by one factor of
-%$\gamma$ from the mass scaling. 
-%
-%For the parallel component of the velocity
-%dispersion in the non-relativistic limit, $\left|\betaCM\right| \ll 1$, we find
-%to lowest order in $\betaCM$:
-%\begin{equation}
-% \dbetabarLABpara \approx \frac{\dbetabarCMpara}{\gammabeam^{4}},
-% \label{eq:betapara}
-%\end{equation}
-%Again this scaling follows immediately from the transformation of velocities under Lorentz boosts, and may be simply understood as a consequence of both time dilation and, for the parallel case, length contraction.
-%This scaling of the parallel component of the velocity
-%dispersion has important consequences below.
-%
-
-% To see how this distribution function appears in velocity space, we show four examples of the  Maxwell-J{\"u}ttner distribution in the center-of-mass frame for the nonrelativistic temperatures ($T=10^{-2}$ and $ 0.1$) in Figure \ref{fig:distribution_non_rel} and for relativistic temperatures ( $T=1$ and $10$) in Figure \ref{fig:distribution_rel}, where $T$ is listed in the natural relativistic units: $m_e c^2/k_B$.
-% We plot these in terms of $\beta_z$, which is parallel to the beam and $\beta_x$, which is perpendicular to the beam direction though, in the center-of-mass frame, an isotropic distribution makes no distinction between these directions.
-We note that not all combinations of $\beta_x$ and $\beta_z$ are allowed as $\beta_x^2 + \beta_z^2 < 1$.
-\ac{[Could you blank out the portions for which $\beta^2>=1$ in the figure to both drive this point home as well has emphasize the scaling difference?]}
-% 
-% Also plotted in Figures \ref{fig:distribution_non_rel} and \ref{fig:distribution_rel} on the right side is the distribution function that is boosted by $\gammabeam = 10^6$.
-Motivated by our discussion in Appendix~\ref{sec:lorentz}, we have scaled the perpendicular component of the velocity by $\gammabeam^{-1}$ and the parallel component by $\gammabeam^{-2}$.
-% When $T\ll1$, as anticipated above, the parallel and transverse velocity dispersions are well approximated by the scaling in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}).
-As seen explicitly in Figure \ref{fig:distribution_non_rel}, the distribution functions remain roughly isotropic in the rescaled and shifted velocity units.
-% 
-% For relativistic temperatures, the conditions under which the expressions in Equations (\ref{eq:betaperp}) and (\ref{eq:betapara}) were derived no longer hold.
-Nevertheless, despite the importance of higher-order and anisotropic terms in the distribution function, the magnitudes of the velocity dispersions remain well approximated by these scalings.
-This may be seen explicitly in the right panels of Figure \ref{fig:distribution_rel}.
-% While these scaling will necessarily break down as $T \rightarrow \gammabeam$, for the problem of interest this is never the case.
-Hence, for both the nonrelativistic and mildly relativistic case, the velocity dispersion generally scales like $\gammabeam^{-1}$ in the perpendicular direction and like $\gammabeam^{-2}$ in the parallel direction provided the boost is much larger than the center-of-mass temperature ($\gammabeam \gg T$).
-% 
-% \begin{figure*}
-%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_01.pdf}
-%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_01.pdf}\\
-%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_0_1.pdf}
-%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_0_1.pdf}\\
-%  \caption{Distribution function in velocity space in the center-of-mass frame (left) and the boosted (lab) frame (right) for nonrelativistic temperatures: $T=0.01$ (top) and $0.1$ (bottom).
-Note that while we plot $\beta_z$ and $\beta_x$ from -1 to 1, not all values of $\beta_z$ and $\beta_x$ are allowed.
-\ac{[Can this be shown explicitly, e.g., by making those values NaN, or some other method that results in no plot in that instance?
-Also, how are these normalized?]}  On the right plot, we show the distribution in a boosted frame by a Lorentz factor of $10^6$.
-As these plots would be compressed near $\beta_z \approx 1$, we have chosen to plot $\beta_z-\beta_b$ on the x-axis.
-We have also multiplied the $\beta$ factors by a factor of $\gammabeam^2$ for the z (parallel) component and $\gammabeam$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
-These plots show that a non-relativistic velocity dispersion transforms as $\gammabeam^{-2}$ and $\gammabeam^{-1}$ for the parallel and perpendicular direction, respectively.\label{fig:distribution_non_rel}}
-% \end{figure*}
-% \begin{figure*}
-% \includegraphics[width=0.5\textwidth]{distribution_cm_E_1.pdf}
-%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_1.pdf}\\
-%  \includegraphics[width=0.5\textwidth]{distribution_cm_E_10.pdf}
-%  \includegraphics[width=0.5\textwidth]{distribution_boosted_E_10.pdf}
-%  \caption{Same as Figure \ref{fig:distribution_non_rel} but for relativistic temperatures: $T=1$ (top) and $10$ (bottom).
-Again, we have multiplied the $\beta$ factors on the right plot by a factor of $\gammabeam^2$ for the z (parallel) component and $\gammabeam$ for the x (perpendicular) component inspired by our discussion in Appendix~\ref{sec:lorentz}.
-Even here the dominant effect of the boost remains the rescaling of the velocity dispersions obtained formally for $T\ll1$.
-Nevertheless, higher-order corrections are beginning to spread the distribution function over a velocity range larger than that anticipated by these scalings.
-% \ac{[Can you make the righthand panels with square aspect ratios?
-Right now, I think that they extend a factor of two more in the parallel direction, which appears to give the impression that they are more isotropic than they really are!
-Also, how are the normlizations obtained?
-Specifically, are they unit normalized?]}
-% \label{fig:distribution_rel}}
-% \end{figure*}
-% 
-% Motivated by these scalings, to probe the dependence of the oblique and beam instabilities upon $\gammabeam$ and $T$, we adopt the following simplified distribution function\footnote{This form for $f$ may be explicitly derived in the $T\ll1$ limit by Taylor expanding $\gammabeam(E-\beta_b p_\parallel)$ around $\bmath{\beta}-\bmath{\beta}_{\rm B}=0$.}:
-% \begin{equation}
-%  f = \frac{n_b}{\sqrt{\pi^3\alpha_\parallel\alpha_\perp^2}}\exp\left(-\frac{\left(\beta_{\parallel}-\beta_b\right)^2}{\dbetabarpara}-\frac{\beta_{\perp}^2}{\dbetabarperp}\right),
-% \end{equation}
-% where $\alpha_\perp\equiv T/\gammabeam^2$ and $\alpha_\parallel \equiv T/\gammabeam^4 = \alpha_\perp/\gammabeam^2$ and here and henceforth we have dropped the subscript L, with all quanties are to be interpreted in the lab frame unless otherwise indicated.
-% This form of the distribution function has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.
+{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the distribution function that are produced by VHEGR photo annihilation is different \citep[see for instance][]{2012ApJ...758..101S}.
+In particular, the the parallel and perpendicular momentum spread will likely be different due to the distribution of VHEGR photons and their respective mean free paths.
+However, a Maxwell-J{\"u}ttner distribution is still useful as it has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).
+In addition, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.
+It also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.
+Finally, we note that the analytic methodology use to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}
 
@@
 Here, $s$ is the species label, $+$ for positrons and $-$ for electrons,
-with $q_{\pm} = \pm e$.  %\ac{[$v$ vs. $\vel$ vs. $\beta$ problems above.]}
+with $q_{\pm} = \pm e$. 
 Upon linearizing this in small perturbations about a background distribution, i.e., setting 
@@
 Linearizing the Vlasov-Maxwell equations then leads to the dispersion relation:
-%\begin{equation}\label{eq:dispersion relation}
-% \epsilon = 1 - \frac{\omega_{p}^2}{k}\int \frac{\kvec \cdot \gradp F}{\omega - \kvec\cdot\vel} d^3\!p = 0,
-%\end{equation}
 \begin{equation}\label{eq:dispersion relation}
@@
 Upon integrating by parts, Equation (\ref{eq:dispersion relation}) becomes
-%\begin{eqnarray}
-% 1 - \frac{4\pi q_s^2}{m_s k^2}\int F_s\kvec \cdot \gradp\frac{1}{\omega - \kvec\cdot\vel} d^3p &=& 0\\
-% 1 - \frac{4\pi q_s^2}{m_s k^2}\int F_s \frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &=& 0. \label{eq:dispersion relation by parts}
-%\end{eqnarray}
 \begin{multline}
@@
 \end{multline}
-%\ac{[$\beta$ vs. $v$ issue above.]}
 
@@
 Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.
-%This unstable wave corresponds to a destabilizing beam and does not correspond to any wave in the unperturbed target plasma.
 
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{off_res.pdf}
+\includegraphics[width=0.5\textwidth]{f1.pdf}
 \caption{Beam-plasma growth rate (solid line) and unstable wave frequency (dashed-line) as a function of
@@
 For the beam plasma, we again adopt the Maxwell-J{\"u}ttner distribution (Equation~\ref{eq:distribution function beam}).
-%\begin{equation}\label{eq:distribution function beam}
-%F_b = \frac{n_b}{4\pi\gammabeam m^2 T K_2(m/T_{\rm L})} \exp\left(-\frac {\gammabeam(E - \betabeam \cdot \pmom)} {T}\right),
-%\end{equation}
 Inserting these into the dispersion relation (Equation \ref{eq:dispersion relation}), we find
-%\begin{equation}\label{eq:dispersion kinetic}
-% 1 - \frac{\omega_{p,t}^2}{k}\int F_t\frac{k^2 - (\kvec\cdot\betavec)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p - \frac{\omega_{p,t}^2}{k}\int \frac{\kvec\cdot\gradp F_b}{\omega - \kvec\cdot\vel}d^3p  = 0,
-%\end{equation}
 \begin{equation}\label{eq:dispersion kinetic}
@@
 - \Gamma_0
-%\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3 }\\
 \frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}\\
@@
 \end{equation}
-where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity %\ab{[While this is true, I think $\cG'$ is a little different -- it relates to the transverse boosted frame.]}\phil{[Hopefully, this captures it.]}.
+where $\mu = m_e c^2/k_B T_b$, $\gph = \left(1 - \vph^2/c^2\right)^{-1/2}$, $\vph = \omega/k$ is the phase velocity of the wave, $v_{b,z'}$ is the velocity of the beam oriented along the wavevector, $v_{b,x'}$ is the velocity of the beam perpendicular to the wavevector, $\gamma_w = \left(1 - w^2/c^2\right)^{-1/2}$, $w = \gamma_{\rm ph}^{-1}v_{b,x'}/(1 - v_{b,z'}v_{\rm ph}/c^2)$ is the beam velocity transverse to the wavevector in a frame that is comoving with the wave at its phase velocity, and $\cG'\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)/\gamma_w$ is the Lorentz factor of a beam particle in a frame that is comoving with the wave at the phase velocity and the transverse (to the wavevector) beam bulk velocity
 Finally, we define the typical maximum growth rate, $\Gamma_0$, as 
@@
 - \Gamma_0
-%\frac{\pi \gamma_w^2 \gph^3 m_e (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
 \frac{\pi \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG^3 c}
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{pp1.pdf}
+\includegraphics[width=0.5\textwidth]{f2.pdf}
 \caption{Oblique growth rate maximized over $\vph$ as a function of
@@
 \begin{figure*}
-\includegraphics[width=0.5\textwidth]{pp2.pdf}
-\includegraphics[width=0.5\textwidth]{pp3.pdf}
-
-%\plottwo{pp2.pdf}{pp3.pdf}
+\includegraphics[width=0.5\textwidth]{f3a.pdf}
+\includegraphics[width=0.5\textwidth]{f3b.pdf}
 \caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right),
@@
 
-% We discuss the solution to Equation (\ref{eq:dispersion kinetic}) in Appendix \ref{sec:solution kinetic}.  The associate growth rate for the kinetic beam plasma instability is  
-% \begin{equation}\label{eq:growth-rate-bp}
-%  \Gamma = -\frac{\sqrt{\pi} n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\parallel}}\alpha'_{\parallel}}\exp\left(-\frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}}\right),
-% \end{equation}
-% where $\alpha'_{\parallel} = \gamma^4\alpha_{\parallel}$ is the rescaled factor such that it is order unity.
-% 
-% In Figure \ref{fig:bp-growth-rate}, we plot the growth rate for the beam-plasma instability (eq.[\ref{eq:growth-rate-bp}]) for $\gamma^4\alpha_{\parallel} = 0.1, 0.3, 1,$ and 3.
-The scaling of $\alpha_{\parallel}$ with $\gamma$ removes the overall scaling of $\alpha_{\parallel}$ with $\gamma$.
-We have also parametrized the wavevector in terms of $\Delta k = k - k_0$, where $k_0 = \omega_{p,t}/\vbeam$ is the wavevector of the wave whose phase speed $v_{\rm ph} = \vbeam$ is equal to the speed of the beam.
-Here in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$, so the positive values of $\Delta k$ results in negative values of $\Delta v_{\rm ph}$, which puts these Langmuir waves in the rising part of the distribution function, i.e., $\partial F_2/\partial\Delta\beta_z > 0$, which is required for growth.
-Note as well that the maximum growth rate scales like $1/\gamma^4\alpha_{\parallel}$ in analogy with the scaling of $\Delta v^{-2}$ scaling in the non-relativistic case, where $\Delta v$ is the velocity dispersion \citep{Boyd}.
-% 
-% The general solution for arbitrary $\kvec$ produces a growth rate of 
-% \begin{eqnarray}
-% \Gamma &=&  -\sqrt{\frac{\pi k_z}{k_x}}\frac{n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\perp}}}\frac{k_z}{k}\left(\frac{1}{\alpha'_{\perp}} + \frac{1}{\alpha'_{\parallel}} - \frac{k_z}{k_x\alpha'_{\parallel}}\right)\nonumber\\
-% && \exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right),\label{eq:growth-rate-oblique}
-% \end{eqnarray}
-% where $\alpha'_{\perp} = \gamma^2\alpha_{\perp}$ is an O(1) factor.
-Again, we find that the overall scaling of the kinetic oblique instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.
-Also we have made the assumption that $k_x/k_z \gg 1/\gamma$.
-For $k_x/k_z < 1/\gamma$, the dispersion relation reverts back to the beam plasma case given in Equation (\ref{eq:growth-rate-bp})
-% 
-% % In Figure \ref{fig:oblique-growth-rate} we plot the growth rate for the oblique instability (eq.[\ref{eq:growth-rate-oblique}]) for $\alpha=\alpha'_{\perp} = \alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3.
-Here we have assumed that the perpendicular and parallel velocity dispersions are the same once the scaling for $\gamma$ has been removed.
-This is true for small values of $\alpha$ that result from a rest-frame nonrelativistic distribution in Figure \ref{fig:distribution_non_rel}, but would not be the case for a rest frame relativistic distribution as shown in Figure \ref{fig:distribution_rel} where larger $v/c$ corrections come into play.
-% 
-% Figure \ref{fig:oblique-growth-rate} also demonstrates that the oblique mode follows its name in that the fastest 
-% growing mode is neither perfectly parallel ($k_x/k_z = 0$) nor perpendicular ($k_x/k_z = \infty$), but rather at a k-vector 
-% where $k_x/k_z \sim O(1)$. This qualitatively matches the results of a full analysis of \citet{Bret-Grem-Beni:10}, who also found the greatest growth for nontrivial $k_x/k_z$. 
-% 
-% % \begin{figure}
-% %  \includegraphics[width=0.5\textwidth]{bp-growth-rate.pdf}
-% %  \caption{Growth rate for the relativistic kinetic beam-plasma instability (or bump on tail instability) for four different $\alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3.
-We choose to parametrized $\alpha_{\parallel}$ in this way to remove the scaling of $\alpha_{\rm parallel}$ with $\gamma$.
-On the x-axis, we have plotted in terms of $\gamma^2\Delta k/k_0$, where $k_0 = \omega_{p,t}/\vbeam$ and $\Delta k = k - k_0$.
-A Langmuir wave with $k=k_0$ moves at the exact speed of the beam.
-For growth, Langmuir waves must move with a phase velocity $v_{\rm ph} < \vbeam$ to be on the rising part of the distribution function.
-Hence, in the limit of large $\gamma$, $\Delta v_{\rm ph}/v_{\rm ph} =  -\Delta k/k_0$ where the extra factor of $\gamma^2$ is remove the scaling of $\Delta v_{\rm phase}$ with $\gamma$.
-% %  \label{fig:bp-growth-rate}}
-% % \end{figure}
-% 
-% \begin{figure*}
-%  \includegraphics[width=0.5\textwidth]{oblique-0_1.pdf}
-%  \includegraphics[width=0.5\textwidth]{oblique-0_3.pdf}\\
-%  \includegraphics[width=0.5\textwidth]{oblique-1.pdf}
-%  \includegraphics[width=0.5\textwidth]{oblique-3.pdf}
-%  \caption{Growth rate of the oblique instability for four different $\alpha=\alpha'_{\perp} = \alpha'_{\parallel} = 0.1, 0.3, 1,$ and 3.
-1.
-As in Figure \ref{fig:bp-growth-rate}, we choose to parametrized $\alpha_{\parallel}$ and $\alpha_{\perp}$ in this way to remove the scaling with $\gamma$.
-On the x-axis, we have plotted in terms of $\gamma \Delta k/k_0$, for the same reasons as in Figure \ref{fig:bp-growth-rate}, but scaling with $\gamma$ instead of $\gamma^2$.
-\label{fig:oblique-growth-rate}}
-% \end{figure*}
-%
-
 \subsection{The Transition between the Kinetic and Hydrodynamic Instability}\label{sec:transition}
@@
 For $k_\perp \approx \omega_p/v_b$, this gives
-%\ab{[While it is true that $\omega-\omega_{p,t}\approx\Gamma$, this is not fully fleshed out in the main text; rather it is an inline comment in Appendix B.
-Thus, this is likely to be a confusing part.
-Also, if $\kvec=\omega_{p,t}/\vel_b$ then isn't the LHS $\Delta v_\parallel/v_b$.]}
 \begin{equation}\label{eq:oblique reactive regime}
  \frac{\Delta v_{\perp}}{\vbeam} \ll \left(\frac{n_b}{\gammabeam n_t}\right)^{1/3}
-%\left(1 - \beta_b^2\cos^2\Psi\right)^{1/3}, 
 \end{equation}
@@
 \end{equation}
-%\begin{eqnarray}
-%\frac{\Delta v_{\perp}}{\vbeam} &\sim&  \gammabeam^{-1}\sqrt{\frac{k_BT_b}{m_ec^2}},\\
-%\frac{\Delta v_{\parallel}}{\vbeam} &\sim& \gammabeam^{-2}\sqrt{\frac{k_BT_b}{m_ec^2}}.
-%\end{eqnarray}
 Hence, the conditions for the reactive regime for the oblique mode (Equation~\ref{eq:oblique reactive regime}) and beam-plasma mode (Equation~\ref{eq:bp reactive regime}) can be reduced to 
@@
 \end{equation}
-where $r_e = e^2/m_e c^2$ is the classical electron radius and $v_e$ is the electron velocity in the COM frame.
-To find $v_e$, we use the energy of the electron in the COM frame, $E_{\rm e, COM}$, which is
+where $r_e = e^2/m_e c^2$ is the classical electron radius and $v_e$ is the electron velocity in the COM frame of the generated pair.\footnote{In this section, the COM frame and subscript ``COM'' refer to the center of momentum frame of the pair that is produced by a single $\gamma-\gamma$ annihilation.}  To find $v_e$, we use the energy of the electron in the COM frame, $E_{\rm e, COM}$, which is
 \begin{equation}
@@
 
-In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron), $E_{e,COM}$ for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
-%We use the COM energy in this case because in the COM frame, the energy of the EBL and TeV photon is the same and is $E_{\rm ph, COM} = \sqrt{E_{\rm EBL}E_{\rm ph}}$ and $\theta_{\rm COM} \approx \pi - \sqrt{E_{\rm EBL}/E_{\rm ph}} \approx \pi$; in the COM frame the VHEGR and EBL photon have the same energy but opposite momentum.
-\phil{A rough order of magnitude calculation bears this out.
-We begin by noting that $\tan\theta = p_{\rm EBL,\perp}/p_{\rm EBL, \parallel}$, where $p_{\rm EBL,\perp}$ and $p_{\rm EBL,\parallel}$ is the perpendicular and parallel components of the EBL photon's momentum relative to the TeV photon's momentum.
-Boosting to the COM frame, the parallel component of the EBL's photons momentum is $p_{\rm EBL, COM,\parallel} \approx -E_{\rm ph, COM}/c$, while the perpendicular momentum is conserved, $p_{\rm EBL, COM,\perp} = p_{\rm EBL, \perp}$.
-As a result, the angle in the COM frame is
-%\begin{equation}
-%\tan\theta_{\rm COM} \approx \frac{p_{\rm EBL, COM,\perp}}{p_{\rm EBL, COM,\parallel}} \sim -\sqrt{\frac{E_{\rm EBL}}{E_{\rm ph}}} \approx -10^{-3} \left(\frac{E_{\rm EBL}} {1\,{\rm eV}}\right)^{1/2}\left(\frac{E_{\rm ph}}{1\,{\rm TeV}}\right)^{-1/2}
-%\end{equation}
-%}
-%As a result, we can ignore the dependence on $\theta$ on the energy of the electron in the COM frame (Equation~\ref{eq:threshold}) and set $1-\cos\theta = 2$.
-\ep{[Not sure I understand this.
-The $E_{\rm EBL}, E_{\rm ph}$ we get from the spectra are not in the COM frame.
-So how do we get rid of $\theta$?
-Or is this an additional assumption here that head-on collisions will dominate pair production?
-Was that assumption used in Fig.
-4?]}\phil{[Assumption i $E_{\rm EBL} \ll E_{\rm ph}$ and so in the COM frame the photons have equal energy, but opposite momentum.]}
+In Figure \ref{fig:rates}, we plot the differential rate of pair production as a function of the COM energy of the electron (and positron), $E_{\rm e,COM}$ for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
 Note the distribution of COM energy for the electrons (and positrons) depends on the initial photon energy.
@@
 Hence we expect that these pairs are in the sub-relativistic to mildly-relativistic regimes in their COM frame.
-%Second, the effect of the EBL spectrum on the distribution of electron COM energies is also seen in the breaks in the distribution.
-For instance, the cutoff in $d\Gamma_{\pm}/dE_{\rm EBL}$ above $E_{\rm e, COM}/m_e c^2 \approx 3$ for $E_{\rm ph} = 1\,\TeV$ is due to the cutoff in the EBL spectrum above $10\,\eV$.
-Additionally, the break in $d\Gamma_{\pm}/dE_{\rm EBL}$ at the same position for $E_{\rm ph} = 10\,\TeV$ is due to change in the EBL spectrum at $1\,\eV$.
 
 \begin{figure}
- \includegraphics[width=0.5\textwidth]{rates.pdf}
+ \includegraphics[width=0.5\textwidth]{f4.pdf}
  \caption{Differential rate of pair production as a function of the COM energy of the electron (and positron) for a photon energy of $E_{\rm ph} = 0.3$ (dotted line), $1$ (solid line), $3$ (dash-dotted line) and $10\,\TeV$ (dashed line).
@@
 \end{multline}
-%\begin{eqnarray}
-%\frac{\gammabeam^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
-%&&\ab{\times}\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
-%\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
-%\end{eqnarray}
 and
@@
 \end{multline}
-%\begin{eqnarray}
-%\frac{\gammabeam}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT_b/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
-%&&\ab{\times}\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
-%\left(\frac{E}{\TeV}\right)^{1/3}\label{eq:beam bp condition}
-%\end{eqnarray}
 where $\gamma_6 = \gamma/10^6$. 
@@
 \begin{itemize}
-%\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.
-We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam.
-We also show that this can be identified with a drifting relativistic Maxwell-J{\"u}ttner distribution.
-}
 \item{We analytically calculated growth rate of the beam-plasma and oblique instabilities in both the reactive and kinetic regimes.
@@
 The average COM energies of the generated pairs implies that the oblique instability is in the kinetic regime, validating our results from BCP12.}
-%\item{We demonstrate that the oblique mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime.
-In any case, the use of either regime produces roughly the same growth rate.
-Moreover, the inclusion of plasma instability losses on the number density of beam particles may place the beam plasma mode firmly in the kinetic regime.}
 \end{itemize}
-
-%The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.
-In particular, by developing a simple distribution function, whose modes can be analytically calculated, we will develop a quasilinear theory for these modes in a forthcoming publication.
-Such a quasilinear calculation will capture the dominant wave-particle interaction and capture the initial nonlinear evolution of this instability.
 
@@
 
-%\onecolappendix
-%\appendix%[onecolumn]
-%\onecolumn
 \begin{appendix}
@@
 \end{equation}
-%\begin{eqnarray}\label{eq:betabarLAB}
-% \betabarLAB &=& N^{-1}\int \frac {\pmom_{\rm L}}{\gamma_{\rm L} m_ec} \fLAB \dVLAB  \nonumber\\
-%&=& N^{-1}\int \frac {\pmom_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM})}{\gamma_{\rm L}(\pmom_{\rm COM}, \ppos_{\rm COM}) m_ec} \fCM \dVCM.
-%\end{eqnarray}
 Breaking the components of \betabarLAB\ into components parallel and
@@
 \end{equation}
-%\begin{eqnarray}
-%\betabarLABpara &=& N^{-1}\int \frac{\betaCMpara +\betaBOOST}{1+\betaBOOST\betaCMpara} \fCM \dVCM,\\
-%\betabarLABperp &=& N^{-1}\int \frac{\betaCMperp}{\gammabeam\left(1+\betaBOOST\betaCMpara\right)} \fCM \dVCM,
-%\end{eqnarray}
 where $\gammabeam = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
@@
 \end{equation}
-%\begin{eqnarray}
-%  \dbetabarLABpara &=& N^{-1}\int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
-% &=& N^{-1}\int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\
-% &=&N^{-1}\int \left(\frac{\betaCMpara + \betaBOOST}{1+\betaCMpara\betaBOOST}\right)^2 \fCM \dVCM - \betabarLABpara^2,\label{eq:dbetabarLABpara}
-%\end{eqnarray}
 while the perpendicular component becomes
@@
 \end{equation}
-%\begin{eqnarray}
-% \dbetabarLABperp &=& N^{-1}\int \beta_{\rm L,\perp}^2 \fLAB \dVLAB, \nonumber\\
-%&=& N^{-1}\gammabeam^{-2} \int \frac{\betaCMperp^2}{\left(1+\betaCMpara\betaBOOST\right)^2} \fCM \dVCM.
-%\label{eq:dbetabarLABperp}
-%\end{eqnarray}
 
@@
 \end{equation}
-%\begin{eqnarray}
-% \int F_t\frac{k^2c^2 - (\kvec\cdot\vel)^2}{\gamma(\omega - \kvec\cdot\vel)^2} d^3p &\approx& k^2c^2
-% \int F_t \left(\frac {1}{\omega^2} + \frac {2\kvec\cdot\vel}{\omega^3} + \frac{3 (\kvec\cdot\vel)^2}{\omega^4}\right) d^3p \nonumber \\
-%&\approx& \frac {k^2c^2}{\omega^2}\left(1 + 3 {k^2\lambda_D^2}\right),
-%\label{eq:dispersion-real-part}
-%\end{eqnarray}
 where the second term is zero because it is odd, $\lambda_D^2 = k_B T_t/m_e \omega_p^2$ is the Debye length, and we have assumed that $k^2\lambda_D^2 \ll 1$ and $\omega \approx \omega_p$ in the last term on the RHS\footnote{A direct solution to Equation (\ref{eq:dispersion-real-part}) without approximating $\omega\approx \omega_p$ will reveal waves with nontrivial growth or damping rates.
@@
 \end{equation}
-%where we have used
-%\begin{equation}
-%\left.
-%\gammabeam(E-v_{bz} p_z -v_{bx} p_x)/T
-%\right|_{\rm pole}
-%=
-%\gammabeam \gph ( 1-v_{bz}\vph ) E_\perp/T - \gammabeam v_{bx} p_x/T\,.
-%\label{eq:Eopole}
-%\end{equation}
-
-%and therefore,
-%where $\vbeam\rightarrow v_{bz}$ and the only new term is the final
-%exponential factor which depends upon $p_x$.
 
@@
 \end{equation}
-%\begin{equation}\label{eq:I-integral}
-%\kI \equiv \int d^2\!p_\perp E_\perp \exp\left(-\frac{\cG(E_\perp-w p_x)}{k_B T_b}\right)
-%\end{equation}
 $\cG\equiv\gammabeam\gph(1-v_{b,z'}\vph/c^2)$, and
@@
 \end{equation}
-%\begin{eqnarray}
-%%E'_\perp &=& \gamma_w(E_\perp - w p_x),\\
-%E_\perp &=& \gamma_w(E'_\perp + w p'_x),\\
-%d p_x d p_y &=& \frac{E_\perp}{E'_\perp} d p'_x d p'_y.
-%\end{eqnarray}
 Inserting this into Equation (\ref{eq:Rogen I-integral}), we find
@@
 \int d^2\!p'_\perp \frac{E_\perp^2}{E'_\perp} \exp\left(-\frac{\cG'E'_\perp}{k_B T_b}\right) %\\
-%&=
-%\gamma_w^2 \int d^2\!p'_\perp \left(
- % E'_\perp + 2 w 
-%\qc{p'_x}
-%+ w^2 \frac{p'^2_x}{E'_\perp}
-%\right) e^{-\cG'E'_\perp/m}\\
-%&
 =
@@
 where $\cG'\equiv\cG/\gamma_w$. Note in the second line that we have used isotropy in $\bp_{\perp}'$ to eliminate terms linear in $\bp_{\perp}'$. 
-% and in the second line we used the fact
-%that  $E'_\perp e^{-\cG'E'_\perp/m}$ is isotropic in
-%$\bp'_\perp$.  
 Using the following integrals: 
@@
 \end{equation}
-%\begin{equation}
-%\int_0^\infty dx \frac{x}{\sqrt{1+x}} e^{-a\sqrt{1+x}}
-%=
-%\frac{4(a+1)}{a^3} e^{-a}\,,
-%\end{equation}
 we find
@@
 R =
-%\frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu m_ec^2 \cG'^3 K_2(\mu)}
 \frac{\gph^3 \gamma_w^2 (\vph-v_{b,z'})}{ 2 \mu  \cG'^3 K_2(\mu) c} 
@@
 \Gamma \approx
-%\simeq
 - \Gamma_0
-%\frac{\pi \gamma_w^2 \gph^3 m (\vph-v_{b,z'})}{4 \gammabeam \mu K_2(\mu) \cG'^3}
 \frac{\pi \gamma_w^2 \gph^3 (\vph-v_{b,z'})}{4 \gammabeam \mu^2 K_2(\mu) \cG'^3 c}
@@
 where $\Gamma_0 \equiv \omega_p \gammabeam (n_b/n_t) (m_e v_B^2/k_B T_b)$ is the typical maximum growth rate.
-%The above differs from the beam plasma instability only due to the
-%minor difference in the definition of $\cG'$ and the second term.  When
-%$v_{bx}=0$, it is precisely the beam plasma instability growth rate.
-
-%With an eye toward expanding to lowest order in $v_{bx}\ll1/\gammabeam\simeq1/\gph$, we find
-%\begin{equation}
-%\gamma_w 
-%\simeq
-%1 + \frac{v_{bx}^2}{2 \gph^2(1-v_{bz}\vph)^2}
-%=
-%1 + \frac{\gammabeam^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
-%\quad\Rightarrow\quad
-%\cG'\simeq \cG - \frac{\gammabeam^2 v_{bx}^2}{2 \cG}\frac{m^2}{T^2}.
-%\end{equation}  
-%Thus, to lowest order in $\gammabeam v_{bx}$, equation (\ref{eq:OGgen}) becomes
-%\begin{equation}
-%\begin{aligned}
-%\Gamma 
-%&\simeq 
-%- \Gamma_0
-%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gammabeam T K_2(m/T)}
-%\frac{e^{-\cG}}{\cG^3}
-%\left(
-%  1
-%  +
-%  \frac{\gammabeam^2 v_{bx}^2}{\cG^2} \frac{m^2}{T^2}
-%\right)\\
-%&\qquad\times
-%\left[
-%\left( 
-%  \cG^2 + 2\cG + 2 - \gammabeam^2 v_{bx}^2\frac{m^2}{T^2}  - \frac{\gammabeam^2 v_{bx}\qc{^2}}{\cG}\frac{m^2}{T^2}
-%\right)
-%+
-%\frac{\gammabeam^2 v_{bx}^2}{2\cG^2} \frac{m^2}{T^2} \left(2 \cG + 2\right)
-%\right]
-%\\
-%&\qquad\qquad\qquad\qquad\times 
-%\left(
-%  1 + 3 \frac{\gammabeam^2 v_{bx}^2}{2\cG^2}\frac{m^2}{T^2}
-%\right)
-%\left(
-%  1 + \frac{\gammabeam^2 v_{bx}^2}{2\cG}\frac{m^2}{T^2}
-%\right)\\
-%&\simeq 
-%- \Gamma_0
-%\frac{\pi \gph^3 m (\vph-v_{bz})}{4 \gammabeam T K_2(m/T)}
-%e^{-\cG}\\
-%&\qquad\qquad\times 
-%\left( 
-%  \frac{\cG^2 + 2\cG + 2}{\cG^3}
-%  +
-%  \frac{\gammabeam^2 v_{bx}^2}{2}\frac{m^2}{T^2}
-%  \frac{
-%    \cG^3
-%    +
-%    5\cG^2
-%    +
-%    12\cG
-%    +
-%    12
-%  }{\cG^5}
-%\right)\,,
-%\end{aligned}
-%\end{equation}
-%where $\Gamma_0\equiv \omega_{Pt} (n_b/n_t) \gammabeam m v_{bz}^2/T$, i.e.,
-%again $v_{bz}$.
-
-%However, obtaining asymptotic expansions for the growth rate in the high and
-%low temperature limits that can be easily solved for the maximum
-%growth rate is difficult.  Even with the tremendous numerical insight
-%provided by Figure \ref{fig:ObliqueGMZ}, it is unclear how to make
-%significant progress.  However, some general remarks can be made.  It
-%appears at the maximum growth rate, we have to leading order:
-%\begin{equation}
-%\begin{gathered}
-%\vph\simeq\sqrt{2\delta Z_x}\,,\quad
-%v_{bx}=(1-\delta Z_x) \vbeam\,,\quad
-%v_{bz}\simeq\sqrt{2\delta Z_x}\,\vbeam\,,\quad\\
-%v_{bz}-\vph\simeq\left\{
-%\begin{aligned}
-%&\frac{\sqrt{T/m}}{\gammabeam} & T\ll m\\
-%&\frac{1}{2\gammabeam} & T\gg m
-%\end{aligned}
-%\right.\,,\quad
-%\gph\simeq1+\delta Z_x\,,\quad
-%\cG' \sim \frac{m}{T}\,,\\
-%\text{and}\quad w \sim 1\,,
-%\end{gathered}
-%\end{equation}
-%\qc{[can you add more details which equations you approximate under which conditions? e.g., I get $1/2\gammabeam^2$ for the hot beam case \ldots]}\\
-%where $\delta Z_x\equiv 1 - v_{bx}/\vbeam$.  Inserting these into Equation
-%(\ref{eq:OGgen}) suffices to show that there exists growth rates that
-%are $\propto \Gamma_0$, with
-
-%\begin{equation}
-%\Gamma_M
-%\simeq
-%\left\{
-%\begin{aligned}
-%& 0.38 \Gamma_0 & T\ll m\\
-%& 0.34 \Gamma_0 & T\gg m
-%\end{aligned}
-%\right.\,,
-%\end{equation}
-%for $\gammabeam\gtrsim10$, as seen in Figure (\ref{fig:ObliqueGMZ}).  
-%
-
-% We now proceed to the third term.
-Without loss of generality, we orient our axes such that the beam is along the z-axis and the wavevector k sits in the x-z plane, i.e., $\kvec = k_x \xhat + k_z \zhat$.
-The beam distribution function (\ref{eq:distribution function beam}) now becomes
-% \begin{equation}
-% df=\frac{n_b}{\sqrt{\pi^3\dbetabarpara\dbetabarperp^2}}\exp\left(-\frac{\Delta \beta_{z}^2}{\dbetabarpara} - \frac{\Delta \beta_{x}^2 + \Delta\beta_{y}^2}{\dbetabarperp}\right) d^3\Delta\beta,
-% \end{equation}
-% We may integrate over $d\Delta\beta_y$ leaving an integral over $d\Delta\beta_z, d\Delta\beta_x$.  
-% \begin{equation}\label{eq:integral1}
-% \mathcal{I} = \int \frac{(k_x \partial_{p_x} + k_z\partial_{p_z}) F_1}{\omega - \kvec\cdot\vel}d\Delta\beta_z\, d\Delta\beta_x,
-% \end{equation}
-% where $F_1$ is the distribution function after has been integrated over $d\Delta\beta_y$.
-Note that we have integrated the dispersion relation (\ref{eq:dispersion relation}) rather that the one that was integrated by parts (eq.[\ref{eq:dispersion relation by parts}]) because the Landau contour that we perform below is more explicit with this version.
-Equation (\ref{eq:integral1}) can be rewritten as
-% \begin{equation}\label{eq:integral2}
-% \mathcal{I} \approx \int \frac{(\gamma^2k_x\partial_{\Delta\beta_x} + k_z\partial_{\Delta\beta_z} ) F_1 d\Delta\beta_z\, d\Delta\beta_x}{k_z\gamma^3m_e c^2(\omega/k_z c - \beta_z - k_x\Delta\beta_x/k_z - \Delta\beta_z)},
-% \end{equation}
-% where we have converted the derivatives on $p_x$ and $p_z$ to derivatives on $\Delta \beta_x, \Delta \beta_z$ by assume that they are $ \ll 1/\gamma$ and $\ll 1/\gamma^2$, and take the following expansions:
-% \begin{eqnarray}
-%  p_x &\approx& 0 + \frac{\partial p_x}{\partial \Delta\beta_x} \Delta\beta_x \approx \gamma_{0} m_e c \Delta\beta_x\\
-%  p_z &\approx& p_{z,0} + \frac{\partial p_z}{\partial \Delta\beta_z} \Delta\beta_z \approx p_{z,0} + \gamma_{0}^3 m_e c\Delta\beta_z,
-% \end{eqnarray}
-% where $\gamma_0 = \sqrt{1 + p_{z,0}^2/m_e^2c^2}$. 
-%  
-% \ac{[I had some difficulty with the above, in part because I think many approximations are being utilized at the same time.
-Thus perhaps it would be better to explain each step in a bit more dtail.
-% The thing I had the most difficulty with was the $\bmath{p}\rightarrow\vel$ transformation in the integrands.
-This is because the expansions in C6 \& C7 do not imply the transformation used in C5.
-Rather, we should have
-% \begin{equation*}
-% \frac{\partial}{\partial p^j} 
-% =
-% \frac{\partial \beta^k}{\partial p^j} \frac{\partial}{\partial \beta^k}
-% =
-% \frac{\delta^k_j - \beta_j\beta^k}{\gamma m} \frac{\partial}{\partial \beta_k}\,,
-% \end{equation*}
-% and thus,
-% \begin{equation*}
-% \begin{aligned}
-% \kvec\cdot\nabla_p F_1
-% &=
-% \frac{k^k - (k^j \beta_j)\beta^k}{\gamma m} \frac{\partial F_1}{\partial \beta_k}\\
-% &=
-% \frac{1}{\gamma m}
-% \bigg\{
-% \left[k_x - (k_x\beta_x+k_z\beta_z)\beta_x\right] \frac{\partial F_1}{\partial \beta_x}\\
-% &\qquad\qquad
-% +
-% \left[k_z - (k_x\beta_x+k_z\beta_z)\beta_z\right] \frac{\partial F_1}{\partial \beta_z}
-% \bigg\}\,.
-% \end{aligned}
-% \end{equation*}
-% 
-% When $k_x=0$, this simplifies a little bit, and we have
-% \begin{equation*}
-% \kvec\cdot\nabla_p F_1
-% =
-% \frac{1}{\gamma m}
-% \left[
-% -k_z\beta_z\beta_x\frac{\partial F_1}{\partial \beta_x}
-% +
-% k_z \left(1-\beta_z^2\right) \frac{\partial F_1}{\partial \beta_z}
-% \right]\,,
-% \end{equation*}
-% and with $\beta_x\ll\beta_z\sim1$ we have
-% \begin{equation*}
-% \kvec\cdot\nabla_p F_1
-% \simeq
-% \frac{k_z}{\gamma^3 m}
-% \left[
-% -\gamma^2 \beta_z\beta_x\frac{\partial F_1}{\partial \beta_x}
-% +
-% \frac{\partial F_1}{\partial \beta_z}
-% \right]\,,
-% \end{equation*}
-% with the important difference being the first term.  With the form of $F_1$ we have $\partial F_1/\partial \beta_j = 2\beta_j F_1 / \alpha_j$, and may evaluate each term to lowest order:
-% \begin{equation*}
-% \begin{aligned}
-% &I_{x} \equiv -\int \frac{1}{\omega-k_z \beta_z} \frac{k_z}{\gamma^3 m} \gamma^2 \beta_z \beta_x 
-% \left(\frac{2\beta_x}{\alpha_x} F_1\right)
-% d\beta_x d\beta_z\\
-% &\qquad\qquad= 
-% -\int \frac{k_z \beta_z  F_2}{\gamma m (\omega-k_z \beta_z)} d\beta_z
-% \simeq
-% \left.-\frac{i \pi F_2}{\gamma m}\right|_{\beta_z=\omega/k_z}
-% \\
-% &I_{z} \equiv \int \frac{1}{\omega-k_z \beta_z} \frac{k_z}{\gamma^3 m} 
-% \left[\frac{2(\beta_z-\beta_b)}{\alpha_z} F_1\right]
-% d\beta_x d\beta_z\\
-% &\qquad\qquad= 
-% \int \frac{2 k_z (\beta_z-\beta_b) F_2}{\gamma^3\alpha_z m (\omega-k_z \beta_z)} d\beta_z
-% \simeq
-% \left.\frac{2 i \pi (\beta_z-\beta_b) F_2}{\gamma^3 \alpha_z m}\right|_{\beta_z=\omega/k_z}\\
-% &\qquad\qquad\simeq
-% - \frac{2(\omega/k_z-\beta_b)}{\gamma^2 \alpha_z} I_x\,.
-% \end{aligned}
-% \end{equation*} 
-% Note that $I_{z}$ is just your expression C9.
-% Now, since $\alpha_z\sim T/\gamma^4$, and $F_2$ is large for $|\omega/k_z-\beta_b|\lesssim\sqrt{\alpha_z}\sim T^{1/2}/\gamma^2$, $|I_z|\sim |I_x|/T^{1/2}$ and thus the neglected term is not generally small.
-% 
-% More problematic, I would like to understand why the $k_x=0$ case doesn't reproduce the standard cold plasma instability rate when $\alpha_z$ is very small.
-% ]}
-% 
-% 
-% 
-% 
-% 
-% To make a connection with the the beam plasma instability, we take $k_x = 0$ and integrate Equation (\ref{eq:integral2}) over $d\Delta\beta_x$ to find:
-% \begin{equation}\label{eq:integral-bp}
-% \mathcal{I} \approx \frac{1}{\gamma^3 m_e c^2}\int \frac{\partial_{\Delta\beta_z} F_2}{(\omega/k c - \beta_z - \Delta\beta_z)}d\Delta\beta_z, 
-%  \end{equation}
-% where $F_2 = \sqrt{1/\pi\alpha_{\parallel}}\exp(-\alpha_{\parallel}^{-1} \Delta\beta_z^2)$ is the normalized distribution function after it has been integrated over $d\Delta\beta_x$.
-Equation (\ref{eq:integral-bp}) can be evaluated using the Landau prescription, which yields a principle value (which is small and real and hence can be neglected as the background is dominant) and an imaginary component from the contribution of the pole
-% \begin{equation}\label{eq:imaginary-bp}
-%  \Im(\mathcal{I}) = \frac{i\pi}{\gamma^3 m_e c^2}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)}
-% \end{equation}
-% 
-% Combining equations (\ref{eq:dispersion kinetic}), (\ref{eq:dispersion-real-part}) and (\ref{eq:imaginary-bp}), we find
-% \begin{equation}
-%  \omega^2 - \omega_{p,t}^2\left(1 + 3k^2\lambda_D^2\right) - \frac{i\pi\omega_{p,b}^2\omega^2}{\gamma^3 k^2c^2}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)} = 0.
-% \end{equation}
-% Expanding about $\omega = \omega_r - i\Gamma$ \cc{for $kc \approx \omega_r \approx \omega_{p,t}$ and assuming the imaginary parts is $O(\Gamma)$}, we find 
-% \begin{eqnarray}\label{eq:growth-rate-bp-appendix}
-%  \Gamma &=& \frac{\pi n_b}{2\gamma^3n_t}\left.\frac{\partial F_2}{\partial\Delta\beta_z}\right|_{\Delta\beta_z = (\omega/k_z c - \beta_z)}\omega_{p,t} \nonumber\\
-%  &=& -\frac{\sqrt{\pi} n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\parallel}}\alpha'_{\parallel}}\exp\left(-\frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}}\right),
-% \end{eqnarray}
-% where $\alpha'_{\parallel} = \gamma^4\alpha_{\parallel}$ is the rescaled factor such that it is order unity.
-% As $\Delta\beta_z \sim \sqrt{\alpha_{\parallel}}$, Equation (\ref{eq:growth-rate-bp}) has an overall scaling like $\gamma\nb/n_t$ in line scalings found for the kinetic oblique mode quoted in Paper I (also see below).
-% 
-% We now assume $k_x \neq 0$ and perform the integral over $\Delta\beta_z$ in Equation (\ref{eq:integral2}) to find for the imaginary part
-% \begin{equation}\label{eq:integral-kinetic}
-%  \Im(\mathcal{I}) = \frac{i\pi}{\gamma^3 m_ec^2}\int \left(\gamma^2\frac{k_x}{k_z}\frac{\partial F_1}{\partial\Delta\beta_x} + \frac{\partial F_1}{\partial\Delta\beta_z}\right)_{\Delta\beta_z = \Delta\beta_{\rm res}} d\Delta\beta_x,
-% \end{equation}
-% where $\Delta\beta_{\rm res} = \omega/k_z c - \beta_z - k_x\Delta\beta_x/k_z$.  
-% 
-% To perform this integral, we write the exponent in $F_1$ as 
-% \begin{eqnarray}
-%  F_1 \sim \exp\left[\left(\frac 1 {\alpha_{\perp}} + \frac {(k_x/k_z)^2} {\alpha_{\parallel}}\right)\left(\Delta\beta_x - \left(1 - \frac{k_x^2\alpha_{\parallel}}{k_z^2\alpha_{\perp}}\right)\frac{\Delta\beta_{z,r}}{k_x/k_z}\right)^2 - \frac{\Delta\beta_{z,r}^2}{\alpha_{\parallel}} \right],
-% \end{eqnarray}
-% where the normalization $1/\pi\sqrt{\alpha_{\perp}\alpha_{\parallel}}$ has been left out and $\beta_{z,r} = \omega/k_zc - \beta_z$.  
-% Making an appropriate substitution for $\Delta\beta_x \rightarrow \beta_x - \left(k_z/k_x - {k_x\alpha_{\parallel}}/{k_z\alpha_{\perp}}\right)\Delta\beta_{z,r}$, we find for Equation (\ref{eq:integral-kinetic})
-% \begin{eqnarray}
-% \Im(\mathcal{I}) = -\frac{i2}{\gamma^3 c}\sqrt{\frac{\pi k_z}{k_x\alpha_{\perp}}}k_z\left(\gamma^2\frac{1}{\alpha_{\perp}} + \frac{1}{\alpha_{\parallel}} - \frac{k_z}{k_x\alpha_{\parallel}}\right)\Delta\beta_{z,r}\exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right)
-% \end{eqnarray}
-% The resulting growth rate is then
-% \begin{eqnarray}
-% \Gamma = -\sqrt{\frac{\pi k_z}{k_x}}\frac{n_b\gamma}{n_t}\frac{\Delta\beta_{z,r}}{\sqrt{\alpha_{\perp}}}\frac{k_z}{k}\left(\frac{1}{\alpha'_{\perp}} + \frac{1}{\alpha'_{\parallel}} - \frac{k_z}{k_x\alpha'_{\parallel}}\right) \exp\left(-\frac{k_z^2\Delta\beta_{z,r}^2}{k_x^2\alpha_{\perp}}\right),\label{eq:growth-rate-oblique-appendix}
-% \end{eqnarray}
-% where $\alpha'_{\perp} = \gamma^2\alpha_{\perp}$ is an O(1) factor.
-Again, we find that the overall scaling of the kinetic oblique instability is $\sim \gamma n_b/n_t$ like the kinetic beam plasma instability above.
-Also we have made the assumption that $k_x/k_z \gg 1/\gamma$.
-For $k_x/k_z < 1/\gamma$, the dispersion relation reverts back to the beam plasma case given in Equation (\ref{eq:growth-rate-bp})
 
@@
 
-\bibliography{bigmh}
+\bibliography{ms}
 \bibliographystyle{apj}
@@
 \end{document}
-
-\ac{[The paragraph below is somewhat redundant now as the primary justifications have all been given above and in the appendix.
-Plus there is the rather confusing redefinition of $\beta$ as well as the introduction of $v$.
-I leave it here because I'm not quite sure what the point of it is, though believing that the main message is conveyed with my suggestion above.
-Perhaps this could be an after-the-fact, more rigorous motivation for the above?
-Is such a motivation necessary?
-Could it be moved efficiently into the footnote (a potential minimal footnote is already now provided)?]
-> > > > > > > > > > > > > > > > > >}
-\begin{equation}
- f\propto\exp\left(-\frac{\left(v_{\parallel}-v_{\parallel,0}\right)^2}{\dbetabarpara}-\frac{v_{\perp}^2}{\dbetabarperp}\right),
-\end{equation}
-where $v_{\parallel} = p_{\parallel}/\gamma m_e c$ is the parallel velocity, $v_{\parallel,0} = p_{\parallel,0}/\gamma_0 m_e c \approx c$ is the velocity of the beam, and $v_{\perp} = p_{\perp}/\gamma m_e c$ is the perpendicular velocity.
-To make further progress, we make the assumption that the spread in momentum $\ll p_{\parallel,0}$.
-In other words, we assume that $\alpha_{\parallel}, \alpha_{\perp} \ll 1$.
-In this limit, we can perform a Taylor expansion for $p_{\parallel}$ about $p_{\parallel,0}$ and $p_{\perp}$ about $0$.
-This gives: 
-\begin{eqnarray}
- p_{\perp} &\approx& 0 + \frac{\partial p_{\perp}}{\partial v_{\perp}} \Delta v_{\perp} \approx \gamma_{0} m_e \Delta v_{\perp},\\
- p_{\parallel} &\approx& p_{\parallel,0} + \frac{\partial p_{\parallel}}{\partial v_{\parallel}} \Delta v_{\parallel} \approx p_{\parallel,0} + \gamma_{0}^3 m_e \Delta v_{\parallel},\\
- \gamma &\approx& \gamma_0 + \gamma_0^3 \frac{\Delta v_{\parallel}}{c}
-\end{eqnarray}
-where $\gamma_0 = \sqrt{1 + p_{\parallel,0}^2/m_e^2c^2}$. Hence this gives an approximate distribution function for a beam in terms of the perturbed velocities as 
-\begin{equation}\label{eq:distribution function beam}
-df=\frac{n_b}{\sqrt{\pi^3\dbetabarpara\dbetabarperp^2}}\exp\left(-\frac{\Delta \beta_{\parallel}^2}{\dbetabarpara} - \frac{\Delta \beta_{\perp}^2}{\dbetabarperp}\right) d\Delta\beta_{\parallel} d^2\Delta\beta_{\perp},
-\end{equation}
-where $\Delta\beta = \Delta v/c$ and we have assumed a 3-D geometry so that the perpendicular direction spans two dimensions: $\Delta\beta_{\perp}^2 = \Delta\beta_{\perp,1}^2 + \Delta\beta_{\perp,2}^2$.
-This form of the distributions function (\ref{eq:distribution function beam}) has many of the properties of the Maxwell-Boltzmann distribution function, which we exploit below in our analytic calculations.
-\ac{< < < < < < < < < < < < < < <}

### 2016-09-21 apjRevised/response.txt
updated and massaged the text

@@
 2. It has all the features that allows us the elucidate the physics (continuous spread of momenta in the parallel and perpendicular directions)
-3. It is relatively not ad-hoc — by this we mean that it is a distribution that can be realized in physical situation.
+3. It is physically motivated — by this we mean that it is a distribution that can be realized in physical situation.
 
+We added note that we use the Maxwell-Juttner distribution in the abstract.
 We have added an additional paragraph to the end of section 2.1 elucidating this choice and saying that a correct calculation involves using a distribution function calculated from the \gamma\gamma process such as the ones mentioned by the referee.

### 2016-09-21 apjRevised/ms.tex
updated and massaged the text

@@
 
+\usepackage{ulem}
 \newcommand{\mathsc}[1]{\text{\textsc{#1}}}
@@
 The annihilation of TeV photons from extragalactic TeV sources and the extragalactic background light produces ultrarelativistic $\epm$ beams, which are subject to powerful plasma instabilities that sap their kinetic energy.
-Here we study the initial conditions of these \epm\ beams and the linear phase of the plasma instabilities that they drive.
-Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes.
+Here we  study {\bf \sout{the initial conditions of these e± beams and}} the linear phase of the plasma instabilities that these pair beams drive.
+Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes{\bf,  assuming a Maxwell-J{\"u}ttner distribution for the pair beam}.
 We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode.
@@
 
-{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the distribution function that are produced by VHEGR photo annihilation is different \citep[see for instance][]{2012ApJ...758..101S}.
-In particular, the the parallel and perpendicular momentum spread will likely be different due to the distribution of VHEGR photons and their respective mean free paths.
-However, a Maxwell-J{\"u}ttner distribution is still useful as it has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).
-In addition, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.
-It also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.
-Finally, we note that the analytic methodology use to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}
+{\bf While we have modeled the pair distribution function as a Maxwell-J{\"u}ttner distribution, the physical distribution function that is produced by VHEGR photo annihilation is likely different \citep[see for instance][]{2012ApJ...758..101S}.
+In particular, the parallel and perpendicular momentum spread will be influenced by the distribution of VHEGR photons and their respective mean free paths.
+However, a Maxwell-J{\"u}ttner distribution is still useful.
+First, it also is sufficiently simple to allows us to calculate the kinetic instability exactly in the electrostatic approximation.
+Second, its instability growth rates has been calculated without approximation previously by \citet{Bret-Grem-Beni:10}, allowing a point of comparison for our calculation using the electrostatic approximation (as mentioned below).
+Third, it possess a continuous (small) distribution of parallel and perpendicular momenta that allow us to elucidate the physics.
+Finally, the analytic methodology used to calculate the Maxwell-J{\"u}ttner distribution may be useful for the full calculation using the physical distribution function.}
 

### 2016-09-21 apjRevised/response.txt
small edits

@@
 We were aware of these subtleties and focus just on the linear physics in this paper avoid any quantitative comparison with the ICC timescales.
-We merely confirmed the linear results uses in earlier papers, clarified the physics, especially in regard to the broad support of modes that are unstable, and address the transition between the kinetic and reactive mode for the oblique instability.
+We merely confirmed the linear results used in earlier papers, clarified the physics, especially in regard to the broad support of modes that are unstable, and address the transition between the kinetic and reactive mode for the oblique instability.
 

### 2016-09-21 apjRevised/ms.tex
small edits

@@
 Here we  study {\bf \sout{the initial conditions of these e± beams and}} the linear phase of the plasma instabilities that these pair beams drive.
-Here we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes{\bf,  assuming a Maxwell-J{\"u}ttner distribution for the pair beam}.
+{\bf \sout{Here} To this end,} we calculate the linear growth rate of the beam plasma and oblique instability in the electrostatic approximation in both the reactive and kinetic regimes{\bf,  assuming a Maxwell-J{\"u}ttner distribution for the pair beam}.
 We reproduce the well-known reactive and kinetic growth rates for both the beam plasma and oblique mode.
-We demonstrate for the oblique instability that there is broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wavevector orientations relative to the beam.
+We demonstrate for the oblique instability that there is a broad spectrum of unstable modes that grow at the maximum rate for a wide range of beam temperatures and wave vector orientations relative to the beam.
 We also delineate the conditions for applicability for the reactive and kinetic regimes and find that the beam plasma mode transitions to the reactive regime at a lower Lorentz factor than the oblique mode due to a combination of their different scalings and the anisotropy of the velocity dispersions.
-Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to the both the kinetic oblique mode and the reactive beam-plasma mode.
+Applying these results to the ultrarelativistic $\epm$ beams from TeV blazars, we confirm that these beams are unstable to both, the kinetic oblique mode and the reactive beam-plasma mode.
 These results are important in understanding how powerful plasma instabilities may sap the energy of the ultrarelativistic $\epm$ beams as they propagate through intergalactic space.
@@
  IC halo emission from TeV blazars does not imply the existence of the IGMF as previous groups have argued
-(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}). {\bf We note that this effectiveness of this plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.}
+(BCP12;\citealt{Schlickeiser+12,Schlickeiser+13}). {\bf We note that the effectiveness of these plasma instabilities is complicated by nonlinear effects, which we briefly discuss below.}
 
